### 2015-10-20 halo/text/abc.tex
first draft

\documentclass[12pt, preprint]{aastex}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{bbm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue }

\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}

\begin{document}

\title{Likelihood free inference with measurements of galaxy clustering, group richness, and number denisty }

\begin{abstract}

The Halo Occupation Distribution model (HOD) is a powerful prescription for 
understanding how galaxies populate dark matter halos.
In this work, we use the Approximate Bayesian Computation method, 
ABC, in order to explore the parameters of the halo model with mock
measurements of galaxy number denisty, galaxy clustering, 
and group richness. These observables are forward-modeled in the ABC
by populating halo catalogs in the Bolshoi N-body dark matter simulation 
with the HOD model and making the above measurements on the generated
galaxy catalogs. We use the publicly available software, 
HALOTOOLS to perform the forward-modeling and make the mock measurements.
We find that including group richness in our set of observables leads to stronger
constraints on the HOD parameters that govern the population of satellite galaxies.
Our ABC analysis also successfully constrains HOD parameters for these same
measurements made in SDSS data and our constrained values are consistent with
previous attempts in the literature.

\end{abstract}

\section{Introduction}

Parameter estimation in cosomlogy is commonly done with Bayesian inference. 
In the Bayesian framework, constraints on cosmological parameters are achieved by 
sampling from the posterior distribution, the probability of the parameters of the 
model given the data. Posteriors are evaluated by multiplying the prior probability 
distribution of the paramters of the model with the likelihood function, the probability 
of the observed data given the parameters of the model.
Therefore, Bayesian inference relies on the evaluation of the likelihood function.

Evaluation of the likelihood requires on making assumption about the functional 
form of the likelihood. It is often assumed that the likelihood is Gaussian distribution 
with a given covariance matrix that needs to be evaluated from simulating a large set of 
mock catalogs, or other techniques such as bootstrap. In order to 
compute the likelihood funciton, one needs to invert a high dimensional covariance 
matrix which makes the likelihood evaluation and therefore sampling from the posterior 
computationally intractable.

Approximate Bayesian Computation, ABC, is a method that eliminates the need for likelihood. 
In recent years, this technique has been used in a number of cosmological studies, such as 
the inference of cosmological parameters from modelling of the distances of supernovae type Ia, 
and modelling of weak lensing peak maps. Exploring the parameter space with ABC algorithm 
requires three essential ingredients: (1) a prior pdf on the parameters that we wish to estimate, 
(2) a forward model (simulator) of the observations, and (3) a distance metric for comparing 
the model and the observed data. In problems where likelihood evaluation is not analytically or 
computationally tractable, this algorithm bypasses the need for likelihood evaluation.

To date, different variations of ABC algorithm have been introduced  and implemented 
in the literature. One of the earliest methods is likelihood free Markov Chain Monte 
Carlo. In this algorithm, an initial draw from the prior is kept, if the distance between 
the simulated model corresponding to that parameter and the observed data is less than a threshold. 
In the consequent steps, the proposed parameters are accepted if \emph{first} they satisfy the 
distance requirement, and \emph{second} the Metropolis-Hasting (MH) acceptance ratio evaluated for the 
proposal is less than one. In the MH acceptance ratio, the likelihood is replaced with some transition 
kernel.

Based on importance sampling, several modifications to this algorithm have been proposed. 
These methods are called ABC Sequential Monte Carlo (ABC-SMC), or ABC Population Monte Carlo 
(ABC-PMC). These algorithms rely on sampling a large set of parameters from the prior pdf, and assigning 
equal weights to them as an initial step. In the subsequent steps, sampling is modified by 
proposing intermediate distribution built from the previous sample of parameters and their 
corresponding weights. Parameters are updated by sampling from the intermediate distribution 
and are accepted if they pass the distance requirements. Furthermore, weights are updated according 
to the intermediate distribution.

Talk about importance sampling here ...
Sampling from the parameter space in ABC algorithm can become more efficient by 
employing adaptive importance sampling...

\section{Method}

\subsection{Approximate Bayesian Computation}

\subsubsection{basics}

Talk about generic abc ...

\subsubsection{ABC with adaptive importance sampling}

\section{Data}

\section{Results}

\begin{thebibliography}{70}

\end{thebibliography}

\end{document}

### 2015-10-20 demo/three_hod_params.py

@@ -1,373 +1,0 @@
-import numpy as np
-import abcpmc
-import matplotlib.pyplot as plt
-from interruptible_pool import InterruptiblePool
-import time
-plt.switch_backend("Agg")
-
-
-from halotools.empirical_models import Zheng07
-
-
-model = Zheng07(threshold = -21.)
-print 'Data HOD Parameters ', model.param_dict
-
-n_avg = 5
-avg_nz, avg_corr = 0., 0.
-
-for i in xrange(n_avg): 
-    
-    model.populate_mock()
-    
-    # number density
-    avg_nz += model.mock.number_density
-    
-    # 6th element of xi(r) array
-    r, xi_r = model.mock.compute_galaxy_clustering(N_threads=1)
-
-    try:
-    	avg_xi += xi_r
-    except NameError:
-	avg_xi = xi_r
-
-    avg_corr += xi_r[6]
-
-avg_nz /= np.float(n_avg)
-avg_corr /= np.float(n_avg)
-avg_xi /=np.float(n_avg)
-
-data = [avg_nz, avg_xi]
-data_hod = np.array([1.15 , 0.39, 12.79])
-
-"""simulator"""
-
-class HODsim(object): 
-    
-    def __init__(self): 
-        self.model = Zheng07(threshold = -21.)
-    
-    def sum_stat(self, theta_star):
-
-	#print theta_star
-        
-        self.model.param_dict['alpha'] = theta_star[0]
-        self.model.param_dict['logMmin'] = theta_star[2]     
-        self.model.param_dict['sigma_logM'] = theta_star[1]     
-
-	#print self.model.param_dict
-        self.model.populate_mock()
-
-        nz = self.model.mock.number_density
-	        
-        r, xi_r = self.model.mock.compute_galaxy_clustering(N_threads = 1)
-
-        #xi = xi_r[6]
-        #xi = xi_r[7]
-        
-        return [nz, xi_r]
-
-ourmodel = HODsim()
-simz = ourmodel.sum_stat
-
-"""distance"""
-
-def distance(d_data, d_model, type = 'sum_stat_L2'): 
-    
-    if type == 'sum_stat_L1': 
-        dist_nz = np.abs(d_data[0] - d_model[0])/d_data[0]
-        dist_xi = np.sum(np.abs(d_data[1] - d_model[1])/d_data[1])
-        
-        dist = dist_nz + dist_xi 
-    elif type == 'sum_stat_L2':
[…]

### 2015-10-20 halo/text/abc.tex
edited abstract

@@
 previous attempts in the literature.
+
+The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies populate 
+dark matter halos that has had much success in reproducing large scale galaxy clustering and understanding
+galaxy evolution within the context of the halo model. Constraints on the parameters of the HOD model have 
+so far been imposed using the standard practice of Bayesian inference with likelihood estimation that 
+assumes a functional (often times Gaussian) form of the likelihood. Likelihood free inference methods such 
+as the Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever having to calculate the likelihood.
+
+In this work, we present ABC in the context of constraining the parameters of the HOD model. For a mock 
+galaxy catalog with known HOD parameters, we use ABC on measurements of a set of observables generated
+through HALOTOOLs, a publicly available ..., to infer tight constraints that reproduce the known HOD paramters. 
+Since ABC does not require computing the likelihood, we include observables beyond galaxy clustering such as 
+the galaxy number density and the galaxy group richness, for which accurately constructing a likelihood is intractable. 
+We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters, 
+specifically the ones that govern the satellite galaxy population. Moreover, the constraints we infer that 
+reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.
+
+We further extend our ABC method with the above set of observables to constrain HOD parameters for the Sloan 
+Digital Sky Survey Data Release 7 observed galaxy catalog. The constraints we find on the HOD parameters are consistent with 
+(or stronger than) previous constraints in the literature. (Ultimately include more interesting results).
 

### 2015-10-20 halo/code/richness_hod.py
preliminary hod with richness and multiple distance threshold implemented

@@ -1,0 +1,408 @@
+import numpy as np
+import abcpmc
+import matplotlib.pyplot as plt
+from interruptible_pool import InterruptiblePool
+import time
+plt.switch_backend("Agg")
+from halotools.empirical_models import Zheng07
+
+from astropy.table import Table
+model = Zheng07(threshold = -21.)
+print 'Data HOD Parameters ', model.param_dict
+
+
+
+def richness(group_id): 
+    gals = Table() 
+    gals['groupid'] = group_id
+    gals['dummy'] = 1
+    grouped_table = gals.group_by('groupid')
+    grp_richness = grouped_table['dummy'].groups.aggregate(np.sum)
+    return grp_richness
+
+
+
+
+n_mocks = 1000
+n_bins = 12
+hist_bins = np.rint(3 * np.logspace(0., 1.2, n_bins+1 ))
+hist_bins[-1] = 10000
+print hist_bins
+
+#histograms = np.zeros((n_mocks , n_bins))
+
+#avg_nz = []
+
+#for i in xrange(n_mocks): 
+    
+#    model.populate_mock()
+    
+    # number density
+#    avg_nz.append(model.mock.number_density)
+    
+#     # richness histogram
+#     group_id = model.mock.compute_fof_group_ids()
+#     group_richness = richness(group_id)
+# 
+#     #print 'Group Richness computation takes ', time.time() - hod_time, ' seconds'
+#     hist, bin_edge = np.histogram(group_richness, bins=hist_bins)
+# 
+#     #bin_mid = 0.5 * (bin_edge[1:] + bin_edge[:-1])
+#    
+#     histograms[i,:] = hist
+
+# np.savetxt("group_rich.dat", histograms)
+
+#np.savetxt("nz.dat", avg_nz)
+
+"""data and covariance """
+
+nz = np.loadtxt("nz.dat")
+histograms = np.loadtxt("group_rich.dat")
+
+covar_nz = np.cov(nz)
+covar_gr = np.cov(histograms)
+
+snr_gr   = 1./np.diag(covar_gr)
+
+avg_nz = np.mean(nz)
+avg_gr  = np.mean(histograms.T , axis = 0)
+
+data = [avg_nz, avg_gr]
+
+print "data=" , data
+
+data_hod = np.array([1.15 , 0.39, 12.79])
+
+"""simulator"""
+
+class HODsim(object): 
+    
+    def __init__(self): 
[…]

### 2015-10-20 halo/code/richness_hod.py
fixed weighted covariance matrix for ABC in richness HOD

@@ -63,6 +63,9 @@ histograms = np.loadtxt("group_rich.dat")
 covar_nz = np.cov(nz)
 covar_gr = np.cov(histograms)
 
+#print covar_nz
+#print np.diag(covar_gr)
+
 snr_gr   = 1./np.diag(covar_gr)
 
 avg_nz = np.mean(nz)
@@ -70,7 +73,7 @@ avg_gr  = np.mean(histograms.T , axis = 0)
 
 data = [avg_nz, avg_gr]
 
-print "data=" , data
+#print "data=" , data
 
 data_hod = np.array([1.15 , 0.39, 12.79])
 
@@ -133,9 +136,11 @@ def covariance(theta , w , type = 'weighted'):
       return np.cov(theta)
 
     if type == 'weighted':
+      ww = w.sum() / (w.sum()**2 - (w**2).sum()) 
       mean = np.sum(theta*w[None,:] , axis = 1)/ np.sum(w)
       tmm  = theta - mean.reshape(theta.shape[0] , 1)
-      sigma2 = 1./(w.sum()) * (tmm*w[None,:]).dot(tmm.T)
+      sigma2 = ww * (tmm*w[None,:]).dot(tmm.T)
+      
       return sigma2  
 
 
@@ -174,12 +179,6 @@ class Prior(object):
         return priorz
 
 
-#prior_dict = { 
-#    'alpha': {'shape': 'uniform', 'min': 0.8, 'max': 1.2},
-#    'm_1'  : {'shape': 'uniform', 'min': 13., 'max': 14.},   
-#    'm_min': {'shape': 'uniform', 'min': 11.5, 'max': 12.5}
-#}
-
 prior_dict = { 
     'alpha': {'shape': 'uniform', 'min': 1.05,  'max': 1.25},
     'm_min'  : {'shape': 'uniform', 'min': 12.5,  'max': 13.},   
@@ -215,8 +214,8 @@ import seaborn as seabreeze
 plot_range = []
 for key in ['alpha', 'sigma', 'm_min']: 
 	plot_range.append([prior_dict[key]['min'], prior_dict[key]['max']])
-print plot_range
-#[(0.7, 1.6), (11.0 , 13.0), (0.1, 0.5) , (11. , 13.0) , (12.5 , 14.)], 
+
+print "prior range is = " plot_range
 
 def plot_thetas(theta , w , t): 
     fig = corner.corner(
@@ -233,14 +232,12 @@ def plot_thetas(theta , w , t):
     
[…]

### 2015-10-20 ccppabc/pmc_abc.py
fixing minor bugs here and there

@@ -1,5 +1,5 @@
 """
-PMC-ABC code
+ABC code
 
 you can turn off
 multiprocessing by setting the parameter N_threads to 1 and run the code
@@ -31,7 +31,8 @@ class _Initialpoolsampling(object):
         
         rho = 1.e37
         prior_obj = Prior(self.prior_dict)
-        while rho > self.eps_0:
+        while np.all(rho > self.eps_0):
+
             theta_star = prior_obj.sampler()
             model_theta = self.model(theta_star)
             rho = self.distance(self.data, model_theta)
@@ -40,9 +41,10 @@ class _Initialpoolsampling(object):
         for i_param in xrange(self.n_params):
              pool_list.append(theta_star[i_param])
         pool_list.append(1.)
-        pool_list.append(rho)
+        for r in rho:
+            pool_list.append(r)
         
-        return pool_list
+        return np.array(pool_list)
 
 
 
@@ -64,7 +66,7 @@ class _Importancepoolsampling(object):
 
         rho = 1.e37
         prior_obj = Prior(self.prior_dict)
-        while rho > self.eps_t:
+        while np.all(rho > self.eps_t):
            theta_star = utils.weighted_sampling(self.theta_t, self.w_t)
            np.random.seed()
            theta_starstar = np.random.multivariate_normal(theta_star, self.sig_t, 1)[0]
@@ -82,10 +84,11 @@ class _Importancepoolsampling(object):
         theta_starstar = np.atleast_1d(theta_starstar)
         for i_param in xrange(self.n_params):
            pool_list.append(theta_starstar[i_param])
-           pool_list.append(w_starstar)
-           pool_list.append(rho)
+        pool_list.append(w_starstar)
+        for r in rho:
+            pool_list.append(r)
 
-        return pool_list
+        return np.array(pool_list)
 
 class ABC(object):
 
@@ -104,7 +107,7 @@ class ABC(object):
         self.model = model
         self.distance = distance
         self.prior_dict = prior_dict
-        self.eps0 = eps0
+        self.eps0 = np.array(eps0)
         self.N_threads = N_threads
         self.N_particles = N_particles
         self.T = T
@@ -139,8 +142,8 @@ class ABC(object):
 
         results = np.array(results).T
         self.theta_t = results[1:self.n_params+1,:]
-        self.w_t = results[self.n_params+1,:]
-        self.rhos = results[self.n_params+2,:]
[…]

### 2015-10-20 ccppabc/utils.py

@@ -1,5 +1,5 @@
 import numpy as np
-
+from scipy.linalg import cho_factor, cho_solve
 
 def weighted_sampling(theta, w):
 
@@ -28,7 +28,7 @@ def better_multinorm(theta_stst, theta_before, cov):
     return multinorm
 
 
-def covariance(theta , w , type = 'neutral'):
+def covariance(theta , w , type = 'weighted'):
  
     if type == 'neutral':
       if len(theta) == 1:
@@ -37,9 +37,9 @@ def covariance(theta , w , type = 'neutral'):
         covar =  np.cov(theta)
 
     if type == 'weighted':
+      ww = w.sum() / (w.sum()**2. - (w**2.).sum())
       mean = np.sum(theta*w[None,:] , axis = 1)/ np.sum(w)
       tmm  = theta - mean.reshape(theta.shape[0] , 1)
-      sigma2 = 1./(w.sum()) * (tmm*w[None,:]).dot(tmm.T)
-      covar = sigma2
-    print covar
+      covar = ww * (tmm*w[None,:]).dot(tmm.T)
+  
     return covar

### 2015-10-20 scripts/1D_Gaussian_with_unknown_mean/test_normal.py

@@ -32,7 +32,7 @@ class ABC_class(object):
         return np.random.normal(post_mean , post_stdev , size=size)
 
     def dist(self,x,y):
-        return np.abs(np.mean(x) - np.mean(y)) + np.abs(np.std(x) - np.std(y))
+        return np.array([np.abs((np.mean(x) - np.mean(y))/np.mean(x)) , np.abs((np.std(x) - np.std(y))/np.std(x))])
     
 
 n = 500
@@ -56,7 +56,7 @@ prior_dict = {
              }
 
 
-test_abc = ABC(test_data, test_simulator, test_dist, prior_dict , eps0 = .2 , T = 20)
+test_abc = ABC(test_data, test_simulator, test_dist, prior_dict , eps0 = [10 ,10] , T = 1)
 
 def run_serial():
 

### 2015-12-18 halo/text/abc.tex
words

--- 2015-10-20 halo/text/abc.tex
@@
 \usepackage{amsmath}	% for \iint
-\usepackage{bbm}
+\usepackage{bm}
 \usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
@@
 \newcommand{\eeq}{\end{equation}}
+\newcommand{\la}{\langle}
+\newcommand{\ra}{\rangle}
+\newcommand{\vep}{\bm{\epsilon}}
+\newcommand{\ep}{\epsilon}
 
@@
 
-\title{Likelihood free inference with measurements of galaxy clustering, group richness, and number denisty }
+\title{Likelihood free inference of the halo model parameters with simulations of galaxy mock catalogs}
 
@@
 
-The Halo Occupation Distribution model (HOD) is a powerful prescription for 
-understanding how galaxies populate dark matter halos.
-In this work, we use the Approximate Bayesian Computation method, 
-ABC, in order to explore the parameters of the halo model with mock
-measurements of galaxy number denisty, galaxy clustering, 
-and group richness. These observables are forward-modeled in the ABC
-by populating halo catalogs in the Bolshoi N-body dark matter simulation 
-with the HOD model and making the above measurements on the generated
-galaxy catalogs. We use the publicly available software, 
-HALOTOOLS to perform the forward-modeling and make the mock measurements.
-We find that including group richness in our set of observables leads to stronger
-constraints on the HOD parameters that govern the population of satellite galaxies.
-Our ABC analysis also successfully constrains HOD parameters for these same
-measurements made in SDSS data and our constrained values are consistent with
-previous attempts in the literature.
+The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies populate 
+dark matter halos that has had much success in reproducing large scale galaxy clustering and understanding
+galaxy evolution within the context of the halo model. Constraints on the parameters of the HOD model have 
+so far been infered using a likelihood function that assumes, \emph{First} a functional (often times Gaussian) form of the likelihood, 
+\emph{Second} analytical model for the galaxy clustering statistics. We use Approximate Bayesian Computation (ABC) 
+in order to infer the HOD parameters by assuming a prior over the parameters and a distance function between 
+statistical summaries of the data and statistical summaries of the simulations.
+Using a mock galaxy catalog created with known HOD parameters, we obtain precise constraints on the HOD parameters that 
+reproduces the summary statistics of the mock galaxy catalog accurately. 
+Since ABC does not require computing the likelihood, we include observables beyond galaxy clustering such as 
+the galaxy number density and the galaxy group richness, for which accurately constructing a likelihood is intractable. 
+We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters, 
+specifically the ones that govern the satellite galaxy population. Moreover, the constraints we infer that 
+reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.
 
@@
 
-Parameter estimation in cosomlogy is commonly done with Bayesian inference. 
-In the Bayesian framework, constraints on cosmological parameters are achieved by 
-sampling from the posterior distribution, the probability of the parameters of the 
-model given the data. Posteriors are evaluated by multiplying the prior probability 
-distribution of the paramters of the model with the likelihood function, the probability 
-of the observed data given the parameters of the model.
-Therefore, Bayesian inference relies on the evaluation of the likelihood function.
-
-Evaluation of the likelihood requires on making assumption about the functional 
 form of the likelihood. It is often assumed that the likelihood is Gaussian distribution 
@@
 
-\subsubsection{basics}
+ABC is a likelihood free method for inference of a set of parameters that are required for the simulation of a 
+physical system, through a forward model, a distance function, and a prior probability distribution over the parameters. 
+Let us denote the the data by $\mathcal{D}$. We assume that there exist a forward model $\mathcal{M}$ that 
+produces a simulation of the data $\bar{\mathcal{D}}$ for a given set of parameters $\bm{\theta}$
+\beq
+\bar{\mathcal{D}}_{\theta} = \mathcal{M}(\bm{\theta})
+\eeq
 
-Talk about generic abc ...
+Furthermore, ABC requires adopting a distance function between the data $\mathcal{D}$ and the simulation $\bar{\mathcal{D}}_{\theta}$.
+The distance function, in general, can be a vector $\bm{d}(\mathcal{D},\bar{\mathcal{D}}_{\theta})$
+\begin{eqnarray}
+\bm{d}(\mathcal{D},\bar{\mathcal{D}}_{\theta}) &=& [d_{1,\theta}, ... , d_{M,\theta}], \\
+d_{i,\theta} &=& d_{i}(\mathcal{S}_{i}(\mathcal{D}), \mathcal{S}_{i}(\bar{\mathcal{D}}_{\theta}))  ~~\mbox{for $i=\{1,...,M\}$},
+\end{eqnarray}
+where $d_{i}(\dot,\dot)$ is a scalar function that defines a distance 
+between $i$-$th$ summary statistics of the data $\mathcal{S}_{i}(\mathcal{D})$ and the same summary statistics of 
+the simulation $\mathcal{S}_{i}(\bar{\mathcal{D}}_{\theta})$, and $M$ is the total number of statistical summaries 
+used in the analysis.
 
-\subsubsection{ABC with adaptive importance sampling}
+ABC algorithm draws a large number of samples from the prior distribution over parameter space, 
+and rejects every sample for which the distance between the data and the simulation corresponding to that 
+sample is not less than a certain threshold $\bm{\epsilon}$. This requirement is satisfied \emph{if and only if}
+every component of the distance vector is less than its corresponding threshold. Note that, each summary statistics 
+has a separate threshold. This procedure is deone iteratively untill, an investigator is left with a certain number 
+of samples that satisfy the distance criteria. The final ensemble approximates the posterior pdf over 
+the paramters of the forward model of the data. This sampling method is called \emph{rejected sampling}.
 
-\section{Data}
+In practice however, rejected sampling is not desirable because $(i)$ it requires fine tuning of the threshold 
+$\bm{\epsilon}$ whose elements are small, and $(ii)$ finding a large number of samples that pass the threshold 
+criteria is computationally expensive. Alternatively, we draw a large number of proposals from the prior pdf over 
+the HOD parameters with the only requirement that they pass the distance criterion. But since we select an 
+arbitrarily large initial threhold, all th proposals pass the test. Moreover, we assign simillar weights to all 
+the proposals in the initial pool. In the next steps, we draw new proposals from the previos pool through 
+weighted sampling, and we perturb the particles by a kernel that is twice the covariance matrix of the proposals 
+in the previous pool.
+
+blaalalalal
+
+\subsection{Halo Occupation Distribution}
+
+Halo Occupation distribution is a powerful tool for description of galaxy/halo relation.
+
+\section{Summary statistics}
+
+\section{Forward modeling of the galaxy population with HALOTOOLS}
+
+\section{Mock data}
+
+We use HALOTOOLS to populate halos with galaxies above a certain luminosity threshold $M_r>-20$.
 
 \section{Results}
+
+Hello World!
 

### 2015-12-19 halo/code/alternative_clustering_nbar_Mr20_hod.py
new gmf

@@ -22,7 +22,7 @@ print 'Data HOD Parameters ', model.param_dict
 """data and covariance"""
 #We shouldn't treat a mean of mocks as a data!
 
-#mock_nbar = np.loadtxt("nbar_Mr20.dat")
+mock_nbar = np.loadtxt("nbar_Mr20.dat")
 #data_nbar = np.mean(mock_nbar)
 #mocks_xir = np.loadtxt("xir.dat")
 #data_xir = np.mean(mocks_xir , axis = 0)
@@ -39,19 +39,19 @@ covar_nbar = np.var(mock_nbar)
 
 """list of true parameters"""
 
-data_hod = np.array([11.92 , 0.39 , 12.79 , 1.15 , 13.94])
+data_hod = np.array([11.38 , 0.26 , 12.02 , 1.06 , 13.31])
 
 
 """Prior"""
 
-prior = abcpmc.TophatPrior([9.,.1,12.5,.9,13.6],[15.,1.,13.09,1.45,14.25])
+prior = abcpmc.TophatPrior([10.,.1,11.02,.8,13.],[13.,.5,13.02,1.3,14.])
 prior_dict = {
  
-    'logM0'  : {'shape': 'uniform', 'min': 9.  ,  'max': 15.},
-    'sigma_logM': {'shape': 'uniform', 'min': 0. ,  'max': 1.},
-    'logMmin': {'shape': 'uniform', 'min': 12.5,  'max': 13.09},   
-    'alpha': {'shape': 'uniform', 'min': .9 ,  'max': 1.45},
-    'logM1'  : {'shape': 'uniform', 'min': 13.6  ,  'max': 14.25},
+    'logM0'  : {'shape': 'uniform', 'min': 10.  ,  'max': 13.},
+    'sigma_logM': {'shape': 'uniform', 'min': .1 ,  'max': .5},
+    'logMmin': {'shape': 'uniform', 'min': 11.02,  'max': 13.02},   
+    'alpha': {'shape': 'uniform', 'min': .8 ,  'max': 1.3},
+    'logM1'  : {'shape': 'uniform', 'min': 13.  ,  'max': 14.},
 }
 
 """Plot range"""
@@ -66,7 +66,7 @@ prior_range = np.array(plot_range)
 class HODsim(object): 
     
     def __init__(self): 
-        self.model = Zheng07(threshold = -21.)
+        self.model = Zheng07()
     
     def sum_stat(self, theta):
         
[…]

### 2015-12-19 halo/code/alternative_richness_hod.py

@@ -27,9 +27,15 @@ def richness(group_id):
 
 
 """Load data and variance"""
-data = np.loadtxt("gmf_Mr20.dat")
+data_gmf = np.loadtxt("gmf_Mr20.dat")
 sigma = np.loadtxt("gmf_noise_Mr20.dat")
 bins = np.loadtxt("gmf_bins_Mr20.dat")
+
+mock_nbar = np.loadtxt("nbar_Mr20.dat")
+data_nbar = np.mean(mock_nbar)
+covar_nz = np.var(mock_nbar)
+
+data = [data_nbar , data_gmf]
 
 """True HOD parameters"""
 
@@ -75,21 +81,23 @@ class HODsim(object):
         if np.all((prior_range[:,0] < theta)&(theta < prior_range[:,1])):
           try:
             self.model.populate_mock()
+	    nbar = self.model.mock.number_density
             group_id =self. model.mock.compute_fof_group_ids()
             group_richness = richness(group_id)
             y = plt.hist(group_richness , bins)[0] / 250.**3.
-            return y
+            plt.close()
+            return [nbar , y]       
           except ValueError:
-            return np.ones_like(bins)[:-1]*1000.
+            return [10. , np.ones_like(bins)[:-1]*1000.]
         else:
-            return np.ones_like(bins)[:-1]*1000.
+            return [10. , np.ones_like(bins)[:-1]*1000.]
 
 ourmodel = HODsim()
 simz = ourmodel.sum_stat
 
-"""distance"""
 
-def distance(data, model, type = 'chisq distance'): 
+
+def distance(data, model, type = 'multiple distance'): 
     
     if type == 'added distance': 
         dist_nz = np.abs(d_data[0] - d_model[0])/d_data[0]
@@ -97,17 +105,17 @@ def distance(data, model, type = 'chisq distance'):
         
         dist = dist_nz + dist_xi 
 
-    elif type == 'separate distance':
+    elif type == 'multiple distance':
         
-        dist_nz = (d_data[0] - d_model[0])**2. / covar_nz
-        dist_xi = np.sum((d_data[1] - d_model[1])**2. * snr_gr)
+        dist_nz = (data[0] - model[0])**2. / covar_nz
+        dist_gr = np.sum((data[1] - model[1])**2. / sigma**2.)
         
[…]

### 2015-12-19 halo/code/wp_covariance.py

@@ -1,0 +1,52 @@
+from halotools.empirical_models import Zheng07 , model_defaults
+from halotools.mock_observables import wp
+from halotools.mock_observables.clustering import tpcf
+from halotools.empirical_models.mock_helpers import (three_dim_pos_bundle,
+                                                     infer_mask_from_kwargs)
+from halotools.mock_observables.clustering import wp
+from halotools.sim_manager import supported_sims
+import matplotlib.pyplot as plt
+plt.switch_backend("Agg")
+import time
+import numpy as np
+model = Zheng07()
+
+xir = []
+for i in range(500):
+ model.populate_mock()
+ xir.append(model.mock.compute_galaxy_clustering()[1])
+
+covar = np.cov(np.array(xir).T)
+np.savetxt("clustering_covariance_Mr20.dat" , covar)
+
+"""
+a = time.time()
+model.mock.compute_galaxy_clustering()
+print time.time()  - a
+rbins = model_defaults.default_rbins
+rbin_centers  = (rbins[1:] + rbins[:-1])/2.
+cat = supported_sims.HaloCatalog()
+l = cat.Lbox
+print l
+p_bins = np.linspace(0,l/2,200)
+mask = infer_mask_from_kwargs(model.mock.galaxy_table)
+
+
+pos = three_dim_pos_bundle(table=model.mock.galaxy_table,
+                               key1='x', key2='y', key3='z', mask=mask,
+                               return_complement=False)
+figure = plt.figure(figsize=(10,10))
+
+cl = wp(pos , rbins, p_bins , period = l , estimator = 'Landy-Szalay')
+
+for n_pbins in np.array([2,8,16]):
+
+  p_bins = np.linspace(0 , l/2 , n_pbins)
+  a = time.time()
+  clustering = wp(pos, rbins, p_bins , period = l , estimator = 'Landy-Szalay')
+  print time.time() - a
+  plt.plot(rbin_centers , (clustering)/cl , label = "$N\pi_{bin}$="+str(n_pbins) , lw = 2)
+  plt.xscale("Log")
+  plt.yscale("Log")
+  plt.legend() 
+plt.savefig("/home/mj/public_html/wpex.png")"""

### 2015-12-20 halo/text/abc.tex
polished up the abstract

@@
 
-The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies populate 
+The Halo Occupation Distribution model is a powerful prescription of how galaxies populate 
 dark matter halos that has had much success in reproducing large scale galaxy clustering and understanding
 galaxy evolution within the context of the halo model. Constraints on the parameters of the HOD model have 
-so far been infered using a likelihood function that assumes, \emph{First} a functional (often times Gaussian) form of the likelihood, 
-\emph{Second} analytical model for the galaxy clustering statistics. We use Approximate Bayesian Computation (ABC) 
+so far been infered using a likelihood function that assumes, first a functional---often times Gaussian---form of the likelihood, 
+and Second, an analytical model for the galaxy clustering statistics. We use Approximate Bayesian Computation 
 in order to infer the HOD parameters by assuming a prior over the parameters and a distance function between 
@@
 Since ABC does not require computing the likelihood, we include observables beyond galaxy clustering such as 
-the galaxy number density and the galaxy group richness, for which accurately constructing a likelihood is intractable. 
+the galaxy number density and the galaxy group multiplicity function. 
 We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters, 
-specifically the ones that govern the satellite galaxy population. Moreover, the constraints we infer that 
-reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.
+specifically the ones that govern the population of the satelite galaxies.
 
@@
 \section{Introduction}
-
-form of the likelihood. It is often assumed that the likelihood is Gaussian distribution 
-with a given covariance matrix that needs to be evaluated from simulating a large set of 
-mock catalogs, or other techniques such as bootstrap. In order to 
-compute the likelihood funciton, one needs to invert a high dimensional covariance 
-matrix which makes the likelihood evaluation and therefore sampling from the posterior 
-computationally intractable.
-
-Approximate Bayesian Computation, ABC, is a method that eliminates the need for likelihood. 
-In recent years, this technique has been used in a number of cosmological studies, such as 
-the inference of cosmological parameters from modelling of the distances of supernovae type Ia, 
-and modelling of weak lensing peak maps. Exploring the parameter space with ABC algorithm 
-requires three essential ingredients: (1) a prior pdf on the parameters that we wish to estimate, 
-(2) a forward model (simulator) of the observations, and (3) a distance metric for comparing 
-the model and the observed data. In problems where likelihood evaluation is not analytically or 
-computationally tractable, this algorithm bypasses the need for likelihood evaluation.
-
-To date, different variations of ABC algorithm have been introduced  and implemented 
-in the literature. One of the earliest methods is likelihood free Markov Chain Monte 
-Carlo. In this algorithm, an initial draw from the prior is kept, if the distance between 
-the simulated model corresponding to that parameter and the observed data is less than a threshold. 
-In the consequent steps, the proposed parameters are accepted if \emph{first} they satisfy the 
-distance requirement, and \emph{second} the Metropolis-Hasting (MH) acceptance ratio evaluated for the 
-proposal is less than one. In the MH acceptance ratio, the likelihood is replaced with some transition 
-kernel.
-
-Based on importance sampling, several modifications to this algorithm have been proposed. 
-These methods are called ABC Sequential Monte Carlo (ABC-SMC), or ABC Population Monte Carlo 
-(ABC-PMC). These algorithms rely on sampling a large set of parameters from the prior pdf, and assigning 
-equal weights to them as an initial step. In the subsequent steps, sampling is modified by 
-proposing intermediate distribution built from the previous sample of parameters and their 
-corresponding weights. Parameters are updated by sampling from the intermediate distribution 
-and are accepted if they pass the distance requirements. Furthermore, weights are updated according 
-to the intermediate distribution.
-
-Talk about importance sampling here ...
-Sampling from the parameter space in ABC algorithm can become more efficient by 
-employing adaptive importance sampling...
 
@@
 Let us denote the the data by $\mathcal{D}$. We assume that there exist a forward model $\mathcal{M}$ that 
-produces a simulation of the data $\bar{\mathcal{D}}$ for a given set of parameters $\bm{\theta}$
+simulates the data $\bar{\mathcal{D}}$ for a given set of parameters $\bm{\theta}$
 \beq
@@
 
-ABC algorithm draws a large number of samples from the prior distribution over parameter space, 
-and rejects every sample for which the distance between the data and the simulation corresponding to that 
-sample is not less than a certain threshold $\bm{\epsilon}$. This requirement is satisfied \emph{if and only if}
+We draw a large number of proposals from the prior distribution over the parameter space, 
+and reject the samples for which the distance between the data and the simulation corresponding to that 
+sample is more than a certain threshold $\bm{\epsilon}$. This requirement is satisfied \emph{if and only if}
 every component of the distance vector is less than its corresponding threshold. Note that, each summary statistics 
-has a separate threshold. This procedure is deone iteratively untill, an investigator is left with a certain number 
+has a separate threshold. This procedure is done iteratively untill, we are left with a certain number
 of samples that satisfy the distance criteria. The final ensemble approximates the posterior pdf over 
@@
 
-In practice however, rejected sampling is not desirable because $(i)$ it requires fine tuning of the threshold 
-$\bm{\epsilon}$ whose elements are small, and $(ii)$ finding a large number of samples that pass the threshold 
-criteria is computationally expensive. Alternatively, we draw a large number of proposals from the prior pdf over 
-the HOD parameters with the only requirement that they pass the distance criterion. But since we select an 
-arbitrarily large initial threhold, all th proposals pass the test. Moreover, we assign simillar weights to all 
-the proposals in the initial pool. In the next steps, we draw new proposals from the previos pool through 
-weighted sampling, and we perturb the particles by a kernel that is twice the covariance matrix of the proposals 
-in the previous pool.
+Recected sampling is expensive. Instead we use population monte carlo, adaptive importance sampling, ....
 
-blaalalalal
-
-\subsection{Halo Occupation Distribution}
+\subsection{Halo Model}
 
@@
 
-\section{Summary statistics}
+...
+
+\subsection{Summary statistics}
+
+Talk about 2pcf, group multiplicity function, and number density ...
 
@@
 
+Hello World!
+Talk about why it is better to simulate the mock and compute the summary statistics of the mock instead of 
+using analytical formulas...
+
+Talk about all the assumptions that go into analytic formulas...
+
+Talk about why we do not want to write down likelihood function for gmf, 2pcf, ...
+
+...
+
 \section{Mock data}
 
-We use HALOTOOLS to populate halos with galaxies above a certain luminosity threshold $M_r>-20$.
+Hello World!
+
+Plots showing 2pcf, gmf, ... of the mock data.
+How we compute the covariance matrix from mocks...
+Talk about error bars. Talk about poisson errors in group populations ...
 
@@
 
-Hello World!
+Show the posteriors...
+Show evolution of the posterior errorbars with iteration. Show epsilon. Convice that we are converging ...
+Draw samples from the posterior, populate the mock catalogs with those samples, and compute the summary statistcs with them 
+and demonstrate that they match the data...
+
+\section{Discussion}
+
+talk about cosmological infeence and why this is good first step for likelihood free inference in large scale ...
+
+talk about why it is hard to do cosmology ...
+
+talk about halo model ...
+
+talk about HOD and why it is too simple to be the right model ...
+
+talk about assembly bias perhaps ... and how we can use the same machinery for detection of assemblt bias ...
+
+talk about writing down a likelihood for gmf, tpcf, ... , and the gaussianity assumption ....
+
+talk about are results ...
+
+talk about future directions ...
 

### 2016-01-05 halo/code/alternative_richness_hod.py
gmf hod inference with new priors

@@ -7,13 +7,11 @@ from astropy.table import Table
 import corner
 from scipy.stats import norm , gamma 
 from scipy.stats import multivariate_normal
-from scipy.spatial import cKDTree
 import abcpmc
 import seaborn as sns
 sns.set_style("white")
 np.random.seed()
 from abcpmc import mpi_util
-from scipy.stats import ks_2samp
 
 
 def richness(group_id): 
@@ -28,8 +26,8 @@ def richness(group_id):
 
 """Load data and variance"""
 data_gmf = np.loadtxt("gmf_Mr20.dat")
-sigma = np.loadtxt("gmf_noise_Mr20.dat")
-bins = np.loadtxt("gmf_bins_Mr20.dat")
+sigma = np.loadtxt("gmf_noise_Mr20_2.dat")
+bins = np.loadtxt("gmf_bins_Mr20_2.dat")
 
 mock_nbar = np.loadtxt("nbar_Mr20.dat")
 data_nbar = np.mean(mock_nbar)
@@ -39,18 +37,18 @@ data = [data_nbar , data_gmf]
 
 """True HOD parameters"""
 
-data_hod = np.array([11.38 , 0.26 , 12.02 , 1.06 , 13.31])
+data_hod = np.array([11.38 , np.log(0.26) , 12.02 , 1.06 , 13.31])
 
 """Prior"""
 
-prior = abcpmc.TophatPrior([10.,.1,11.02,.8,13.],[13.,.5,13.02,1.3,14.])
+prior = abcpmc.TophatPrior([10.,np.log(.1),11.02,.8,13.],[13.,np.log(.7),13.02,1.3,14.])
 
 #prior = abcpmc.TophatPrior([11.91,.38,12.78,1.1,13.9],[11.92,.4,12.8,1.2,14.])
 
 prior_dict = {
  
     'logM0'  : {'shape': 'uniform', 'min': 10.  ,  'max': 13.},
-    'sigma_logM': {'shape': 'uniform', 'min': .1 ,  'max': .5},
+    'sigma_logM': {'shape': 'uniform', 'min': np.log(.1) ,  'max': np.log(.7)},
     'logMmin': {'shape': 'uniform', 'min': 11.02,  'max': 13.02},   
     'alpha': {'shape': 'uniform', 'min': .8 ,  'max': 1.3},
     'logM1'  : {'shape': 'uniform', 'min': 13.  ,  'max': 14.},
@@ -74,7 +72,7 @@ class HODsim(object):
[…]

### 2016-01-05 halo/text/abc.tex
started writing into the intro section

@@
 
-\title{Likelihood free inference of the halo model parameters with simulations of galaxy mock catalogs}
+\title{Likelihood free inference of the halo model parameters with group and clustering 
+       statistics of the simulations of galaxy mock catalogs}
 
@@
 galaxy evolution within the context of the halo model. Constraints on the parameters of the HOD model have 
-so far been infered using a likelihood function that assumes, first a functional---often times Gaussian---form of the likelihood, 
+so far been infered using a likelihood function that assumes, first a functional---Gaussian---form of the likelihood, 
 and Second, an analytical model for the galaxy clustering statistics. We use Approximate Bayesian Computation 
-in order to infer the HOD parameters by assuming a prior over the parameters and a distance function between 
-statistical summaries of the data and statistical summaries of the simulations.
+in order to infer the HOD parameters by assuming a prior over the parameters, and a distance function between 
+statistical summaries of the data and those of the simulations.
 Using a mock galaxy catalog created with known HOD parameters, we obtain precise constraints on the HOD parameters that 
@@
 the galaxy number density and the galaxy group multiplicity function. 
-We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters, 
-specifically the ones that govern the population of the satelite galaxies.
+We demonstrate that adding these measurements leads to stronger constraints on the HOD parameters, specifically the parameters 
+that control the population of the satellite galaxies.
 
@@
 \section{Introduction}
+
+Hogg writes here...
+
+Inference of the cosmological parameters and understanding the formation of galaxies whitihn the context of large scale 
+structure, requires understanding of the Galaxy-Halo connection. Halos are biased tracers of the large scale structure 
+and galaxies reside dark matter Halos. Halo occupation distribution---HOD---is a probabilistic framework that describes how 
+galaxies populate dark matter halos. The central tenet of HOD modeling is that the probability of $N$ galaxies residing a 
+dark matter halo is determined by the mass of the host halo. This statistical prescription for connection of galaxies and halos 
+has been successful in reproducing the galaxy two-point clustring as well as the galaxy-matter cross correlation---galaxy-galaxy 
+lensing---within the context of the halo model, and has been used in understanding teh co-evolution of galaxies and halos and putting 
+constraints on th relation between the stellar mass and halo mass of galaxies.
 

### 2016-01-05 halo/text/abc.tex
more stuff in the text

@@
 
-Inference of the cosmological parameters and understanding the formation of galaxies whitihn the context of large scale 
+Inference of the cosmological parameters and understanding the formation of galaxies within the context of large scale 
 structure, requires understanding of the Galaxy-Halo connection. Halos are biased tracers of the large scale structure 
@@
 constraints on th relation between the stellar mass and halo mass of galaxies.
+
+In the context of large scale clustering of galaxies, inference of the cosmological and halo model parameters are performed assuming 
+an analytical model for the two-point correlation function of the galaxies, and a gaussian likehood. This Gaussian likelihood approximates 
+the probability of observing the data---observed positions of the galaxies in the configuration space---given the halo model by computing 
+the squared difference of the two-point correlation function measured from the observations and the analytical model for the galaxy 
+clustering, weighted by a covariance matrix. The weighted difference of the number density of the observed galaxies and the analytical 
+number density is often added to this.
+
+Hogg writes why we do not want to do this .....
+
+Approximate bayesian computation (ABC) provides a framework that alleviates the issues arising from approximating the likelihood of 
+observing the large scale population of galaxies given the cosmological and halo model parameters. ABC approximates the posterior 
+probability distribution function over the model parameters given the data---the angular positions of the galaxies on the sky and 
+their redshifts---by drawing proposals from the prior over the model parameters, simulating the data with the proposals, and rejecting 
+the proposals for which the distance between the data and the simulation (measured by a metric) is less than a certain threshold. The 
+procedure described here is rejected sampling. In practice rejected sampling is expensive and instead, investigators change start 
+by sampling from the prior with a large ..........
 

### 2016-01-05 halo/code/abc_pemcee.py
code clean up/review

@@ -1,0 +1,87 @@
+import time
+import numpy as np
+
+from scipy.stats import norm , gamma 
+from scipy.stats import multivariate_normal
+
+import abcpmc
+from abcpmc import mpi_util
+
+# --- Local --- 
+import data as Data
+#from distance import rho 
+from hod_sim import HODsim
+from group_richness import richness
+
+# --- Plotting ---
+from plotting import plot_thetas
+
+def abcpmc_nbar_gmf(T, eps_val, N_part=1000): 
+    '''
+    '''
+    # data observables
+    data_gmf, data_gmf_sigma = Data.data_gmf(Mr=20)
+    data_nbar, data_nbar_var = Data.data_nbar(Mr=20)
+    data = [data_nbar , data_gmf]   # nbar, GMF
+    
+    # True HOD parameters
+    data_hod = np.array([11.38 , np.log(0.26) , 12.02 , 1.06 , 13.31])
+    
+    # Priors
+    prior_min = [10., np.log(0.1), 11.02, 0.8, 13.]
+    prior_max = [13., np.log(0.7), 13.02, 1.3, 14.]
+    prior = abcpmc.TophatPrior(prior_min, prior_max)
+    prior_range = np.zeros((len(prior_min),2))
+    prior_range[:,0] = prior_min
+    prior_range[:,1] = prior_max
+
+    # simulator
+    our_model = HODsim()    # initialize model
+    kwargs = {'prior_range': prior_range, 'observables': ['nbar', 'gmf']}
+    def simz(theta): 
+        return our_model.sum_stat(theta, **kwargs)
+
+    def multivariate_rho(datum, model): 
+        dist_nz = (datum[0] - model[0])**2. / data_nbar_var 
+        dist_gr = np.sum((datum[1] - model[1])**2. / data_gmf_sigma**2.)
+        
+        dist = np.array([dist_nz , dist_gr])
+        return dist
+
+    mpi_pool = mpi_util.MpiPool()
+    abcpmc_sampler = abcpmc.Sampler(
+            N=N_part,       # N_particles
+            Y=data,         # data
+            postfn=simz,    # simulator 
+            dist=multivariate_rho,       # distance function  
+            pool=mpi_pool)  
+    abcpmc_sampler.particle_proposal_cls = abcpmc.ParticleProposal
+
+    eps = abcpmc.ConstEps(T, [1.e13,1.e13])
+    pools = []
[…]

### 2016-01-05 halo/code/alternative_richness_hod.py

@@ -1,28 +1,27 @@
+'''
+
+Documentation here
+
+Authors : MJ, Chang
+
+'''
+import time
 import numpy as np
 import matplotlib.pyplot as plt
-import time
 plt.switch_backend("Agg")
 from halotools.empirical_models import Zheng07
-from astropy.table import Table
+
 import corner
+
 from scipy.stats import norm , gamma 
 from scipy.stats import multivariate_normal
+
 import abcpmc
-import seaborn as sns
-sns.set_style("white")
-np.random.seed()
 from abcpmc import mpi_util
 
-
-def richness(group_id): 
-    gals = Table() 
-    gals['groupid'] = group_id
-    gals['dummy'] = 1
-    grouped_table = gals.group_by('groupid')
-    grp_richness = grouped_table['dummy'].groups.aggregate(np.sum)
-    return grp_richness
-
-
+# --- Local --- 
+from group_richness import richness
+from distance import rho 
 
 """Load data and variance"""
 data_gmf = np.loadtxt("gmf_Mr20.dat")
@@ -41,12 +40,9 @@ data_hod = np.array([11.38 , np.log(0.26) , 12.02 , 1.06 , 13.31])
 
 """Prior"""
 
-prior = abcpmc.TophatPrior([10.,np.log(.1),11.02,.8,13.],[13.,np.log(.7),13.02,1.3,14.])
-
 #prior = abcpmc.TophatPrior([11.91,.38,12.78,1.1,13.9],[11.92,.4,12.8,1.2,14.])
 
 prior_dict = {
- 
     'logM0'  : {'shape': 'uniform', 'min': 10.  ,  'max': 13.},
     'sigma_logM': {'shape': 'uniform', 'min': np.log(.1) ,  'max': np.log(.7)},
     'logMmin': {'shape': 'uniform', 'min': 11.02,  'max': 13.02},   
@@ -62,94 +58,20 @@ for key in ['logM0', 'sigma_logM', 'logMmin','alpha','logM1']:
 prior_range = np.array(plot_range)
 print prior_range[:,0] , prior_range[:,1]
 
-"""simulator"""
-
-class HODsim(object): 
-    
-    def __init__(self): 
-        self.model = Zheng07()
-    
-    def sum_stat(self, theta):
-        
-        self.model.param_dict['logM0'] = theta[0]
-        self.model.param_dict['sigma_logM'] = np.exp(theta[1])
-        self.model.param_dict['logMmin'] = theta[2]     
[…]

### 2016-01-05 halo/code/data.py

@@ -1,0 +1,44 @@
+'''
+
+Modules for data and observables in ABC-PMC
+
+'''
+import numpy as np
+
+def data_gmf(Mr=20): 
+    '''
+    Observed GMF from 'data'
+    '''
+
+    gmf_dat_file = ''.join([
+        '../dat/', 
+        'gmf_Mr', str(Mr), '.dat'
+        ])
+    
+    gmf_sig_file = ''.join([
+        '../dat/', 
+        'gmf_noise_Mr20_2.dat'
+        ])
+    sigma = np.loadtxt(gmf_sig_file)
+    
+    return [np.loadtxt(gmf_dat_file), sigma]
+
+def data_gmf_bins(Mr=20):
+    '''
+    Observed GMF bins
+    '''
+    bins = np.array([3.000000000000000000e+00, 4.000000000000000000e+00, 5.000000000000000000e+00, 6.000000000000000000e+00, 7.000000000000000000e+00, 8.000000000000000000e+00, 9.000000000000000000e+00, 1.000000000000000000e+01, 1.100000000000000000e+01, 1.200000000000000000e+01, 1.300000000000000000e+01, 1.400000000000000000e+01, 1.500000000000000000e+01, 1.600000000000000000e+01, 1.700000000000000000e+01, 1.800000000000000000e+01, 1.900000000000000000e+01, 2.000000000000000000e+01, 2.100000000000000000e+01, 2.200000000000000000e+01, 2.300000000000000000e+01, 2.500000000000000000e+01, 2.900000000000000000e+01, 3.100000000000000000e+01, 3.500000000000000000e+01, 4.300000000000000000e+01, 6.100000000000000000e+01]) # same hardcoded bins as data 
+
+    return bins
+
+def data_nbar(Mr=20): 
+    '''
+    Observed nbar from 'data'
+    '''
+    nbar_dat_file = ''.join([
+        '../dat/',
+        'nbar_Mr', str(Mr),'.dat'
+        ])
+    mock_nbar = np.loadtxt(nbar_dat_file)
+
+    return [np.mean(mock_nbar), np.var(mock_nbar)]

### 2016-01-05 halo/code/group_richness.py

@@ -1,0 +1,31 @@
+'''
+
+
+Modules for Galaxy Group Richness
+
+
+'''
+import numpy as np
+from astropy.table import Table
+
+
+def richness(group_id): 
+    '''
+    Calculate the richness of a group given group_ids of galaxies. Uses astropy.table module
+    '''
+    gals = Table() 
+    gals['groupid'] = group_id
+    gals['dummy'] = 1
+    grouped_table = gals.group_by('groupid')
+    grp_richness = grouped_table['dummy'].groups.aggregate(np.sum)
+
+    return grp_richness
+
+def gmf(group_richness): 
+    '''
+    Calculate Group Multiplicity Function (GMF) given group richness (see above module). 
+    '''
+    bins = np.array([3.000000000000000000e+00, 4.000000000000000000e+00, 5.000000000000000000e+00, 6.000000000000000000e+00, 7.000000000000000000e+00, 8.000000000000000000e+00, 9.000000000000000000e+00, 1.000000000000000000e+01, 1.100000000000000000e+01, 1.200000000000000000e+01, 1.300000000000000000e+01, 1.400000000000000000e+01, 1.500000000000000000e+01, 1.600000000000000000e+01, 1.700000000000000000e+01, 1.800000000000000000e+01, 1.900000000000000000e+01, 2.000000000000000000e+01, 2.100000000000000000e+01, 2.200000000000000000e+01, 2.300000000000000000e+01, 2.500000000000000000e+01, 2.900000000000000000e+01, 3.100000000000000000e+01, 3.500000000000000000e+01, 4.300000000000000000e+01, 6.100000000000000000e+01]) # same hardcoded bins as data 
+
+    gmeff = np.histogram(group_richness , bins)[0] / 250.**3.   # calculate GMF
+    return gmeff

### 2016-01-06 halo/code/abc_pemcee.py
getting abc_pemcee.py working along with the plotting module

@@ -64,8 +64,8 @@ def abcpmc_nbar_gmf(T, eps_val, N_part=1000):
         print eps(pool.t)
 
         plot_thetas(pool.thetas , pool.ws , pool.t, truths=data_hod, plot_range=prior_range)
-        np.savetxt("/../dat/nbar_gmf5_Mr20_theta_t"+str(t)+".dat" , theta)
-        np.savetxt("/../dat/nbar_gmf5_Mr20_w_t"+str(t)+".dat" , w)
+        np.savetxt("/../dat/nbar_gmf5_Mr20_theta_t"+str(pool.t)+".dat" , theta)
+        np.savetxt("/../dat/nbar_gmf5_Mr20_w_t"+str(pool.t)+".dat" , w)
 
         if pool.t < 3: 
             eps.eps = np.percentile(np.atleast_2d(pool.dists), 50 , axis = 0)
@@ -84,4 +84,4 @@ def abcpmc_nbar_gmf(T, eps_val, N_part=1000):
     return pools
 
 if __name__=="__main__": 
-    abcpmc_nbar_gmf(40, 60, N_part=10)
+    abcpmc_nbar_gmf(2, 60, N_part=10)

### 2016-01-06 halo/code/plotting.py

@@ -5,6 +5,9 @@ Plotting modules
 '''
 import corner
 
+import matplotlib.pyplot as plt
+plt.switch_backend("Agg")
+
 def plot_thetas(theta, w , t, truths=None, plot_range=None): 
     '''
     Corner plots of input theta values 
@@ -26,7 +29,7 @@ def plot_thetas(theta, w , t, truths=None, plot_range=None):
             bins=20,
             smooth=1.0)
     
-    plt.savefig("/home/mj/public_html/nbar_gmf5_Mr20_t"+str(t)+".png")
+    plt.savefig("nbar_gmf5_Mr20_t"+str(t)+".png")
     plt.close()
     
     # not weighted
@@ -44,5 +47,5 @@ def plot_thetas(theta, w , t, truths=None, plot_range=None):
             color='b', 
             bins=16, 
             smooth=1.0)
-    plt.savefig("/home/mj/public_html/nbar_gmf5_Mr20_now_t"+str(t)+".png")
+    plt.savefig("nbar_gmf5_Mr20_now_t"+str(t)+".png")
     plt.close()

### 2016-01-06 halo/code/abc_pemcee.py
hod_sim adjusted to account for new version of Halotools; Zheng07() no longer supported

@@ -1,9 +1,6 @@
 import time
 import numpy as np
 
-from scipy.stats import norm , gamma 
-from scipy.stats import multivariate_normal
-
 import abcpmc
 from abcpmc import mpi_util
 
@@ -16,7 +13,7 @@ from group_richness import richness
 # --- Plotting ---
 from plotting import plot_thetas
 
-def abcpmc_nbar_gmf(T, eps_val, N_part=1000): 
+def abcpmc_nbar_gmf(T, eps_val, N_part=1000, threads=1): 
     '''
     '''
     # data observables
@@ -54,6 +51,7 @@ def abcpmc_nbar_gmf(T, eps_val, N_part=1000):
             Y=data,         # data
             postfn=simz,    # simulator 
             dist=multivariate_rho,       # distance function  
+            threads=threads,
             pool=mpi_pool)  
     abcpmc_sampler.particle_proposal_cls = abcpmc.ParticleProposal
 
@@ -84,4 +82,4 @@ def abcpmc_nbar_gmf(T, eps_val, N_part=1000):
     return pools
 
 if __name__=="__main__": 
-    abcpmc_nbar_gmf(2, 60, N_part=10)
+    abcpmc_nbar_gmf(10, 60, N_part=100)

### 2016-01-06 halo/code/hod_sim.py

@@ -4,7 +4,7 @@ HaloTools HOD Simulation
 
 '''
 import numpy as np
-from halotools.empirical_models import Zheng07
+from halotools.empirical_models import PrebuiltHodModelFactory
 
 from data import data_gmf_bins
 from group_richness import gmf 
@@ -19,7 +19,8 @@ class HODsim(object):
         Our model forward models the galaxy catalog using HOD parameters using HaloTools. 
         '''
 
-        self.model = Zheng07()      # Zheng et al. (2007) model
+        #self.model = zheng07()      # Zheng et al. (2007) model
+        self.model = PrebuiltHodModelFactory('zheng07')
     
     def sum_stat(self, theta, prior_range=None, observables=['nbar', 'gmf']):
         '''

### 2016-01-06 halo/code/abc_pemcee.py
abc code being updated; much more pythonic

@@ -13,16 +13,27 @@ from group_richness import richness
 # --- Plotting ---
 from plotting import plot_thetas
 
-def abcpmc_nbar_gmf(T, eps_val, N_part=1000, threads=1): 
+def abcpmc_nbar_gmf(T, eps_val, Mr=20, N_part=1000, threads=1, observables=['nbar', 'gmf']): 
     '''
     '''
     # data observables
-    data_gmf, data_gmf_sigma = Data.data_gmf(Mr=20)
-    data_nbar, data_nbar_var = Data.data_nbar(Mr=20)
-    data = [data_nbar , data_gmf]   # nbar, GMF
+    data = []       # list of observables 
+    for obv in observables: 
+        if obv == 'nbar': 
+            data_nbar, data_nbar_var = Data.data_nbar(Mr=Mr)
+            data.append(data_nbar)
+        if obv == 'gmf': 
+            data_gmf, data_gmf_sigma = Data.data_gmf(Mr=Mr)
+            data.append(data_gmf)
+        if obv == 'xi': 
+            data_xi, data_cov_ii = Data.data_xi(Mr=Mr)
+            data.append(data_xi)
     
     # True HOD parameters
-    data_hod = np.array([11.38 , np.log(0.26) , 12.02 , 1.06 , 13.31])
+    if Mr == 20: 
+        data_hod = np.array([11.38 , np.log(0.26) , 12.02 , 1.06 , 13.31])
+    else: 
+        raise NotImplementedError
     
     # Priors
     prior_min = [10., np.log(0.1), 11.02, 0.8, 13.]
@@ -34,16 +45,22 @@ def abcpmc_nbar_gmf(T, eps_val, N_part=1000, threads=1):
 
     # simulator
     our_model = HODsim()    # initialize model
-    kwargs = {'prior_range': prior_range, 'observables': ['nbar', 'gmf']}
+    kwargs = {'prior_range': prior_range, 'observables': observables}
     def simz(theta): 
         return our_model.sum_stat(theta, **kwargs)
 
     def multivariate_rho(datum, model): 
-        dist_nz = (datum[0] - model[0])**2. / data_nbar_var 
-        dist_gr = np.sum((datum[1] - model[1])**2. / data_gmf_sigma**2.)
-        
-        dist = np.array([dist_nz , dist_gr])
-        return dist
+        dists = [] 
+        for i_obv, obv in enumerate(observables): 
[…]

### 2016-01-06 halo/code/hod_sim.py

@@ -46,9 +46,12 @@ class HODsim(object):
                 if obv == 'nbar': 
                     obvs.append(self.model.mock.number_density)       # nbar of the galaxy catalog
                 elif obv == 'gmf': 
-                    group_id = self. model.mock.compute_fof_group_ids() 
+                    group_id = self.model.mock.compute_fof_group_ids() 
                     group_richness = richness(group_id)         # group richness of the galaxies
                     obvs.append(gmf(group_richness))                 # calculate GMF
+                elif obv == 'xi': 
+                    r, xi_r = self.model.mock.compute_galaxy_clustering()
+                    obvs.append(xi_r)
                 else: 
                     raise NotImplementedError('Only nbar and GMF implemented so far')
 
@@ -68,6 +71,9 @@ class HODsim(object):
                             group_id = self. model.mock.compute_fof_group_ids()
                             group_richness = richness(group_id)         # group richness of the galaxies
                             obvs.append(gmf(group_richness))                 # calculate GMF
+                        elif obv == 'xi': 
+                            r, xi_r = self.model.mock.compute_galaxy_clustering()
+                            obvs.append(xi_r)
                         else: 
                             raise NotImplementedError('Only nbar and GMF implemented so far')
 
@@ -79,4 +85,3 @@ class HODsim(object):
                 else:
                     bins = data_gmf_bins(Mr=20)
                     return [10. , np.ones_like(bins)[:-1]*1000.]
-

### 2016-01-06 halo/code/plotting.py

@@ -3,12 +3,12 @@
 Plotting modules 
 
 '''
+import util
 import corner
-
 import matplotlib.pyplot as plt
 plt.switch_backend("Agg")
 
-def plot_thetas(theta, w , t, truths=None, plot_range=None): 
+def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=None): 
     '''
     Corner plots of input theta values 
     '''
@@ -29,7 +29,10 @@ def plot_thetas(theta, w , t, truths=None, plot_range=None):
             bins=20,
             smooth=1.0)
     
-    plt.savefig("nbar_gmf5_Mr20_t"+str(t)+".png")
+    fig_file = ''.join([util.fig_dir(), 
+        util.observable_id_flag(observables), 
+        '_Mr', str(Mr), '_t', str(t), '.png'])
+    plt.savefig(fig_file)
     plt.close()
     
     # not weighted
@@ -47,5 +50,8 @@ def plot_thetas(theta, w , t, truths=None, plot_range=None):
             color='b', 
             bins=16, 
             smooth=1.0)
-    plt.savefig("nbar_gmf5_Mr20_now_t"+str(t)+".png")
+    fig_file = ''.join([util.fig_dir(), 
+        util.observable_id_flag(observables), 
+        '_Mr', str(Mr), '_now_t', str(t), '.png'])
+    plt.savefig(fig_file)
     plt.close()

### 2016-01-06 halo/code/util.py

@@ -1,0 +1,23 @@
+'''
+
+Utility modules 
+
+'''
+
+def observable_id_flag(observables): 
+    '''
+    Observable identification flag string given list of observables
+    '''
+    return '_'.join(observables) 
+
+def fig_dir(): 
+    '''
+    figure directory
+    '''
+    return ''
+
+def dat_dir(): 
+    '''
+    Dat directory
+    '''
+    return '/../dat/'

### 2016-01-06 halo/text/abc.tex
text so far

--- 2015-12-20 halo/text/abc.tex
@@
 \documentclass[12pt, preprint]{aastex}
+\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
 \usepackage{graphicx}	% For figures
@@
 
+\newcommand{\todo}[1]{{\bf \textcolor{red}{ #1}}}
 \newcommand{\beq}{\begin{equation}}
 \newcommand{\eeq}{\end{equation}}
-\newcommand{\la}{\langle}
+\newcommand{\lang}{\langle}
 \newcommand{\ra}{\rangle}
@@
 
-\title{Likelihood free inference of the halo model parameters with simulations of galaxy mock catalogs}
+\title{Likelihood free inference of the halo model parameters with group and clustering 
+       statistics of the simulations of galaxy mock catalogs}
 
 \begin{abstract}
+{\bf other abstract for reference}\\
+The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies populate dark matter halos that has had much success in reproducing large scale galaxy clustering and understanding galaxy evolution within the context of the halo model.
+Constraints on the parameters of the HOD model have so far been imposed using the standard practice of Bayesian inference with likelihood estimation that assumes a functional (often times Gaussian) form of the likelihood.
+Likelihood free inference methods such as the Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever having to calculate the likelihood.
+
+In this work, we present ABC in the context of constraining the parameters of the HOD model.
+For a mock galaxy catalog with known HOD parameters, we use ABC on measurements of a set of observables generated through {\it Halotools}, a publicly available ..., to infer tight constraints that reproduce the known HOD paramters.
+Since ABC does not require computing the likelihood, we include observables beyond galaxy clustering such as the galaxy number density and the galaxy group richness, for which accurately constructing a likelihood is intractable.
+We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters,
+ specifically the ones that govern the satellite galaxy population.
+Moreover, the constraints we infer that reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.
 
@@
 galaxy evolution within the context of the halo model. Constraints on the parameters of the HOD model have 
-so far been infered using a likelihood function that assumes, first a functional---often times Gaussian---form of the likelihood, 
-and Second, an analytical model for the galaxy clustering statistics. We use Approximate Bayesian Computation 
-in order to infer the HOD parameters by assuming a prior over the parameters and a distance function between 
-statistical summaries of the data and statistical summaries of the simulations.
+so far been inferred using a likelihood function that assumes a functional--Gaussian--form of the likelihood
+and an analytic model for the galaxy clustering statistics.
+Likelihood free inference methods such as the Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever having to calculate the likelihood.
+
+In this work, we use ABC in order to infer the HOD parameters by assuming a prior over the parameters, and a distance function between 
+statistical summaries of the data and those of the simulations.
 Using a mock galaxy catalog created with known HOD parameters, we obtain precise constraints on the HOD parameters that 
 reproduces the summary statistics of the mock galaxy catalog accurately. 
-Since ABC does not require computing the likelihood, we include observables beyond galaxy clustering such as 
+Since ABC does not require computing the likelihood, we include observable beyond galaxy clustering such as 
 the galaxy number density and the galaxy group multiplicity function. 
-We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters, 
-specifically the ones that govern the population of the satelite galaxies.
+We demonstrate that adding these measurements leads to stronger constraints on the HOD parameters, specifically the parameters 
+that control the population of the satellite galaxies.
+
+\todo{Hogg writes here too}
 
@@
 \section{Introduction}
+
+\todo{Hogg writes here ...}
+
+Inference of the cosmological parameters and understanding the formation of galaxies within the context of large scale 
+structure, requires understanding of the Galaxy-Halo connection. Halos are biased tracers of the large scale structure 
+and galaxies reside in dark matter Halos.
+Halo occupation distribution---HOD---is a probabilistic framework that describes how galaxies populate dark matter halos.
+The central tenet of halo occupation model is that the probability that a certain number of galaxies populate a dark matter halo is determined by the mass of the host halo.
+This statistical prescription for connection of galaxies and halos has been successful in reproducing the galaxy two-point clustering as well as the galaxy-matter cross correlation---galaxy-galaxy
+lensing---within the context of the halo model, and has been used in understanding the co-evolution of galaxies and halos and putting constraints on the relation between the stellar mass and halo mass of galaxies.
+
+In the context of large scale clustering of galaxies, inference of the cosmological and halo model parameters are performed assuming an analytic model for the two-point correlation function of the galaxies, and a Gaussian likelihood.
+The accuracy of the analytic model of the galaxy clustering statistics hinges upon the accuracy of the fitting functions such as halo bias, halo mass function that are accurate to ten percent level.
+There are other simplifying assumptions encoded in these analytic models such as the scale independence of the halo bias.
+
+Furthermore, the likelihood of observing the data---observed positions of the galaxies in the configuration space---given the cosmological and halo model parameters is approximated by a Gaussian function whose argument is the squared difference of the two-point correlation function measured from the observations and the analytic model for the galaxy clustering, weighted by a covariance matrix.
+The weighted difference of the number density of the observed galaxies and the analytic number density is also added to this.
+In this work, we investigate the how accurately we can infer the model parameters by relaxing the underlying assumptions of the parameter inference methods.
+
+Approximate bayesian computation (ABC) provides a \emph{rejected sampling} framework framework that alleviates the issues arising from approximating the likelihood of observing the large scale population of galaxies given the cosmological and halo model parameters.
+ABC approximates the posterior probability distribution function over the model parameters given the data---the angular positions of the galaxies on the sky and their redshifts---by drawing proposals from the prior over the model parameters, simulating the data with the proposals, and rejecting the proposals for which the distance between the data and the simulation (measured by a metric) is less than a certain small threshold.
+
+Furthermore, This sampling formalism permits us to sample the posterior of the model by simulating the data via a forward generative model that accounts for all the complications and uncertainties associated with the survey data that are difficult to take into account in approximating the likelihood function.
+In the context of the large galaxy surveys, these complications include masks, nontrivial survey geometry, and missing data as a result of fiber collision.
+
+Rejected sampling in ABC requires a distance function that quantifies the closeness of the simulation (e.g.
+fiber collided galaxies) to the data (observed positions of the galaxies in the survey).
+In principle, this distance metric is a positive definite function  that compares various summary statistics of the data and those of the simulations.
+In this investigation, we focus on the two-point correlation, number density, and the group statistics (more specifically the abundance of the galaxy group richnesses).
+
+Performing a full cosmological inference using this sampling method requires a simulator that, (I) takes a set of cosmological parameters as input and generates a snapshot of dark matter particles at a given redshift using an N-body simulation, (II) builds a catalog of dark matter halos for that snapshot of the N-body simulation using a halo finder algorithm, and (III) paints galaxies into host halos given a set of halo occupation model parameters.
+
+Computationally, this process is extremely demanding.
+One way to make this problem more tractable is to split this into to separate this problem into two parts: \emph{emulation}, and \emph{simulation}.
+That is, one can use a set of N-body simulations with known cosmologies, and perform halo occupation simulations using the pre-built halo catalogs of those simulations.
+In order to compute the summary statistics corresponding to the samples from the parameter space for which there is no N-body simulation, one can interpolate between the summary statistics of the parameters for which we have available N-body simulations.
+This procedure is called emulation.
+Emulation has been introduced in the literature in coyote universe [CITE COYOTE].
+
+In this work however, we narrow our focus on inferring the posterior probability over the parameters of the halo model using an already existing halo catalog of a dark matter particle N-body simulation with a fixed cosmology.
+That is we simplify the problem by leaving the problem of emulation for future investigations.
+
+In practice, this rejected sampling method is replaced by a more efficient procedure.
+Instead, we start by sampling from the prior with a relatively large threshold.
+In the consequent steps, we iteratively perform importance sampling from the previous samples, and lower the threshold adaptively.
+
+This method requires a simulator or a forward model of the positions of galaxies given the model parameters.
+We use the publicly available code Halotools in order to populate the halo catalogs with galaxies brighter than a certain luminosity threshold.
 
@@
 has a separate threshold. This procedure is done iteratively untill, we are left with a certain number
-of samples that satisfy the distance criteria. The final ensemble approximates the posterior pdf over 
-the paramters of the forward model of the data. This sampling method is called \emph{rejected sampling}.
+of samples that satisfy the distance criteria.
+The final ensemble approximates the posterior probability distribution over the parameters of the generative model of the data.
+This sampling method is called \emph{rejected sampling}.
 
-Recected sampling is expensive. Instead we use population monte carlo, adaptive importance sampling, ....
+Rejected sampling is expensive. Instead we use Population Monte Carlo, adaptive importance sampling, ....
 
@@
 
-Halo Occupation distribution is a powerful tool for description of galaxy/halo relation.
+Halo Occupation distribution is a powerful tool for description of galaxy/halo relation. \todo{Here we explain HOD modeling of galaxy positions, etc.}
 
@@
 \subsection{Summary statistics}
+\begin{itemize}
 
-Talk about 2pcf, group multiplicity function, and number density ...
+\item \todo{Talk about the observables: 2pcf, group multiplicity function, and number density ...}
+
+\end{itemize}
 
@@
 
-Hello World!
+\begin{itemize}
+\item \todo{Halotools allows us to forward model galaxy positions from HOD parameters.
+From the galaxy positions, we can calculate observable summary statistics such as the 2PCF, Group Multiplicity Function (GMF) and the number density $\bar{n}$.}
+
+\end{itemize}
 Talk about why it is better to simulate the mock and compute the summary statistics of the mock instead of 
-using analytical formulas...
+using analytic formulas...
 
@@
 \section{Mock data}
-
-Hello World!
-
 Plots showing 2pcf, gmf, ... of the mock data.
@@
 \section{Results}
-
 Show the posteriors...
@@
 
+\subsection{Comparison to MCMC}
+\todo{Apples to apples comparison between ABC and MCMC.}
+\begin{itemize}
+\item ABC-PMC using nbar and xi vs MCMC
+\cite{Ishida:2015aa}
+\end{itemize}
+
+\subsection{ABC Including GMF}
+ABC-PMC results using observables beyond galaxy clustering and nbar(z) GMF
+
 \section{Discussion}
+\todo{talk about cosmological infeence and why this is good first step for likelihood free inference in large scale ...}
 
-talk about cosmological infeence and why this is good first step for likelihood free inference in large scale ...
-
-talk about why it is hard to do cosmology ...
+\todo{talk about why it is hard to do cosmology ...}
 
@@
 talk about future directions ...
-
-\begin{thebibliography}{70}
-
-\end{thebibliography}
-
+\bibliographystyle{yahapj}
+\bibliography{ccppabc}
 \end{document}

### 2016-01-07 halo/text/abc.tex
text so far

@@
 urlcolor=darkred,
-citecolor=darkblue }
+citecolor=darkblue}
 
@@
 \begin{abstract}
-{\bf other abstract for reference}\\
-The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies populate dark matter halos that has had much success in reproducing large scale galaxy clustering and understanding galaxy evolution within the context of the halo model.
-Constraints on the parameters of the HOD model have so far been imposed using the standard practice of Bayesian inference with likelihood estimation that assumes a functional (often times Gaussian) form of the likelihood.
-Likelihood free inference methods such as the Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever having to calculate the likelihood.
+The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies
+populate dark matter halos that has had much success in reproducing large scale galaxy
+clustering and understanding galaxy evolution within the context of the halo model.
+Constraints on the parameters that describe the HOD model have so far been inferred using 
+the standard practice of Bayesian inference with likelihood estimation that assumes a
+functional Gaussian form of the likelihood.\todo{(Do we want to include analytic models for clustering statistics here?)} Likelihood free inference methods such as the
+Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever 
+having to calculate the likelihood.
 
 In this work, we present ABC in the context of constraining the parameters of the HOD model.
-For a mock galaxy catalog with known HOD parameters, we use ABC on measurements of a set of observables generated through {\it Halotools}, a publicly available ..., to infer tight constraints that reproduce the known HOD paramters.
-Since ABC does not require computing the likelihood, we include observables beyond galaxy clustering such as the galaxy number density and the galaxy group richness, for which accurately constructing a likelihood is intractable.
-We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters,
- specifically the ones that govern the satellite galaxy population.
-Moreover, the constraints we infer that reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.
-
-The Halo Occupation Distribution model is a powerful prescription of how galaxies populate 
-dark matter halos that has had much success in reproducing large scale galaxy clustering and understanding
-galaxy evolution within the context of the halo model. Constraints on the parameters of the HOD model have 
-so far been inferred using a likelihood function that assumes a functional--Gaussian--form of the likelihood
-and an analytic model for the galaxy clustering statistics.
-Likelihood free inference methods such as the Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever having to calculate the likelihood.
+For a mock galaxy catalog created with HOD parameters selected from literature
+\todo{(Should we specify Zheng07?)}, we use ABC on measurements of a set of observables
+generated through {\it Halotools}, a publicly available \todo{insert description about
+forward modeling of galaxy positions}, to infer precise constraints that reproduce the 
+known HOD parameters.
 
@@
 reproduces the summary statistics of the mock galaxy catalog accurately.
+
+Furthermore, since ABC does not require computing the likelihood, we include observables
+beyond galaxy clustering such as the galaxy number density and the galaxy group 
+multiplicity function, for which accurately constructing a likelihood is intractable.
+We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters,
+ specifically the ones that govern the satellite galaxy population.
+Moreover, the constraints we infer that reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.
+
 Since ABC does not require computing the likelihood, we include observable beyond galaxy clustering such as 
@@
 
+In practice, the rejected sampling method is replaced by a more efficient adaptive importance sampling operation called Population Monte Carlo (PMC).
+Initially, PMC rejects the draws from prior pdf with a relatively large threshold.
+In the consequent steps, importance sampling from the previous samples that passed the distance criteria is performed, and the threshold is updated adaptively.
+That is, after each iteration, the threshold is lowered according to the distribution of the distances.
+
 Furthermore, This sampling formalism permits us to sample the posterior of the model by simulating the data via a forward generative model that accounts for all the complications and uncertainties associated with the survey data that are difficult to take into account in approximating the likelihood function.
-In the context of the large galaxy surveys, these complications include masks, nontrivial survey geometry, and missing data as a result of fiber collision.
+In the context of the large galaxy surveys, these complications may include masks, nontrivial survey geometry, and missing data as a result of fiber collision.
 
@@
 
-In practice, this rejected sampling method is replaced by a more efficient procedure.
-Instead, we start by sampling from the prior with a relatively large threshold.
-In the consequent steps, we iteratively perform importance sampling from the previous samples, and lower the threshold adaptively.
+The calculations regarding populating the dark matter halos, computing the clustering statistics and finding groups in the galaxy mock catalogs are performed using ({\fontfamily{cmss}\selectfont Halotools}) which is an open-source package tailored for modeling galaxy-halo connection in large scale structure modeling ({\fontfamily{cmss}\selectfont http://halotools.readthedocs.org}).
 
-This method requires a simulator or a forward model of the positions of galaxies given the model parameters.
-We use the publicly available code Halotools in order to populate the halo catalogs with galaxies brighter than a certain luminosity threshold.
+This paper is structured as follows. In Section \ref{sec:method},
+we discuss the components of the Halo occupation modeling, and the ABC sampling method.
+We also give a brief overview of the summary statistics used in this investigation.
+In Section \ref{sec:halo} we give a brief overview of the forward modeling of the galaxy mock catalogs using Halotools as well as generating a mock galaxy catalog as data.
+In Section \ref{sec:results} we explore the results of our parameter inference using different summary statistics.
+Finally, we discuss and conclude in Section \ref{sec:discussion}.
 
-\section{Method}
+\section{Method}\label{sec:method}
 
@@
 
-\section{Forward modeling of the galaxy population with HALOTOOLS}
+\section{Forward modeling of the galaxy population with HALOTOOLS}\label{sec:halo}
 
@@
 
-\section{Mock data}
+\subsection{Mock data}
 Plots showing 2pcf, gmf, ... of the mock data.
@@
 
-\section{Results}
+\section{Results}\label{sec:results}
 Show the posteriors...
@@
 
-\section{Discussion}
-\todo{talk about cosmological infeence and why this is good first step for likelihood free inference in large scale ...}
+\section{Discussion}\label{sec:discussion}
+\todo{talk about cosmological inference and why this is good first step for likelihood free inference in large scale ...}
 
@@
 
-talk about writing down a likelihood for gmf, tpcf, ... , and the gaussianity assumption ....
+talk about writing down a likelihood for gmf, tpcf, ... , and the Gaussianity assumption ....
 

### 2016-01-19 halo/text/abc.tex
more text

@@
 \newcommand{\ep}{\epsilon}
+\newcommand{\pars}{\vec{\theta}}
+\newcommand{\dev}{\mathrm{d}}
 
 \begin{document}
-
-\title{Likelihood free inference of the halo model parameters with group and clustering 
-       statistics of the simulations of galaxy mock catalogs}
-
+% We need a better definitely shorter title...
+\title{Approximate bayesian computation in large scale structure modeling : constraining the galaxy-halo connection}
 \begin{abstract}
-The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies
+\todo{This is not a paper about hod, let's not start with describing hod.
+I suggest that we talk about parameter estimation in large scale structure in general and then move on to hod.
+Constraining hod is a just the first step in testing our methodology.} The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies
 populate dark matter halos that has had much success in reproducing large scale galaxy
@@
 the standard practice of Bayesian inference with likelihood estimation that assumes a
-functional Gaussian form of the likelihood.\todo{(Do we want to include analytic models for clustering statistics here?)} Likelihood free inference methods such as the
+functional Gaussian form of the likelihood.\todo{(Do we want to include analytic models for clustering statistics here?
+Yes, also we should say we use summary statistics that are not independent: clustering + group.
+Also there is no analytic model for the abundance of galaxy group richnesses)}  % I think we should keep it simple.
+There's no need to go into detail about everything in the abstract.
+This is already a pretty sizable abstract.
+Likelihood free inference methods such as the
 Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever 
@@
 
-In this work, we present ABC in the context of constraining the parameters of the HOD model.
-For a mock galaxy catalog created with HOD parameters selected from literature
-\todo{(Should we specify Zheng07?)}, we use ABC on measurements of a set of observables
-generated through {\it Halotools}, a publicly available \todo{insert description about
-forward modeling of galaxy positions}, to infer precise constraints that reproduce the 
-known HOD parameters.
-
-In this work, we use ABC in order to infer the HOD parameters by assuming a prior over the parameters, and a distance function between 
-statistical summaries of the data and those of the simulations.
-Using a mock galaxy catalog created with known HOD parameters, we obtain precise constraints on the HOD parameters that 
-reproduces the summary statistics of the mock galaxy catalog accurately.
-
-Furthermore, since ABC does not require computing the likelihood, we include observables
-beyond galaxy clustering such as the galaxy number density and the galaxy group 
-multiplicity function, for which accurately constructing a likelihood is intractable.
-We demonstrate that adding these measurements to our set of observables leads to stronger constraints on the HOD parameters,
- specifically the ones that govern the satellite galaxy population.
-Moreover, the constraints we infer that reproduce the known HOD parameters illustrate the success of ABC in the context of constraining HOD parameters.
-
-Since ABC does not require computing the likelihood, we include observable beyond galaxy clustering such as 
-the galaxy number density and the galaxy group multiplicity function. 
-We demonstrate that adding these measurements leads to stronger constraints on the HOD parameters, specifically the parameters 
-that control the population of the satellite galaxies.
+In this work, we present ABC in the context of constraining the parameters of the HOD model
+using {\it Halotools}, a publicly available ... First, we set a mock galaxy catalog created
+with HOD parameters from \cite{Zheng:2007aa} \todo{why do we site here?} as our data catalog. Then we use ABC on a set 
+% To give it a false sense of legitimacy... 
+of observables measured from forward modeled mock catalogs generated with {\it Halotools} 
+in order to accurately infer constraints on the HOD parameters that 
+reproduce the original parameters. \todo{Remark about the comparison between our ABC 
+method constraints and the standard MCMC constraints.} Furthermore, since ABC does 
+not require computing the likelihood, we include observables beyond galaxy clustering 
+such as the galaxy number density and the galaxy group multiplicity function, for which
+accurately constructing a likelihood is intractable. 
+We demonstrate that adding these measurements to our set of observables leads to stronger
+constraints on the HOD parameters, specifically the ones that govern the satellite galaxy
+population. Moreover, the constraints we infer illustrate the feasibility of ABC in 
+cosmological contexts.
 
@@
 
-\todo{Hogg writes here ...}
+Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
+pro\-ba\-bil\-iites for the data given the theoretical model and parameter settings---%
+for combining data from different surveys and performing principled inferences of
+the cosmological parameters (CITE CITE).
+Nowhere has this been more true than in cosmic microwave background studies,
+where it is nearly possible to analytically evaluate a likelihood function that
+involves no (or minimal) approximations (CITE Wandelt and others).
+
+Fundamentally, the tractability of likelihood functions in cosmology flows from
+the fact that the initial conditions are exceedingly close to Gaussian in form,
+and many sources of measurement noise are also Gaussian.
+Likelihood functions are easier to write down and evaluate when things are closer to Gaussian,
+so they get easier as the length scales get larger and the structure formation gets
+younger.
+Hence the suitability of CMB data for likelihood analyses.
+
+In large-scale structure studies, which have galaxies, quasars, or quasar absorption
+systems as tracers, the likelihood cannot be Gaussian:
+Even if the initial conditions are perfectly Gaussian, the growth of structure
+creates non-linearities which are non-Gaussian, and the galaxies form
+within the density field in some complex manner that is unknown (or modeled only effectively).
+Even if the galaxies were a Poisson sampling of the density field (which they are demonstrably
+not; CITE PEOPLE), it would be tremendously hard to write down even an approximate
+likelihood function.
+
+The standard approach is to make two strong assumptions, both demonstrably wrong:
+The first is that the (density and) two-point function measurements constitute
+sufficient statistics for the cosmological parameters.
+That is, the assumption is that all of the information about the cosmological model is
+encoded in the mean and variance statistics of the tracer population.
+This first assumption is equivalent to assuming that the density field is a Gaussian process,
+and is therefore closer to true on larger scales.
+
+The second strong and wrong assumption is that the likelihood functon for the data can
+be approximated by a pseudo-likelihood function that is a Gaussian probability density
+in the space of the two-point correlation function estimate.
+That is, the assumption is that a pseudo-likelihood of the data given by a squared distance
+(scaled by an inverse covariance matrix) between an estimate of the correlation function
+(the data) and a theoretical prediction of that same function will return sensible posterior
+inferences.
+This assumption cannot be correct (in detail) at any scale, since a correlation function,
+being related to the variance of a continuous field,
+must satisfy non-trivial positive-definiteness
+requirements (it must take the form of a valid kernel function; CITE SOMETHING).
+These requirements truncate function space such that the likelihood in that functon space
+could never be Gaussian.
+This second assumption becomes truer as the correlation function gets better measured, so
+it is most true on intermediate scales (where neither shot noise nor cosmic variance blow up).
+We expect it to be very strongly violated at the baryon acoustic scale in particular!
+
+Neither of these two assumptions are required for cosmological inferences, however:
+We have very sophisticated simulations and emulations of structure and galaxy formation.
+Indeed, the theory of large-scale structure \emph{is} the output of these simulations;
+we can simulate the data in great detail.
+Therefore, we can simulate not just the one- and two-point statistics of the galaxies,
+but also any higher order statistics that might be relevant.
+Furthermore, we can see the distributions of these statistics, and where the data lie within
+those distributions.
+That is---in principle---we don't need to make these strong and wrong assumptions;
+we can create---in principle---a likelihood function directly from (an enormous number of)
+simulation outputs.
+
+Of course, any naive approach to sufficiently simulating the data would be ruinously
+expensive.
+Fortunately, there are principled, (relatively) efficient methods for minimizing computation
+and delivering correct posterior inferences,
+using only a data simulator and some choices about statistics.
+Here we use \todo{Introduce the names ABC and LFI with references}.
+
+As we will see, the LFI approach---despite cleverness---will be very computationally expensive.
+For context, it is worth remembering that even the traditional approach to large-scale structure,
+outlined above, requires enormous numbers of structure- and galaxy-formation simulations as well,
+to construct the covariance matrices that go into the pseudo-likelihood.
+So although our costs will be high, there are no inexpensive methods to compare to.
+
+\todo{MJ: What Hogg wrote above doesn't mesh with (that is, repeats differently) what you have written below, so can you go through and fix that?
+Lots of what is below should remain, but some should get woven into what's above.}
+
+% style note: Here and in what follows: --- separates parenthetical phrases, -- connects words that appear together in a relationship (like galaxy--galaxy) and (galaxy--halo), and - connects pairs of words into an adjective (like two-point).
 
 Inference of the cosmological parameters and understanding the formation of galaxies within the context of large scale 
-structure, requires understanding of the Galaxy-Halo connection. Halos are biased tracers of the large scale structure 
+structure, requires understanding of the galaxy--halo connection. Halos are biased tracers of the large scale structure 
 and galaxies reside in dark matter Halos.
@@
 The central tenet of halo occupation model is that the probability that a certain number of galaxies populate a dark matter halo is determined by the mass of the host halo.
-This statistical prescription for connection of galaxies and halos has been successful in reproducing the galaxy two-point clustering as well as the galaxy-matter cross correlation---galaxy-galaxy
+This statistical prescription for connection of galaxies and halos has been successful in reproducing the galaxy two-point clustering as well as the galaxy--matter cross correlation---galaxy--galaxy
 lensing---within the context of the halo model, and has been used in understanding the co-evolution of galaxies and halos and putting constraints on the relation between the stellar mass and halo mass of galaxies.
 
-In the context of large scale clustering of galaxies, inference of the cosmological and halo model parameters are performed assuming an analytic model for the two-point correlation function of the galaxies, and a Gaussian likelihood.
+In the context of large scale clustering of galaxies, inference of the cosmological and halo model parameters are performed by treating some estimate of the galaxy--galaxy two-point correlation function as the data,
+assuming an analytic model for the theoretical two-point correlation function prediction,
+and a Gaussian likelihood connecting the two.
 The accuracy of the analytic model of the galaxy clustering statistics hinges upon the accuracy of the fitting functions such as halo bias, halo mass function that are accurate to ten percent level.
@@
 Furthermore, the likelihood of observing the data---observed positions of the galaxies in the configuration space---given the cosmological and halo model parameters is approximated by a Gaussian function whose argument is the squared difference of the two-point correlation function measured from the observations and the analytic model for the galaxy clustering, weighted by a covariance matrix.
+\todo{Why MUST this be wrong?}
 The weighted difference of the number density of the observed galaxies and the analytic number density is also added to this.
@@
 
-ABC is a likelihood free method for inference of a set of parameters that are required for the simulation of a 
-physical system, through a forward model, a distance function, and a prior probability distribution over the parameters. 
-Let us denote the the data by $\mathcal{D}$. We assume that there exist a forward model $\mathcal{M}$ that 
-simulates the data $\bar{\mathcal{D}}$ for a given set of parameters $\bm{\theta}$
+ABC is a generative simulation-based inference based on rejection sampling.
+That is, the parameters of the forward model of the data are drawn from the prior probability, and are rejected if the simulated data is not sufficiently close to the observations according to some distance metric $\rho$.
+The method requires a prior probability over the parameters of interest $\pi(\theta)$, a generative forward model of the data given a set of parameters $f(\theta)$, and a distance metric $\rho(.,.)$ which measures the distance between the simulation $X=f(\theta)$ and the observed data $D$.
+
+In a large scale structure study however, the observed data---which encompasses the angular positions of galaxies on the sky and their redshifts---is a high dimensional object.
+Therefore the distance metric measures the closeness of statistical summaries of the data and those of the simulation.
+When the observed data-set under study is the positions of all galaxies observed in the survey, possible statistical summaries of the data are the number density of galaxies, two-point correlation function, three-point correlation function, galaxy group statistics (counts in cells), and so on.
+
+Here, we provide a brief overview of rejection sampling algorithm and how it is used in a simulation-based inference.
+We want to draw samples from a probability distribution $p(\alpha)$.
+Now, let us assume that: (i) there exists an auxiliary density $q(\alpha)$ from which we can directly sample from, and (ii) there exist $M>1$ which is an appropriate bound on $p(\alpha)/q(\alpha)$.
+Given these assumptions, rejection sampling works in the following way.
+We draw a sample $\alpha$ from the auxiliary density $q(\alpha)$.
+Afterwards, we draw a random number $u$ from a uniform distribution between zero and one.
+Then we accept the draw $\alpha$ if the $u$ is less than the ratio $p(\alpha)/Mq(\alpha)$.
+This procedure is repeated until sufficient number of draws are accepted.
+
+In the context of a simulation-based inference, we aim to sample from the joint probability of a simulation $X$, and parameters $\pars$ given the observed data $D$
+
 \beq
-\bar{\mathcal{D}}_{\theta} = \mathcal{M}(\bm{\theta})
+p(\pars,X | D)
 \eeq
 
-Furthermore, ABC requires adopting a distance function between the data $\mathcal{D}$ and the simulation $\bar{\mathcal{D}}_{\theta}$.
-The distance function, in general, can be a vector $\bm{d}(\mathcal{D},\bar{\mathcal{D}}_{\theta})$
+which according to the Bayes rule can be written as
+
 \begin{eqnarray}
-\bm{d}(\mathcal{D},\bar{\mathcal{D}}_{\theta}) &=& [d_{1,\theta}, ... , d_{M,\theta}], \\
-d_{i,\theta} &=& d_{i}(\mathcal{S}_{i}(\mathcal{D}), \mathcal{S}_{i}(\bar{\mathcal{D}}_{\theta}))  ~~\mbox{for $i=\{1,...,M\}$},
+p(\pars,X | D)  &=& \frac{p(D|X)p(X|\pars)\pi(\pars)}{Z},  \\
+Z &=& \int \dev \pars \; \dev X \; p(D|X)p(X|\pars)\pi(\pars),
 \end{eqnarray}
-where $d_{i}(\dot,\dot)$ is a scalar function that defines a distance 
-between $i$-$th$ summary statistics of the data $\mathcal{S}_{i}(\mathcal{D})$ and the same summary statistics of 
-the simulation $\mathcal{S}_{i}(\bar{\mathcal{D}}_{\theta})$, and $M$ is the total number of statistical summaries 
-used in the analysis.
+where $Z$ is the evidence. There exists a constant $M>1$ 
+\beq 
+M = \frac{\mathrm{max}_{X} \; p(D|X)}{Z},
+\eeq 
+and auxiliary density $q(\pars ,X)$
+\beq 
+q(\pars ,X)=p(X|\pars )\pi(\pars ),
+\eeq 
+such that:
+\beq 
+p(\pars ,X|D)<Mq(\pars ,X).
+\eeq 
+Now in a simulation-based inference, we first draw a pair $\{\pars , X \}$ from the density $q(\pars , X)$.
+This requires drawing a sample $\pars$ from the prior $\pi(\pars)$, and generating a simulation $X=f(\pars)$ via the forward model.
+Then, we accept the pair $\{\pars , X \}$, if the following condition is met
+
+\beq 
+u \sim \mathrm{Uniform}[0,1] < \frac{p(\pars,X|D)}{Mq(\pars,X)}=\frac{p(D|X)}{\mathrm{max}_{X}p(D|X)}.
+\eeq
+
+ABC postulates that the probability of observing data $D$ given simulation $X$, $p(D|X)$ is controlled by a threshold $\epsilon$, and that this probability is one if the distance between the data and the simulation, $\rho(D,X)$ is less than $\epsilon$.
+That is
+
+\beq 
+p(D|X) = p_{\epsilon}(D|X) \equiv \begin{cases}
+                                    1 \ \ \  \mathrm{if} \ \rho(D,X)<\epsilon \; , \\
+                                    0 \ \ \ \mathrm{otherwise} \; .
+                                    \end{cases}
+\eeq
+
+Thus the set of parameters $\pars$ is accepted if $\rho(D,f(\pars))<\epsilon$. In principle, the distance function measures the distance 
+between the statistical summaries of the data and those of the simulation. 
+The distance, in general, can be a vector, with each component representing the distance between one summary statistics of the data (e.g.
+variance) and the same summary statistics of the simulation.
+That is
+\beq 
+\rho(D,X) = \Big[\rho_{1}(S_{1}(D) , S_{1}(X)), ... , \rho_{M}(S_{M}(D) , S_{M}(X)) \Big],
+\eeq 
+where $M$ is the number of statistical summaries used in the inference, $S_{i}$ is the i-th summary statistics, and $\rho_{i}$ is a scalar function that measures the distance between the i-th summary statistics of the data and that of the simulation.
+In this case, the threshold $\epsilon$ is an $M$ dimensional vector.
 
@@
 every component of the distance vector is less than its corresponding threshold. Note that, each summary statistics 
-has a separate threshold. This procedure is done iteratively untill, we are left with a certain number
-of samples that satisfy the distance criteria.
-The final ensemble approximates the posterior probability distribution over the parameters of the generative model of the data.
-This sampling method is called \emph{rejected sampling}.
+has a separate threshold. This procedure is done iteratively until, we are left with a certain number
+of samples that satisfy the distance criteria. The final ensemble approximates the posterior probability distribution over the parameters of the generative model of the data.
 
-Rejected sampling is expensive. Instead we use Population Monte Carlo, adaptive importance sampling, ....
+For a sufficiently small threshold, this algorithm samples from the posterior probability, $p(\pars | D)$, and for an arbitrarily large threshold, the algorithm samples from the prior $\pi(\pars)$.
+This rejection sampling technique requires repeated sampling from the prior, and a very small threshold to succeed.
 
-\subsection{Halo Model}
+In this investigation instead, we use Population Monte Carlo Algorithm (PMC) as our sampling technique.
+PMC is an iterative method that performs rejected sampling from a sequence of distributions $\{p_{1}(\pars),...,p_{T}(\pars)\}$.These distributions are paired with a sequence of decreasing distance thresholds $\{\epsilon_{1} , ...
+, \epsilon_{T}\}$.
 
+At each iteration, $N$ proposals are drawn from a distribution over parameters, and a weight is assigned to each one.
+In the first step, the threshold is set to an arbitrarily large value.
+Therefore the algorithm makes $N$ draws from the prior and assigns equal weights to all the proposals.
+
+In the consequent iterations, the distance threshold is set to a smaller value.
+A proposal is weighted-sampled from the previous sample.
+It is perturbed by a kernel which is set to the covariance of the parameters in the previous sample.
+A simulation is then generated by feeding the perturbed parameter into the forward model.
+The perturbed parameter is then accepted if the distance between the data and the generated simulation is less than the value of the threshold in that iteration.
+This procedure is repeated until $N$ particles are sampled.
+Then the weights are updated according to the kernel, the prior, and the previous set of weights.
+
+\subsection{Forward model of the large scale galaxy populations}\label{sec:halo}
+\subsubsection{halo model}
 Halo Occupation distribution is a powerful tool for description of galaxy/halo relation. \todo{Here we explain HOD modeling of galaxy positions, etc.}
+\subsubsection{Generating mock galaxy populations with halotools}
 
-...
-
-\subsection{Summary statistics}
-\begin{itemize}
-
-\item \todo{Talk about the observables: 2pcf, group multiplicity function, and number density ...}
-
-\end{itemize}
-
-\section{Forward modeling of the galaxy population with HALOTOOLS}\label{sec:halo}
-
-\begin{itemize}
-\item \todo{Halotools allows us to forward model galaxy positions from HOD parameters.
+\todo{Halotools allows us to forward model galaxy positions from HOD parameters.
 From the galaxy positions, we can calculate observable summary statistics such as the 2PCF, Group Multiplicity Function (GMF) and the number density $\bar{n}$.}
 
-\end{itemize}
 Talk about why it is better to simulate the mock and compute the summary statistics of the mock instead of 
@@
 Talk about why we do not want to write down likelihood function for gmf, 2pcf, ...
+\subsubsection{summary statistics of galaxy populations}
 
-...
+\subsubsection{distance metrics}
 
-\subsection{Mock data}
+\todo{Talk about the observable: 2pcf, group multiplicity function, and number density ...}
+
+\section{Mock data}\label{sec:data}
+
 Plots showing 2pcf, gmf, ... of the mock data.
@@
 \todo{Apples to apples comparison between ABC and MCMC.}
-\begin{itemize}
-\item ABC-PMC using nbar and xi vs MCMC
-\cite{Ishida:2015aa}
-\end{itemize}
+ABC-PMC using nbar and xi vs MCMC
 

### 2016-01-20 halo/code/data.py
reduce covariance plot for xi in test_data.py; xi r_bins coded in data.py

@@ -3,6 +3,7 @@
 Modules for data and observables in ABC-PMC. 
 
 '''
+import os 
 import numpy as np
 from group_richness import gmf_bins
 from group_richness import richness 
@@ -43,6 +44,12 @@ def data_xi(Mr=20, Nmock=500):
 
     return [xi, cii]
 
+def data_xi_bin(Mr=20):
+    ''' r bins for xi(r)
+    '''
+    rbin_file = ''.join(['../dat/xir_rbin.Mr', str(Mr), '.dat'])
+    return np.loadtxt(rbin_file)
+
 def data_xi_cov(Mr=20, Nmock=500): 
     '''
     Observed xi covariance. The entire covariance matrix
@@ -55,7 +62,6 @@ def data_xi_cov(Mr=20, Nmock=500):
     
 
 # Build observables ---------------
-
 def build_xi_nbar_gmf(Mr=20): 
     '''
     Build "data" xi, nbar, GMF values and write to file 
@@ -80,6 +86,17 @@ def build_xi_nbar_gmf(Mr=20):
     np.savetxt(output_file, gmf) 
 
     return None 
+
+def build_xi_bin(Mr=20): 
+    ''' Write out xi r bins
+    '''
+    model = PrebuiltHodModelFactory('zheng07', threshold = -1.0*np.float(Mr))
+    model.populate_mock() # population mock realization 
+
+    # write xi 
+    r_bin  = model.mock.compute_galaxy_clustering()[0]
+    output_file = ''.join(['../dat/xir_rbin.Mr', str(Mr), '.dat'])
+    np.savetxt(output_file, r_bin)
 
 def build_xi_cov(Mr=20, Nmock=500): 
     '''
@@ -147,6 +164,7 @@ def build_observations(Mr=20, Nmock=500):
     # xi, nbar, gmf
     print 'Building xi(r), nbar, GMF ... ' 
     build_xi_nbar_gmf(Mr=Mr)
+    build_xi_bin(Mr=Mr)
     
     # covariances
     print 'Building xi covariance ... ' 

### 2016-01-20 halo/code/data.py
stupid bug in data.py fixed

@@ -98,65 +98,60 @@ def build_xi_bin(Mr=20):
     output_file = ''.join(['../dat/xir_rbin.Mr', str(Mr), '.dat'])
     np.savetxt(output_file, r_bin)
 
-def build_xi_cov(Mr=20, Nmock=500): 
+def build_xi_nbar_gmf_cov(Mr=20, Nmock=500): 
     '''
-    Build covariance matrix for xi, using Nmock simulations 
+    Build covariance matrix for xi, variance for nbar, and a bunch of stuff for gmf 
+    ...  
+    using Nmock realizations of halotool mocks  
     ''' 
     xir = [] 
-    for i in xrange(Nmock): 
-        model = PrebuiltHodModelFactory('zheng07', threshold = -1.0*np.float(Mr))
-        model.populate_mock() 
-
-        xir.append(model.mock.compute_galaxy_clustering()[1])
-
-    covar = np.cov(np.array(xir).T)
-    output_file = ''.join(['../dat/xir_covariance.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
-
-    np.savetxt(output_file, covar)
-    
-    return None
-
-def build_nbar_cov(Mr=20, Nmock=500): 
-    ''' Build observed nbar value
-    '''
-    model = PrebuiltHodModelFactory('zheng07', threshold = -1.0*np.float(Mr))
     nbars = [] 
-    for i in xrange(Nmock): 
-        model.populate_mock() 
-        nbars.append(model.mock.number_density)
-    
-    nbar_cov = np.var(nbars, axis=0) 
-
-    # save nbar values 
-    output_file = ''.join(['../dat/nbar_cov.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
-    np.savetxt(output_file, [nbar_cov]) 
-    return None
-
-    
-def build_gmf_sigma(Mr=20, Nmock=500): 
-    ''' Build 'observed' uncertainty in GMF from Nmock simulated realizations. 
-    Uncertainty is the quadruture of poisson errors and stddev. 
-    '''
-    model = PrebuiltHodModelFactory('zheng07', threshold = -1.0*np.float(Mr))
-    
     gmfs = [] 
     gmf_counts = [] 
     for i in xrange(Nmock): 
+        model = PrebuiltHodModelFactory('zheng07', threshold = -1.0*np.float(Mr))
[…]

### 2016-01-20 halo/code/data.py
no significant change; committing as a formality

@@ -166,5 +166,5 @@ def build_observations(Mr=20, Nmock=500):
     build_xi_nbar_gmf_cov(Mr=Mr, Nmock=Nmock)
 
 if __name__=='__main__': 
-    build_observations(Nmock=2)
+    build_observations(Nmock=5)
 

### 2016-01-21 halo/code/abc_pemcee.py
ABC-PMC updated

@@ -13,20 +13,30 @@ from group_richness import richness
 # --- Plotting ---
 from plotting import plot_thetas
 
-def abcpmc_nbar_gmf(T, eps_val, Mr=20, N_part=1000, threads=1, observables=['nbar', 'gmf']): 
+def abcpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data_dict={'Mr':20, 'Nmock': 500}):
     '''
+    ABC-PMC implementation. 
+
+    Parameters
+    ----------
+    - T : Number of iterations 
+    - eps_val : ???? (@mjv: what does this do?)
+    - N_part : Number of particles
+    - Threads : Number of MPI threads (not sure if this keyword actually works or not 
+    - observables : list of observables. Options are 'nbar', 'gmf', 'xi'
+    - data_dict : dictionary that specifies the observation keywords 
     '''
     # data observables
     data = []       # list of observables 
     for obv in observables: 
         if obv == 'nbar': 
-            data_nbar, data_nbar_var = Data.data_nbar(Mr=Mr)
+            data_nbar, data_nbar_var = Data.data_nbar(**data_dict)
             data.append(data_nbar)
         if obv == 'gmf': 
-            data_gmf, data_gmf_sigma = Data.data_gmf(Mr=Mr)
+            data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
             data.append(data_gmf)
         if obv == 'xi': 
-            data_xi, data_cov_ii = Data.data_xi(Mr=Mr)
+            data_xi, data_cov_ii = Data.data_xi(**data_dict)
             data.append(data_xi)
     
     # True HOD parameters
@@ -46,8 +56,8 @@ def abcpmc_nbar_gmf(T, eps_val, Mr=20, N_part=1000, threads=1, observables=['nba
     # simulator
     our_model = HODsim()    # initialize model
     kwargs = {'prior_range': prior_range, 'observables': observables}
-    def simz(theta): 
-        return our_model.sum_stat(theta, **kwargs)
+    def simz(tt): 
+        return our_model.sum_stat(tt, **kwargs)
 
     def multivariate_rho(datum, model): 
         dists = [] 
@@ -60,6 +70,7 @@ def abcpmc_nbar_gmf(T, eps_val, Mr=20, N_part=1000, threads=1, observables=['nba
[…]

### 2016-01-21 halo/code/abc_pemcee.py
fixed a minor bug in the new abcpmc

@@ -79,12 +79,11 @@ def abcpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data
             Y=data,         # data
             postfn=simz,    # simulator 
             dist=multivariate_rho,       # distance function  
-            threads=threads,
             pool=mpi_pool)  
     abcpmc_sampler.particle_proposal_cls = abcpmc.ParticleProposal
 
     #eps = abcpmc.ConstEps(T, [1.e13,1.e13])
-    eps = abcpmc.ConstEps(T, eps_val)       # (@mjv: what does this do?)
+    eps = abcpmc.MultiConstEps(T, eps_val)
     pools = []
     for pool in abcpmc_sampler.sample(prior, eps):
         print("T:{0},ratio: {1:>.4f}".format(pool.t, pool.ratio))
@@ -117,4 +116,4 @@ def abcpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data
     return pools
 
 if __name__=="__main__": 
-    abcpmc(10, 60, N_part=100)
+    abcpmc(10, [1.e10,1.e10], N_part=100)

### 2016-01-21 halo/code/abc_pemcee.py
fixed some bugs on abc-pemcee code

@@ -13,7 +13,7 @@ from group_richness import richness
 # --- Plotting ---
 from plotting import plot_thetas
 
-def abcpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data_dict={'Mr':20, 'Nmock': 500}):
+def ABCpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data_dict={'Mr':20, 'Nmock': 500}):
     '''
     ABC-PMC implementation. 
 
@@ -117,4 +117,4 @@ def abcpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data
     return pools
 
 if __name__=="__main__": 
-    abcpmc(10, 60, N_part=100)
+    ABCpmc(10, 60, N_part=100)

### 2016-01-21 halo/text/abc.tex

@@
 \usepackage{hyperref}
+\usepackage{algorithmic,algorithm}
 \hypersetup{colorlinks}
@@
 
+\begin{algorithm}
+\caption{The procedure for ABC-PMC}
+\begin{algorithmic}[1]
+%\STATE \DATA: D
+%\STATE \RESULT: ABC posterior sample of $\pars$
+\IF{$t=1:$}
+%\STATE $\epsilon_t \gets \infty$
+\FOR{$i=1,...,N$}
+   \STATE // \emph{This loop can now be done in parallel for all i}
+   \WHILE{$\rho(X,D)>\epsilon_t$}
+   \STATE $\pars^{*}_{t} \gets \pi(\pars)$
+   \STATE $X = f(\pars^{*}_{t})$
+   \ENDWHILE
+   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
+   \STATE $w^{(i)}_{t} \gets 1/N$
+\ENDFOR
+\ENDIF
+\IF{$t=2,...,T:$}
+\FOR{$i=1,...,N$}
+   \STATE // \emph{This loop can now be done in parallel for all i}
+   \WHILE{$\rho(X,D)>\epsilon_t$}
+   \STATE Draw $\pars^{*}_{t}$ from $\{\pars^{j}_{t-1}\}$ with probabilities $\{w^{j}_{t-1}\}$
+   \STATE $\pars^{*}_{t} \gets K(\pars^{*}_{t},.)$
+   \STATE $X = f(\pars^{*}_{t})$
+   \ENDWHILE
+   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
+   \STATE $w^{(i)}_{t} \gets \frac{\pi(\pars^{(i)}_{t})}{\sum_{j=1}^{N}w_{t-1}^{(i)}K(\pars^{(j)}_{t-1},\pars^{(i)}_{t})}$
+\ENDFOR
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+%\begin{algorithm}[H]
+% \Data{this text}
+% \Result{how to write algorithm with \LaTeX2e }
+% initialization\;
+% \While{not at end of this document}{
+%  read current\;
+%  \eIf{understand}{
+%   go to next section\;
+%   current section becomes this one\;
+%   }{
+%   go back to the beginning of current section\;
+%  }
+% }
+% \caption{How to write algorithms}
+%\end{algorithm}
+
 \subsection{Forward model of the large scale galaxy populations}\label{sec:halo}
 \subsubsection{halo model}
-Halo Occupation distribution is a powerful tool for description of galaxy/halo relation. \todo{Here we explain HOD modeling of galaxy positions, etc.}
+Dark matter halos are biased tracers of the underlying matter distribution of the universe.
+The halo model postulates that galaxies reside in dark matter halos.
+Halo Occupation distribution---HOD---is a powerful tool for description of galaxy--halo connection.
+HOD provides a probabilistic prescription for how galaxies populate halos.
+According to HOD modeling, the number of galaxies that populate a dark matter halo is dictated by the host halo mass $M_h$.
+The probability of a dark matter halo of mass $M_h$ being populated by $N_g$ galaxies---$P(N_g|M_h)$--- is a Poisson distribution, and is separable to two parts
+\beq
+P(N_g|M_h) = P(N_\mathrm{cen}|M_h) + P(N_\mathrm{sat}|M_h),
+\eeq
+where $N_\mathrm{cen}$ is the number of central galaxies, and $N_\mathrm{sat}$ is the number of satellites.
+
 \subsubsection{Generating mock galaxy populations with halotools}
 
-\todo{Halotools allows us to forward model galaxy positions from HOD parameters.
+%\todo{Halotools allows us to forward model galaxy positions from HOD parameters.
 From the galaxy positions, we can calculate observable summary statistics such as the 2PCF, Group Multiplicity Function (GMF) and the number density $\bar{n}$.}
 
-Talk about why it is better to simulate the mock and compute the summary statistics of the mock instead of 
-using analytic formulas...
+%Talk about why it is better to simulate the mock and compute %the summary statistics of the mock instead of using analytic formulas...
 
-Talk about all the assumptions that go into analytic formulas...
+%Talk about all the assumptions that go into analytic %formulas...
 
-Talk about why we do not want to write down likelihood function for gmf, 2pcf, ...
+%Talk about why we do not want to write down likelihood %function for gmf, 2pcf, ...
 \subsubsection{summary statistics of galaxy populations}
@@
 
-\todo{Talk about the observable: 2pcf, group multiplicity function, and number density ...}
+%\todo{Talk about the observable: 2pcf, group multiplicity %function, and number density ...}
 
-\section{Mock data}\label{sec:data}
-
-Plots showing 2pcf, gmf, ... of the mock data.
-How we compute the covariance matrix from mocks...
-Talk about error bars. Talk about poisson errors in group populations ...
+\section{Mock Observations}\label{sec:data}
+\todo{The fake observations we use as "data" in our inference.}
+%Plots showing 2pcf, gmf, ... of the mock data.
+%How we compute the covariance matrix from mocks...
+%Talk about error bars. Talk about poisson errors in group %populations ...
 
@@
 
-\todo{talk about why it is hard to do cosmology ...}
-
 talk about halo model ...

### 2016-01-21 halo/code/emcee_emcee.py
MC-MC code for updated format of package

@@ -1,0 +1,80 @@
+import sys
+import numpy as np
+import emcee
+from halotools.empirical_models import Zheng07
+from numpy.linalg import solve
+from emcee.utils import MPIPool
+
+#import matplotlib.pyplot as plt
+#import time
+#plt.switch_backend("Agg")
+
+"""data, covariance matrix, and the inverse covariance """
+
+xir_data = np.loadtxt("xir_Mr20.dat")
+covariance = np.loadtxt("clustering_covariance_Mr20.dat")
+N_mocks = 500
+N_bins  = len(xir_data)
+inv_c =  solve(np.eye(len(xir_data)) , covariance)*(N_mocks - 2 - N_bins)/(N_mocks - 1)
+
+
+
+"""likelihood"""
+
+def lnlike(theta):
+    
+    model = Zheng07(threshold = -20.)
+    model.param_dict['logM0'] = theta[0]
+    model.param_dict['sigma_logM'] = np.exp(theta[1])
+    model.param_dict['logMmin'] = theta[2]     
+    model.param_dict['alpha'] = theta[3]
+    model.param_dict['logM1'] = theta[4]
+    model.populate_mock()
+    r , xir_model = model.mock.compute_galaxy_clustering()
+    res = xir_data - xir_model
+    return -0.5*np.sum(np.dot(np.dot(res , inv_c) , res))
+
+"""prior"""
+
+def lnprior(theta):
+    a , b , c , d , e = theta
+    if 10. < a < 13. and np.log(.1) < b < np.log(.7) and  11.02< c < 13.02 and .8< d < 1.3 and 13.< e < 14.:
+        return 0.0
+    return -np.inf
+
+"""posterior"""
+
+def lnprob(theta):
+    lp = lnprior(theta)
+    if not np.isfinite(lp):
+        return -np.inf
+    return lp + lnlike(theta)
+
+"""Initializing Walkers"""
+
+ndim, nwalkers = 5, 100
+pos = [np.array([11.5 , np.log(.4) , 12.02 , 1.03 , 13.5]) + 1e-4*np.random.randn(ndim) for i in range(nwalkers)]
+
+"""Initializing MPIPool"""
+
+pool = MPIPool()
+if not pool.is_master():
+    pool.wait()
+    sys.exit(0)
+
+sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob , pool = pool)
+
[…]

### 2016-01-21 halo/code/abc_pemcee.py
updates to data.py to include inverse covariance matrix calculation

@@ -1,3 +1,11 @@
+'''
+
+Module for ABC-PMC inference
+
+Author(s): Chang, MJ
+
+
+'''
 import time
 import numpy as np
 
@@ -6,7 +14,6 @@ from abcpmc import mpi_util
 
 # --- Local --- 
 import data as Data
-#from distance import rho 
 from hod_sim import HODsim
 from group_richness import richness
 

### 2016-01-21 halo/code/data.py

@@ -59,6 +59,23 @@ def data_xi_cov(Mr=20, Nmock=500):
     cov = np.loadtxt(cov_dat_file)
 
     return cov 
+
+def data_xi_inv_cov(Mr=20, Nmock=500, unbias_str=True):
+    '''
+    Observed inverse covariance of xi with/without the unbias estimator 
+    factor. Default multiplies by the unbias estimator factor. 
+    '''
+    if unbias_str: 
+        unbias_str = '.unbias'
+    else: 
+        unbias_str = ''
+
+    inv_cov_file = ''.join(['../dat/', 
+        'xi_inv_cov', unbias_str, '.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
+
+    inv_cov = np.loadtxt(inv_cov_file) 
+
+    return cov
     
 
 # Build observables ---------------
@@ -148,10 +165,33 @@ def build_xi_nbar_gmf_cov(Mr=20, Nmock=500):
     sigma_tot = (sigma_gmf**2 + poisson_gmf**2)**0.5    # total sigma
     output_file = ''.join(['../dat/gmf_sigma.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
     np.savetxt(output_file, sigma_tot) 
-
-    
     return None
 
+def build_xi_inv_cov(Mr=20, Nmock=500, unbias=True): 
+    '''
+    Calculate the inverse covariance of xi multiplied by the unbiased
+    estimator factor (Nmocks - 2 - Nbins)/(Nmocks - 1). 
+
+    Mainly used in MCMC inference
+    '''
+    xi_cov = data_xi_cov(Mr=Mr, Nmock=Nmock)    # covariance matrix of xi
+
+    N_bins = int(np.sqrt(len(xi_cov)))          # cov matrix is N_bin x N_bin
+    
+    if unbias: 
+        f_unbias = np.float(Nmocks - 2. - N_bins)/np.float(Nmocks - 1.)
+        unbias_str = '.unbias'
+    else: 
+        f_unbias = 1.0 
+        unbias_str = ''
+
+    inv_c = solve(np.eye(len(xir_data)) , covariance) * f_unbias 
+
+    output_file = ''.join(['../dat/', 
+        'xi_inv_cov', unbias_str, '.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
+    np.savetxt(output_file, inv_c) 
+
+    return None
 
[…]

### 2016-01-21 halo/code/emcee_emcee.py

@@ -1,13 +1,23 @@
+'''
+
+
+Module for standard MCMC inference
+
+Author(s): Chang, MJ
+
+
+'''
 import sys
 import numpy as np
 import emcee
-from halotools.empirical_models import Zheng07
+
 from numpy.linalg import solve
 from emcee.utils import MPIPool
 
-#import matplotlib.pyplot as plt
-#import time
-#plt.switch_backend("Agg")
+# --- Local ---
+import data as Data
+from hod_sim import HODsim
+from group_richness import richness
 
 """data, covariance matrix, and the inverse covariance """
 

### 2016-01-21 halo/code/hod_sim.py
threshold keyword option included in HODsim

@@ -13,14 +13,14 @@ from group_richness import richness
 
 class HODsim(object): 
     
-    def __init__(self): 
+    def __init__(self, Mr=20): 
         '''
         Class object that describes our forward model used in AMC-PMC inference.
         Our model forward models the galaxy catalog using HOD parameters using HaloTools. 
         '''
 
         #self.model = zheng07()      # Zheng et al. (2007) model
-        self.model = PrebuiltHodModelFactory('zheng07')
+        self.model = PrebuiltHodModelFactory('zheng07', threshold=-1*Mr)
     
     def sum_stat(self, theta, prior_range=None, observables=['nbar', 'gmf']):
         '''

### 2016-01-21 halo/code/data.py
inverse covariance calcuation tested on Nmock=5

@@ -5,6 +5,8 @@ Modules for data and observables in ABC-PMC.
 '''
 import os 
 import numpy as np
+from numpy.linalg import solve
+
 from group_richness import gmf_bins
 from group_richness import richness 
 from group_richness import gmf as GMF
@@ -175,17 +177,16 @@ def build_xi_inv_cov(Mr=20, Nmock=500, unbias=True):
     Mainly used in MCMC inference
     '''
     xi_cov = data_xi_cov(Mr=Mr, Nmock=Nmock)    # covariance matrix of xi
-
-    N_bins = int(np.sqrt(len(xi_cov)))          # cov matrix is N_bin x N_bin
+    N_bins = int(np.sqrt(xi_cov.size))          # cov matrix is N_bin x N_bin
     
     if unbias: 
-        f_unbias = np.float(Nmocks - 2. - N_bins)/np.float(Nmocks - 1.)
+        f_unbias = np.float(Nmock - 2. - N_bins)/np.float(Nmock - 1.)
         unbias_str = '.unbias'
     else: 
         f_unbias = 1.0 
         unbias_str = ''
-
-    inv_c = solve(np.eye(len(xir_data)) , covariance) * f_unbias 
+    
+    inv_c = solve(np.eye(N_bins) , xi_cov) * f_unbias 
 
     output_file = ''.join(['../dat/', 
         'xi_inv_cov', unbias_str, '.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])

### 2016-01-21 halo/code/abc_pemcee.py
modified the new abcpmc, but still needs some work

@@ -16,11 +16,11 @@ from abcpmc import mpi_util
 import data as Data
 from hod_sim import HODsim
 from group_richness import richness
-
+import util
 # --- Plotting ---
 from plotting import plot_thetas
 
-def ABCpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data_dict={'Mr':20, 'Nmock': 500}):
+def ABCpmc(T, eps_val, N_part=1000, observables=['nbar', 'gmf'], data_dict={'Mr':20, 'Nmock': 500}):
     '''
     ABC-PMC implementation. 
 
@@ -34,20 +34,21 @@ def ABCpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data
     - data_dict : dictionary that specifies the observation keywords 
     '''
     # data observables
-    data = []       # list of observables 
+    fake_obs = []       # list of observables 
     for obv in observables: 
         if obv == 'nbar': 
             data_nbar, data_nbar_var = Data.data_nbar(**data_dict)
-            data.append(data_nbar)
+            print data_nbar
+            fake_obs.append(data_nbar)
         if obv == 'gmf': 
             data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
-            data.append(data_gmf)
+            fake_obs.append(data_gmf)
         if obv == 'xi': 
             data_xi, data_cov_ii = Data.data_xi(**data_dict)
-            data.append(data_xi)
-    
+            fake_obs.append(data_xi)
+
     # True HOD parameters
-    if Mr == 20: 
+    if data_dict["Mr"] == 20: 
         data_hod = np.array([11.38 , np.log(0.26) , 12.02 , 1.06 , 13.31])
     else: 
         raise NotImplementedError
@@ -69,6 +70,8 @@ def ABCpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data
     def multivariate_rho(datum, model): 
         dists = [] 
         for i_obv, obv in enumerate(observables): 
+            print "datum=" , datum
+	    print "model=" , model
             if obv == 'nbar': 
                 dist_nz = (datum[i_obv] - model[i_obv])**2. / data_nbar_var 
                 dists.append(dist_nz)
@@ -83,7 +86,7 @@ def ABCpmc(T, eps_val, N_part=1000, threads=1, observables=['nbar', 'gmf'], data
[…]

### 2016-01-21 halo/code/data.py
data.py and test_data.py edited in order to produce covariance matrix plots

@@ -24,6 +24,13 @@ def data_gmf_bins():
     '''
     return gmf_bins()
 
+def data_gmf_cov(Mr=20, Nmock=500): 
+    '''
+    GMF covariance matrix 
+    '''
+    cov_file = ''.join(['../dat/gmf_cov.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
+    return np.loadtxt(cov_file)
+
 def data_nbar(Mr=20, Nmock=500): 
     '''
     Observed nbar from 'data'

### 2016-01-21 halo/code/test_data.py

@@ -7,6 +7,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 
 import data 
+import util 
 
 def xi_cov(Mr=20, Nmock=500):
     '''Plot the reduced xi covariance
@@ -34,7 +35,42 @@ def xi_cov(Mr=20, Nmock=500):
 
     sub.set_xlabel(r'$r\;(\mathrm{Mpc}/h)$', fontsize=25)
     sub.set_ylabel(r'$r\;(\mathrm{Mpc}/h)$', fontsize=25)
-    plt.show()
+    fig_file = ''.join([util.fig_dir(),
+        'xi_covariance.Mr', str(Mr), '.Nmock', str(Nmock), '.png'])
+    fig.savefig(fig_file, bbox_inches='tight') 
+    plt.close()
+
+
+def gmf_cov(Mr=20, Nmock=500):
+    '''Plot the reduced xi covariance
+    '''
+    # covariance  
+    gmf_cov = data.data_gmf_cov(Mr=Mr, Nmock=Nmock)
+    n_bin = int(np.sqrt(gmf_cov.size))
+    x_bin = data.data_gmf_bins()
+    
+    # calculate the reduced covariance for plotting
+    gmf_cov_red = np.zeros([n_bin, n_bin])
+    for ii in range(n_bin): 
+        for jj in range(n_bin): 
+            gmf_cov_red[ii][jj] = gmf_cov[ii][jj]/np.sqrt(gmf_cov[ii][ii] * gmf_cov[jj][jj])
+
+    fig = plt.figure()
+    sub = fig.add_subplot(111)
+    cont = sub.pcolormesh(x_bin, x_bin, gmf_cov_red, cmap=plt.cm.afmhot_r)
+    plt.colorbar(cont)
+
+    sub.set_xlim([x_bin[0], x_bin[-1]])
+    sub.set_ylim([x_bin[0], x_bin[-1]])
+
+    sub.set_xlabel(r'$\mathtt{Group\;\;Richness}$', fontsize=25)
+    sub.set_ylabel(r'$\mathtt{Group\;\;Richness}$', fontsize=25)
+    
+    fig_file = ''.join([util.fig_dir(),
+        'gmf_covariance.Mr', str(Mr), '.Nmock', str(Nmock), '.png'])
+    fig.savefig(fig_file, bbox_inches='tight') 
+    plt.close()
 
[…]

### 2016-01-21 halo/code/data.py
test_data and data adjusted to include plotting fake obs xi; prettyplot and prettycolor added to plotting.py

@@ -49,7 +49,7 @@ def data_xi(Mr=20, Nmock=500):
     # load covariance of xi 
     cov_dat_file = ''.join(['../dat/xir_covariance.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
     cov = np.loadtxt(cov_dat_file)
-    cii = np.diag(covariance)   # diagonal elements
+    cii = np.diag(cov)   # diagonal elements
 
     return [xi, cii]
 

### 2016-01-21 halo/code/plotting.py

@@ -3,9 +3,15 @@
 Plotting modules 
 
 '''
-import util
 import corner
+import numpy as np
+import matplotlib as mpl
 import matplotlib.pyplot as plt
+from matplotlib import rc
+from matplotlib.ticker import AutoMinorLocator, MultipleLocator
+
+# --- local ---
+import util
 plt.switch_backend("Agg")
 
 def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=None): 
@@ -56,4 +62,57 @@ def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=N
     plt.savefig(fig_file)
     plt.close()
 
+def prettyplot():
+    '''
+    Some settings to make pretty plots
+    '''
+    # Use Latex
+    mpl.rcParams['text.usetex']=True
+    mpl.rcParams['text.latex.preamble']=[r"\usepackage[T1]{fontenc}",
+    r"\usepackage{cmbright}",]
+    # Set Major tick size and width
+    mpl.rcParams['xtick.major.size']=10
+    mpl.rcParams['xtick.major.width']=2.5
+    mpl.rcParams['ytick.major.size']=10
+    mpl.rcParams['ytick.major.width']=2.5
+    # Set minor tick size and wdith
+    mpl.rcParams['ytick.minor.size']=3
+    mpl.rcParams['ytick.minor.width']=1.5
+    mpl.rcParams['xtick.minor.size']=3
+    mpl.rcParams['xtick.minor.width']=1.5
+    # Set space between axes and tick label
+    mpl.rcParams['xtick.major.pad']=12
+    mpl.rcParams['ytick.major.pad']=12
+    mpl.rcParams['xtick.labelsize']='large'
+    mpl.rcParams['ytick.labelsize']='large'
+    mpl.rcParams['axes.linewidth']=2.5
+
+    mpl.rcParams['font.family']='monospace'
+    mpl.rcParams['font.monospace']='Courier'
+    mpl.rcParams['font.weight']=800
+    mpl.rcParams['font.size']=16
+    # legend settings
+    mpl.rcParams['legend.frameon']=False
+    mpl.rcParams['legend.markerscale']=5.0
+    mpl.rcParams['axes.xmargin']=1
+    return None
+
+def prettycolors(): 
+    ''' preset pretty colors
+    '''
+    # Tableau20 colors 
+    pretty_colors = [
[…]

### 2016-01-21 halo/code/test_data.py

@@ -6,16 +6,50 @@ Test data.py module
 import numpy as np 
 import matplotlib.pyplot as plt
 
-import data 
 import util 
+import data as Data
+from plotting import prettyplot
+from plotting import prettycolors
+
+def xi(Mr=20, Nmock=500): 
+    '''
+    Plot xi(r) of the fake observations
+    '''
+    prettyplot() 
+    pretty_colors = prettycolors() 
+
+    xir, cii = Data.data_xi(Mr=Mr, Nmock=Nmock)
+    rbin = Data.data_xi_bin(Mr=Mr) 
+    
+    fig = plt.figure(1) 
+    sub = fig.add_subplot(111)
+    sub.plot(rbin, rbin*xir, c='k', lw=1)
+    sub.errorbar(rbin, rbin*xir, yerr = rbin*cii**0.5 , fmt="ok", ms=1, capsize=2, alpha=1.)
+
+    sub.set_xlim([0.1, 15])
+    sub.set_ylim([1, 10])
+    sub.set_yscale("log")
+    sub.set_xscale("log")
+
+    sub.set_xlabel(r'$\mathtt{r}\; (\mathtt{Mpc})$', fontsize=25)
+    sub.set_ylabel(r'$\mathtt{r} \xi_{\rm gg}$', fontsize=25)
+
+    fig_file = ''.join([util.fig_dir(),
+        'xi.Mr', str(Mr), '.Nmock', str(Nmock), '.png'])
+    fig.savefig(fig_file, bbox_inches='tight')
+    plt.close()
+    return None
 
 def xi_cov(Mr=20, Nmock=500):
     '''Plot the reduced xi covariance
     '''
+    prettyplot() 
+    pretty_colors = prettycolors() 
+
     # covariance  
-    xi_cov = data.data_xi_cov(Mr=Mr, Nmock=Nmock)
+    xi_cov = Data.data_xi_cov(Mr=Mr, Nmock=Nmock)
     n_bin = int(np.sqrt(xi_cov.size))
-    r_bin = data.data_xi_bin(Mr=Mr)
+    r_bin = Data.data_xi_bin(Mr=Mr)
     
     # calculate the reduced covariance for plotting
     xi_cov_red = np.zeros([n_bin, n_bin])
@@ -40,14 +74,16 @@ def xi_cov(Mr=20, Nmock=500):
     fig.savefig(fig_file, bbox_inches='tight') 
     plt.close()
[…]

### 2016-01-21 halo/code/util.py

@@ -14,7 +14,7 @@ def fig_dir():
     '''
     figure directory
     '''
-    return ''
+    return '../fig/'
 
 def dat_dir(): 
     '''

### 2016-01-21 halo/code/data.py
GMF plot in test_data.py

@@ -12,6 +12,13 @@ from group_richness import richness
 from group_richness import gmf as GMF
 from halotools.empirical_models import PrebuiltHodModelFactory
 
+def data_hod_param(Mr=20):
+    '''
+    HOD parameters of 'observations'. Returns dictionary with hod parameters. 
+    '''
+    model = PrebuiltHodModelFactory('zheng07', threshold = -1.0*np.float(Mr))
+    return model.param_dict
+
 def data_gmf(Mr=20, Nmock=500): 
     ''' Observed GMF from 'data'
     '''

### 2016-01-21 halo/code/test_data.py

@@ -74,6 +74,35 @@ def xi_cov(Mr=20, Nmock=500):
     fig.savefig(fig_file, bbox_inches='tight') 
     plt.close()
 
+def gmf(Mr=20, Nmock=500): 
+    '''
+    Plot Group Multiplicty Function of fake observations 
+    '''
+    prettyplot() 
+    pretty_colors = prettycolors() 
+    
+    # import fake obs GMF
+    gmf, sig_gmf = Data.data_gmf(Mr=Mr, Nmock=Nmock)
+    # group richness bins
+    gmf_bin = Data.data_gmf_bins()
+
+    fig = plt.figure(1) 
+    sub = fig.add_subplot(111)
+
+    sub.errorbar(
+            0.5*(gmf_bin[:-1]+gmf_bin[1:]), gmf, yerr=sig_gmf,
+            fmt="ok", capsize=1.0
+            )
+    sub.set_xlim([1, 60])
+    sub.set_yscale('log')
+    sub.set_ylabel(r"Group Multiplicity Function (h$^{3}$ Mpc$^{-3}$)", fontsize=20)
+    sub.set_xlabel(r"$\mathtt{Group\;\;Richness}$", fontsize=20)
+    # save to file 
+    fig_file = ''.join([util.fig_dir(), 
+        'gmf.Mr', str(Mr), '.Nmock', str(Nmock), '.png'])
+    fig.savefig(fig_file, bbox_inches='tight')
+    return None
+
 def gmf_cov(Mr=20, Nmock=500):
     '''Plot the reduced xi covariance
     '''
@@ -108,6 +137,4 @@ def gmf_cov(Mr=20, Nmock=500):
     plt.close()
 
 if __name__=='__main__': 
-    xi(Mr=20, Nmock=500)
-    xi_cov(Mr=20, Nmock=500)
-    gmf_cov(Mr=20, Nmock=500)
+    gmf(Mr=20, Nmock=500)

### 2016-01-21 halo/code/test_data.py
test data corrected for prettier plot

@@ -67,8 +67,8 @@ def xi_cov(Mr=20, Nmock=500):
     sub.set_xscale('log')
     sub.set_yscale('log')
 
-    sub.set_xlabel(r'$r\;(\mathrm{Mpc}/h)$', fontsize=25)
-    sub.set_ylabel(r'$r\;(\mathrm{Mpc}/h)$', fontsize=25)
+    sub.set_xlabel(r'$\mathtt{r}\;(\mathtt{Mpc}/$h)', fontsize=25)
+    sub.set_ylabel(r'$\mathtt{r}\;(\mathtt{Mpc}/$h)', fontsize=25)
     fig_file = ''.join([util.fig_dir(),
         'xi_covariance.Mr', str(Mr), '.Nmock', str(Nmock), '.png'])
     fig.savefig(fig_file, bbox_inches='tight') 
@@ -137,4 +137,4 @@ def gmf_cov(Mr=20, Nmock=500):
     plt.close()
 
 if __name__=='__main__': 
-    gmf(Mr=20, Nmock=500)
+    xi_cov(Mr=20, Nmock=500)

### 2016-01-21 halo/code/data.py
bug in data_nbar()

@@ -43,7 +43,7 @@ def data_nbar(Mr=20, Nmock=500):
     Observed nbar from 'data'
     '''
     nbar = np.loadtxt(''.join(['../dat/nbar.Mr', str(Mr), '.dat']))
-    nbar_cov = np.loadtxt(''.join(['../dat/nbar_cov.Mr', str(Mr), '.Nock', str(Nmock), '.dat']))
+    nbar_cov = np.loadtxt(''.join(['../dat/nbar_cov.Mr', str(Mr), '.Nmock', str(Nmock), '.dat']))
     return [nbar, nbar_cov]
 
 def data_xi(Mr=20, Nmock=500): 

### 2016-01-21 halo/code/util.py
util edited

@@ -20,4 +20,4 @@ def dat_dir():
     '''
     Dat directory
     '''
-    return '/../dat/'
+    return '../dat/'

### 2016-01-22 halo/code/abc_pemcee.py
abc_pemcee.py now runs! still needs to be checked

@@ -7,6 +7,7 @@ Author(s): Chang, MJ
 
 '''
 import time
+import pickle
 import numpy as np
 
 import abcpmc
@@ -27,9 +28,8 @@ def ABCpmc(T, eps_val, N_part=1000, observables=['nbar', 'gmf'], data_dict={'Mr'
     Parameters
     ----------
     - T : Number of iterations 
-    - eps_val : ???? (@mjv: what does this do?)
+    - eps_val : 
     - N_part : Number of particles
-    - Threads : Number of MPI threads (not sure if this keyword actually works or not 
     - observables : list of observables. Options are 'nbar', 'gmf', 'xi'
     - data_dict : dictionary that specifies the observation keywords 
     '''
@@ -38,7 +38,6 @@ def ABCpmc(T, eps_val, N_part=1000, observables=['nbar', 'gmf'], data_dict={'Mr'
     for obv in observables: 
         if obv == 'nbar': 
             data_nbar, data_nbar_var = Data.data_nbar(**data_dict)
-            print data_nbar
             fake_obs.append(data_nbar)
         if obv == 'gmf': 
             data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
@@ -65,13 +64,19 @@ def ABCpmc(T, eps_val, N_part=1000, observables=['nbar', 'gmf'], data_dict={'Mr'
     our_model = HODsim()    # initialize model
     kwargs = {'prior_range': prior_range, 'observables': observables}
     def simz(tt): 
-        return our_model.sum_stat(tt, **kwargs)
+        sim = our_model.sum_stat(tt, **kwargs)
+        if sim is None: 
+            print 'Simulator is giving NoneType.'
+            pickle.dump(tt, open("simz_crash_theta.p", 'wb'))
+            print 'The input parameters are', tt
+            pickle.dump(kwargs, open('simz_crash_kwargs.p', 'wb'))
+            print 'The kwargs are', kwargs
+            raise ValueError
+        return sim
 
     def multivariate_rho(datum, model): 
         dists = [] 
         for i_obv, obv in enumerate(observables): 
-            print "datum=" , datum
-	    print "model=" , model
             if obv == 'nbar': 
                 dist_nz = (datum[i_obv] - model[i_obv])**2. / data_nbar_var 
                 dists.append(dist_nz)
[…]

### 2016-01-22 halo/code/hod_sim.py

@@ -80,8 +80,8 @@ class HODsim(object):
                     return obvs 
 
                 except ValueError:
-                    bins = data_gmf_bins(Mr=20)
+                    bins = data_gmf_bins()
                     return [10. , np.ones_like(bins)[:-1]*1000.]
-                else:
-                    bins = data_gmf_bins(Mr=20)
-                    return [10. , np.ones_like(bins)[:-1]*1000.]
+            else:
+                bins = data_gmf_bins()
+                return [10. , np.ones_like(bins)[:-1]*1000.]

### 2016-01-22 halo/code/test_abc_pemcee.py

@@ -1,0 +1,24 @@
+'''
+
+Tests for abc_pemcee module
+
+'''
+import pickle
+
+from hod_sim import HODsim
+
+def simz_crash_test(): 
+    '''
+    ########################## 
+    ######## RESOLVED ######## 
+    ########################## 
+    Simz crashes and gives None as output. Why? 
+
+    Because of an indentation error in the sum_stat 
+    moduel in the HODsim class.
+    '''
+    theta = pickle.load(open('simz_crash_theta.p', 'rb'))
+    kwargs = pickle.load(open('simz_crash_kwargs.p', 'rb'))
+
+    ourmodel = HODsim()
+    return ourmodel.sum_stat(theta, **kwargs)

### 2016-01-22 halo/code/abc_pemcee.py
simulator bug with out of prior range xi distance fixed

@@ -131,4 +131,4 @@ def ABCpmc(T, eps_val, N_part=1000, observables=['nbar', 'gmf'], data_dict={'Mr'
     return pools
 
 if __name__=="__main__": 
-    ABCpmc(10, [1.e10,1.e10], N_part=100)
+    ABCpmc(2, [1.e10,1.e10], N_part=10, observables=['nbar', 'xi'])

### 2016-01-22 halo/code/hod_sim.py

@@ -6,6 +6,7 @@ HaloTools HOD Simulation
 import numpy as np
 from halotools.empirical_models import PrebuiltHodModelFactory
 
+from data import data_xi_bin
 from data import data_gmf_bins
 from group_richness import gmf 
 from group_richness import richness
@@ -20,6 +21,7 @@ class HODsim(object):
         '''
 
         #self.model = zheng07()      # Zheng et al. (2007) model
+        self.Mr = Mr
         self.model = PrebuiltHodModelFactory('zheng07', threshold=-1*Mr)
     
     def sum_stat(self, theta, prior_range=None, observables=['nbar', 'gmf']):
@@ -80,8 +82,27 @@ class HODsim(object):
                     return obvs 
 
                 except ValueError:
-                    bins = data_gmf_bins()
-                    return [10. , np.ones_like(bins)[:-1]*1000.]
+
+                    obvs = []
+                    for obv in observables: 
+                        if obv == 'nbar': 
+                            obvs.append(10.)
+                        elif obv == 'gmf': 
+                            bins = data_gmf_bins()
+                            obvs.append(np.ones_like(bins)[:-1]*1000.)
+                        elif obv == 'xi': 
+                            bins = data_xi_bin(Mr=self.Mr)
+                            obvs.append(np.zeroes(len(bins)))
+                    return obvs 
             else:
-                bins = data_gmf_bins()
-                return [10. , np.ones_like(bins)[:-1]*1000.]
+                obvs = []
+                for obv in observables: 
+                    if obv == 'nbar': 
+                        obvs.append(10.)
+                    elif obv == 'gmf': 
+                        bins = data_gmf_bins()
+                        obvs.append(np.ones_like(bins)[:-1]*1000.)
+                    elif obv == 'xi': 
+                        bins = data_xi_bin(Mr=self.Mr)
+                        obvs.append(np.zeroes(len(bins)))
+                return obvs

### 2016-01-22 halo/code/abc_pemcee.py
hod_sim corrected

@@ -131,4 +131,5 @@ def ABCpmc(T, eps_val, N_part=1000, observables=['nbar', 'gmf'], data_dict={'Mr'
     return pools
 
 if __name__=="__main__": 
-    ABCpmc(2, [1.e10,1.e10], N_part=10, observables=['nbar', 'xi'])
+    ABCpmc(20, [1.e10,1.e10], N_part=100, observables=['nbar', 'xi'])
+    ABCpmc(20, [1.e10,1.e10], N_part=100, observables=['nbar', 'xi', 'gmf'])

### 2016-01-22 halo/code/emcee_emcee.py
coded up mcmc inference using xi , nbar

@@ -10,7 +10,7 @@ Author(s): Chang, MJ
 import sys
 import numpy as np
 import emcee
-
+from numpy.linalg import solve
 from emcee.utils import MPIPool
 
 # --- Local ---
@@ -19,77 +19,131 @@ from hod_sim import HODsim
 from group_richness import richness
 
 
-def McMc(data_dict={'Mr':20, 'Nmock':500}): 
+def mcmc(Nwalkers, Nchains_burn, Nchains_pro , Ndim, observables=['nbar', 'xi'], data_dict={'Mr':20, 'Nmock':500}): 
     '''
-    Standard MCMC inference
-    '''
-    # import fake observables
-"""data, covariance matrix, and the inverse covariance """
-
-xir_data = np.loadtxt("xir_Mr20.dat")
-covariance = np.loadtxt("clustering_covariance_Mr20.dat")
-N_mocks = 500
-
-
-"""Initializing Walkers"""
-
-ndim, nwalkers = 5, 100
-pos = [np.array([11.5 , np.log(.4) , 12.02 , 1.03 , 13.5]) + 1e-4*np.random.randn(ndim) for i in range(nwalkers)]
-
-"""Initializing MPIPool"""
-
-pool = MPIPool()
-if not pool.is_master():
-    pool.wait()
-    sys.exit(0)
-
-sampler = emcee.EnsembleSampler(nwalkers, ndim, lnprob , pool = pool)
-
-sampler.run_mcmc(pos, 500)
-
-pool.close()
-
-
-
-"""likelihood"""
-
-def lnlike(theta):
+    Standard MCMC implementaion
     
-    model = Zheng07(threshold = -20.)
-    model.param_dict['logM0'] = theta[0]
-    model.param_dict['sigma_logM'] = np.exp(theta[1])
-    model.param_dict['logMmin'] = theta[2]     
-    model.param_dict['alpha'] = theta[3]
-    model.param_dict['logM1'] = theta[4]
-    model.populate_mock()
-    r , xir_model = model.mock.compute_galaxy_clustering()
-    res = xir_data - xir_model
-    return -0.5*np.sum(np.dot(np.dot(res , inv_c) , res))
 
-"""prior"""
+    Parameters
+    -----------
+    - Nwalker : Number of walkers
+    - Nchains_burn : Number of burn-in chains
+    - Nchains_pro : Number of production chains   
+    - ndim    : Dimensionality of the parameter space
+    - observables : list of observables. Options are 'nbar', 'gmf', 'xi'
[…]

### 2016-01-22 halo/code/data.py
added a function that returns the full covariance for mcmc

@@ -60,6 +60,22 @@ def data_xi(Mr=20, Nmock=500):
 
     return [xi, cii]
 
+
+def data_xi_full_cov(Mr=20, Nmock=500): 
+    '''
+    Observed xi (2PCF) from 'data' and the diagonal elements of the xi covariance matrix
+    '''
+    xi_dat_file = ''.join(['../dat/xir.Mr', str(Mr), '.dat'])
+    xi = np.loadtxt(xi_dat_file, unpack=True)
+
+    # load covariance of xi 
+    cov_dat_file = ''.join(['../dat/xir_covariance.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
+    cov = np.loadtxt(cov_dat_file)
+    #cii = np.diag(cov)   # diagonal elements
+
+    return [xi, cov]
+
+
 def data_xi_bin(Mr=20):
     ''' r bins for xi(r)
     '''

### 2016-01-22 halo/code/emcee_emcee.py
fixed a few bugs in emcee_emcee.py

@@ -56,7 +56,6 @@ def mcmc(Nwalkers, Nchains_burn, Nchains_pro , Ndim, observables=['nbar', 'xi'],
     # Priors
     prior_min = [10., np.log(0.1), 11.02, 0.8, 13.]
     prior_max = [13., np.log(0.7), 13.02, 1.3, 14.]
-    prior = abcpmc.TophatPrior(prior_min, prior_max)
     prior_range = np.zeros((len(prior_min),2))
     prior_range[:,0] = prior_min
     prior_range[:,1] = prior_max
@@ -81,9 +80,9 @@ def mcmc(Nwalkers, Nchains_burn, Nchains_pro , Ndim, observables=['nbar', 'xi'],
         """log-likelihood without the term -.5*log(det(cov))"""
 
         nbar_model , xi_model = simz(theta)
-        res_nbar = fake_obs['nbar'] - nbar_model
-        res_xi   = fake_obs['xi'] - xi_model
-	chi_nbar = -0.5*(res_nbar)**2. / 
+        res_nbar = fake_obs[0] - nbar_model
+        res_xi   = fake_obs[1] - xi_model
+	chi_nbar = -0.5*(res_nbar)**2. / data_nbar_var 
         chi_xi   = -0.5*np.sum(np.dot(np.dot(res_xi , data_xi_invcov) , res_xi))
 
         return chi_nbar + chi_xi
@@ -94,10 +93,10 @@ def mcmc(Nwalkers, Nchains_burn, Nchains_pro , Ndim, observables=['nbar', 'xi'],
         """log-prior"""
 
         a , b , c , d , e = theta
-        if prior_min[0] < a < prior_max[0] and \\
-           prior_min[1] < b < prior_max[1] and \\
-           prior_min[2] < c < prior_max[2] and \\
-           prior_min[3] < d < prior_max[3] and \\
+        if prior_min[0] < a < prior_max[0] and \
+           prior_min[1] < b < prior_max[1] and \
+           prior_min[2] < c < prior_max[2] and \
+           prior_min[3] < d < prior_max[3] and \
            prior_min[4] < e < prior_max[4]:
             return 0.0
         else:
@@ -121,14 +120,14 @@ def mcmc(Nwalkers, Nchains_burn, Nchains_pro , Ndim, observables=['nbar', 'xi'],
 
     """Initializing MPIPool"""
 
-    pool = MPIPool()
[…]

### 2016-01-22 halo/code/emcee_emcee.py
fixed a few bugs in mcmc

@@ -115,8 +115,7 @@ def mcmc(Nwalkers, Nchains_burn, Nchains_pro , Ndim, observables=['nbar', 'xi'],
 
     """Initializing Walkers"""
 
-    ndim, nwalkers = 5, 100
-    pos = [np.array([11.5 , np.log(.4) , 12.02 , 1.03 , 13.5]) + 1e-6*np.random.randn(Ndim) for i in range(Nwalkers)]
+    pos = [np.array([11. , np.log(.4) , 11.5 , 1.0 , 13.5]) + 1e-3*np.random.randn(Ndim) for i in range(Nwalkers)]
 
     """Initializing MPIPool"""
 
@@ -126,23 +125,20 @@ def mcmc(Nwalkers, Nchains_burn, Nchains_pro , Ndim, observables=['nbar', 'xi'],
     #   sys.exit(0)
 
     """Initializing the emcee sampler"""
-
-    sampler = emcee.EnsembleSampler(Nwalkers, Ndim, lnprob)# , pool = pool)
-
-    # Burn in.
-    pos, _, _ = sampler.run_mcmc(pos, Nchains_burn)
-    sampler.reset()
+    sampler = emcee.EnsembleSampler(Nwalkers, Ndim, lnprob)
+    #sampler = emcee.EnsembleSampler(Nwalkers, Ndim, lnprob, pool = pool)
+    
+    # Burn in + Production
+    sampler.run_mcmc(pos, Nchains_burn + Nchains_pro)
 
     # Production.
-    pos, _, _ = sampler.run_mcmc(pos, Nchains_pro)
-    
+    samples = sampler.chain[:, Nchains_burn:, :].reshape((-1, Ndim))
     #closing the pool 
     #pool.close()
-
     #saving the mcmc samples
-    np.savetxt("mcmc_sample.dat" , sampler.flat_chain)
-
+    np.savetxt("mcmc_sample.dat" , samples)
+    
 if __name__=="__main__": 
     
-    mcmc(10, 10, 10, 5, observables=['nbar', 'xi'], data_dict={'Mr':20, 'Nmock':500})
+    mcmc(20, 100, 600, 5, observables=['nbar', 'xi'], data_dict={'Mr':20, 'Nmock':500})
 

### 2016-01-25 halo/code/plotting.py
mcmc chain plotting; default prior implemented

@@ -9,6 +9,8 @@ import matplotlib.pyplot as plt
 
 # --- local ---
 import util
+import data as Data
+from prior import PriorRange
 plt.switch_backend("Agg")
 
 def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=None, filename=None): 
@@ -64,3 +66,60 @@ def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=N
         '_Mr', str(Mr), '_now_t', str(t), '.png'])
     plt.savefig(fig_file)
     plt.close()
+
+
+def plot_mcmc_chains(Nchains_burn, Mr=20, truths=None, observables=['nbar', 'xi'], 
+        plot_range=None): 
+    '''
+    Plot MCMC chains
+    '''
+    if truths is None: 
+        data_hod_dict = Data.data_hod_param(Mr=Mr)
+        truths = np.array([
+            data_hod_dict['logM0'],                 # log M0 
+            np.log(data_hod_dict['sigma_logM']),    # log(sigma)
+            data_hod_dict['logMmin'],               # log Mmin
+            data_hod_dict['alpha'],                 # alpha
+            data_hod_dict['logM1']                  # log M1
+            ])
+    if plot_range is None: 
+        prior_min, prior_max = PriorRange(None)
+        plot_range = np.zeros((len(prior_min),2))
+        plot_range[:,0] = prior_min
+        plot_range[:,1] = prior_max
+    
+    # chain files 
+    chain_file = ''.join([util.dat_dir(), 
+        util.observable_id_flag(observables), 
+        '_Mr', str(Mr), '_theta.mcmc_chain.dat'])
+
+    sample = np.loadtxt(chain_file)
+        
+    fig = corner.corner(
+            sample[Nchains_burn:], 
+            truths=truths,
+            truth_color='#ee6a50', 
+            labels=[
+                r'$\logM_{0}$',r'$\sigma_{\logM}$',r'$\logM_{min}$',r'$\alpha$',r'$\logM_{1}$'
+                ],
+            range=plot_range, 
+            quantiles=[0.16,0.5,0.84],
+            show_titles=True, 
+            title_args={"fontsize": 12},
+            plot_datapoints=True, 
+            fill_contours=True, 
+            levels=[0.68, 0.95], 
+            color='b', 
+            bins=16, 
+            smooth=1.0)
+
[…]

### 2016-01-25 halo/code/plotting.py
plotting for mcmc chian fixed

@@ -68,7 +68,7 @@ def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=N
     plt.close()
 
 
-def plot_mcmc_chains(Nchains_burn, Mr=20, truths=None, observables=['nbar', 'xi'], 
+def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables=['nbar', 'xi'], 
         plot_range=None): 
     '''
     Plot MCMC chains
@@ -94,9 +94,10 @@ def plot_mcmc_chains(Nchains_burn, Mr=20, truths=None, observables=['nbar', 'xi'
         '_Mr', str(Mr), '_theta.mcmc_chain.dat'])
 
     sample = np.loadtxt(chain_file)
+    Nchain = len(sample) / Nwalkers 
         
     fig = corner.corner(
-            sample[Nchains_burn:], 
+            sample[Nchains_burn*Nwalkers:], 
             truths=truths,
             truth_color='#ee6a50', 
             labels=[
@@ -115,11 +116,11 @@ def plot_mcmc_chains(Nchains_burn, Mr=20, truths=None, observables=['nbar', 'xi'
 
     fig_file = ''.join([util.fig_dir(), 
         util.observable_id_flag(observables), 
-        '_Mr', str(Mr), '.Nsample', str(len(sample)),
+        '_Mr', str(Mr), '.Nchain', str(Nchain),
         '.Nburn', str(Nchains_burn), '.mcmc_chain.png'])
     plt.savefig(fig_file)
     plt.close()
 
 
 if __name__=='__main__':
-    plot_mcmc_chains(100, Mr=20, observables=['nbar', 'xi'])
+    plot_mcmc_chains(20, Nchains_burn=100, Mr=20, observables=['nbar', 'xi'])

### 2016-01-25 halo/code/plotting.py
implemented plotting the evolution of walkers

@@ -95,13 +95,63 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
 
     sample = np.loadtxt(chain_file)
     Nchain = len(sample) / Nwalkers 
+    
+    chain_ensemble = sample.reshape(Nchain , Nwalkers, 5)
+    fig , axes = plt.subplot(5, 1 , sharex=True, figsize=(10, 12))
+
+    labels=[
+        r'$\logM_{0}$',r'$\log \sigma_{\logM}$',r'$\logM_{min}$',r'$\alpha$',r'$\logM_{1}$'
+           ]
+  
+    for i in xrange(5):
+        
+        axes[i].plot(chain_ensemble[:, :, i], color="k", alpha=0.1)
+	axes[i].yaxis.set_major_locator(MaxNLocator(5))
+        axes[i].axhline(thruths[i], color="#888888", lw=2)
+        axes[i].set_ylabel(labels[i])
+
+    fig.tight_layout(h_pad=0.0) 
+    fig_file = ''.join([util.fig_dir(), 
+        util.observable_id_flag(observables), 
+        '_Mr', str(Mr), '.Nchain', str(Nchain),
+        '.Nburn', str(Nchains_burn), '.mcmc_time.png'])
+    plt.savefig(fig_file)
+    plt.close()
+
+def plot_mcmc_samples(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables=['nbar', 'xi'], 
+        plot_range=None): 
+    '''
+    Plot MCMC chains
+    '''
+    if truths is None: 
+        data_hod_dict = Data.data_hod_param(Mr=Mr)
+        truths = np.array([
+            data_hod_dict['logM0'],                 # log M0 
+            np.log(data_hod_dict['sigma_logM']),    # log(sigma)
+            data_hod_dict['logMmin'],               # log Mmin
+            data_hod_dict['alpha'],                 # alpha
+            data_hod_dict['logM1']                  # log M1
+            ])
+    if plot_range is None: 
+        prior_min, prior_max = PriorRange(None)
+        plot_range = np.zeros((len(prior_min),2))
+        plot_range[:,0] = prior_min
+        plot_range[:,1] = prior_max
+    
+    # chain files 
+    chain_file = ''.join([util.dat_dir(), 
+        util.observable_id_flag(observables), 
[…]

### 2016-01-25 halo/code/plotting.py
fixed a bug in plotting the evolution of walkers

@@ -6,6 +6,7 @@ Plotting modules
 import corner
 import numpy as np
 import matplotlib.pyplot as plt
+from matplotlib.ticker import MaxNLocator 
 
 # --- local ---
 import util
@@ -97,7 +98,7 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
     Nchain = len(sample) / Nwalkers 
     
     chain_ensemble = sample.reshape(Nchain , Nwalkers, 5)
-    fig , axes = plt.subplot(5, 1 , sharex=True, figsize=(10, 12))
+    fig , axes = plt.subplots(5, 1 , sharex=True, figsize=(10, 12))
 
     labels=[
         r'$\logM_{0}$',r'$\log \sigma_{\logM}$',r'$\logM_{min}$',r'$\alpha$',r'$\logM_{1}$'
@@ -107,7 +108,7 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
         
         axes[i].plot(chain_ensemble[:, :, i], color="k", alpha=0.1)
 	axes[i].yaxis.set_major_locator(MaxNLocator(5))
-        axes[i].axhline(thruths[i], color="#888888", lw=2)
+        axes[i].axhline(truths[i], color="#888888", lw=2)
         axes[i].set_ylabel(labels[i])
 
     fig.tight_layout(h_pad=0.0) 

### 2016-01-25 halo/code/plotting.py
fixed a few things in the chain plot

@@ -106,11 +106,11 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
   
     for i in xrange(5):
         
-        axes[i].plot(chain_ensemble[:, :, i], color="k", alpha=0.1)
+        axes[i].plot(chain_ensemble[:, :, i], color="k", alpha=0.4)
 	axes[i].yaxis.set_major_locator(MaxNLocator(5))
         axes[i].axhline(truths[i], color="#888888", lw=2)
         axes[i].set_ylabel(labels[i])
-
+    axes[5].set_xlabel("step number")    
     fig.tight_layout(h_pad=0.0) 
     fig_file = ''.join([util.fig_dir(), 
         util.observable_id_flag(observables), 

### 2016-01-25 halo/text/abc.tex

@@
+ss[12pt, preprint]{aastex}
+\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
+\usepackage{graphicx}	% For figures
+\usepackage{natbib}	% For citep and citep
+\usepackage{amsmath}	% for \iint
+\usepackage{bm}
+\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
+\usepackage{hyperref}
+\usepackage{algorithmic,algorithm}
+\hypersetup{colorlinks}
+\usepackage{color}
+\usepackage{morefloats}
+\definecolor{darkred}{rgb}{0.5,0,0}
+\definecolor{darkgreen}{rgb}{0,0.5,0}
+\definecolor{darkblue}{rgb}{0,0,0.5}
+\hypersetup{ colorlinks,
+linkcolor=darkblue,
+filecolor=darkgreen,
+urlcolor=darkred,
+citecolor=darkblue}
+
+\DeclareMathOperator*{\argmax}{arg\,max}
+
+\newcommand{\todo}[1]{{\bf \textcolor{red}{ #1}}}
+\newcommand{\beq}{\begin{equation}}
+\newcommand{\eeq}{\end{equation}}
+\newcommand{\lang}{\langle}
+\newcommand{\ra}{\rangle}
+\newcommand{\vep}{\bm{\epsilon}}
+\newcommand{\ep}{\epsilon}
+\newcommand{\pars}{\vec{\theta}}
+\newcommand{\dev}{\mathrm{d}}
+
+\begin{document}
+% We need a better definitely shorter title...
+\title{Approximate bayesian computation in large scale structure modeling : constraining the galaxy-halo connection}
+\begin{abstract}
+\todo{This is not a paper about hod, let's not start with describing hod.
+I suggest that we talk about parameter estimation in large scale structure in general and then move on to hod.
+Constraining hod is a just the first step in testing our methodology.} The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies
+populate dark matter halos that has had much success in reproducing large scale galaxy
+clustering and understanding galaxy evolution within the context of the halo model.
+Constraints on the parameters that describe the HOD model have so far been inferred using 
+the standard practice of Bayesian inference with likelihood estimation that assumes a
+functional Gaussian form of the likelihood.\todo{(Do we want to include analytic models for clustering statistics here?
+Yes, also we should say we use summary statistics that are not independent: clustering + group.
+Also there is no analytic model for the abundance of galaxy group richnesses)}  % I think we should keep it simple.
+There's no need to go into detail about everything in the abstract.
+This is already a pretty sizable abstract.
+Likelihood free inference methods such as the
+Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever 
+having to calculate the likelihood.
+
+In this work, we present ABC in the context of constraining the parameters of the HOD model
+using {\it Halotools}, a publicly available ... First, we set a mock galaxy catalog created
+with HOD parameters from \cite{Zheng:2007aa} \todo{why do we site here?} as our data catalog. Then we use ABC on a set 
+% To give it a false sense of legitimacy... 
+of observables measured from forward modeled mock catalogs generated with {\it Halotools} 
+in order to accurately infer constraints on the HOD parameters that 
+reproduce the original parameters. \todo{Remark about the comparison between our ABC 
+method constraints and the standard MCMC constraints.} Furthermore, since ABC does 
+not require computing the likelihood, we include observables beyond galaxy clustering 
+such as the galaxy number density and the galaxy group multiplicity function, for which
+accurately constructing a likelihood is intractable. 
+We demonstrate that adding these measurements to our set of observables leads to stronger
+constraints on the HOD parameters, specifically the ones that govern the satellite galaxy
+population. Moreover, the constraints we infer illustrate the feasibility of ABC in 
+cosmological contexts.
+
+\todo{Hogg writes here too}
+
+\end{abstract}
+
+\section{Introduction}
+
+Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
+pro\-ba\-bil\-iites for the data given the theoretical model and parameter settings---%
+for combining data from different surveys and performing principled inferences of
+the cosmological parameters (CITE CITE).
+Nowhere has this been more true than in cosmic microwave background studies,
+where it is nearly possible to analytically evaluate a likelihood function that
+involves no (or minimal) approximations (CITE Wandelt and others).
+
+Fundamentally, the tractability of likelihood functions in cosmology flows from
+the fact that the initial conditions are exceedingly close to Gaussian in form,
+and many sources of measurement noise are also Gaussian.
+Likelihood functions are easier to write down and evaluate when things are closer to Gaussian,
+so they get easier as the length scales get larger and the structure formation gets
+younger.
+Hence the suitability of CMB data for likelihood analyses.
+
+In large-scale structure studies, which have galaxies, quasars, or quasar absorption
+systems as tracers, the likelihood cannot be Gaussian:
+Even if the initial conditions are perfectly Gaussian, the growth of structure
+creates non-linearities which are non-Gaussian, and the galaxies form
+within the density field in some complex manner that is unknown (or modeled only effectively).
+Even if the galaxies were a Poisson sampling of the density field (which they are demonstrably
+not; CITE PEOPLE), it would be tremendously hard to write down even an approximate
+likelihood function.
+
+The standard approach is to make two strong assumptions, both demonstrably wrong:
+The first is that the (density and) two-point function measurements constitute
+sufficient statistics for the cosmological parameters.
+That is, the assumption is that all of the information about the cosmological model is
+encoded in the mean and variance statistics of the tracer population.
+This first assumption is equivalent to assuming that the density field is a Gaussian process,
+and is therefore closer to true on larger scales.
+
+The second strong and wrong assumption is that the likelihood functon for the data can
+be approximated by a pseudo-likelihood function that is a Gaussian probability density
+in the space of the two-point correlation function estimate.
+That is, the assumption is that a pseudo-likelihood of the data given by a squared distance
+(scaled by an inverse covariance matrix) between an estimate of the correlation function
+(the data) and a theoretical prediction of that same function will return sensible posterior
+inferences.
+This assumption cannot be correct (in detail) at any scale, since a correlation function,
+being related to the variance of a continuous field,
+must satisfy non-trivial positive-definiteness
+requirements (it must take the form of a valid kernel function; CITE SOMETHING).
+These requirements truncate function space such that the likelihood in that functon space
+could never be Gaussian.
+This second assumption becomes truer as the correlation function gets better measured, so
+it is most true on intermediate scales (where neither shot noise nor cosmic variance blow up).
+We expect it to be very strongly violated at the baryon acoustic scale in particular!
+
+Neither of these two assumptions are required for cosmological inferences, however:
+We have very sophisticated simulations and emulations of structure and galaxy formation.
+Indeed, the theory of large-scale structure \emph{is} the output of these simulations;
+we can simulate the data in great detail.
+Therefore, we can simulate not just the one- and two-point statistics of the galaxies,
+but also any higher order statistics that might be relevant.
+Furthermore, we can see the distributions of these statistics, and where the data lie within
+those distributions.
+That is---in principle---we don't need to make these strong and wrong assumptions;
+we can create---in principle---a likelihood function directly from (an enormous number of)
+simulation outputs.
+
+Of course, any naive approach to sufficiently simulating the data would be ruinously
+expensive.
+Fortunately, there are principled, (relatively) efficient methods for minimizing computation
+and delivering correct posterior inferences,
+using only a data simulator and some choices about statistics.
+Here we use \todo{Introduce the names ABC and LFI with references}.
+
+As we will see, the LFI approach---despite cleverness---will be very computationally expensive.
+For context, it is worth remembering that even the traditional approach to large-scale structure,
+outlined above, requires enormous numbers of structure- and galaxy-formation simulations as well,
+to construct the covariance matrices that go into the pseudo-likelihood.
+So although our costs will be high, there are no inexpensive methods to compare to.
+
+\todo{MJ: What Hogg wrote above doesn't mesh with (that is, repeats differently) what you have written below, so can you go through and fix that?
+Lots of what is below should remain, but some should get woven into what's above.}
+
+% style note: Here and in what follows: --- separates parenthetical phrases, -- connects words that appear together in a relationship (like galaxy--galaxy) and (galaxy--halo), and - connects pairs of words into an adjective (like two-point).
+
+Inference of the cosmological parameters and understanding the formation of galaxies within the context of large scale 
+structure, requires understanding of the galaxy--halo connection. Halos are biased tracers of the large scale structure 
+and galaxies reside in dark matter Halos.
+Halo occupation distribution---HOD---is a probabilistic framework that describes how galaxies populate dark matter halos.
+The central tenet of halo occupation model is that the probability that a certain number of galaxies populate a dark matter halo is determined by the mass of the host halo.
+This statistical prescription for connection of galaxies and halos has been successful in reproducing the galaxy two-point clustering as well as the galaxy--matter cross correlation---galaxy--galaxy
+lensing---within the context of the halo model, and has been used in understanding the co-evolution of galaxies and halos and putting constraints on the relation between the stellar mass and halo mass of galaxies.
+
+In the context of large scale clustering of galaxies, inference of the cosmological and halo model parameters are performed by treating some estimate of the galaxy--galaxy two-point correlation function as the data,
+assuming an analytic model for the theoretical two-point correlation function prediction,
+and a Gaussian likelihood connecting the two.
+The accuracy of the analytic model of the galaxy clustering statistics hinges upon the accuracy of the fitting functions such as halo bias, halo mass function that are accurate to ten percent level.
+There are other simplifying assumptions encoded in these analytic models such as the scale independence of the halo bias.
+
+Furthermore, the likelihood of observing the data---observed positions of the galaxies in the configuration space---given the cosmological and halo model parameters is approximated by a Gaussian function whose argument is the squared difference of the two-point correlation function measured from the observations and the analytic model for the galaxy clustering, weighted by a covariance matrix.
+\todo{Why MUST this be wrong?}
+The weighted difference of the number density of the observed galaxies and the analytic number density is also added to this.
+In this work, we investigate the how accurately we can infer the model parameters by relaxing the underlying assumptions of the parameter inference methods.
+
+Approximate bayesian computation (ABC) provides a \emph{rejected sampling} framework framework that alleviates the issues arising from approximating the likelihood of observing the large scale population of galaxies given the cosmological and halo model parameters.
+ABC approximates the posterior probability distribution function over the model parameters given the data---the angular positions of the galaxies on the sky and their redshifts---by drawing proposals from the prior over the model parameters, simulating the data with the proposals, and rejecting the proposals for which the distance between the data and the simulation (measured by a metric) is less than a certain small threshold.
+
+In practice, the rejected sampling method is replaced by a more efficient adaptive importance sampling operation called Population Monte Carlo (PMC).
+Initially, PMC rejects the draws from prior pdf with a relatively large threshold.
+In the consequent steps, importance sampling from the previous samples that passed the distance criteria is performed, and the threshold is updated adaptively.
+That is, after each iteration, the threshold is lowered according to the distribution of the distances.
+
+Furthermore, This sampling formalism permits us to sample the posterior of the model by simulating the data via a forward generative model that accounts for all the complications and uncertainties associated with the survey data that are difficult to take into account in approximating the likelihood function.
+In the context of the large galaxy surveys, these complications may include masks, nontrivial survey geometry, and missing data as a result of fiber collision.
+
+Rejected sampling in ABC requires a distance function that quantifies the closeness of the simulation (e.g.
+fiber collided galaxies) to the data (observed positions of the galaxies in the survey).
+In principle, this distance metric is a positive definite function  that compares various summary statistics of the data and those of the simulations.
+In this investigation, we focus on the two-point correlation, number density, and the group statistics (more specifically the abundance of the galaxy group richnesses).
+
+Performing a full cosmological inference using this sampling method requires a simulator that, (I) takes a set of cosmological parameters as input and generates a snapshot of dark matter particles at a given redshift using an N-body simulation, (II) builds a catalog of dark matter halos for that snapshot of the N-body simulation using a halo finder algorithm, and (III) paints galaxies into host halos given a set of halo occupation model parameters.
+
+Computationally, this process is extremely demanding.
+One way to make this problem more tractable is to split this into to separate this problem into two parts: \emph{emulation}, and \emph{simulation}.
+That is, one can use a set of N-body simulations with known cosmologies, and perform halo occupation simulations using the pre-built halo catalogs of those simulations.
+In order to compute the summary statistics corresponding to the samples from the parameter space for which there is no N-body simulation, one can interpolate between the summary statistics of the parameters for which we have available N-body simulations.
+This procedure is called emulation.
+Emulation has been introduced in the literature in coyote universe [CITE COYOTE].
+
+In this work however, we narrow our focus on inferring the posterior probability over the parameters of the halo model using an already existing halo catalog of a dark matter particle N-body simulation with a fixed cosmology.
+That is we simplify the problem by leaving the problem of emulation for future investigations.
+
+The calculations regarding populating the dark matter halos, computing the clustering statistics and finding groups in the galaxy mock catalogs are performed using ({\fontfamily{cmss}\selectfont Halotools}) which is an open-source package tailored for modeling galaxy-halo connection in large scale structure modeling ({\fontfamily{cmss}\selectfont http://halotools.readthedocs.org}).
+
+This paper is structured as follows. In Section \ref{sec:method},
+we discuss the components of the Halo occupation modeling, and the ABC sampling method.
+We also give a brief overview of the summary statistics used in this investigation.
+In Section \ref{sec:halo} we give a brief overview of the forward modeling of the galaxy mock catalogs using Halotools as well as generating a mock galaxy catalog as data.
+In Section \ref{sec:results} we explore the results of our parameter inference using different summary statistics.
+Finally, we discuss and conclude in Section \ref{sec:discussion}.
+
+\section{Method}\label{sec:method}
+
+\subsection{Approximate Bayesian Computation}
+
+ABC is a generative simulation-based inference based on rejection sampling.
+That is, the parameters of the forward model of the data are drawn from the prior probability, and are rejected if the simulated data is not sufficiently close to the observations according to some distance metric $\rho$.
+The method requires a prior probability over the parameters of interest $\pi(\theta)$, a generative forward model of the data given a set of parameters $f(\theta)$, and a distance metric $\rho(.,.)$ which measures the distance between the simulation $X=f(\theta)$ and the observed data $D$.
+
+In a large scale structure study however, the observed data---which encompasses the angular positions of galaxies on the sky and their redshifts---is a high dimensional object.
+Therefore the distance metric measures the closeness of statistical summaries of the data and those of the simulation.
+When the observed data-set under study is the positions of all galaxies observed in the survey, possible statistical summaries of the data are the number density of galaxies, two-point correlation function, three-point correlation function, galaxy group statistics (counts in cells), and so on.
+
+Here, we provide a brief overview of rejection sampling algorithm and how it is used in a simulation-based inference.
+We want to draw samples from a probability distribution $p(\alpha)$.
+Now, let us assume that: (i) there exists an auxiliary density $q(\alpha)$ from which we can directly sample from, and (ii) there exist $M>1$ which is an appropriate bound on $p(\alpha)/q(\alpha)$.
+Given these assumptions, rejection sampling works in the following way.
+We draw a sample $\alpha$ from the auxiliary density $q(\alpha)$.
+Afterwards, we draw a random number $u$ from a uniform distribution between zero and one.
+Then we accept the draw $\alpha$ if the $u$ is less than the ratio $p(\alpha)/Mq(\alpha)$.
+This procedure is repeated until sufficient number of draws are accepted.
+
+In the context of a simulation-based inference, we aim to sample from the joint probability of a simulation $X$, and parameters $\pars$ given the observed data $D$
+
+\beq
+p(\pars,X | D)
+\eeq
+
+which according to the Bayes rule can be written as
+
+\begin{eqnarray}
+p(\pars,X | D)  &=& \frac{p(D|X)p(X|\pars)\pi(\pars)}{Z},  \\
+Z &=& \int \dev \pars \; \dev X \; p(D|X)p(X|\pars)\pi(\pars),
+\end{eqnarray}
+where $Z$ is the evidence. There exists a constant $M>1$ 
+\beq 
+M = \frac{\mathrm{max}_{X} \; p(D|X)}{Z},
+\eeq 
+and auxiliary density $q(\pars ,X)$
+\beq 
+q(\pars ,X)=p(X|\pars )\pi(\pars ),
+\eeq 
+such that:
+\beq 
+p(\pars ,X|D)<Mq(\pars ,X).
+\eeq 
+Now in a simulation-based inference, we first draw a pair $\{\pars , X \}$ from the density $q(\pars , X)$.
+This requires drawing a sample $\pars$ from the prior $\pi(\pars)$, and generating a simulation $X=f(\pars)$ via the forward model.
+Then, we accept the pair $\{\pars , X \}$, if the following condition is met
+
+\beq 
+u \sim \mathrm{Uniform}[0,1] < \frac{p(\pars,X|D)}{Mq(\pars,X)}=\frac{p(D|X)}{\mathrm{max}_{X}p(D|X)}.
+\eeq
+
+ABC postulates that the probability of observing data $D$ given simulation $X$, $p(D|X)$ is controlled by a threshold $\epsilon$, and that this probability is one if the distance between the data and the simulation, $\rho(D,X)$ is less than $\epsilon$.
+That is
+
+\beq 
+p(D|X) = p_{\epsilon}(D|X) \equiv \begin{cases}
+                                    1 \ \ \  \mathrm{if} \ \rho(D,X)<\epsilon \; , \\
+                                    0 \ \ \ \mathrm{otherwise} \; .
+                                    \end{cases}
+\eeq
+
+Thus the set of parameters $\pars$ is accepted if $\rho(D,f(\pars))<\epsilon$. In principle, the distance function measures the distance 
+between the statistical summaries of the data and those of the simulation. 
+The distance, in general, can be a vector, with each component representing the distance between one summary statistics of the data (e.g.
+variance) and the same summary statistics of the simulation.
+That is
+\beq 
+\rho(D,X) = \Big[\rho_{1}(S_{1}(D) , S_{1}(X)), ... , \rho_{M}(S_{M}(D) , S_{M}(X)) \Big],
+\eeq 
+where $M$ is the number of statistical summaries used in the inference, $S_{i}$ is the i-th summary statistics, and $\rho_{i}$ is a scalar function that measures the distance between the i-th summary statistics of the data and that of the simulation.
+In this case, the threshold $\epsilon$ is an $M$ dimensional vector.
+
+We draw a large number of proposals from the prior distribution over the parameter space, 
+and reject the samples for which the distance between the data and the simulation corresponding to that 
+sample is more than a certain threshold $\bm{\epsilon}$. This requirement is satisfied \emph{if and only if}
+every component of the distance vector is less than its corresponding threshold. Note that, each summary statistics 
+has a separate threshold. This procedure is done iteratively until, we are left with a certain number
+of samples that satisfy the distance criteria. The final ensemble approximates the posterior probability distribution over the parameters of the generative model of the data.
+
+For a sufficiently small threshold, this algorithm samples from the posterior probability, $p(\pars | D)$, and for an arbitrarily large threshold, the algorithm samples from the prior $\pi(\pars)$.
+This rejection sampling technique requires repeated sampling from the prior, and a very small threshold to succeed.
+
+In this investigation instead, we use Population Monte Carlo Algorithm (PMC) as our sampling technique.
+PMC is an iterative method that performs rejected sampling from a sequence of distributions $\{p_{1}(\pars),...,p_{T}(\pars)\}$.These distributions are paired with a sequence of decreasing distance thresholds $\{\epsilon_{1} , ...
+, \epsilon_{T}\}$.
+
+At each iteration, $N$ proposals are drawn from a distribution over parameters, and a weight is assigned to each one.
+In the first step, the threshold is set to an arbitrarily large value.
+Therefore the algorithm makes $N$ draws from the prior and assigns equal weights to all the proposals.
+
+In the consequent iterations, the distance threshold is set to a smaller value.
+A proposal is weighted-sampled from the previous sample.
+It is perturbed by a kernel which is set to the covariance of the parameters in the previous sample.
+A simulation is then generated by feeding the perturbed parameter into the forward model.
+The perturbed parameter is then accepted if the distance between the data and the generated simulation is less than the value of the threshold in that iteration.
+This procedure is repeated until $N$ particles are sampled.
+Then the weights are updated according to the kernel, the prior, and the previous set of weights.
+
+\begin{algorithm}
+\caption{The procedure for ABC-PMC}
+\begin{algorithmic}[1]
+%\STATE \DATA: D
+%\STATE \RESULT: ABC posterior sample of $\pars$
+\IF{$t=1:$}
+%\STATE $\epsilon_t \gets \infty$
+\FOR{$i=1,...,N$}
+   \STATE // \emph{This loop can now be done in parallel for all i}
+   \WHILE{$\rho(X,D)>\epsilon_t$}
+   \STATE $\pars^{*}_{t} \gets \pi(\pars)$
+   \STATE $X = f(\pars^{*}_{t})$
+   \ENDWHILE
+   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
+   \STATE $w^{(i)}_{t} \gets 1/N$
+\ENDFOR
+\ENDIF
+\IF{$t=2,...,T:$}
+\FOR{$i=1,...,N$}
+   \STATE // \emph{This loop can now be done in parallel for all i}
+   \WHILE{$\rho(X,D)>\epsilon_t$}
+   \STATE Draw $\pars^{*}_{t}$ from $\{\pars^{j}_{t-1}\}$ with probabilities $\{w^{j}_{t-1}\}$
+   \STATE $\pars^{*}_{t} \gets K(\pars^{*}_{t},.)$
+   \STATE $X = f(\pars^{*}_{t})$
+   \ENDWHILE
+   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
+   \STATE $w^{(i)}_{t} \gets \frac{\pi(\pars^{(i)}_{t})}{\sum_{j=1}^{N}w_{t-1}^{(i)}K(\pars^{(j)}_{t-1},\pars^{(i)}_{t})}$
+\ENDFOR
+\ENDIF
+\end{algorithmic}
+\end{algorithm}
+
+%\begin{algorithm}[H]
+% \Data{this text}
+% \Result{how to write algorithm with \LaTeX2e }
+% initialization\;
+% \While{not at end of this document}{
+%  read current\;
+%  \eIf{understand}{
+%   go to next section\;
+%   current section becomes this one\;
+%   }{
+%   go back to the beginning of current section\;
+%  }
+% }
+% \caption{How to write algorithms}
+%\end{algorithm}
+
+\subsection{Forward model of the large scale galaxy populations}\label{sec:halo}
+\subsubsection{halo occupation modeling}
+
+\newcommand{\lcdm}{\Lambda {\rm CDM}}
+\newcommand{\dd}{\mathrm{d}}
+\newcommand{\mean}[2]{\left\langle#1 \vert {#2}\right\rangle}
+
+%%% HALO OCCUPATION DISTRIBUTION %%%
+\newcommand{\ngal}{N_{\mathrm{g}}}
+\newcommand{\nsat}{N_\mathrm{s}}
+\newcommand{\ncen}{N_\mathrm{c}}
+\newcommand{\pnm}[2]{P(#1|#2)}
+
+\newcommand{\mhalo}{M_{\rm h}}
+\newcommand{\mvir}{M_\mathrm{vir}}
+
+\newcommand{\dndmvir}{\frac{\dd n}{\dd\mvir}}
+\newcommand{\dndmhalo}{\frac{\dd n}{\dd\mhalo}}
+\newcommand{\dndmvirprime}{\frac{\dd n}{\dd\mvir'}}
+
+%%% CLUSTERING STATISTICS %%%
+\newcommand{\xigg}{\xi_{\mathrm{gg}}}
+\newcommand{\xihh}{\xi_{\mathrm{hh}}}
+\newcommand{\xiggr}{\xi_{\mathrm{gg}}(r)}
+\newcommand{\xiggroneh}{\xi^{1h}_{\mathrm{gg}}(r)}
+\newcommand{\xiggrtwoh}{\xi^{2h}_{\mathrm{gg}}(r)}
+\newcommand{\ngalaxy}{\bar{n}_{\mathrm{g}}}
+
+One of the most well-studied statistics of the galaxy distribution 
+is $\xigg,$ the galaxy-galaxy correlation function. Mathematically, $\xigg$ encodes the
+excess probability that a pair of galaxies in a sample are 
+separated by three-dimensional distance $r.$ 
+Galaxy clustering is straightforward to measure observationally, 
+and it is a highly sensitive probe of both the physics of galaxy evolution 
+as well as the fundamental parameters of $\lcdm$ cosmology.
+
+The assumption that galaxies reside in dark matter halos is the bedrock underlying 
+all contemporary theoretical predictions for galaxy clustering. Since the 
+clustering of dark matter halos can be predicted to high precision with N-body simulations, 
+then knowledge of how galaxies populate and are distributed within halos 
+is sufficient to predict $\xigg.$ Under this assumption, $\xigg$ can be formally decomposed
+into a ``one-halo term" and a ``two-halo term'', 
+referring to galaxy pairs that reside in the same halo vs. distinct halos, respectively. 
+The two-halo term of $\xigg$ is a pair-number weighted average of $\xihh,$ 
+the halo-halo correlation function, whereas the one-halo term is the pair-number 
+weighted average of the intra-halo distribution of galaxies. 
+Models for the galaxy-halo connection aiming to predict $\xigg$ must therefore 
+provide theoretical means by which to calculate these weights and spatial distributions.
+
+The Halo Occupation Distribution (HOD) is one of the most widely used 
+ approaches to characterizing the galaxy-halo connection. 
+The central quantity In the HOD is $\pnm{\ngal}{\mhalo},$ the probability that a halo of mass
+$\mhalo$ hosts $\ngal$ galaxies. The first two moments of $\pnm$ contain the 
+necessary information to calculate the one- and two-halo terms of the galaxy correlation function:
+
+\begin{eqnarray}
+\label{eq:onehaloterm}
+1+\xiggroneh \simeq \frac{1}{4\pi{}r^{2}\ngalaxy^{2}}\int\dd\mhalo\dndmhalo\Xi_{\rm gg}(r|\mhalo) \times \mean{\ngal(\ngal-1)}{\mhalo},
+\end{eqnarray} and
+
+\begin{eqnarray}
+\label{eq:twohaloterm}
+\xiggrtwoh \simeq \xi_{\mathrm{mm}}(r)\left[\frac{1}{\ngalaxy}\int\dd\mhalo\dndmhalo \mean{\ngal}{\mhalo}b_{\mathrm{h}}(\mhalo)\right]^{2}
+\end{eqnarray}
+
+In Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}), $\ngalaxy$
+is the galaxy number density,
+$\dd\mathrm{n}/\dd\mhalo$ is the halo mass function, 
+the spatial bias of dark matter halos is $b_{\mathrm{h}}(\mhalo),$ and
+$\xi_{\rm mm}$ is the correlation function of dark matter.  
+If we represent the spherically symmetric intra-halo distribution of galaxies 
+by a unit-normalized $n_{\rm g}(r),$
+then the quantity $\Xi_{\rm gg}(r)$ appearing in the above two equations 
+is the convolution of $n_{\rm g}(r)$ with itself.
+
+Employing Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) to 
+calculate $\xigg(r)$ is the most common technical method used in HOD 
+studies of galaxy clustering. 
+However, this technique requires appeal to many simplifying assumptions 
+in addition to the supposition that galaxies reside in halos. For example, 
+Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) presume that 
+the intra-halo galaxy distribution is spherically symmetric, and 
+that the halo bias $b_{\mathrm{h}}(\mhalo)$ has no radial dependence. 
+Moreover, these equations does not take halo exclusion into account. 
+We refer the interested reader to \citet{cooray02,mo_vdb_white10,vdBosch13}, 
+and references therein, for further information.
+
+In addition to the above, in the vast majority of HOD studies the actual calculation of 
+Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) relies on further assumption still. 
+The quantities $b_{\mathrm{h}}(\mhalo),$ $\xi_{\mathrm{mm}}(r)$ and $\dndmhalo$ 
+are commonly replaced by analytical fitting functions that have been calibrated 
+against a suite of N-body simulations \citep[e.g.,][]{sheth_tormen01,smith03,tinker05,tinker10}.   
+And yet, it is well established that the precision of these fitting functions is 
+limited to the $\sim10\%$ level \citep[e.g.][]{tinker08}.
+
+\todo{APH: Ok, so the above level of detail may be overkill for this paper.
+I figured that part of the argument we'd make is this: it is now time to start taking the precision of our predictions seriously, the ten-percent game is over.
+ABC is a big contribution to going beyond ten percent, and now that Halotools makes it possible to circumvent the need for these fitting functions, this opens up new limiting factors such as the hard Gaussianity assumption.
+So I see the above argument as part of the argument for using ABC.
+However, I'm not married to any of the above text, so just let me know if you want me to scrap some of it.
+Otherwise, I will next pick up from here and start talking about how Halotools relaxes the reliance on these approximations, and I will make explicit the point about how this new level of precision motivates the exploration of ABC.}
+
+\todo{MJ: I think all of the text written above is necessary for making a case for simulation-based inference in large-scale structure modeling. So I'm in favor of keeping it for now.}
+
+Dark matter halos are biased tracers of the underlying matter distribution of the universe.
+The halo model postulates that galaxies reside in dark matter halos.
+Halo Occupation distribution---HOD---is a powerful tool for description of galaxy--halo connection.
+HOD provides a probabilistic prescription for how galaxies populate halos.
+According to HOD modeling, the number of galaxies that populate a dark matter halo is dictated by the host halo mass $M_h$.
+The probability of a dark matter halo of mass $M_h$ being populated by $N_g$ galaxies---$P(N_g|M_h)$--- is a Poisson distribution, and is separable to two parts
+\beq
+P(N_g|M_h) = P(N_\mathrm{cen}|M_h) + P(N_\mathrm{sat}|M_h),
+\eeq
+where $N_\mathrm{cen}$ is the number of central galaxies, and $N_\mathrm{sat}$ is the number of satellites.
+
+\subsubsection{Generating mock galaxy populations with halotools}
+
+%\todo{Halotools allows us to forward model galaxy positions from HOD parameters.
+From the galaxy positions, we can calculate observable summary statistics such as the 2PCF, Group Multiplicity Function (GMF) and the number density $\bar{n}$.}
+
+%Talk about why it is better to simulate the mock and compute %the summary statistics of the mock instead of using analytic formulas...
+
+%Talk about all the assumptions that go into analytic %formulas...
+
+%Talk about why we do not want to write down likelihood %function for gmf, 2pcf, ...
+\subsubsection{summary statistics of galaxy populations}
+
+\subsubsection{distance metrics}
+
+%\todo{Talk about the observable: 2pcf, group multiplicity %function, and number density ...}
+
+\section{Mock Observations}\label{sec:data}
+\todo{The fake observations we use as "data" in our inference.}
+%Plots showing 2pcf, gmf, ... of the mock data.
+%How we compute the covariance matrix from mocks...
+%Talk about error bars. Talk about poisson errors in group %populations ...
+
+\section{Results}\label{sec:results}
+Show the posteriors...
+Show evolution of the posterior errorbars with iteration. Show epsilon. Convice that we are converging ...
+Draw samples from the posterior, populate the mock catalogs with those samples, and compute the summary statistcs with them 
+and demonstrate that they match the data...
+
+\subsection{Comparison to MCMC}
+\todo{Apples to apples comparison between ABC and MCMC.}
+ABC-PMC using nbar and xi vs MCMC
+
+\subsection{ABC Including GMF}
+ABC-PMC results using observables beyond galaxy clustering and nbar(z) GMF
+
+\section{Discussion}\label{sec:discussion}
+\todo{talk about cosmological inference and why this is good first step for likelihood free inference in large scale ...}
+
+talk about halo model ...
+
+talk about HOD and why it is too simple to be the right model ...
+
+talk about assembly bias perhaps ... and how we can use the same machinery for detection of assemblt bias ...
+
+talk about writing down a likelihood for gmf, tpcf, ... , and the Gaussianity assumption ....
+
+talk about are results ...
+talk about future directions ...
+
+%%%%%%%%%% FIGURES %%%%%%%%%%
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{xi_covariance_Mr20_Nmock500.png}
+\caption{eft panel:
+Correlation coefficients of the error covariance matrix for the Mock galaxy two-point correlation function signal in Fig. 1, where the covariance matrix is computed from the 500 mock catalogs}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{gmf_covariance_Mr20_Nmock500.png}
+\caption{eft panel:
+Correlation coefficients of the error covariance matrix for the Mock galaxy group multiplicity function signal in Fig. 2, where the covariance matrix
+is computed from the 500 mock catalogs}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{tpcf_sigma.pdf}
+\caption{1-$\sigma$(dark-shaded) and two-$\sigma$ ABC posterior predictions
+for the two-point correlation function of galaxies. The black lines show the data-points.}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{tpcf_res_sigma3.pdf}
+\caption{1-$\sigma$(dark-shaded) and two-$\sigma$ ABC posterior model prediction residuals
+for the two-point correlation function of galaxies.}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{gmf3.png}
+\caption{{\bf Top Panel:}1-$\sigma$(dark-shaded) and 2-$\sigma$ ABC posterior model prediction 
+for the group multiplicity function of galaxies in the mock catalog.{\bf Bottom Panel:}Same as the top panel, but showing the residuals.}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\section*{acknowledgments}
+
+We would like to thank Mike Blanton, Boris Leidstadt, Micheal Schneider, Uros Seljak, Ross Fadely, Jeremy Tinker, Alex Malz, Patrick McDonald, Dan Foreman-Mackey, for productive and insightful discussions.
+MJV was supported by NSF grant AST-1517237. CHH was supported by James Arthur Graduate Award Grant.
+DWH was supported by NSF (grants IIS-1124794 and AST-1517237), NASA (grant NNX12AI50G), and the Moore-Sloan Data Science Environment at NYU.
+
+\bibliographystyle{yahapj}
+\bibliography{ccppabc}
+\end{document}
 \documentclass[12pt, preprint]{aastex}
@@
 \subsection{Forward model of the large scale galaxy populations}\label{sec:halo}
-\subsubsection{halo model}
+\subsubsection{halo occupation modeling}
+
+\newcommand{\lcdm}{\Lambda {\rm CDM}}
+\newcommand{\dd}{\mathrm{d}}
+\newcommand{\mean}[2]{\left\langle#1 \vert {#2}\right\rangle}
+
+%%% HALO OCCUPATION DISTRIBUTION %%%
+\newcommand{\ngal}{N_{\mathrm{g}}}
+\newcommand{\nsat}{N_\mathrm{s}}
+\newcommand{\ncen}{N_\mathrm{c}}
+\newcommand{\pnm}[2]{P(#1|#2)}
+
+\newcommand{\mhalo}{M_{\rm h}}
+\newcommand{\mvir}{M_\mathrm{vir}}
+
+\newcommand{\dndmvir}{\frac{\dd n}{\dd\mvir}}
+\newcommand{\dndmhalo}{\frac{\dd n}{\dd\mhalo}}
+\newcommand{\dndmvirprime}{\frac{\dd n}{\dd\mvir'}}
+
+%%% CLUSTERING STATISTICS %%%
+\newcommand{\xigg}{\xi_{\mathrm{gg}}}
+\newcommand{\xihh}{\xi_{\mathrm{hh}}}
+\newcommand{\xiggr}{\xi_{\mathrm{gg}}(r)}
+\newcommand{\xiggroneh}{\xi^{1h}_{\mathrm{gg}}(r)}
+\newcommand{\xiggrtwoh}{\xi^{2h}_{\mathrm{gg}}(r)}
+\newcommand{\ngalaxy}{\bar{n}_{\mathrm{g}}}
+
+One of the most well-studied statistics of the galaxy distribution 
+is $\xigg,$ the galaxy-galaxy correlation function. Mathematically, $\xigg$ encodes the
+excess probability that a pair of galaxies in a sample are 
+separated by three-dimensional distance $r.$ 
+Galaxy clustering is straightforward to measure observationally, 
+and it is a highly sensitive probe of both the physics of galaxy evolution 
+as well as the fundamental parameters of $\lcdm$ cosmology.
+
+The assumption that galaxies reside in dark matter halos is the bedrock underlying 
+all contemporary theoretical predictions for galaxy clustering. Since the 
+clustering of dark matter halos can be predicted to high precision with N-body simulations, 
+then knowledge of how galaxies populate and are distributed within halos 
+is sufficient to predict $\xigg.$ Under this assumption, $\xigg$ can be formally decomposed
+into a ``one-halo term" and a ``two-halo term'', 
+referring to galaxy pairs that reside in the same halo vs. distinct halos, respectively. 
+The two-halo term of $\xigg$ is a pair-number weighted average of $\xihh,$ 
+the halo-halo correlation function, whereas the one-halo term is the pair-number 
+weighted average of the intra-halo distribution of galaxies. 
+Models for the galaxy-halo connection aiming to predict $\xigg$ must therefore 
+provide theoretical means by which to calculate these weights and spatial distributions.
+
+The Halo Occupation Distribution (HOD) is one of the most widely used 
+ approaches to characterizing the galaxy-halo connection. 
+The central quantity In the HOD is $\pnm{\ngal}{\mhalo},$ the probability that a halo of mass
+$\mhalo$ hosts $\ngal$ galaxies. The first two moments of $\pnm$ contain the 
+necessary information to calculate the one- and two-halo terms of the galaxy correlation function:
+
+\begin{eqnarray}
+\label{eq:onehaloterm}
+1+\xiggroneh \simeq \frac{1}{4\pi{}r^{2}\ngalaxy^{2}}\int\dd\mhalo\dndmhalo\Xi_{\rm gg}(r|\mhalo) \times \mean{\ngal(\ngal-1)}{\mhalo},
+\end{eqnarray} and
+
+\begin{eqnarray}
+\label{eq:twohaloterm}
+\xiggrtwoh \simeq \xi_{\mathrm{mm}}(r)\left[\frac{1}{\ngalaxy}\int\dd\mhalo\dndmhalo \mean{\ngal}{\mhalo}b_{\mathrm{h}}(\mhalo)\right]^{2}
+\end{eqnarray}
+
+In Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}), $\ngalaxy$
+is the galaxy number density,
+$\dd\mathrm{n}/\dd\mhalo$ is the halo mass function, 
+the spatial bias of dark matter halos is $b_{\mathrm{h}}(\mhalo),$ and
+$\xi_{\rm mm}$ is the correlation function of dark matter.  
+If we represent the spherically symmetric intra-halo distribution of galaxies 
+by a unit-normalized $n_{\rm g}(r),$
+then the quantity $\Xi_{\rm gg}(r)$ appearing in the above two equations 
+is the convolution of $n_{\rm g}(r)$ with itself.
+
+Employing Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) to 
+calculate $\xigg(r)$ is the most common technical method used in HOD 
+studies of galaxy clustering. 
+However, this technique requires appeal to many simplifying assumptions 
+in addition to the supposition that galaxies reside in halos. For example, 
+Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) presume that 
+the intra-halo galaxy distribution is spherically symmetric, and 
+that the halo bias $b_{\mathrm{h}}(\mhalo)$ has no radial dependence. 
+Moreover, these equations does not take halo exclusion into account. 
+We refer the interested reader to \citet{cooray02,mo_vdb_white10,vdBosch13}, 
+and references therein, for further information.
+
+In addition to the above, in the vast majority of HOD studies the actual calculation of 
+Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) relies on further assumption still. 
+The quantities $b_{\mathrm{h}}(\mhalo),$ $\xi_{\mathrm{mm}}(r)$ and $\dndmhalo$ 
+are commonly replaced by analytical fitting functions that have been calibrated 
+against a suite of N-body simulations \citep[e.g.,][]{sheth_tormen01,smith03,tinker05,tinker10}.   
+And yet, it is well established that the precision of these fitting functions is 
+limited to the $\sim10\%$ level \citep[e.g.][]{tinker08}.
+
+\todo{APH: Ok, so the above level of detail may be overkill for this paper.
+I figured that part of the argument we'd make is this: it is now time to start taking the precision of our predictions seriously, the ten-percent game is over.
+ABC is a big contribution to going beyond ten percent, and now that Halotools makes it possible to circumvent the need for these fitting functions, this opens up new limiting factors such as the hard Gaussianity assumption.
+So I see the above argument as part of the argument for using ABC.
+However, I'm not married to any of the above text, so just let me know if you want me to scrap some of it.
+Otherwise, I will next pick up from here and start talking about how Halotools relaxes the reliance on these approximations, and I will make explicit the point about how this new level of precision motivates the exploration of ABC.}
+
+\todo{MJ: I think all of the text written above is necessary for making a case for simulation-based inference in large-scale structure modeling. So I'm in favor of keeping it for now.}
+
 Dark matter halos are biased tracers of the underlying matter distribution of the universe.
@@
 talk about are results ...
+talk about future directions ...
 
-talk about future directions ...
+%%%%%%%%%% FIGURES %%%%%%%%%%
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{xi_covariance_Mr20_Nmock500.png}
+\caption{eft panel:
+Correlation coefficients of the error covariance matrix for the Mock galaxy two-point correlation function signal in Fig. 1, where the covariance matrix is computed from the 500 mock catalogs}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{gmf_covariance_Mr20_Nmock500.png}
+\caption{eft panel:
+Correlation coefficients of the error covariance matrix for the Mock galaxy group multiplicity function signal in Fig. 2, where the covariance matrix
+is computed from the 500 mock catalogs}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{tpcf_sigma.pdf}
+\caption{1-$\sigma$(dark-shaded) and two-$\sigma$ ABC posterior predictions
+for the two-point correlation function of galaxies. The black lines show the data-points.}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{tpcf_res_sigma3.pdf}
+\caption{1-$\sigma$(dark-shaded) and two-$\sigma$ ABC posterior model prediction residuals
+for the two-point correlation function of galaxies.}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\begin{figure*}
+\begin{center}
+  \includegraphics[width=1.\textwidth]{gmf3.png}
+\caption{{\bf Top Panel:}1-$\sigma$(dark-shaded) and 2-$\sigma$ ABC posterior model prediction 
+for the group multiplicity function of galaxies in the mock catalog.{\bf Bottom Panel:}Same as the top panel, but showing the residuals.}
+\label{fig:2pcf-model}
+\end{center}
+\end{figure*}
+
+\section*{acknowledgments}
+
+We would like to thank Mike Blanton, Boris Leidstadt, Micheal Schneider, Uros Seljak, Ross Fadely, Jeremy Tinker, Alex Malz, Patrick McDonald, Dan Foreman-Mackey, for productive and insightful discussions.
+MJV was supported by NSF grant AST-1517237. CHH was supported by James Arthur Graduate Award Grant.
+DWH was supported by NSF (grants IIS-1124794 and AST-1517237), NASA (grant NNX12AI50G), and the Moore-Sloan Data Science Environment at NYU.
+
 \bibliographystyle{yahapj}

### 2016-01-25 halo/code/plotting.py
fixed a minor bug in plotting

@@ -110,7 +110,7 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
 	axes[i].yaxis.set_major_locator(MaxNLocator(5))
         axes[i].axhline(truths[i], color="#888888", lw=2)
         axes[i].set_ylabel(labels[i])
-    axes[5].set_xlabel("step number")    
+    axes[4].set_xlabel("step number")    
     fig.tight_layout(h_pad=0.0) 
     fig_file = ''.join([util.fig_dir(), 
         util.observable_id_flag(observables), 

### 2016-01-25 halo/code/emcee_emcee.py
mcmc multiprocessing running on sirocco'

@@ -249,5 +249,5 @@ def mcmc_multi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'xi'],
         f.close()
 
 if __name__=="__main__": 
-    mcmc_multi(20, 100, 1000, observables=['nbar', 'xi'], threads=1)
+    mcmc_multi(200, 100, 10000, observables=['nbar', 'xi'], threads=22)
     #mcmc_mpi(10, 1, 1, observables=['nbar', 'xi'])

### 2016-01-25 halo/code/emcee_emcee.py
new hdf5 output scheme

@@ -8,6 +8,7 @@ Author(s): Chang, MJ
 
 '''
 import sys
+import h5py
 import numpy as np
 import emcee
 from numpy.linalg import solve
@@ -165,7 +166,7 @@ def mcmc_mpi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'xi'],
     #sampler = emcee.EnsembleSampler(Nwalkers, Ndim, lnPost)
     pool.close()
     
-def mcmc_multi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'xi'], 
+def mcmc_multi(Nwalkers, Niter, observables=['nbar', 'xi'], 
         data_dict={'Mr':20, 'Nmock':500}, prior_name = 'first_try', threads=1): 
     '''
     Standard MCMC implementaion
@@ -175,10 +176,8 @@ def mcmc_multi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'xi'],
     -----------
     - Nwalker : 
         Number of walkers
-    - Nchains_burn : 
-        Number of burn-in chains
-    - Nchains_pro : 
-        Number of production chains   
+    - Niter : 
+        Number of chain iterations 
     - observables : 
         list of observables. Options are 'nbar', 'gmf', 'xi'
     - data_dict : dictionary that specifies the observation keywords
@@ -236,18 +235,21 @@ def mcmc_multi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'xi'],
 
     chain_file = ''.join([util.dat_dir(), 
         util.observable_id_flag(observables), 
-        '_Mr', str(data_dict["Mr"]), '_theta.mcmc_chain.dat'])
-    f = open(chain_file, "w")
+        '_Mr', str(data_dict["Mr"]), '_theta.Niter', str(Niter), '.mcmc_chain.hdf5'])
+    
+    f = h5py.File(chain_file, "w")
+    f.create_dataset('positions', data=np.zeros((Nwalkers*Niter, Ndim)))
     f.close()
-    for result in sampler.sample(pos0, iterations=Nchains_burn + Nchains_pro, storechain=False):
-        #print result 
+    i_start = 0 
+    for result in sampler.sample(pos0, iterations=Niter, storechain=False):
         position = result[0]
-        f = open(chain_file, "a")
-        for k in range(position.shape[0]):
[…]

### 2016-01-25 halo/code/util.py
util file updated to have directory stuff dealt with

@@ -3,6 +3,8 @@
 Utility modules 
 
 '''
+import os 
+__file__='util.py'
 
 def observable_id_flag(observables): 
     '''
@@ -14,10 +16,11 @@ def fig_dir():
     '''
     figure directory
     '''
-    return '../fig/'
+    fig_dir = os.path.dirname(os.path.realpath(__file__)).split('code')[0]+'fig/'
+    return fig_dir 
 
 def dat_dir(): 
     '''
     Dat directory
     '''
-    return '../dat/'
+    return os.path.dirname(os.path.realpath(__file__)).split('code')[0]+'dat/'

### 2016-01-25 halo/code/util.py
util directory fixed

@@ -4,7 +4,6 @@ Utility modules
 
 '''
 import os 
-__file__='util.py'
 
 def observable_id_flag(observables): 
     '''

### 2016-01-25 halo/code/emcee_emcee.py
hdf5 disconintued

@@ -8,7 +8,6 @@ Author(s): Chang, MJ
 
 '''
 import sys
-import h5py
 import numpy as np
 import emcee
 from numpy.linalg import solve
@@ -235,19 +234,21 @@ def mcmc_multi(Nwalkers, Niter, observables=['nbar', 'xi'],
 
     chain_file = ''.join([util.dat_dir(), 
         util.observable_id_flag(observables), 
-        '_Mr', str(data_dict["Mr"]), '_theta.Niter', str(Niter), '.mcmc_chain.hdf5'])
+        '_Mr', str(data_dict["Mr"]), '_theta.Niter', str(Niter), '.mcmc_chain.dat'])
     
-    f = h5py.File(chain_file, "w")
-    f.create_dataset('positions', data=np.zeros((Nwalkers*Niter, Ndim)))
+    #f = h5py.File(chain_file, "w")
+    #f.create_dataset('positions', data=np.zeros((Nwalkers*Niter, Ndim)))
+    #f.close()
+    f.open(chain_file, 'w')
     f.close()
     i_start = 0 
     for result in sampler.sample(pos0, iterations=Niter, storechain=False):
         position = result[0]
-        i_end = i_start + position.shape[0]
-        f = h5py.File(chain_file, "r+")
-        f['positions'][i_start:i_end] = position
+        f.open(chain_file, 'a')
+        for k in range(position.shape[0]): 
+            output_str = '\t'.join(position[k].astype('str')) + '\n'
+            f.write(output_str)
         f.close()
-        i_start = i_end
 
 if __name__=="__main__": 
     mcmc_multi(200, 10000, observables=['nbar', 'xi'], threads=10)

### 2016-01-25 halo/code/plotting.py

@@ -91,15 +91,13 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
         plot_range[:,1] = prior_max
     
     # chain files 
-    chain_file = ''.join([util.dat_dir(), 
-        util.observable_id_flag(observables), 
-        '_Mr', str(Mr), '_theta.mcmc_chain.hdf5'])
 
     chain_file = ''.join([util.dat_dir(), 
         util.observable_id_flag(observables), 
-        '_Mr', str(Mr), '_theta.mcmc_chain.hdf5'])
-    f = h5py.File(chain_file, 'r')
-    sample = f['positions'][:]
+        '_Mr', str(data_dict["Mr"]), '_theta.Niter', str(Nchain), '.mcmc_chain.dat'])
+    #f = h5py.File(chain_file, 'r')
+    #sample = f['positions'][:]
+    sample = np.loadtxt(chain_file)
     Ndim = len(sample[0])
     Nchain = len(sample) / Nwalkers 
     
@@ -148,9 +146,11 @@ def plot_mcmc_samples(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observable
     # chain files 
     chain_file = ''.join([util.dat_dir(), 
         util.observable_id_flag(observables), 
-        '_Mr', str(Mr), '_theta.mcmc_chain.hdf5'])
-    f = h5py.File(chain_file, 'r')
-    sample = f['positions'][:]
+        '_Mr', str(data_dict["Mr"]), '_theta.Niter', str(Nchain), '.mcmc_chain.dat'])
+    
+    #f = h5py.File(chain_file, 'r')
+    #sample = f['positions'][:]
+    sample = np.loadtxt(chain_file)
     Nchain = len(sample) / Nwalkers 
         
     fig = corner.corner(

### 2016-01-25 halo/code/emcee_emcee.py
emcee_emcee.py bug fixed

@@ -239,12 +239,12 @@ def mcmc_multi(Nwalkers, Niter, observables=['nbar', 'xi'],
     #f = h5py.File(chain_file, "w")
     #f.create_dataset('positions', data=np.zeros((Nwalkers*Niter, Ndim)))
     #f.close()
-    f.open(chain_file, 'w')
+    f = open(chain_file, 'w')
     f.close()
     i_start = 0 
     for result in sampler.sample(pos0, iterations=Niter, storechain=False):
         position = result[0]
-        f.open(chain_file, 'a')
+        f = open(chain_file, 'a')
         for k in range(position.shape[0]): 
             output_str = '\t'.join(position[k].astype('str')) + '\n'
             f.write(output_str)

### 2016-01-26 halo/code/emcee_emcee.py
playing around with mcmc to get it to work

@@ -241,7 +241,6 @@ def mcmc_multi(Nwalkers, Niter, observables=['nbar', 'xi'],
     #f.close()
     f = open(chain_file, 'w')
     f.close()
-    i_start = 0 
     for result in sampler.sample(pos0, iterations=Niter, storechain=False):
         position = result[0]
         f = open(chain_file, 'a')
@@ -251,6 +250,6 @@ def mcmc_multi(Nwalkers, Niter, observables=['nbar', 'xi'],
         f.close()
 
 if __name__=="__main__": 
-    mcmc_multi(200, 10000, observables=['nbar', 'xi'], threads=10)
+    mcmc_multi(100, 10000, observables=['nbar', 'xi'], threads=10)
     #mcmc_multi(10, 1, 5, observables=['nbar', 'xi'], threads=10)
     #mcmc_mpi(10, 1, 1, observables=['nbar', 'xi'])

### 2016-01-26 halo/code/plotting.py
plotting changes

@@ -70,7 +70,7 @@ def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=N
     plt.close()
 
 
-def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables=['nbar', 'xi'], 
+def plot_mcmc_chains(Nwalkers, Niter=10000, Nchains_burn=100, Mr=20, truths=None, observables=['nbar', 'xi'], 
         plot_range=None): 
     '''
     Plot MCMC chains
@@ -94,14 +94,14 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
 
     chain_file = ''.join([util.dat_dir(), 
         util.observable_id_flag(observables), 
-        '_Mr', str(data_dict["Mr"]), '_theta.Niter', str(Nchain), '.mcmc_chain.dat'])
+        '_Mr', str(Mr), '_theta.Niter', str(Niter), '.mcmc_chain.dat'])
     #f = h5py.File(chain_file, 'r')
     #sample = f['positions'][:]
     sample = np.loadtxt(chain_file)
     Ndim = len(sample[0])
-    Nchain = len(sample) / Nwalkers 
+    Nchain = len(sample)/Nwalkers
     
-    chain_ensemble = sample.reshape(Nchain , Nwalkers, Ndim)
+    chain_ensemble = sample.reshape(Nchain, Nwalkers, Ndim)
     fig , axes = plt.subplots(5, 1 , sharex=True, figsize=(10, 12))
 
     labels=[
@@ -109,7 +109,6 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
            ]
   
     for i in xrange(5):
-        
         axes[i].plot(chain_ensemble[:, :, i], color="k", alpha=0.4)
 	axes[i].yaxis.set_major_locator(MaxNLocator(5))
         axes[i].axhline(truths[i], color="#888888", lw=2)
@@ -118,13 +117,13 @@ def plot_mcmc_chains(Nwalkers, Nchains_burn=100, Mr=20, truths=None, observables
     fig.tight_layout(h_pad=0.0) 
     fig_file = ''.join([util.fig_dir(), 
         util.observable_id_flag(observables), 
-        '_Mr', str(Mr), '.Nchain', str(Nchain),
[…]

### 2016-01-26 halo/code/abc_pemcee.py
abc_pemcee.py for mercer

@@ -143,4 +143,4 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
     return pools
 
 if __name__=="__main__": 
-    ABCpmc_HOD(20, [1.e10,1.e10], N_part=1000, observables=['nbar', 'xi'])
+    ABCpmc_HOD(20, [1.e10,1.e10], N_part=1000, observables=['nbar', 'xi', 'gmf'])

### 2016-01-26 halo/code/abc_pemcee.py
abc_pemcee.py mercer flag

@@ -120,9 +120,9 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
                 Mr=data_dict["Mr"], truths=data_hod, plot_range=prior_range, observables=observables)
         # write theta and w to file 
         theta_file = ''.join([util.dat_dir(), util.observable_id_flag(observables), 
-            '_Mr', str(data_dict["Mr"]), '_theta_t', str(pool.t), '.dat'])
+            '_Mr', str(data_dict["Mr"]), '_theta_t', str(pool.t), '.mercer.dat'])
         w_file = ''.join([util.dat_dir(), util.observable_id_flag(observables), 
-            '_Mr', str(data_dict["Mr"]), '_w_t', str(pool.t), '.dat'])
+            '_Mr', str(data_dict["Mr"]), '_w_t', str(pool.t), '.mercer.dat'])
         np.savetxt(theta_file, pool.thetas)
         np.savetxt(w_file, pool.ws)
 

### 2016-01-26 halo/code/abc_pemcee.py
abc_pemcee.py: eps error fixed

@@ -143,4 +143,4 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
     return pools
 
 if __name__=="__main__": 
-    ABCpmc_HOD(20, [1.e10,1.e10], N_part=1000, observables=['nbar', 'xi', 'gmf'])
+    ABCpmc_HOD(20, [1.e10,1.e10,1.e10], N_part=1000, observables=['nbar', 'xi', 'gmf'])

### 2016-01-27 halo/code/abc_pemcee.py
abc_pemcee

@@ -23,7 +23,8 @@ from group_richness import richness
 # --- Plotting ---
 from plotting import plot_thetas
 
-def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nbar', 'gmf'], data_dict={'Mr':20, 'Nmock': 500}):
+def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nbar', 'gmf'], 
+        data_dict={'Mr':20, 'Nmock': 500}, output_dir=None):
     '''
     ABC-PMC implementation. 
 
@@ -35,6 +36,8 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
     - observables : list of observables. Options are 'nbar', 'gmf', 'xi'
     - data_dict : dictionary that specifies the observation keywords 
     '''
+    if output_dir is None: 
+        output_dir = util.dat_dir()
     # data observables
     fake_obs = []       # list of observables 
     for obv in observables: 
@@ -103,6 +106,7 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
     eps = abcpmc.MultiConstEps(T, eps_val)
     pools = []
     f = open("abc_tolerance.dat" , "w")
+    f.close()
     eps_str = ''
     for pool in abcpmc_sampler.sample(prior, eps):
         #while pool.ratio > 0.01:
@@ -119,7 +123,7 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
         plot_thetas(pool.thetas, pool.ws , pool.t, 
                 Mr=data_dict["Mr"], truths=data_hod, plot_range=prior_range, observables=observables)
         # write theta and w to file 
-        theta_file = ''.join([util.dat_dir(), util.observable_id_flag(observables), 
+        theta_file = ''.join([, util.observable_id_flag(observables), 
             '_Mr', str(data_dict["Mr"]), '_theta_t', str(pool.t), '.mercer.dat'])
         w_file = ''.join([util.dat_dir(), util.observable_id_flag(observables), 
             '_Mr', str(data_dict["Mr"]), '_w_t', str(pool.t), '.mercer.dat'])
[…]

### 2016-01-27 halo/code/abc_pemcee.py
directory fix for abc_pemcee.py

@@ -123,9 +123,9 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
         plot_thetas(pool.thetas, pool.ws , pool.t, 
                 Mr=data_dict["Mr"], truths=data_hod, plot_range=prior_range, observables=observables)
         # write theta and w to file 
-        theta_file = ''.join([, util.observable_id_flag(observables), 
+        theta_file = ''.join([output_dir, util.observable_id_flag(observables), 
             '_Mr', str(data_dict["Mr"]), '_theta_t', str(pool.t), '.mercer.dat'])
-        w_file = ''.join([util.dat_dir(), util.observable_id_flag(observables), 
+        w_file = ''.join([output_dir, util.observable_id_flag(observables), 
             '_Mr', str(data_dict["Mr"]), '_w_t', str(pool.t), '.mercer.dat'])
         np.savetxt(theta_file, pool.thetas)
         np.savetxt(w_file, pool.ws)

### 2016-01-27 halo/code/abc_pemcee.py
call sequence for arguments implemented for abc_pemcee.py so that code doesn't have to be adjusted for runs

@@ -4,8 +4,24 @@ Module for ABC-PMC inference
 
 Author(s): Chang, MJ
 
+Commandline call sequence : 
+
+python abc_pemcee.py Niter Npart ObsStr OutputDir
+
+- Niter : int
+    Number of iterations 
+- Npart : int
+    Number of particles 
+- ObsStr : string
+    String that specifies the observables. 
+    'nbarxi' is ['nbar', 'xi']
+    'nbargmf' is ['nbar', 'gmf']
+    'nbarxigmf' is ['nbar', 'xi', 'gmf']
+- OutputDir : string
+    String that specifies the output directory
 
 '''
+import sys 
 import time
 import pickle
 import numpy as np
@@ -147,4 +163,24 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
     return pools
 
 if __name__=="__main__": 
-    ABCpmc_HOD(20, [1.e10,1.e10,1.e10], N_part=10, observables=['nbar', 'xi', 'gmf'])
+
+    Niter = sys.argv[1]
+    print 'N iterations = ', Niter
+    Npart = sys.argv[2]
+    print 'N particles = ', Npart
+    obv_flag = sys.argv[3]
+    if obv_flag == 'nbarxi': 
+        obv_list = ['nbar', 'xi']
+    elif obv_flag == 'nbargmf':  
+        obv_list = ['nbar', 'gmf']
+    elif obv_flag == 'nbarxigmf':
+        obv_list = ['nbar', 'xi', 'gmf']
+    else: 
+        raise ValueError
+    print 'Observables: ', ', '.join(obv_list)
+    eps_list = [1.e10 for i in range(len(obv_list))]
+    if len(sys.argv) > 4: 
+        out_dir = sys.argv[4]
+        print 'Output to ', out_dir
+
+    #ABCpmc_HOD(Niter, eps_list, N_part=Npart, observables=obv_list)

### 2016-01-27 halo/code/abc_pemcee.py
small fix abc_pemcee.py

@@ -179,8 +179,10 @@ if __name__=="__main__":
         raise ValueError
     print 'Observables: ', ', '.join(obv_list)
     eps_list = [1.e10 for i in range(len(obv_list))]
+
     if len(sys.argv) > 4: 
         out_dir = sys.argv[4]
         print 'Output to ', out_dir
-
-    #ABCpmc_HOD(Niter, eps_list, N_part=Npart, observables=obv_list)
+        ABCpmc_HOD(Niter, eps_list, N_part=Npart, observables=obv_list, output_dir=out_dir)
+    else: 
+        ABCpmc_HOD(Niter, eps_list, N_part=Npart, observables=obv_list, output_dir=None)

### 2016-01-27 halo/code/abc_pemcee.py
call sequence variable type error abc_pemcee.py

@@ -164,9 +164,9 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
 
 if __name__=="__main__": 
 
-    Niter = sys.argv[1]
+    Niter = int(sys.argv[1])
     print 'N iterations = ', Niter
-    Npart = sys.argv[2]
+    Npart = int(sys.argv[2])
     print 'N particles = ', Npart
     obv_flag = sys.argv[3]
     if obv_flag == 'nbarxi': 

### 2016-01-27 halo/code/abc_pemcee.py
more fudging with abc_pemcee.py

@@ -182,6 +182,8 @@ if __name__=="__main__":
 
     if len(sys.argv) > 4: 
         out_dir = sys.argv[4]
+        if out_dir[-1] != '/': 
+            out_dir += '/'
         print 'Output to ', out_dir
         ABCpmc_HOD(Niter, eps_list, N_part=Npart, observables=obv_list, output_dir=out_dir)
     else: 

### 2016-01-27 halo/code/abc_pemcee.py
even more fudging. f'ck

@@ -54,6 +54,8 @@ def ABCpmc_HOD(T, eps_val, N_part=1000, prior_name='first_try', observables=['nb
     '''
     if output_dir is None: 
         output_dir = util.dat_dir()
+    else: 
+        pass
     # data observables
     fake_obs = []       # list of observables 
     for obv in observables: 

### 2016-01-27 halo/code/plot_posterior_model.py
plotting posterior model predictions

@@ -1,0 +1,100 @@
+'''
+
+Module for plotting model posterior predictions 
+for different observables
+Author(s): Chang, MJ
+
+'''
+import matplotlib.pyplot as plt
+import sys
+import numpy as np
+from numpy.linalg import solve
+plt.switch_backend("Agg")
+from halotools.empirical_models import PrebuiltHodModelFactory
+
+# --- Local ---
+import util
+import data as Data
+from hod_sim import HODsim
+from hod_sim import HODsimulator 
+from group_richness import richness
+from prior import PriorRange
+import corner
+    
+def plot_xi_model(theta, observables=['xi'], Mr=20):
+    
+    #load the data   
+ 
+    fake_obs = []
+    fake_obs_cov = [] 
+    for obv in observables: 
+        if obv == 'nbar': 
+            data_nbar, data_nbar_var = Data.data_nbar(**data_dict)
+            fake_obs.append(data_nbar)
+            fake_obs_cov.append(data_nbar_var)
+        if obv == 'gmf': 
+            data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
+            fake_obs.append(data_gmf)
+            fake_obs_cov.append(data_gmf)
+        if obv == 'xi': 
+            # import xir and full covariance matrix of xir
+            data_xi, data_xi_cov = Data.data_xi_full_cov(**data_dict)   
+            fake_obs.append(data_xi)
+            fake_obs_cov.append(data_xi_invcov)
+    
+    xi_gg = []
+
+    for i in xrange(len(theta)):
+        
+        mod = PrebuiltHodModelFactory('zheng07', threshold=-1*Mr)
+        mod.param_dict["logM0"] = theta[i][0]
+        mod.param_dict["sigma_logM"] = np.exp(theta[i][1])
+    	mod.param_dict["logMmin"] = theta[i][2]
+    	mod.param_dict["alpha"] = theta[i][3]
+    	mod.param_dict["logM1"] = theta[i][4]  
+        
+        mod.populate_mock()
+        """ if we want to make a smooth plot, we do this:"""
[…]

### 2016-01-27 halo/text/abc.tex
text update

@@
-ss[12pt, preprint]{aastex}
+\documentclass[12pt, preprint]{aastex}
 \usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
@@
 % We need a better definitely shorter title...
-\title{Approximate bayesian computation in large scale structure modeling : constraining the galaxy-halo connection}
+\title{Approximate Bayesian Computation in Large Scale Structure: Constraining the Galaxy-Halo Connection}
 \begin{abstract}
@@
 Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
-pro\-ba\-bil\-iites for the data given the theoretical model and parameter settings---%
+pro\-ba\-bil\-ities for the data given the theoretical model and parameter settings---%
 for combining data from different surveys and performing principled inferences of
@@
 the fact that the initial conditions are exceedingly close to Gaussian in form,
-and many sources of measurement noise are also Gaussian.
+and many sources of measurement noise are also Gaussian. (may as well cite evidence for this too)
 Likelihood functions are easier to write down and evaluate when things are closer to Gaussian,
@@
 
-%\todo{Talk about the observable: 2pcf, group multiplicity %function, and number density ...}
-
-\section{Mock Observations}\label{sec:data}
-\todo{The fake observations we use as "data" in our inference.}
+\section{Experiments}\label{sec:data}
+\subsection{Mock observations}
+\subsection{}
+In order to test the va
 %Plots showing 2pcf, gmf, ... of the mock data.
 %How we compute the covariance matrix from mocks...
-%Talk about error bars. Talk about poisson errors in group %populations ...
+%Talk about error bars. Talk about poisson errors in group %populations .
+%--------------------------------------------------------
+% Covariance matrix of xi  
+%--------------------------------------------------------
+\begin{figure}
+\includegraphics[width=0.5\textwidth]{xi_covariance_Mr20_Nmock500.png}
+\caption{The reduce covariance matrix $C_{i,j}/\sqrt{C_{i,i} C_{j,j}}$ of 
+the two-point correlation function, $\xi(r)$, estimated from 500 Halotools 
+realizations. We describe the calculation of $\xi(r)$ in Section \todo{reference section}. \todo{binning and range of $r$}.}
+\label{fig:2pcf_cov}
+\end{figure}
+%--------------------------------------------------------
+% Covariance matrix of GMF 
+%--------------------------------------------------------
+\begin{figure}
+  \includegraphics[width=0.5\textwidth]{gmf_covariance_Mr20_Nmock500.png}
+\caption{The reduce covariance matrix $C_{i,j}/\sqrt{C_{i,i} C_{j,j}}$ of 
+the Galaxy Multiplicity Function (GMF) estimated from 500 Halotools 
+realizations. \todo{reference to GMF calculation in text}. 
+\todo{binning and range of $r$}.}
+\label{fig:gmf_cov}
+\end{figure}
 
+%%%%%%%%%%%%%%%%%%%%%%% SECTION: %%%%%%%%%%%%%%%%%%%%%%%
+% SECTION: RESULTS 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}\label{sec:results}
@@
 ABC-PMC using nbar and xi vs MCMC
+%--------------------------------------------------------
+% ABC-PMC Posterior likelihoods using nbar and xi
+%--------------------------------------------------------
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{nbar_xi_Mr20_Niter10000_Nburn100_mcmc_samples_test.png}
+\caption{Posterior likelihood of HOD parameters $\log M_0$, 
+$\sigma_{\log M}$, $\log M_{min}$, $\alpha$, and $\log M_1$ using the
+standard Markov Chain Monte Carlo method as described in Section \todo{CITECITE}. 
+\todo{THIS IS A SPACE MARKER NOT THE FINAL FIGURE.}}
+\label{fig:post_abc_nbarxi}
+\end{center}
+\end{figure*}
+%--------------------------------------------------------
+% MC-MC Posterior likelihoods using nbar and xi
+%--------------------------------------------------------
+\begin{figure*}
+\begin{center}
+\includegraphics[width=\textwidth]{nbar_xi_Mr20_now_t11.png}
+\caption{Posterior likelihood of HOD parameters $\log M_0$, 
+$\sigma_{\log M}$, $\log M_{min}$, $\alpha$, and $\log M_1$ using ABC-PMC 
+with distance criteria defined by the observables $\bar{n}$ and $\xi(r)$.
+\todo{This is the 12th iteration. NOT THE FINAL.}}
+\label{fig:post_abc_nbarxi}
+\end{center}
+\end{figure*}
 
+%%%%%%%%%%%%%%%%%%%%%%% SECTION: %%%%%%%%%%%%%%%%%%%%%%%
+% ABC INCLUDING GMF
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{ABC Including GMF}
 ABC-PMC results using observables beyond galaxy clustering and nbar(z) GMF
+%--------------------------------------------------------
+% ABC-PMC Posterior likelihoods using nbar, xi and GMF 
+%--------------------------------------------------------
+\todo{Insert ABC-PMC Posterior Likelihoods using $\bar{n}$, $\xi(r)$ and GMF 
+once it finishes running}
 
@@
 
-%%%%%%%%%% FIGURES %%%%%%%%%%
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABC-PMC Posterior likelihoods using nbar and GMF 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure*}
 \begin{center}
-  \includegraphics[width=1.\textwidth]{xi_covariance_Mr20_Nmock500.png}
-\caption{eft panel:
-Correlation coefficients of the error covariance matrix for the Mock galaxy two-point correlation function signal in Fig. 1, where the covariance matrix is computed from the 500 mock catalogs}
-\label{fig:2pcf-model}
+\includegraphics[width=\textwidth]{nbar_gmf5_Mr20_now_t18.png}
+\caption{Posterior likelihood of HOD parameters $\log M_0$, 
+$\sigma_{\log M}$, $\log M_{min}$, $\alpha$, and $\log M_1$ using ABC-PMC 
+with distance criteria defined by the observables $\bar{n}$ and GMF.}
+\label{fig:post_abc_nbargmf}
 \end{center}
 \end{figure*}
-
-\begin{figure*}
-\begin{center}
-  \includegraphics[width=1.\textwidth]{gmf_covariance_Mr20_Nmock500.png}
-\caption{eft panel:
-Correlation coefficients of the error covariance matrix for the Mock galaxy group multiplicity function signal in Fig. 2, where the covariance matrix
-is computed from the 500 mock catalogs}
-\label{fig:2pcf-model}
-\end{center}
-\end{figure*}
-
 \begin{figure*}
@@
 for the two-point correlation function of galaxies. The black lines show the data-points.}
-\label{fig:2pcf-model}
+\label{fig:gmf_cov}
 \end{center}
@@
 
-\section*{acknowledgments}
+\section*{Acknowledgments}
 
-We would like to thank Mike Blanton, Boris Leidstadt, Micheal Schneider, Uros Seljak, Ross Fadely, Jeremy Tinker, Alex Malz, Patrick McDonald, Dan Foreman-Mackey, for productive and insightful discussions.
-MJV was supported by NSF grant AST-1517237. CHH was supported by James Arthur Graduate Award Grant.
-DWH was supported by NSF (grants IIS-1124794 and AST-1517237), NASA (grant NNX12AI50G), and the Moore-Sloan Data Science Environment at NYU.
+We would like to thank Mike Blanton, Boris Leidstadt, Micheal Schneider, 
+Uros Seljak, Ross Fadely, Jeremy Tinker, Alex Malz, Patrick McDonald, Dan 
+Foreman-Mackey, for productive and insightful discussions.
+
+MJV was supported by NSF grant AST-1517237. CHH was supported by James 
+Arthur Graduate Award Grant. DWH was supported by NSF (grants IIS-1124794 
+and AST-1517237), NASA (grant NNX12AI50G), and the Moore-Sloan Data Science 
+Environment at NYU.
+
+\todo{We need to acknowledge AstroHackWeek.}
 
@@
 \end{document}
-\documentclass[12pt, preprint]{aastex}
-\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
-\usepackage{graphicx}	% For figures
-\usepackage{natbib}	% For citep and citep
-\usepackage{amsmath}	% for \iint
-\usepackage{bm}
-\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
-\usepackage{hyperref}
-\usepackage{algorithmic,algorithm}
-\hypersetup{colorlinks}
-\usepackage{color}
-\usepackage{morefloats}
-\definecolor{darkred}{rgb}{0.5,0,0}
-\definecolor{darkgreen}{rgb}{0,0.5,0}
-\definecolor{darkblue}{rgb}{0,0,0.5}
-\hypersetup{ colorlinks,
-linkcolor=darkblue,
-filecolor=darkgreen,
-urlcolor=darkred,
-citecolor=darkblue}
-
-\DeclareMathOperator*{\argmax}{arg\,max}
-
-\newcommand{\todo}[1]{{\bf \textcolor{red}{ #1}}}
-\newcommand{\beq}{\begin{equation}}
-\newcommand{\eeq}{\end{equation}}
-\newcommand{\lang}{\langle}
-\newcommand{\ra}{\rangle}
-\newcommand{\vep}{\bm{\epsilon}}
-\newcommand{\ep}{\epsilon}
-\newcommand{\pars}{\vec{\theta}}
-\newcommand{\dev}{\mathrm{d}}
-
-\begin{document}
-% We need a better definitely shorter title...
-\title{Approximate bayesian computation in large scale structure modeling : constraining the galaxy-halo connection}
-\begin{abstract}
-\todo{This is not a paper about hod, let's not start with describing hod.
-I suggest that we talk about parameter estimation in large scale structure in general and then move on to hod.
-Constraining hod is a just the first step in testing our methodology.} The Halo Occupation Distribution (HOD) model is a powerful prescription of how galaxies
-populate dark matter halos that has had much success in reproducing large scale galaxy
-clustering and understanding galaxy evolution within the context of the halo model.
-Constraints on the parameters that describe the HOD model have so far been inferred using 
-the standard practice of Bayesian inference with likelihood estimation that assumes a
-functional Gaussian form of the likelihood.\todo{(Do we want to include analytic models for clustering statistics here?
-Yes, also we should say we use summary statistics that are not independent: clustering + group.
-Also there is no analytic model for the abundance of galaxy group richnesses)}  % I think we should keep it simple.
-There's no need to go into detail about everything in the abstract.
-This is already a pretty sizable abstract.
-Likelihood free inference methods such as the
-Approximate Bayesian Computation (ABC) makes Bayesian inference possible without ever 
-having to calculate the likelihood.
-
-In this work, we present ABC in the context of constraining the parameters of the HOD model
-using {\it Halotools}, a publicly available ... First, we set a mock galaxy catalog created
-with HOD parameters from \cite{Zheng:2007aa} \todo{why do we site here?} as our data catalog. Then we use ABC on a set 
-% To give it a false sense of legitimacy... 
-of observables measured from forward modeled mock catalogs generated with {\it Halotools} 
-in order to accurately infer constraints on the HOD parameters that 
-reproduce the original parameters. \todo{Remark about the comparison between our ABC 
-method constraints and the standard MCMC constraints.} Furthermore, since ABC does 
-not require computing the likelihood, we include observables beyond galaxy clustering 
-such as the galaxy number density and the galaxy group multiplicity function, for which
-accurately constructing a likelihood is intractable. 
-We demonstrate that adding these measurements to our set of observables leads to stronger
-constraints on the HOD parameters, specifically the ones that govern the satellite galaxy
-population. Moreover, the constraints we infer illustrate the feasibility of ABC in 
-cosmological contexts.
-
-\todo{Hogg writes here too}
-
-\end{abstract}
-
-\section{Introduction}
-
-Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
-pro\-ba\-bil\-iites for the data given the theoretical model and parameter settings---%
-for combining data from different surveys and performing principled inferences of
-the cosmological parameters (CITE CITE).
-Nowhere has this been more true than in cosmic microwave background studies,
-where it is nearly possible to analytically evaluate a likelihood function that
-involves no (or minimal) approximations (CITE Wandelt and others).
-
-Fundamentally, the tractability of likelihood functions in cosmology flows from
-the fact that the initial conditions are exceedingly close to Gaussian in form,
-and many sources of measurement noise are also Gaussian.
-Likelihood functions are easier to write down and evaluate when things are closer to Gaussian,
-so they get easier as the length scales get larger and the structure formation gets
-younger.
-Hence the suitability of CMB data for likelihood analyses.
-
-In large-scale structure studies, which have galaxies, quasars, or quasar absorption
-systems as tracers, the likelihood cannot be Gaussian:
-Even if the initial conditions are perfectly Gaussian, the growth of structure
-creates non-linearities which are non-Gaussian, and the galaxies form
-within the density field in some complex manner that is unknown (or modeled only effectively).
-Even if the galaxies were a Poisson sampling of the density field (which they are demonstrably
-not; CITE PEOPLE), it would be tremendously hard to write down even an approximate
-likelihood function.
-
-The standard approach is to make two strong assumptions, both demonstrably wrong:
-The first is that the (density and) two-point function measurements constitute
-sufficient statistics for the cosmological parameters.
-That is, the assumption is that all of the information about the cosmological model is
-encoded in the mean and variance statistics of the tracer population.
-This first assumption is equivalent to assuming that the density field is a Gaussian process,
-and is therefore closer to true on larger scales.
-
-The second strong and wrong assumption is that the likelihood functon for the data can
-be approximated by a pseudo-likelihood function that is a Gaussian probability density
-in the space of the two-point correlation function estimate.
-That is, the assumption is that a pseudo-likelihood of the data given by a squared distance
-(scaled by an inverse covariance matrix) between an estimate of the correlation function
-(the data) and a theoretical prediction of that same function will return sensible posterior
-inferences.
-This assumption cannot be correct (in detail) at any scale, since a correlation function,
-being related to the variance of a continuous field,
-must satisfy non-trivial positive-definiteness
-requirements (it must take the form of a valid kernel function; CITE SOMETHING).
-These requirements truncate function space such that the likelihood in that functon space
-could never be Gaussian.
-This second assumption becomes truer as the correlation function gets better measured, so
-it is most true on intermediate scales (where neither shot noise nor cosmic variance blow up).
-We expect it to be very strongly violated at the baryon acoustic scale in particular!
-
-Neither of these two assumptions are required for cosmological inferences, however:
-We have very sophisticated simulations and emulations of structure and galaxy formation.
-Indeed, the theory of large-scale structure \emph{is} the output of these simulations;
-we can simulate the data in great detail.
-Therefore, we can simulate not just the one- and two-point statistics of the galaxies,
-but also any higher order statistics that might be relevant.
-Furthermore, we can see the distributions of these statistics, and where the data lie within
-those distributions.
-That is---in principle---we don't need to make these strong and wrong assumptions;
-we can create---in principle---a likelihood function directly from (an enormous number of)
-simulation outputs.
-
-Of course, any naive approach to sufficiently simulating the data would be ruinously
-expensive.
-Fortunately, there are principled, (relatively) efficient methods for minimizing computation
-and delivering correct posterior inferences,
-using only a data simulator and some choices about statistics.
-Here we use \todo{Introduce the names ABC and LFI with references}.
-
-As we will see, the LFI approach---despite cleverness---will be very computationally expensive.
-For context, it is worth remembering that even the traditional approach to large-scale structure,
-outlined above, requires enormous numbers of structure- and galaxy-formation simulations as well,
-to construct the covariance matrices that go into the pseudo-likelihood.
-So although our costs will be high, there are no inexpensive methods to compare to.
-
-\todo{MJ: What Hogg wrote above doesn't mesh with (that is, repeats differently) what you have written below, so can you go through and fix that?
-Lots of what is below should remain, but some should get woven into what's above.}
-
-% style note: Here and in what follows: --- separates parenthetical phrases, -- connects words that appear together in a relationship (like galaxy--galaxy) and (galaxy--halo), and - connects pairs of words into an adjective (like two-point).
-
-Inference of the cosmological parameters and understanding the formation of galaxies within the context of large scale 
-structure, requires understanding of the galaxy--halo connection. Halos are biased tracers of the large scale structure 
-and galaxies reside in dark matter Halos.
-Halo occupation distribution---HOD---is a probabilistic framework that describes how galaxies populate dark matter halos.
-The central tenet of halo occupation model is that the probability that a certain number of galaxies populate a dark matter halo is determined by the mass of the host halo.
-This statistical prescription for connection of galaxies and halos has been successful in reproducing the galaxy two-point clustering as well as the galaxy--matter cross correlation---galaxy--galaxy
-lensing---within the context of the halo model, and has been used in understanding the co-evolution of galaxies and halos and putting constraints on the relation between the stellar mass and halo mass of galaxies.
-
-In the context of large scale clustering of galaxies, inference of the cosmological and halo model parameters are performed by treating some estimate of the galaxy--galaxy two-point correlation function as the data,
-assuming an analytic model for the theoretical two-point correlation function prediction,
-and a Gaussian likelihood connecting the two.
-The accuracy of the analytic model of the galaxy clustering statistics hinges upon the accuracy of the fitting functions such as halo bias, halo mass function that are accurate to ten percent level.
-There are other simplifying assumptions encoded in these analytic models such as the scale independence of the halo bias.
-
-Furthermore, the likelihood of observing the data---observed positions of the galaxies in the configuration space---given the cosmological and halo model parameters is approximated by a Gaussian function whose argument is the squared difference of the two-point correlation function measured from the observations and the analytic model for the galaxy clustering, weighted by a covariance matrix.
-\todo{Why MUST this be wrong?}
-The weighted difference of the number density of the observed galaxies and the analytic number density is also added to this.
-In this work, we investigate the how accurately we can infer the model parameters by relaxing the underlying assumptions of the parameter inference methods.
-
-Approximate bayesian computation (ABC) provides a \emph{rejected sampling} framework framework that alleviates the issues arising from approximating the likelihood of observing the large scale population of galaxies given the cosmological and halo model parameters.
-ABC approximates the posterior probability distribution function over the model parameters given the data---the angular positions of the galaxies on the sky and their redshifts---by drawing proposals from the prior over the model parameters, simulating the data with the proposals, and rejecting the proposals for which the distance between the data and the simulation (measured by a metric) is less than a certain small threshold.
-
-In practice, the rejected sampling method is replaced by a more efficient adaptive importance sampling operation called Population Monte Carlo (PMC).
-Initially, PMC rejects the draws from prior pdf with a relatively large threshold.
-In the consequent steps, importance sampling from the previous samples that passed the distance criteria is performed, and the threshold is updated adaptively.
-That is, after each iteration, the threshold is lowered according to the distribution of the distances.
-
-Furthermore, This sampling formalism permits us to sample the posterior of the model by simulating the data via a forward generative model that accounts for all the complications and uncertainties associated with the survey data that are difficult to take into account in approximating the likelihood function.
-In the context of the large galaxy surveys, these complications may include masks, nontrivial survey geometry, and missing data as a result of fiber collision.
-
-Rejected sampling in ABC requires a distance function that quantifies the closeness of the simulation (e.g.
-fiber collided galaxies) to the data (observed positions of the galaxies in the survey).
-In principle, this distance metric is a positive definite function  that compares various summary statistics of the data and those of the simulations.
-In this investigation, we focus on the two-point correlation, number density, and the group statistics (more specifically the abundance of the galaxy group richnesses).
-
-Performing a full cosmological inference using this sampling method requires a simulator that, (I) takes a set of cosmological parameters as input and generates a snapshot of dark matter particles at a given redshift using an N-body simulation, (II) builds a catalog of dark matter halos for that snapshot of the N-body simulation using a halo finder algorithm, and (III) paints galaxies into host halos given a set of halo occupation model parameters.
-
-Computationally, this process is extremely demanding.
-One way to make this problem more tractable is to split this into to separate this problem into two parts: \emph{emulation}, and \emph{simulation}.
-That is, one can use a set of N-body simulations with known cosmologies, and perform halo occupation simulations using the pre-built halo catalogs of those simulations.
-In order to compute the summary statistics corresponding to the samples from the parameter space for which there is no N-body simulation, one can interpolate between the summary statistics of the parameters for which we have available N-body simulations.
-This procedure is called emulation.
-Emulation has been introduced in the literature in coyote universe [CITE COYOTE].
-
-In this work however, we narrow our focus on inferring the posterior probability over the parameters of the halo model using an already existing halo catalog of a dark matter particle N-body simulation with a fixed cosmology.
-That is we simplify the problem by leaving the problem of emulation for future investigations.
-
-The calculations regarding populating the dark matter halos, computing the clustering statistics and finding groups in the galaxy mock catalogs are performed using ({\fontfamily{cmss}\selectfont Halotools}) which is an open-source package tailored for modeling galaxy-halo connection in large scale structure modeling ({\fontfamily{cmss}\selectfont http://halotools.readthedocs.org}).
-
-This paper is structured as follows. In Section \ref{sec:method},
-we discuss the components of the Halo occupation modeling, and the ABC sampling method.
-We also give a brief overview of the summary statistics used in this investigation.
-In Section \ref{sec:halo} we give a brief overview of the forward modeling of the galaxy mock catalogs using Halotools as well as generating a mock galaxy catalog as data.
-In Section \ref{sec:results} we explore the results of our parameter inference using different summary statistics.
-Finally, we discuss and conclude in Section \ref{sec:discussion}.
-
-\section{Method}\label{sec:method}
-
-\subsection{Approximate Bayesian Computation}
-
-ABC is a generative simulation-based inference based on rejection sampling.
-That is, the parameters of the forward model of the data are drawn from the prior probability, and are rejected if the simulated data is not sufficiently close to the observations according to some distance metric $\rho$.
-The method requires a prior probability over the parameters of interest $\pi(\theta)$, a generative forward model of the data given a set of parameters $f(\theta)$, and a distance metric $\rho(.,.)$ which measures the distance between the simulation $X=f(\theta)$ and the observed data $D$.
-
-In a large scale structure study however, the observed data---which encompasses the angular positions of galaxies on the sky and their redshifts---is a high dimensional object.
-Therefore the distance metric measures the closeness of statistical summaries of the data and those of the simulation.
-When the observed data-set under study is the positions of all galaxies observed in the survey, possible statistical summaries of the data are the number density of galaxies, two-point correlation function, three-point correlation function, galaxy group statistics (counts in cells), and so on.
-
-Here, we provide a brief overview of rejection sampling algorithm and how it is used in a simulation-based inference.
-We want to draw samples from a probability distribution $p(\alpha)$.
-Now, let us assume that: (i) there exists an auxiliary density $q(\alpha)$ from which we can directly sample from, and (ii) there exist $M>1$ which is an appropriate bound on $p(\alpha)/q(\alpha)$.
-Given these assumptions, rejection sampling works in the following way.
-We draw a sample $\alpha$ from the auxiliary density $q(\alpha)$.
-Afterwards, we draw a random number $u$ from a uniform distribution between zero and one.
-Then we accept the draw $\alpha$ if the $u$ is less than the ratio $p(\alpha)/Mq(\alpha)$.
-This procedure is repeated until sufficient number of draws are accepted.
-
-In the context of a simulation-based inference, we aim to sample from the joint probability of a simulation $X$, and parameters $\pars$ given the observed data $D$
-
-\beq
-p(\pars,X | D)
-\eeq
-
-which according to the Bayes rule can be written as
-
-\begin{eqnarray}
-p(\pars,X | D)  &=& \frac{p(D|X)p(X|\pars)\pi(\pars)}{Z},  \\
-Z &=& \int \dev \pars \; \dev X \; p(D|X)p(X|\pars)\pi(\pars),
-\end{eqnarray}
-where $Z$ is the evidence. There exists a constant $M>1$ 
-\beq 
-M = \frac{\mathrm{max}_{X} \; p(D|X)}{Z},
-\eeq 
-and auxiliary density $q(\pars ,X)$
-\beq 
-q(\pars ,X)=p(X|\pars )\pi(\pars ),
-\eeq 
-such that:
-\beq 
-p(\pars ,X|D)<Mq(\pars ,X).
-\eeq 
-Now in a simulation-based inference, we first draw a pair $\{\pars , X \}$ from the density $q(\pars , X)$.
-This requires drawing a sample $\pars$ from the prior $\pi(\pars)$, and generating a simulation $X=f(\pars)$ via the forward model.
-Then, we accept the pair $\{\pars , X \}$, if the following condition is met
-
-\beq 
-u \sim \mathrm{Uniform}[0,1] < \frac{p(\pars,X|D)}{Mq(\pars,X)}=\frac{p(D|X)}{\mathrm{max}_{X}p(D|X)}.
-\eeq
-
-ABC postulates that the probability of observing data $D$ given simulation $X$, $p(D|X)$ is controlled by a threshold $\epsilon$, and that this probability is one if the distance between the data and the simulation, $\rho(D,X)$ is less than $\epsilon$.
-That is
-
-\beq 
-p(D|X) = p_{\epsilon}(D|X) \equiv \begin{cases}
-                                    1 \ \ \  \mathrm{if} \ \rho(D,X)<\epsilon \; , \\
-                                    0 \ \ \ \mathrm{otherwise} \; .
-                                    \end{cases}
-\eeq
-
-Thus the set of parameters $\pars$ is accepted if $\rho(D,f(\pars))<\epsilon$. In principle, the distance function measures the distance 
-between the statistical summaries of the data and those of the simulation. 
-The distance, in general, can be a vector, with each component representing the distance between one summary statistics of the data (e.g.
-variance) and the same summary statistics of the simulation.
-That is
-\beq 
-\rho(D,X) = \Big[\rho_{1}(S_{1}(D) , S_{1}(X)), ... , \rho_{M}(S_{M}(D) , S_{M}(X)) \Big],
-\eeq 
-where $M$ is the number of statistical summaries used in the inference, $S_{i}$ is the i-th summary statistics, and $\rho_{i}$ is a scalar function that measures the distance between the i-th summary statistics of the data and that of the simulation.
-In this case, the threshold $\epsilon$ is an $M$ dimensional vector.
-
-We draw a large number of proposals from the prior distribution over the parameter space, 
-and reject the samples for which the distance between the data and the simulation corresponding to that 
-sample is more than a certain threshold $\bm{\epsilon}$. This requirement is satisfied \emph{if and only if}
-every component of the distance vector is less than its corresponding threshold. Note that, each summary statistics 
-has a separate threshold. This procedure is done iteratively until, we are left with a certain number
-of samples that satisfy the distance criteria. The final ensemble approximates the posterior probability distribution over the parameters of the generative model of the data.
-
-For a sufficiently small threshold, this algorithm samples from the posterior probability, $p(\pars | D)$, and for an arbitrarily large threshold, the algorithm samples from the prior $\pi(\pars)$.
-This rejection sampling technique requires repeated sampling from the prior, and a very small threshold to succeed.
-
-In this investigation instead, we use Population Monte Carlo Algorithm (PMC) as our sampling technique.
-PMC is an iterative method that performs rejected sampling from a sequence of distributions $\{p_{1}(\pars),...,p_{T}(\pars)\}$.These distributions are paired with a sequence of decreasing distance thresholds $\{\epsilon_{1} , ...
-, \epsilon_{T}\}$.
-
-At each iteration, $N$ proposals are drawn from a distribution over parameters, and a weight is assigned to each one.
-In the first step, the threshold is set to an arbitrarily large value.
-Therefore the algorithm makes $N$ draws from the prior and assigns equal weights to all the proposals.
-
-In the consequent iterations, the distance threshold is set to a smaller value.
-A proposal is weighted-sampled from the previous sample.
-It is perturbed by a kernel which is set to the covariance of the parameters in the previous sample.
-A simulation is then generated by feeding the perturbed parameter into the forward model.
-The perturbed parameter is then accepted if the distance between the data and the generated simulation is less than the value of the threshold in that iteration.
-This procedure is repeated until $N$ particles are sampled.
-Then the weights are updated according to the kernel, the prior, and the previous set of weights.
-
-\begin{algorithm}
-\caption{The procedure for ABC-PMC}
-\begin{algorithmic}[1]
-%\STATE \DATA: D
-%\STATE \RESULT: ABC posterior sample of $\pars$
-\IF{$t=1:$}
-%\STATE $\epsilon_t \gets \infty$
-\FOR{$i=1,...,N$}
-   \STATE // \emph{This loop can now be done in parallel for all i}
-   \WHILE{$\rho(X,D)>\epsilon_t$}
-   \STATE $\pars^{*}_{t} \gets \pi(\pars)$
-   \STATE $X = f(\pars^{*}_{t})$
-   \ENDWHILE
-   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
-   \STATE $w^{(i)}_{t} \gets 1/N$
-\ENDFOR
-\ENDIF
-\IF{$t=2,...,T:$}
-\FOR{$i=1,...,N$}
-   \STATE // \emph{This loop can now be done in parallel for all i}
-   \WHILE{$\rho(X,D)>\epsilon_t$}
-   \STATE Draw $\pars^{*}_{t}$ from $\{\pars^{j}_{t-1}\}$ with probabilities $\{w^{j}_{t-1}\}$
-   \STATE $\pars^{*}_{t} \gets K(\pars^{*}_{t},.)$
-   \STATE $X = f(\pars^{*}_{t})$
-   \ENDWHILE
-   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
-   \STATE $w^{(i)}_{t} \gets \frac{\pi(\pars^{(i)}_{t})}{\sum_{j=1}^{N}w_{t-1}^{(i)}K(\pars^{(j)}_{t-1},\pars^{(i)}_{t})}$
-\ENDFOR
-\ENDIF
-\end{algorithmic}
-\end{algorithm}
-
-%\begin{algorithm}[H]
-% \Data{this text}
-% \Result{how to write algorithm with \LaTeX2e }
-% initialization\;
-% \While{not at end of this document}{
-%  read current\;
-%  \eIf{understand}{
-%   go to next section\;
-%   current section becomes this one\;
-%   }{
-%   go back to the beginning of current section\;
-%  }
-% }
-% \caption{How to write algorithms}
-%\end{algorithm}
-
-\subsection{Forward model of the large scale galaxy populations}\label{sec:halo}
-\subsubsection{halo occupation modeling}
-
-\newcommand{\lcdm}{\Lambda {\rm CDM}}
-\newcommand{\dd}{\mathrm{d}}
-\newcommand{\mean}[2]{\left\langle#1 \vert {#2}\right\rangle}
-
-%%% HALO OCCUPATION DISTRIBUTION %%%
-\newcommand{\ngal}{N_{\mathrm{g}}}
-\newcommand{\nsat}{N_\mathrm{s}}
-\newcommand{\ncen}{N_\mathrm{c}}
-\newcommand{\pnm}[2]{P(#1|#2)}
-
-\newcommand{\mhalo}{M_{\rm h}}
-\newcommand{\mvir}{M_\mathrm{vir}}
-
-\newcommand{\dndmvir}{\frac{\dd n}{\dd\mvir}}
-\newcommand{\dndmhalo}{\frac{\dd n}{\dd\mhalo}}
-\newcommand{\dndmvirprime}{\frac{\dd n}{\dd\mvir'}}
-
-%%% CLUSTERING STATISTICS %%%
-\newcommand{\xigg}{\xi_{\mathrm{gg}}}
-\newcommand{\xihh}{\xi_{\mathrm{hh}}}
-\newcommand{\xiggr}{\xi_{\mathrm{gg}}(r)}
-\newcommand{\xiggroneh}{\xi^{1h}_{\mathrm{gg}}(r)}
-\newcommand{\xiggrtwoh}{\xi^{2h}_{\mathrm{gg}}(r)}
-\newcommand{\ngalaxy}{\bar{n}_{\mathrm{g}}}
-
-One of the most well-studied statistics of the galaxy distribution 
-is $\xigg,$ the galaxy-galaxy correlation function. Mathematically, $\xigg$ encodes the
-excess probability that a pair of galaxies in a sample are 
-separated by three-dimensional distance $r.$ 
-Galaxy clustering is straightforward to measure observationally, 
-and it is a highly sensitive probe of both the physics of galaxy evolution 
-as well as the fundamental parameters of $\lcdm$ cosmology.
-
-The assumption that galaxies reside in dark matter halos is the bedrock underlying 
-all contemporary theoretical predictions for galaxy clustering. Since the 
-clustering of dark matter halos can be predicted to high precision with N-body simulations, 
-then knowledge of how galaxies populate and are distributed within halos 
-is sufficient to predict $\xigg.$ Under this assumption, $\xigg$ can be formally decomposed
-into a ``one-halo term" and a ``two-halo term'', 
-referring to galaxy pairs that reside in the same halo vs. distinct halos, respectively. 
-The two-halo term of $\xigg$ is a pair-number weighted average of $\xihh,$ 
-the halo-halo correlation function, whereas the one-halo term is the pair-number 
-weighted average of the intra-halo distribution of galaxies. 
-Models for the galaxy-halo connection aiming to predict $\xigg$ must therefore 
-provide theoretical means by which to calculate these weights and spatial distributions.
-
-The Halo Occupation Distribution (HOD) is one of the most widely used 
- approaches to characterizing the galaxy-halo connection. 
-The central quantity In the HOD is $\pnm{\ngal}{\mhalo},$ the probability that a halo of mass
-$\mhalo$ hosts $\ngal$ galaxies. The first two moments of $\pnm$ contain the 
-necessary information to calculate the one- and two-halo terms of the galaxy correlation function:
-
-\begin{eqnarray}
-\label{eq:onehaloterm}
-1+\xiggroneh \simeq \frac{1}{4\pi{}r^{2}\ngalaxy^{2}}\int\dd\mhalo\dndmhalo\Xi_{\rm gg}(r|\mhalo) \times \mean{\ngal(\ngal-1)}{\mhalo},
-\end{eqnarray} and
-
-\begin{eqnarray}
-\label{eq:twohaloterm}
-\xiggrtwoh \simeq \xi_{\mathrm{mm}}(r)\left[\frac{1}{\ngalaxy}\int\dd\mhalo\dndmhalo \mean{\ngal}{\mhalo}b_{\mathrm{h}}(\mhalo)\right]^{2}
-\end{eqnarray}
-
-In Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}), $\ngalaxy$
-is the galaxy number density,
-$\dd\mathrm{n}/\dd\mhalo$ is the halo mass function, 
-the spatial bias of dark matter halos is $b_{\mathrm{h}}(\mhalo),$ and
-$\xi_{\rm mm}$ is the correlation function of dark matter.  
-If we represent the spherically symmetric intra-halo distribution of galaxies 
-by a unit-normalized $n_{\rm g}(r),$
-then the quantity $\Xi_{\rm gg}(r)$ appearing in the above two equations 
-is the convolution of $n_{\rm g}(r)$ with itself.
-
-Employing Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) to 
-calculate $\xigg(r)$ is the most common technical method used in HOD 
-studies of galaxy clustering. 
-However, this technique requires appeal to many simplifying assumptions 
-in addition to the supposition that galaxies reside in halos. For example, 
-Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) presume that 
-the intra-halo galaxy distribution is spherically symmetric, and 
-that the halo bias $b_{\mathrm{h}}(\mhalo)$ has no radial dependence. 
-Moreover, these equations does not take halo exclusion into account. 
-We refer the interested reader to \citet{cooray02,mo_vdb_white10,vdBosch13}, 
-and references therein, for further information.
-
-In addition to the above, in the vast majority of HOD studies the actual calculation of 
-Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) relies on further assumption still. 
-The quantities $b_{\mathrm{h}}(\mhalo),$ $\xi_{\mathrm{mm}}(r)$ and $\dndmhalo$ 
-are commonly replaced by analytical fitting functions that have been calibrated 
-against a suite of N-body simulations \citep[e.g.,][]{sheth_tormen01,smith03,tinker05,tinker10}.   
-And yet, it is well established that the precision of these fitting functions is 
-limited to the $\sim10\%$ level \citep[e.g.][]{tinker08}.
-
-\todo{APH: Ok, so the above level of detail may be overkill for this paper.
-I figured that part of the argument we'd make is this: it is now time to start taking the precision of our predictions seriously, the ten-percent game is over.
-ABC is a big contribution to going beyond ten percent, and now that Halotools makes it possible to circumvent the need for these fitting functions, this opens up new limiting factors such as the hard Gaussianity assumption.
-So I see the above argument as part of the argument for using ABC.
-However, I'm not married to any of the above text, so just let me know if you want me to scrap some of it.
-Otherwise, I will next pick up from here and start talking about how Halotools relaxes the reliance on these approximations, and I will make explicit the point about how this new level of precision motivates the exploration of ABC.}
-
-\todo{MJ: I think all of the text written above is necessary for making a case for simulation-based inference in large-scale structure modeling. So I'm in favor of keeping it for now.}
-
-Dark matter halos are biased tracers of the underlying matter distribution of the universe.
-The halo model postulates that galaxies reside in dark matter halos.
-Halo Occupation distribution---HOD---is a powerful tool for description of galaxy--halo connection.
-HOD provides a probabilistic prescription for how galaxies populate halos.
-According to HOD modeling, the number of galaxies that populate a dark matter halo is dictated by the host halo mass $M_h$.
-The probability of a dark matter halo of mass $M_h$ being populated by $N_g$ galaxies---$P(N_g|M_h)$--- is a Poisson distribution, and is separable to two parts
-\beq
-P(N_g|M_h) = P(N_\mathrm{cen}|M_h) + P(N_\mathrm{sat}|M_h),
-\eeq
-where $N_\mathrm{cen}$ is the number of central galaxies, and $N_\mathrm{sat}$ is the number of satellites.
-
-\subsubsection{Generating mock galaxy populations with halotools}
-
-%\todo{Halotools allows us to forward model galaxy positions from HOD parameters.
-From the galaxy positions, we can calculate observable summary statistics such as the 2PCF, Group Multiplicity Function (GMF) and the number density $\bar{n}$.}
-
-%Talk about why it is better to simulate the mock and compute %the summary statistics of the mock instead of using analytic formulas...
-
-%Talk about all the assumptions that go into analytic %formulas...
-
-%Talk about why we do not want to write down likelihood %function for gmf, 2pcf, ...
-\subsubsection{summary statistics of galaxy populations}
-
-\subsubsection{distance metrics}
-
-%\todo{Talk about the observable: 2pcf, group multiplicity %function, and number density ...}
-
-\section{Mock Observations}\label{sec:data}
-\todo{The fake observations we use as "data" in our inference.}
-%Plots showing 2pcf, gmf, ... of the mock data.
-%How we compute the covariance matrix from mocks...
-%Talk about error bars. Talk about poisson errors in group %populations ...
-
-\section{Results}\label{sec:results}
-Show the posteriors...
-Show evolution of the posterior errorbars with iteration. Show epsilon. Convice that we are converging ...
-Draw samples from the posterior, populate the mock catalogs with those samples, and compute the summary statistcs with them 
-and demonstrate that they match the data...
-
-\subsection{Comparison to MCMC}
-\todo{Apples to apples comparison between ABC and MCMC.}
-ABC-PMC using nbar and xi vs MCMC
-
-\subsection{ABC Including GMF}
-ABC-PMC results using observables beyond galaxy clustering and nbar(z) GMF
-
-\section{Discussion}\label{sec:discussion}
-\todo{talk about cosmological inference and why this is good first step for likelihood free inference in large scale ...}
-
-talk about halo model ...
-
-talk about HOD and why it is too simple to be the right model ...
-
-talk about assembly bias perhaps ... and how we can use the same machinery for detection of assemblt bias ...
-
-talk about writing down a likelihood for gmf, tpcf, ... , and the Gaussianity assumption ....
-
-talk about are results ...
-talk about future directions ...
-
-%%%%%%%%%% FIGURES %%%%%%%%%%
-
-\begin{figure*}
-\begin{center}
-  \includegraphics[width=1.\textwidth]{xi_covariance_Mr20_Nmock500.png}
-\caption{eft panel:
-Correlation coefficients of the error covariance matrix for the Mock galaxy two-point correlation function signal in Fig. 1, where the covariance matrix is computed from the 500 mock catalogs}
-\label{fig:2pcf-model}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-  \includegraphics[width=1.\textwidth]{gmf_covariance_Mr20_Nmock500.png}
-\caption{eft panel:
-Correlation coefficients of the error covariance matrix for the Mock galaxy group multiplicity function signal in Fig. 2, where the covariance matrix
-is computed from the 500 mock catalogs}
-\label{fig:2pcf-model}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-  \includegraphics[width=1.\textwidth]{tpcf_sigma.pdf}
-\caption{1-$\sigma$(dark-shaded) and two-$\sigma$ ABC posterior predictions
-for the two-point correlation function of galaxies. The black lines show the data-points.}
-\label{fig:2pcf-model}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-  \includegraphics[width=1.\textwidth]{tpcf_res_sigma3.pdf}
-\caption{1-$\sigma$(dark-shaded) and two-$\sigma$ ABC posterior model prediction residuals
-for the two-point correlation function of galaxies.}
-\label{fig:2pcf-model}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-  \includegraphics[width=1.\textwidth]{gmf3.png}
-\caption{{\bf Top Panel:}1-$\sigma$(dark-shaded) and 2-$\sigma$ ABC posterior model prediction 
-for the group multiplicity function of galaxies in the mock catalog.{\bf Bottom Panel:}Same as the top panel, but showing the residuals.}
-\label{fig:2pcf-model}
-\end{center}
-\end{figure*}
-
-\section*{acknowledgments}
-
-We would like to thank Mike Blanton, Boris Leidstadt, Micheal Schneider, Uros Seljak, Ross Fadely, Jeremy Tinker, Alex Malz, Patrick McDonald, Dan Foreman-Mackey, for productive and insightful discussions.
-MJV was supported by NSF grant AST-1517237. CHH was supported by James Arthur Graduate Award Grant.
-DWH was supported by NSF (grants IIS-1124794 and AST-1517237), NASA (grant NNX12AI50G), and the Moore-Sloan Data Science Environment at NYU.
-
-\bibliographystyle{yahapj}
-\bibliography{ccppabc}
-\end{document}

### 2016-01-27 halo/code/plot_posterior_model.py
residuals added to the posterior prediction plotting

@@ -71,9 +71,9 @@ def plot_xi_model(theta, observables=['xi'], Mr=20):
     ax.fill_between(rr, a, e, color="k", alpha=0.1, edgecolor="none")
     ax.fill_between(rr, b, d, color="k", alpha=0.3, edgecolor="none")
     ax.plot(rr, c, "k", lw=1)
-    ax.errorbar(rr, data_xi, yerr=np.sqrt(np.diag(data_xi_cov)), fmt=".k",
+    ax.errorbar(rr, data_xi, yerr = np.sqrt(np.diag(data_xi_cov)), fmt=".k",
                 capsize=0)    
-    xlabel = ax.set_xlabel('R [Mpc]', fontsize=20)
+    xlabel = ax.set_xlabel(r'$r[\mathrm{Mpc}h^{-1}]$', fontsize=20)
     ylabel = ax.set_ylabel(r'$\xi_{\rm gg}$', fontsize=25)
     
     plt.xscale('log')
@@ -88,8 +88,8 @@ def plot_xi_model(theta, observables=['xi'], Mr=20):
 
     ax.fill_between(rr, rr*a, rr*e, color="k", alpha=0.1, edgecolor="none")
     ax.fill_between(rr, rr*b, rr*d, color="k", alpha=0.3, edgecolor="none")
-    ax.plot(rr, c, "k", lw=1)
-    ax.errorbar(rr, rr*data_xi, yerr=rr*np.sqrt(np.diag(data_xi_cov)), fmt=".k",
+    #ax.plot(rr, c, "k", lw=1) we don't care about the best fit here
+    ax.errorbar(rr, rr*data_xi, yerr = rr*np.sqrt(np.diag(data_xi_cov)), fmt=".k",
                 capsize=0)
     xlabel = ax.set_xlabel(r'$r[\mathrm{Mpc}h^{-1}]$', fontsize=20)
     ylabel = ax.set_ylabel(r'$\xi_{\rm gg}$', fontsize=25)
@@ -97,4 +97,21 @@ def plot_xi_model(theta, observables=['xi'], Mr=20):
     plt.xscale('log')
     plt.xlim(xmin = .1 , xmax = 15)
     fig2.savefig('../figs/xi_scaled_posterior_prediction'+str(Mr)+'.pdf',
+                bbox_extra_artists=[xlabel, ylabel], bbox_inches='tight')
+
+    
+    fig3 = plt.figure()
+    ax = fig3.add_subplot(111)
+
[…]

### 2016-01-27 halo/code/plot_posterior_model.py
added joint model-residual plot

@@ -92,7 +92,7 @@ def plot_xi_model(theta, observables=['xi'], Mr=20):
     ax.errorbar(rr, rr*data_xi, yerr = rr*np.sqrt(np.diag(data_xi_cov)), fmt=".k",
                 capsize=0)
     xlabel = ax.set_xlabel(r'$r[\mathrm{Mpc}h^{-1}]$', fontsize=20)
-    ylabel = ax.set_ylabel(r'$\xi_{\rm gg}$', fontsize=25)
+    ylabel = ax.set_ylabel(r'$r\xi_{\rm gg}$', fontsize=25)
 
     plt.xscale('log')
     plt.xlim(xmin = .1 , xmax = 15)
@@ -113,5 +113,64 @@ def plot_xi_model(theta, observables=['xi'], Mr=20):
 
     plt.xscale('log')
     plt.xlim(xmin = .1 , xmax = 15)
-    fig2.savefig('../figs/xi_residual_posterior_prediction'+str(Mr)+'.pdf',
+    fig3.savefig('../figs/xi_residual_posterior_prediction'+str(Mr)+'.pdf',
+                bbox_extra_artists=[xlabel, ylabel], bbox_inches='tight')
+
+
+    fig4, axes = pl.subplots(2, 1, figsize=(10, 8) , sharex = True)
+    fig4.subplots_adjust(wspace=0.0 hspace=0.4)
+
+    ax1 = axes[0,0]
+    
+    ax1.fill_between(rr, a, e, color="k", alpha=0.1, edgecolor="none")
+    ax1.fill_between(rr, b, d, color="k", alpha=0.3, edgecolor="none")
+    #ax1.plot(rr, c, "k", lw=1)
+    ax1.errorbar(rr, data_xi, yerr = np.sqrt(np.diag(data_xi_cov)), fmt=".k",
+                capsize=0)    
+    ylabel = ax1.set_ylabel(r'$\xi_{\rm gg}$', fontsize=25)
+    ax1.set_xscale("log")
+    ax1.set_yscale("log")
+    ax1_set.xlim(xmin = .1 , xmax = 15)
+    
+    ax2 = axes[0,0]
+    ax2.fill_between(rr, a - data_xi, e- data_xi, color="k", alpha=0.1, edgecolor="none")
+    ax2.fill_between(rr, b- data_xi, d- data_xi, color="k", alpha=0.3, edgecolor="none")
[…]

### 2016-01-27 halo/code/plot_posterior_model.py
posterior predictions for gmf added

@@ -174,3 +174,278 @@ def plot_xi_model(theta, observables=['xi'], Mr=20):
 
     fig5.savefig('../figs/xi&residual_scaled_posterior_prediction'+str(Mr)+'.pdf',
                 bbox_extra_artists=[xlabel, ylabel], bbox_inches='tight')
+
+
+from group_richness import richness , gmf'''
+
+Module for plotting model posterior predictions 
+for different observables
+Author(s): Chang, MJ
+
+'''
+import matplotlib.pyplot as plt
+import sys
+import numpy as np
+from numpy.linalg import solve
+plt.switch_backend("Agg")
+from halotools.empirical_models import PrebuiltHodModelFactory
+
+# --- Local ---
+import util
+import data as Data
+from hod_sim import HODsim
+from hod_sim import HODsimulator 
+from group_richness import richness
+from prior import PriorRange
+import corner
+    
+def plot_xi_model(theta, observables=['xi'], Mr=20):
+    
+    #load the data   
+ 
+    fake_obs = []
+    fake_obs_cov = [] 
+    for obv in observables: 
+        if obv == 'nbar': 
+            data_nbar, data_nbar_var = Data.data_nbar(**data_dict)
+            fake_obs.append(data_nbar)
+            fake_obs_cov.append(data_nbar_var)
+        if obv == 'gmf': 
+            data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
+            fake_obs.append(data_gmf)
+            fake_obs_cov.append(data_gmf)
+        if obv == 'xi': 
+            # import xir and full covariance matrix of xir
+            data_xi, data_xi_cov = Data.data_xi_full_cov(**data_dict)   
+            fake_obs.append(data_xi)
+            fake_obs_cov.append(data_xi_invcov)
+    
+    xi_gg = []
+
+    for i in xrange(len(theta)):
+        
+        mod = PrebuiltHodModelFactory('zheng07', threshold=-1*Mr)
+        mod.param_dict["logM0"] = theta[i][0]
[…]

### 2016-01-27 halo/code/plot_posterior_model.py
fixed plot_posterior_model.py

@@ -15,350 +15,162 @@ from halotools.empirical_models import PrebuiltHodModelFactory
 # --- Local ---
 import util
 import data as Data
-from hod_sim import HODsim
-from hod_sim import HODsimulator 
-from group_richness import richness
-from prior import PriorRange
-import corner
-    
-def plot_xi_model(theta, observables=['xi'], Mr=20):
-    
-    #load the data   
- 
-    fake_obs = []
-    fake_obs_cov = [] 
-    for obv in observables: 
-        if obv == 'nbar': 
-            data_nbar, data_nbar_var = Data.data_nbar(**data_dict)
-            fake_obs.append(data_nbar)
-            fake_obs_cov.append(data_nbar_var)
-        if obv == 'gmf': 
-            data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
-            fake_obs.append(data_gmf)
-            fake_obs_cov.append(data_gmf)
-        if obv == 'xi': 
-            # import xir and full covariance matrix of xir
-            data_xi, data_xi_cov = Data.data_xi_full_cov(**data_dict)   
-            fake_obs.append(data_xi)
-            fake_obs_cov.append(data_xi_invcov)
-    
-    xi_gg = []
-
-    for i in xrange(len(theta)):
-        
-        mod = PrebuiltHodModelFactory('zheng07', threshold=-1*Mr)
-        mod.param_dict["logM0"] = theta[i][0]
-        mod.param_dict["sigma_logM"] = np.exp(theta[i][1])
-    	mod.param_dict["logMmin"] = theta[i][2]
-    	mod.param_dict["alpha"] = theta[i][3]
-    	mod.param_dict["logM1"] = theta[i][4]  
-        
-        mod.populate_mock()
-        """ if we want to make a smooth plot, we do this:"""
-        #rr , xi = mod.compute_average_compute_average_galaxy_clustering(num_iterations = 6 )
-        """else"""
-        rr , xi = mod.mock.compute_galaxy_clustering()
-        xi_gg.append(xi)
- 
-    xi_gg = np.array(xi_gg)
[…]

### 2016-01-27 halo/code/plot_posterior_model.py
fixed more bugs in model plotting

@@ -18,18 +18,15 @@ import data as Data
 from group_richness import gmf , richness , gmf_bins
 from prior import PriorRange
     
-def plot_xi_model(file_name="infered_hod_file_name", observables=['xi'], Mr=20):
+def plot_xi_model(file_name="infered_hod_file_name", observables=['xi'], 
+                  Mr=20 , data_dict={'Mr':20, 'Nmock':500}):
     
     #load the data   
     theta = np.loadtxt(file_name+".dat") 
-    fake_obs = []
-    fake_obs_cov = [] 
     for obv in observables: 
         if obv == 'xi': 
             # import xir and full covariance matrix of xir
             data_xi, data_xi_cov = Data.data_xi_full_cov(**data_dict)   
-            fake_obs.append(data_xi)
-            fake_obs_cov.append(data_xi_invcov)
     
     xi_gg = []
 
@@ -107,7 +104,7 @@ def plot_xi_model(file_name="infered_hod_file_name", observables=['xi'], Mr=20):
 
 
     fig4, axes = pl.subplots(2, 1, figsize=(10, 8) , sharex = True)
-    fig4.subplots_adjust(wspace=0.0 hspace=0.4)
+    fig4.subplots_adjust(wspace=0.0 , hspace=0.4)
 
     ax1 = axes[0,0]
     
@@ -137,14 +134,14 @@ def plot_xi_model(file_name="infered_hod_file_name", observables=['xi'], Mr=20):
 
 
     fig5, axes = pl.subplots(2, 1, figsize=(10, 8) , sharex = True)
-    fig5.subplots_adjust(wspace=0.0 hspace=0.4)
+    fig5.subplots_adjust(wspace=0.0 , hspace=0.4)
 
     ax1 = axes[0,0]
     
     ax1.fill_between(rr,rr* a, rr* e, color="k", alpha=0.1, edgecolor="none")
     ax1.fill_between(rr, rr* b, rr* d, color="k", alpha=0.3, edgecolor="none")
     #ax1.plot(rr, c, "k", lw=1)
-    ax1.errorbar(rr, rr* data_xi, rr* yerr = np.sqrt(np.diag(data_xi_cov)), fmt=".k",
+    ax1.errorbar(rr, rr* data_xi, yerr = rr*np.sqrt(np.diag(data_xi_cov)), fmt=".k",
                 capsize=0)    
     ylabel = ax1.set_ylabel(r'$r\xi_{\rm gg}$', fontsize=25)
[…]

### 2016-01-27 halo/code/emcee_emcee.py
put gmf gaussian likelihood in mcmc

@@ -66,7 +66,7 @@ def lnPost(theta, **kwargs):
         chi_tot += -0.5*(res_nbar)**2. / fake_obs_cov[ind] 
         ind += 1
     if 'gmf' in observables: 
-        raise NotImplementedError('GMF Likelihood has not yet been implemented')
+        chi_tot += -0.5*(res_gmf)**2. / fake_obs_cov[ind] 
     if 'xi' in observables: 
         chi_tot += -0.5*np.sum(np.dot(np.dot(res_xi , fake_obs_cov[ind]) , res_xi))
     lnLike = chi_tot
@@ -102,7 +102,7 @@ def mcmc_mpi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'xi'],
         if obv == 'gmf': 
             data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
             fake_obs.append(data_gmf)
-            fake_obs_cov.append(data_gmf)
+            fake_obs_cov.append(data_gmf_sigma**2.)
         if obv == 'xi': 
             # import xir and full covariance matrix of xir
             data_xi, data_xi_cov = Data.data_xi_full_cov(**data_dict)   
@@ -130,7 +130,7 @@ def mcmc_mpi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'xi'],
     # Initializing Walkers (@mjv : f@#k for loops)
     random_guess = np.array([11. , np.log(.4) , 11.5 , 1.0 , 13.5])
     pos0 = np.repeat(random_guess, Nwalkers).reshape(Ndim, Nwalkers).T + \
-            1e-1 * np.random.randn(Ndim * Nwalkers).reshape(Nwalkers, Ndim)
+            1e-3 * np.random.randn(Ndim * Nwalkers).reshape(Nwalkers, Ndim)
 
     # Initializing MPIPool
     pool = MPIPool()
@@ -194,7 +194,7 @@ def mcmc_multi(Nwalkers, Niter, observables=['nbar', 'xi'],
         if obv == 'gmf': 
             data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
             fake_obs.append(data_gmf)
-            fake_obs_cov.append(data_gmf)
+            fake_obs_cov.append(data_gmf_sigma**2.)
         if obv == 'xi': 
             # import xir and full covariance matrix of xir
             data_xi, data_xi_cov = Data.data_xi_full_cov(**data_dict)   

### 2016-01-27 halo/code/plot_posterior_model.py
fixed a few more bugs in model plotting for now

@@ -22,7 +22,7 @@ def plot_xi_model(file_name="infered_hod_file_name", observables=['xi'],
                   Mr=20 , data_dict={'Mr':20, 'Nmock':500}):
     
     #load the data   
-    theta = np.loadtxt(file_name+".dat") 
+    theta = np.loadtxt(file_name+".dat")[:3] 
     for obv in observables: 
         if obv == 'xi': 
             # import xir and full covariance matrix of xir
@@ -48,7 +48,6 @@ def plot_xi_model(file_name="infered_hod_file_name", observables=['xi'],
  
     xi_gg = np.array(xi_gg)
     a, b, c, d, e = np.percentile(xi_gg, [2.5, 16, 50, 84, 97.5], axis=0)
-
     
     fig1 = plt.figure()
     ax = fig1.add_subplot(111)
@@ -167,7 +166,7 @@ def plot_gmf_model(file_name="infered_hod_file_name", observables=['gmf'],
                    Mr=20, data_dict={'Mr':20, 'Nmock':500}):
     
     #load the data   
-    theta = np.loadtxt(file_name+".dat") 
+    theta = np.loadtxt(file_name+".dat")[:3] 
     
     for obv in observables: 
         if obv == 'gmf': 

### 2016-01-27 halo/code/plot_posterior_model.py
fixed moar bugs in model plotting

@@ -15,12 +15,21 @@ from halotools.empirical_models import PrebuiltHodModelFactory
 # --- Local ---
 import util
 import data as Data
-from group_richness import gmf , richness , gmf_bins
 from prior import PriorRange
-    
+from hod_sim import HODsimulator    
+from group_richness import gmf_bins
+
+# ---Need this for some smoothing stuff, will most-likely get rid of them
+from data import data_xi_bins
+from data import data_gmf_bins
+from data import hardcoded_xi_bins
+
+
 def plot_xi_model(file_name="infered_hod_file_name", observables=['xi'], 
-                  Mr=20 , data_dict={'Mr':20, 'Nmock':500}):
-    
+                  Mr=20 , data_dict={'Mr':20, 'Nmock':500} , smooth = "False"):
+    """
+    2PCF model 1\sigma & 2\sigma model predictions 
+    """    
     #load the data   
     theta = np.loadtxt(file_name+".dat")[:3] 
     for obv in observables: 
@@ -32,18 +41,18 @@ def plot_xi_model(file_name="infered_hod_file_name", observables=['xi'],
 
     for i in xrange(len(theta)):
         
-        mod = PrebuiltHodModelFactory('zheng07', threshold=-1*Mr)
-        mod.param_dict["logM0"] = theta[i][0]
-        mod.param_dict["sigma_logM"] = np.exp(theta[i][1])
-    	mod.param_dict["logMmin"] = theta[i][2]
-    	mod.param_dict["alpha"] = theta[i][3]
-    	mod.param_dict["logM1"] = theta[i][4]  
+        #mod = PrebuiltHodModelFactory('zheng07', threshold=-1*Mr)
+        #mod.param_dict["logM0"] = theta[i][0]
+        #mod.param_dict["sigma_logM"] = np.exp(theta[i][1])
+    	#mod.param_dict["logMmin"] = theta[i][2]
+    	#mod.param_dict["alpha"] = theta[i][3]
+    	#mod.param_dict["logM1"] = theta[i][4]  
         
-        mod.populate_mock()
+        #mod.populate_mock()
         """ if we want to make a smooth plot, we do this:"""
-        #rr , xi = mod.compute_average_compute_average_galaxy_clustering(num_iterations = 6 )
[…]

### 2016-01-27 halo/code/plotting.py
plotting.py : mcmc plotting fixed and prettied up

@@ -26,8 +26,13 @@ def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=N
             truths=truths,
             truth_color='#ee6a50', 
             labels=[
-                r'$\logM_{0}$',r'$\log \sigma_{\logM}$',r'$\logM_{min}$',r'$\alpha$',r'$\logM_{1}$'
+                r'$\mathtt{\log\;M_{0}}$',
+                r'$\mathtt{\log\;\sigma_{\logM}}$',
+                r'$\mathtt{\log\;M_{min}}$',
+                r'$\mathtt{\alpha}$',
+                r'$\mathtt{\log\;M_{1}}$'
                 ],
+            label_kwargs={'fontsize': 25},
             range=plot_range,
             quantiles=[0.16,0.5,0.84], 
             show_titles=True, 
@@ -69,7 +74,6 @@ def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=N
     plt.savefig(fig_file)
     plt.close()
 
-
 def plot_mcmc_chains(Nwalkers, Niter=10000, Nchains_burn=100, Mr=20, truths=None, observables=['nbar', 'xi'], 
         plot_range=None): 
     '''
@@ -105,15 +109,22 @@ def plot_mcmc_chains(Nwalkers, Niter=10000, Nchains_burn=100, Mr=20, truths=None
     fig , axes = plt.subplots(5, 1 , sharex=True, figsize=(10, 12))
 
     labels=[
-        r'$\logM_{0}$',r'$\log \sigma_{\logM}$',r'$\logM_{min}$',r'$\alpha$',r'$\logM_{1}$'
-           ]
+        r'$\mathtt{\log\;M_{0}}$',
+        r'$\mathtt{\log\;\sigma_{\logM}}$',
+        r'$\mathtt{\log\;M_{min}}$',
+        r'$\mathtt{\alpha}$',
+        r'$\mathtt{\log\;M_{1}}$'
+        ]
   
     for i in xrange(5):
         axes[i].plot(chain_ensemble[:, :, i], color="k", alpha=0.4)
 	axes[i].yaxis.set_major_locator(MaxNLocator(5))
         axes[i].axhline(truths[i], color="#888888", lw=2)
-        axes[i].set_ylabel(labels[i])
-    axes[4].set_xlabel("step number")    
[…]

### 2016-01-27 halo/code/plotting.py
plotting.py : plotting 1sigma 2sigma posterior model implemented for xi. it works for xi

@@ -13,6 +13,7 @@ from matplotlib.ticker import MaxNLocator
 import util
 import data as Data
 from prior import PriorRange
+from hod_sim import HODsimulator    
 plt.switch_backend("Agg")
 
 def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=None, filename=None): 
@@ -74,8 +75,8 @@ def plot_thetas(theta, w , t, Mr=20, truths=None, plot_range=None, observables=N
     plt.savefig(fig_file)
     plt.close()
 
-def plot_mcmc_chains(Nwalkers, Niter=10000, Nchains_burn=100, Mr=20, truths=None, observables=['nbar', 'xi'], 
-        plot_range=None): 
+def plot_mcmc(Nwalkers, Niter=10000, Nchains_burn=100, Mr=20, truths=None, 
+        observables=['nbar', 'xi'], plot_range=None): 
     '''
     Plot MCMC chains
     '''
@@ -95,13 +96,46 @@ def plot_mcmc_chains(Nwalkers, Niter=10000, Nchains_burn=100, Mr=20, truths=None
         plot_range[:,1] = prior_max
     
     # chain files 
-
     chain_file = ''.join([util.dat_dir(), 
         util.observable_id_flag(observables), 
         '_Mr', str(Mr), '_theta.Niter', str(Niter), '.mcmc_chain.dat'])
+    
     #f = h5py.File(chain_file, 'r')
     #sample = f['positions'][:]
     sample = np.loadtxt(chain_file)
+
+    # Posterior Likelihood Corner Plot     
+    fig = corner.corner(
+            sample[Nchains_burn*Nwalkers:], 
+            truths=truths,
+            truth_color='#ee6a50', 
+            labels=[
+                r'$\mathtt{\log\;M_{0}}$',
+                r'$\mathtt{\log\;\sigma_{\logM}}$',
+                r'$\mathtt{\log\;M_{min}}$',
+                r'$\mathtt{\alpha}$',
+                r'$\mathtt{\log\;M_{1}}$'
+                ],
+            label_kwargs={'fontsize': 25},
+            range=plot_range, 
+            quantiles=[0.16,0.5,0.84],
+            show_titles=True, 
+            title_args={"fontsize": 12},
+            plot_datapoints=True, 
+            fill_contours=True, 
+            levels=[0.68, 0.95], 
+            color='b', 
[…]

### 2016-01-27 halo/code/plotting.py
plotting.py : gmf implemented onto the plot posterior model *NOT YET TESTED*

@@ -197,8 +197,10 @@ def plot_posterior_model(observable, abc_theta_file=None, data_dict={'Mr':20, 'N
             model_obv = [obv_i]
     model_obv = np.array(model_obv)
 
-    if observable == 'xi': 
+    if 'xi' in observable: 
         r_bin = Data.data_xi_bins(Mr=data_dict['Mr'])
+    elif observable == 'gmf': 
+        r_bin = Data.data_gmf_bins()
 
     a, b, c, d, e = np.percentile(model_obv, [2.5, 16, 50, 84, 97.5], axis=0)
     a = a[0] 
@@ -235,6 +237,18 @@ def plot_posterior_model(observable, abc_theta_file=None, data_dict={'Mr':20, 'N
         ax.set_ylabel(r'$r \xi_{\rm gg}$', fontsize=25)
         ax.set_xscale('log')
         ax.set_xlim([0.1, 20.])
+
+    elif observable == 'gmf':   # GMF
+        data_gmf, data_gmf_sigma = Data.data_gmf(**data_dict)
+        ax.fill_between(r_bin, a, e, color="k", alpha=0.1, edgecolor="none")
+        ax.fill_between(r_bin, b, d, color="k", alpha=0.3, edgecolor="none")
+        ax.errorbar(r_bin, data_gmf, yerr = data_gmf_sigma, fmt=".k",
+                    capsize=0)    
+        ax.set_xlabel(r'Group Richness', fontsize=25)
+        ax.set_ylabel(r'GMF $[\mathrm{h}^3\mathrm{Mpc}^{-3}]$', fontsize=25)
+        
+        ax.set_yscale('log')
+        ax.set_xlim([1., 50.[)
 
     fig.savefig(
             ''.join([util.fig_dir(), 

### 2016-01-27 halo/code/plotting.py
plotting.py : save obvs(theta)

@@ -3,6 +3,7 @@
 Plotting modules 
 
 '''
+import os 
 import h5py
 import corner
 import numpy as np
@@ -167,7 +168,8 @@ def plot_mcmc(Nwalkers, Niter=10000, Nchains_burn=100, Mr=20, truths=None,
     plt.savefig(fig_file)
     plt.close()
 
-def plot_posterior_model(observable, abc_theta_file=None, data_dict={'Mr':20, 'Nmock':500}):
+def plot_posterior_model(observable, abc_theta_file=None, data_dict={'Mr':20, 'Nmock':500}, 
+        clobber=False):
     '''
     Plot 1\sigma and 2\sigma model predictions from ABC-PMC posterior likelihood 
 
@@ -186,16 +188,23 @@ def plot_posterior_model(observable, abc_theta_file=None, data_dict={'Mr':20, 'N
         obvs_str = 'xi'
     else: 
         obvs_str = observable 
-    for i in xrange(len(theta)):
-        #rr , xi = mod.compute_average_compute_average_galaxy_clustering(rbins = hardcoded_xi_bins , num_iterations = 6 )
-        obv_i  = HODsimulator(
-                theta[i], prior_range=None, 
-                observables=[obvs_str], Mr=data_dict['Mr'])
-        try: 
-            model_obv.append(obv_i)
-        except UnboundLocalError: 
-            model_obv = [obv_i]
-    model_obv = np.array(model_obv)
+
+    obvs_file = ''.join(abc_theta_file.rsplit('.dat')[:-1] + ['.', observable, '.dat'])
+    print obvs_file 
+    if not os.path.isfile(obvs_file) or clobber:
+        for i in xrange(len(theta)):
+            #rr , xi = mod.compute_average_compute_average_galaxy_clustering(rbins = hardcoded_xi_bins , num_iterations = 6 )
+            obv_i  = HODsimulator(
+                    theta[i], prior_range=None, 
+                    observables=[obvs_str], Mr=data_dict['Mr'])
+            try: 
+                model_obv.append(obv_i[0])
+            except UnboundLocalError: 
+                model_obv = [obv_i[0]]
+        model_obv = np.array(model_obv)
+        np.savetxt(obvs_file, model_obv)
+    else: 
[…]

### 2016-01-29 halo/code/data.py
made changes and added emceenoo.py to treat covariances correctly

@@ -1,47 +1,47 @@
 '''
 
-Modules for data and observables in ABC-PMC. 
+Modules for data and observables in ABC-PMC.
 
 '''
-import os 
+import os
 import numpy as np
 from numpy.linalg import solve
 
 import util
 from group_richness import gmf_bins
-from group_richness import richness 
+from group_richness import richness
 from group_richness import gmf as GMF
 from halotools.empirical_models import PrebuiltHodModelFactory
 
 def data_hod_param(Mr=20):
     '''
-    HOD parameters of 'observations'. Returns dictionary with hod parameters. 
+    HOD parameters of 'observations'. Returns dictionary with hod parameters.
     '''
     model = PrebuiltHodModelFactory('zheng07', threshold = -1.0*np.float(Mr))
     return model.param_dict
 
-# --- GMF --- 
-def data_gmf(Mr=20, Nmock=500): 
+# --- GMF ---
+def data_gmf(Mr=20, Nmock=500):
     ''' Observed GMF from 'data'
     '''
     gmf_dat_file = ''.join([util.dat_dir(), 'gmf.Mr', str(Mr), '.dat'])
     gmf_sig_file = ''.join([util.dat_dir(), 'gmf_sigma.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
     return [np.loadtxt(gmf_dat_file), np.loadtxt(gmf_sig_file)]
 
-def data_gmf_bins(): 
-    ''' Just for consistency  
+def data_gmf_bins():
+    ''' Just for consistency
     '''
     return gmf_bins()
 
-def data_gmf_cov(Mr=20, Nmock=500): 
+def data_gmf_cov(Mr=20, Nmock=500):
     '''
-    GMF covariance matrix 
+    GMF covariance matrix
     '''
     cov_file = ''.join([util.dat_dir(), 'gmf_cov.Mr', str(Mr), '.Nmock', str(Nmock), '.dat'])
     return np.loadtxt(cov_file)
 
-# --- nbar --- 
-def data_nbar(Mr=20, Nmock=500): 
+# --- nbar ---
+def data_nbar(Mr=20, Nmock=500):
     '''
     Observed nbar from 'data'
     '''
@@ -49,29 +49,29 @@ def data_nbar(Mr=20, Nmock=500):
[…]

### 2016-01-29 halo/code/emceenoo.py

@@ -1,0 +1,324 @@
+import os
+import sys
+import numpy as np
+import emcee
+from numpy.linalg import solve
+from emcee.utils import MPIPool
+import util
+import data as Data
+from hod_sim import HODsim
+from hod_sim import HODsimulator
+from group_richness import richness
+from prior import PriorRange
+import corner
+
+
+def lnPost(theta, **kwargs):
+    '''
+    log Posterior
+    '''
+    fake_obs = kwargs['data']
+    inv_cov = kwargs['inv_cov']
+    observables = kwargs['observables']
+    prior_range = kwargs['prior_range']
+    prior_min = prior_range[:,0]
+    prior_max = prior_range[:,1]
+
+    # Prior
+    if prior_min[0] < theta[0] < prior_max[0] and \
+       prior_min[1] < theta[1] < prior_max[1] and \
+       prior_min[2] < theta[2] < prior_max[2] and \
+       prior_min[3] < theta[3] < prior_max[3] and \
+       prior_min[4] < theta[4] < prior_max[4]:
+           lnPrior = 0.0
+    else:
+        lnPrior = -np.inf
+
+    if not np.isfinite(lnPrior):
+        return -np.inf
+
+    # Likelihood
+    model_obvs = HODsimulator(theta, **kwargs)
+    ind = 0
+    if 'nbar' in observables:
+        res_nbar = fake_obs[ind] - model_obvs[ind]
+        res = res_nbar
+        ind += 1
+    if 'gmf' in observables:
+        res_gmf = fake_obs[ind] - model_obvs[ind]
+        res = np.hstack((res, res_gmf))
+        ind += 1
+    if 'xi' in observables:
+        res_xi = fake_obs[ind] - model_obvs[ind]
+        res = np.hstack((res, res_xi))
+
+    lnLike = - 0.5 * np.sum(np.dot(np.dot(res, inv_cov), res))
+
+    return lnPrior + lnLike
+
+
+def mcmc_mpi(Nwalkers, Nchains_burn, Nchains_pro,
+             observables=['nbar', 'gmf'],
+             data_dict={'Mr':20, 'Nmock':500},
+             prior_name='first_try',
+             output_dir=None):
+    '''
+    Standard MCMC implementaion
+
+
+    Parameters
+    -----------
+    - Nwalker :
+        Number of walkers
+    - Nchains_burn :
+        Number of burn-in chains
+    - Nchains_pro :
+        Number of production chains
+    - observables :
[…]

### 2016-01-29 halo/code/data.py
include unbias factor in new invcovars

@@ -362,7 +362,8 @@ def build_full_inv_covars(Mr=20, Nmock=500):
 
     full_cov = data_full_cov(Mr=Mr, Nmock=Nmock)
     N_bins = int(np.sqrt(full_cov.size))
-    inv_c = solve(np.eye(N_bins) , full_cov)
+    f_unbias = np.float(Nmock - 2. - N_bins)/np.float(Nmock - 1.)
+    inv_c = solve(np.eye(N_bins) , full_cov) * f_unbias
 
     outfn = ''.join([util.dat_dir(),
                      'nbar_gmf_xir_inv_cov.Mr', str(Mr),
@@ -371,7 +372,8 @@ def build_full_inv_covars(Mr=20, Nmock=500):
 
     nbgmf_cov = data_nbar_gmf_cov(Mr=Mr, Nmock=Nmock)
     N_bins = int(np.sqrt(nbgmf_cov.size))
-    inv_c = solve(np.eye(N_bins) , nbgmf_cov)
+    f_unbias = np.float(Nmock - 2. - N_bins)/np.float(Nmock - 1.)
+    inv_c = solve(np.eye(N_bins) , nbgmf_cov) * f_unbias
 
     outfn = ''.join([util.dat_dir(),
                      'nbar_gmf_inv_cov.Mr', str(Mr),
@@ -380,7 +382,8 @@ def build_full_inv_covars(Mr=20, Nmock=500):
 
     nbxi_cov = data_nbar_xi_cov(Mr=Mr, Nmock=Nmock)
     N_bins = int(np.sqrt(nbxi_cov.size))
-    inv_c = solve(np.eye(N_bins) , nbxi_cov)
+    f_unbias = np.float(Nmock - 2. - N_bins)/np.float(Nmock - 1.)
+    inv_c = solve(np.eye(N_bins) , nbxi_cov) * f_unbias
 
     outfn = ''.join([util.dat_dir(),
                      'nbar_xi_inv_cov.Mr', str(Mr),
@@ -389,7 +392,8 @@ def build_full_inv_covars(Mr=20, Nmock=500):
 
     gmfxi_cov = data_gmf_xi_cov(Mr=Mr, Nmock=Nmock)
     N_bins = int(np.sqrt(gmfxi_cov.size))
-    inv_c = solve(np.eye(N_bins) , gmfxi_cov)
+    f_unbias = np.float(Nmock - 2. - N_bins)/np.float(Nmock - 1.)
[…]

### 2016-01-29 halo/code/emceenoo.py

@@ -103,14 +103,14 @@ def mcmc_mpi(Nwalkers, Nchains_burn, Nchains_pro,
             if xit:
                 inv_cov = Data.data_nb_gmf_xi_fullicov(**data_dict)
             else:
-                inv_cov = Data.data_nbar_gmf_inv_cov()
+                inv_cov = Data.data_nbar_gmf_inv_cov(**data_dict)
         elif xit:
-            inv_cov = Data.data_nbar_xi_inv_cov()
+            inv_cov = Data.data_nbar_xi_inv_cov(**data_dict)
         else:
             inv_cov = 1. / data_nbar_var
     elif gmt:
         if xit:
-            inv_cov = Data.data_gmf_xi_inv_cov()
+            inv_cov = Data.data_gmf_xi_inv_cov(**data_dict)
         else:
             inv_cov = 1. / data_gmf_sigma ** 2
     elif xit:

### 2016-01-30 emcee_emcee.py
mcmc now compatible with MPI

@@ -73,8 +73,8 @@ def lnPost(theta, **kwargs):
 
     return lnPrior + lnLike
 
-def mcmc_mpi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'gmf'], 
-        data_dict={'Mr':20, 'Nmock':500}, prior_name = 'first_try', output_dir=None): 
+def mcmc_mpi(Nwalkers, Niter, observables, continue_chain, output_dir=None, 
+        data_dict={'Mr':20, 'Nmock':500}, prior_name = 'first_try'): 
     '''
     Standard MCMC implementaion
     
@@ -154,7 +154,7 @@ def mcmc_mpi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'gmf'],
         Nchain = Niter
          
         # Initializing Walkers
-        random_guess = np.array([11. , np.log(.4) , 11.5 , 1.0 , 13.5])
+        random_guess = np.array([11.1 , np.log(.27) , 11.5 , 1.03 , 13.1])
         pos0 = np.repeat(random_guess, Nwalkers).reshape(Ndim, Nwalkers).T + \
                          1e-3 * np.random.randn(Ndim * Nwalkers).reshape(Nwalkers, Ndim)
 
@@ -174,12 +174,13 @@ def mcmc_mpi(Nwalkers, Nchains_burn, Nchains_pro, observables=['nbar', 'gmf'],
             }
     sampler = emcee.EnsembleSampler(Nwalkers, Ndim, lnPost, pool=pool, kwargs=hod_kwargs)
     # Initializing Walkers 
-    random_guess = np.array([11. , np.log(.4) , 11.5 , 1.0 , 13.5])
-    pos0 = np.repeat(random_guess, Nwalkers).reshape(Ndim, Nwalkers).T + \
-                1e-3 * np.random.randn(Ndim * Nwalkers).reshape(Nwalkers, Ndim)
+    #random_guess = np.array([11. , np.log(.4) , 11.5 , 1.0 , 13.5])
+    #pos0 = np.repeat(random_guess, Nwalkers).reshape(Ndim, Nwalkers).T + \
+    #            1e-3 * np.random.randn(Ndim * Nwalkers).reshape(Nwalkers, Ndim)
 
 
     for result in sampler.sample(pos0, iterations=Nchain, storechain=False):
+        print "12"
         position = result[0]
         f = open(chain_file, 'a')
         for k in range(position.shape[0]): 
[…]

### 2016-01-30 emcee_emcee.py
mcmc now compatible with MPI

@@ -180,7 +180,7 @@ def mcmc_mpi(Nwalkers, Niter, observables, continue_chain, output_dir=None,
 
 
     for result in sampler.sample(pos0, iterations=Nchain, storechain=False):
-        print "12"
+
         position = result[0]
         f = open(chain_file, 'a')
         for k in range(position.shape[0]): 

### 2016-01-30 halo/text/abc.tex

@@
 Inference of the cosmological parameters and understanding the formation of galaxies within the context of large scale 
-structure, requires understanding of the galaxy--halo connection. Halos are biased tracers of the large scale structure 
-and galaxies reside in dark matter Halos.
+structure typically assumes the existence of dark matter halos (CITE) and therefore requires an understanding of the galaxy--halo connection (CITE).
+Halos are biased tracers of the large scale structure (CITE), the statistical properties of which are well understood from the study of dark matter simulations (12 CHANNELS PEOPLE, 12!),
+and the general assumption is that galaxies reside in dark matter Halos \todo{evidence for this claim or rather, evidence that supports this assumption}.
 Halo occupation distribution---HOD---is a probabilistic framework that describes how galaxies populate dark matter halos.
@@
 This statistical prescription for connection of galaxies and halos has been successful in reproducing the galaxy two-point clustering as well as the galaxy--matter cross correlation---galaxy--galaxy
-lensing---within the context of the halo model, and has been used in understanding the co-evolution of galaxies and halos and putting constraints on the relation between the stellar mass and halo mass of galaxies.
+lensing---within the context of the halo model, and has been used in understanding the co-evolution of galaxies and halos (how?
+CITE) and putting constraints on the relation between the stellar mass and halo mass of galaxies and clusters (M-N ratio?) as well as cosmological parameters.
 
@@
 There are other simplifying assumptions encoded in these analytic models such as the scale independence of the halo bias.
-
-Furthermore, the likelihood of observing the data---observed positions of the galaxies in the configuration space---given the cosmological and halo model parameters is approximated by a Gaussian function whose argument is the squared difference of the two-point correlation function measured from the observations and the analytic model for the galaxy clustering, weighted by a covariance matrix.
-\todo{Why MUST this be wrong?}
 The weighted difference of the number density of the observed galaxies and the analytic number density is also added to this.
@@
 
-Approximate bayesian computation (ABC) provides a \emph{rejected sampling} framework framework that alleviates the issues arising from approximating the likelihood of observing the large scale population of galaxies given the cosmological and halo model parameters.
-ABC approximates the posterior probability distribution function over the model parameters given the data---the angular positions of the galaxies on the sky and their redshifts---by drawing proposals from the prior over the model parameters, simulating the data with the proposals, and rejecting the proposals for which the distance between the data and the simulation (measured by a metric) is less than a certain small threshold.
+Approximate bayesian computation (ABC) provides a \emph{rejected sampling} framework that alleviates the issues arising from approximating the likelihood of observing the large scale population of galaxies given the cosmological and halo model parameters.
+ABC approximates the posterior probability distribution function over the model parameters given the data---the angular positions of the galaxies on the sky and their redshifts---by drawing proposals from the prior over the model parameters, simulating the data with the proposals, and rejecting the proposals for which the distance between the data and the simulation (measured by a metric) is greater than a certain small threshold.
 
@@
 Computationally, this process is extremely demanding.
-One way to make this problem more tractable is to split this into to separate this problem into two parts: \emph{emulation}, and \emph{simulation}.
+One way to make this problem more tractable is to separate it into two parts: \emph{emulation}, and \emph{simulation}.
 That is, one can use a set of N-body simulations with known cosmologies, and perform halo occupation simulations using the pre-built halo catalogs of those simulations.
@@
 This paper is structured as follows. In Section \ref{sec:method},
-we discuss the components of the Halo occupation modeling, and the ABC sampling method.
+we discuss the ABC sampling method and the components of the Halo occupation modeling.
 We also give a brief overview of the summary statistics used in this investigation.
-In Section \ref{sec:halo} we give a brief overview of the forward modeling of the galaxy mock catalogs using Halotools as well as generating a mock galaxy catalog as data.
+In Section \ref{sec:halo} we give a brief overview of the forward modeling of the galaxy mock catalogs using halotools as well as generating a mock galaxy catalog as data.
 In Section \ref{sec:results} we explore the results of our parameter inference using different summary statistics.
@@
 
+%%%%%%%%%%%%%%%%%%%%% SECTION %%%%%%%%%%%%%%%%%%%%%%%%%%
+% Approximate Bayesian Computation  
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{Approximate Bayesian Computation}
 
-ABC is a generative simulation-based inference based on rejection sampling.
-That is, the parameters of the forward model of the data are drawn from the prior probability, and are rejected if the simulated data is not sufficiently close to the observations according to some distance metric $\rho$.
-The method requires a prior probability over the parameters of interest $\pi(\theta)$, a generative forward model of the data given a set of parameters $f(\theta)$, and a distance metric $\rho(.,.)$ which measures the distance between the simulation $X=f(\theta)$ and the observed data $D$.
+ABC is based on rejection sampling, which is a Monte Carlo method 
+used to draw samples from a probability distribution that is difficult 
+to directly sample. For inference, the probability distribution that 
+we want to sample is the posterior probability distribution function.  
+Before we present the theory and algorithm of ABC, we provide a brief
+overview of rejection sampling.
 
-In a large scale structure study however, the observed data---which encompasses the angular positions of galaxies on the sky and their redshifts---is a high dimensional object.
+Briefly, rejection sampling is a method that draws samples from a 
+probability distribution $f(\alpha)$, which is difficult to 
+directly sample, by sampling from the distribution $M g(\alpha)$ that 
+satisfies $f(\alpha) < M g(\alpha)$. $g(\alpha)$ is an instrumental 
+distribution where sampling is easier and $M > 1$.
+
+More specifically, in the context of simulation-based inference, 
+the main goal is to sample from the joint probability of a
+simulation $X$ and parameters $\pars$ given observed data $D$, which 
+from Bayes rule can be written as 
+\beq
+p(\pars, X | D) = \frac{P(D|X)p(X|\pars)\pi(\pars)}{\mathcal{Z}}
+\eeq
+where $\pi(\pars)$ is the prior probabilities over the parameters of 
+interest and $\mathcal{Z}$ is the evidence, 
+\beq
+\mathcal{Z} = \int d\pars \; dX\; p(D|X) p(X|\pars) \pi(\pars). 
+\eeq
+Since $p(\pars, X | D)$ cannot be directly sampled, we use rejection 
+sampling with instrumental distribution 
+\beq
+q(\pars, X) = p(X|\pars) \pi(\pars)
+\eeq
+and 
+\beq
+M = \frac{\mathrm{max}\; p(D|X)}{\mathcal{Z}} > 1.
+\eeq
+The choices of $q(\pars, X)$ and $M$ satisfy the condition 
+\beq
+p(\pars, X | D) < M q(\pars, X)
+\eeq
+so we can sample $p(\pars, X | D)$ by drawing ${\pars, X}$ from 
+the $q(\pars, X)$ distribution. In practice, this is done 
+by first drawing $\pars$ from the prior $\pi(\pars)$ then generating a 
+simulation $X = f(\pars)$ via the forward model. Afterwards, ${\pars, X}$ 
+is accepted if
+\beq
+u < \frac{p(\pars, X | D)}{M q(\pars, X)} = \frac{p(D|X)}{\mathrm{max}\;p(D|X)}
+\eeq
+where $u$ is drawn from $\mathtt{Uniform}[0,1]$.
+
+At this point, ABC postulates that $p(D|X)$
+
+The method requires a prior probability over the parameters of interest 
+($\pi(\theta)$), a generative forward model of the data given a set of 
+parameters ($f(\theta)$) and a distance metric $\rho(x,D)$ which measures 
+the distance between the simulation $X = f(\theta)$ and the observed data $D$.
+
+Technically in large scale structure studies, the observed data is the 
+total angular and redshift positions of galaxies on the sky. However, 
+in practice, this data is an intractably high dimensional object. As a result, we 
+utilize statistical summaries of galaxy positions as our observed 'data' 
+and the distance metric we use measures the closeness of the statistical 
+summaries of observed galaxy positions versus the simulated galaxy 
+positions. Common summary statistics used to encapsulate the galaxy position 
+information include the number density of the galaxy sample, the two-point 
+correlation function (2PCF), three-point correlation function, and galaxy 
+group statistics.
+
 Therefore the distance metric measures the closeness of statistical summaries of the data and those of the simulation.
@@
 
-In the context of a simulation-based inference, we aim to sample from the joint probability of a simulation $X$, and parameters $\pars$ given the observed data $D$
-
+In the context of a simulation-based inference, we main goal is to 
+sample from the joint probability of a simulation $X$ and parameters 
+$\pars$ given the observed data $D$: 
 \beq
@@
 \eeq
+According to Bayes rule $p(\pars, X | D)$ can be written as 
+\beq
+p(\pars,X | D)  = \frac{p(D|X)p(X|\pars)\pi(\pars)}{\mathcal{Z}}
+\eeq
+where $\mathcal{Z}$  is the evidence, 
+\beq
+\mathcal{Z} = \int \dev \pars \; \dev X \; p(D|X)p(X|\pars)\pi(\pars),
+\eeq
 
-which according to the Bayes rule can be written as
-
-\begin{eqnarray}
-p(\pars,X | D)  &=& \frac{p(D|X)p(X|\pars)\pi(\pars)}{Z},  \\
-Z &=& \int \dev \pars \; \dev X \; p(D|X)p(X|\pars)\pi(\pars),
-\end{eqnarray}
-where $Z$ is the evidence. There exists a constant $M>1$ 
+There exists a constant $M>1$ 
 \beq 
@@
 separated by three-dimensional distance $r.$ 
-Galaxy clustering is straightforward to measure observationally, 
+Galaxy clustering is straightforward to measure observationally \todo{(redshift space distortions?)}, 
 and it is a highly sensitive probe of both the physics of galaxy evolution 
@@
 The halo model postulates that galaxies reside in dark matter halos.
-Halo Occupation distribution---HOD---is a powerful tool for description of galaxy--halo connection.
+Halo Occupation distribution---HOD---is a powerful tool for description of the galaxy--halo connection.
 HOD provides a probabilistic prescription for how galaxies populate halos.
-According to HOD modeling, the number of galaxies that populate a dark matter halo is dictated by the host halo mass $M_h$.
+According to (conventional) HOD modeling, the number of galaxies that populate a dark matter halo is dictated by the host halo mass $M_h$.
 The probability of a dark matter halo of mass $M_h$ being populated by $N_g$ galaxies---$P(N_g|M_h)$--- is a Poisson distribution, and is separable to two parts
@@
 where $N_\mathrm{cen}$ is the number of central galaxies, and $N_\mathrm{sat}$ is the number of satellites.
+Given a halo catalog generated from the output of a cosmological N-body simulation, these probability density functions can be used to make draws for each halo to find its occupation number of central and satellite galaxies.
 
@@
 
-%\todo{Halotools allows us to forward model galaxy positions from HOD parameters.
-From the galaxy positions, we can calculate observable summary statistics such as the 2PCF, Group Multiplicity Function (GMF) and the number density $\bar{n}$.}
-
-%Talk about why it is better to simulate the mock and compute %the summary statistics of the mock instead of using analytic formulas...
-
-%Talk about all the assumptions that go into analytic %formulas...
-
-%Talk about why we do not want to write down likelihood %function for gmf, 2pcf, ...
 \subsubsection{summary statistics of galaxy populations}
@@
 \section{Experiments}\label{sec:data}
+
 \subsection{Mock observations}
-\subsection{}
-In order to test the va
+
+\newcommand{\gmf}{\mathcal{\zeta}_{\rm g}}
+
+As a testbed of our method, we create a mock population of galaxies in the following way.
+We set the upper limit of the luminosity of the mock galaxies to -20.
+Then mock observations---positions of galaxies--- are generated by populating the pre-built halo catalogs in the bolshoi simulation according to \cite{Zheng:2007aa} HOD model, and the best-fit HOD parameters found by \citet{Zheng:2007aa} for SDSS galaxy samples with the same luminosity threshold.
+
+Afterwards, summary statistics of the mock observations are computed.
+These summary statistics include the man number density $\ngalaxy$ the galaxy-galaxy correlation function $\xigg$, the group multiplicity function $\gmf$.
+These quantities are measured in the following way.
+
+Co-moving number density of the mock galaxy catalog $\ngalaxy$ is calculated.
+Th radial bins at which $\xigg$ is computed are selected to match the radial bins used in [CITE ZEHAVI ETAL] clustering analysis of the SDSS DR7 galaxy sample.
+The measured $\xigg$ of the mock galaxy catalog is shown in Figure ~\ref{1}.
+The error-bars are measured from the diagonal elements of the error covariance matrix built from 500 mocks.
+That is, with the same set of true HOD parameters, 500 mock galaxy catalogs are generated and after computing the $\xigg$ for every mock catalog, the covariance matrix shown in Figure \ref{2pcf_cov} is calculated in the following way.
+
+\begin{eqnarray} 
+\mathrm{C}_{\xi}[i,j] &=& \frac{1}{N_{\mathrm{mocks}}-1}\sum_{k=1}^{N_{\mathrm{mocks}}} \Big[\xigg^{(k)}(r_{i})-\bar{\xi}_{\rm gg}(r_{i})\Big]\Big[\xigg^{(k)}(r_{j})-\bar{\xi}_{\rm gg}(r_{j}) \Big], \\
+\bar{\xi}_{\rm gg}(r_{i}) &=& \frac{1}{N_{\mathrm{mocks}}}\sum_{k=1}^{N_{\mathrm{mocks}}} \xigg^{(k)}(r_{i}).
+\end{eqnarray}
+
+FoF group IDs of galaxies are derived by setting the projected linking length $b_{\perp}$ to 0.2 and the line-of-sight linking length $b_{\parallel}$ to 0.1.
+Then the galaxies with identical group IDs are grouped together.
+The group IDs are then binned into galaxy group richness bins.
+The group richness bins are selected in away that they match the ones used in [CITE Berlind etal] analysis of SDSS galaxy group identification.
+The resulting product is the histogram of the galaxy group richnesses.
+Group multiplicity function $\gmf$ of the mock observations is computed by divinding the number of galaxies in each richness bin by the volume of the simulation box 250 $\mathrm{Mpc}^3\mathrm{h}^{-3}$.
+
+The measured $\gmf$ of the mock galaxy catalog is shown in figure ~\ref{2}.
+The covariance between the group multiplicity functions between two richness bins has two components.
+The first one is the sample covariance between the richness bins.
+This is computed from the group multiplicity function evaluations of a set of 500 galaxy mock catalogs.
+This covariance is computed in the following way:
+
+\begin{eqnarray} 
+\mathrm{C}_{\zeta}^{\mathrm{sample}}[i,j] &=& \frac{1}{N_{\mathrm{mocks}}-1}\sum_{k=1}^{N_{\mathrm{mocks}}} \Big[\gmf^{(k)}(N_{i})-\bar{\zeta}_{\rm g}(N_{i})\Big]\Big[\gmf^{(k)}(N_{j})-\bar{\zeta}_{\rm g}(N_{j}) \Big], \\
+\bar{\zeta}_{\rm g}(N_{i}) &=& \frac{1}{N_{\mathrm{mocks}}}\sum_{k=1}^{N_{\mathrm{mocks}}} \gmf^{(k)}(N_{i}).
+\end{eqnarray}
+
+The second contribution to the error covariance matrix of $\gmf$ measurements stems from Poisson noise which can be approximated by the number of galaxies within each richness bin devided by the volume of the box.
+That is,
+
+\begin{eqnarray} 
+\mathrm{C}_{\zeta}[i,j] &=& \mathrm{C}_{\zeta}^{\mathrm{sample}}[i,j] + \mathrm{C}_{\zeta}^{\mathrm{poisson}}[i,j], \\
+\mathrm{C}_{\zeta}^{\mathrm{poisson}}[i,j] &=& \delta_{i,j}\frac{\langle N_{i}\rangle^{2}}{V_{\mathrm{box}}},
+\end{eqnarray}
+where $V_{\mathrm{box}} = 250 \mathrm{Mpc}^3\mathrm{h}^{-3}$, and $\langle N_{i}\rangle$ is the estimated number of galaxies in the richness bin $i$.
+
+\subsection{Tests}
+
 %Plots showing 2pcf, gmf, ... of the mock data.

### 2017-02-24 ccppabc/text/archive/abc.tex
updated latex of paper

--- 2016-01-30 halo/text/abc.tex
@@
 
-\section{Method}\label{sec:method}
-
+\section{Methods}\label{sec:method}
 %%%%%%%%%%%%%%%%%%%%% SECTION %%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Approximate Bayesian Computation}
+\subsection{Approximate Bayesian Computation} \label{sec:abc}
 
-ABC is based on rejection sampling, which is a Monte Carlo method 
+ABC is based on rejection sampling, so we begin this section with a brief overview of 
+rejection sampling. Broadly speaking, rejection sampling is a Monte Carlo method 
 used to draw samples from a probability distribution that is difficult 
-to directly sample. For inference, the probability distribution that 
-we want to sample is the posterior probability distribution function.  
-Before we present the theory and algorithm of ABC, we provide a brief
-overview of rejection sampling.
+to directly sample. More explicitly, it samples a probability distribution 
+$f(\alpha)$, which is difficult to directly sample, by drawing samples 
+from the distribution $M g(\alpha)$ that satisfies the condition 
+$f(\alpha) < M g(\alpha)$ where $M > 1$ and $g(\alpha)$ is an 
+instrumental distribution where sampling is easier.
 
-Briefly, rejection sampling is a method that draws samples from a 
-probability distribution $f(\alpha)$, which is difficult to 
-directly sample, by sampling from the distribution $M g(\alpha)$ that 
-satisfies $f(\alpha) < M g(\alpha)$. $g(\alpha)$ is an instrumental 
-distribution where sampling is easier and $M > 1$.
-
-More specifically, in the context of simulation-based inference, 
-the main goal is to sample from the joint probability of a
-simulation $X$ and parameters $\pars$ given observed data $D$, which 
-from Bayes rule can be written as 
+In the context of simulation-based inference, 
+the ultimate goal is to sample from the joint probability of a
+simulation $X$ and parameters $\pars$ given observed data $D$, the
+posterior probability distribution. From Bayes rule this posterior 
+distribution can be written as 
 \beq
@@
 \eeq
-where $\pi(\pars)$ is the prior probabilities over the parameters of 
+where $\pi(\pars)$ is the prior distribution over the parameters of 
 interest and $\mathcal{Z}$ is the evidence, 
@@
 \eeq
-and 
+and  The weighted difference of the number density of the observed galaxies and the analytic number density is also added to this.
+In this work, we investigate the how accurately we can infer the model parameters by relaxing the underlying assumptions of the parameter inference methods.
 \beq
@@
 \eeq
-so we can sample $p(\pars, X | D)$ by drawing ${\pars, X}$ from 
-the $q(\pars, X)$ distribution. In practice, this is done 
-by first drawing $\pars$ from the prior $\pi(\pars)$ then generating a 
-simulation $X = f(\pars)$ via the forward model. Afterwards, ${\pars, X}$ 
+so we can sample $p(\pars, X | D)$ by drawing ${\pars, X}$ from $q(\pars, X)$.
+In practice, this is done 
+by first drawing $\pars$ from the prior $\pi(\pars)$ and then generating a 
+simulation $X = f(\pars)$ via the forward model. ${\pars, X}$ 
 is accepted if
-\beq
-u < \frac{p(\pars, X | D)}{M q(\pars, X)} = \frac{p(D|X)}{\mathrm{max}\;p(D|X)}
+\beq \label{eq:reject_samp}
+\frac{p(\pars, X | D)}{M q(\pars, X)} = \frac{p(D|X)}{\mathrm{max}\;p(D|X)} > u 
 \eeq
-where $u$ is drawn from $\mathtt{Uniform}[0,1]$.
+where $u$ is drawn from $\mathtt{Uniform}[0,1]$. The set of $\pars$ and $X$ 
+accepted through the repeated rejection sampling process, ultimately sample 
+$p(\pars, X |D)$.
 
-At this point, ABC postulates that $p(D|X)$
-
-The method requires a prior probability over the parameters of interest 
-($\pi(\theta)$), a generative forward model of the data given a set of 
-parameters ($f(\theta)$) and a distance metric $\rho(x,D)$ which measures 
-the distance between the simulation $X = f(\theta)$ and the observed data $D$.
-
-Technically in large scale structure studies, the observed data is the 
-total angular and redshift positions of galaxies on the sky. However, 
-in practice, this data is an intractably high dimensional object. As a result, we 
-utilize statistical summaries of galaxy positions as our observed 'data' 
-and the distance metric we use measures the closeness of the statistical 
-summaries of observed galaxy positions versus the simulated galaxy 
-positions. Common summary statistics used to encapsulate the galaxy position 
-information include the number density of the galaxy sample, the two-point 
-correlation function (2PCF), three-point correlation function, and galaxy 
-group statistics.
-
-Therefore the distance metric measures the closeness of statistical summaries of the data and those of the simulation.
-When the observed data-set under study is the positions of all galaxies observed in the survey, possible statistical summaries of the data are the number density of galaxies, two-point correlation function, three-point correlation function, galaxy group statistics (counts in cells), and so on.
-
-Here, we provide a brief overview of rejection sampling algorithm and how it is used in a simulation-based inference.
-We want to draw samples from a probability distribution $p(\alpha)$.
-Now, let us assume that: (i) there exists an auxiliary density $q(\alpha)$ from which we can directly sample from, and (ii) there exist $M>1$ which is an appropriate bound on $p(\alpha)/q(\alpha)$.
-Given these assumptions, rejection sampling works in the following way.
-We draw a sample $\alpha$ from the auxiliary density $q(\alpha)$.
-Afterwards, we draw a random number $u$ from a uniform distribution between zero and one.
-Then we accept the draw $\alpha$ if the $u$ is less than the ratio $p(\alpha)/Mq(\alpha)$.
-This procedure is repeated until sufficient number of draws are accepted.
-
-In the context of a simulation-based inference, we main goal is to 
-sample from the joint probability of a simulation $X$ and parameters 
-$\pars$ given the observed data $D$: 
-\beq
-p(\pars,X | D)
-\eeq
-According to Bayes rule $p(\pars, X | D)$ can be written as 
-\beq
-p(\pars,X | D)  = \frac{p(D|X)p(X|\pars)\pi(\pars)}{\mathcal{Z}}
-\eeq
-where $\mathcal{Z}$  is the evidence, 
-\beq
-\mathcal{Z} = \int \dev \pars \; \dev X \; p(D|X)p(X|\pars)\pi(\pars),
-\eeq
-
-There exists a constant $M>1$ 
-\beq 
-M = \frac{\mathrm{max}_{X} \; p(D|X)}{Z},
-\eeq 
-and auxiliary density $q(\pars ,X)$
-\beq 
-q(\pars ,X)=p(X|\pars )\pi(\pars ),
-\eeq 
-such that:
-\beq 
-p(\pars ,X|D)<Mq(\pars ,X).
-\eeq 
-Now in a simulation-based inference, we first draw a pair $\{\pars , X \}$ from the density $q(\pars , X)$.
-This requires drawing a sample $\pars$ from the prior $\pi(\pars)$, and generating a simulation $X=f(\pars)$ via the forward model.
-Then, we accept the pair $\{\pars , X \}$, if the following condition is met
-
-\beq 
-u \sim \mathrm{Uniform}[0,1] < \frac{p(\pars,X|D)}{Mq(\pars,X)}=\frac{p(D|X)}{\mathrm{max}_{X}p(D|X)}.
-\eeq
-
-ABC postulates that the probability of observing data $D$ given simulation $X$, $p(D|X)$ is controlled by a threshold $\epsilon$, and that this probability is one if the distance between the data and the simulation, $\rho(D,X)$ is less than $\epsilon$.
-That is
-
-\beq 
-p(D|X) = p_{\epsilon}(D|X) \equiv \begin{cases}
-                                    1 \ \ \  \mathrm{if} \ \rho(D,X)<\epsilon \; , \\
-                                    0 \ \ \ \mathrm{otherwise} \; .
+At this stage, ABC distinguishes itself by postulating that $p(D|X)$, 
+the probability of observing data $D$ given simulation $X$, is controlled 
+by a threshold $\epsilon$ and that 
+\beq \label{eq:abc_condition}
+p(D|X)= p_\epsilon(D|X)\equiv 
+\begin{cases}
+    1,		& \text{if}\;\; \rho(D, X) < \epsilon\\
+    0,      & \text{otherwise}
 \end{cases}
 \eeq
+where $\rho(D, X)$ is the distance between the data $D$ and simulation $X$. 
+Plugging Eq. \ref{eq:abc_condition} into the 
+rejection sampling acceptance criteria (Eq. \ref{eq:reject_samp}), we 
+derive the acceptance criteria for ABC, which states that $\pars$ is accepted 
+if $\rho(D, X) < \epsilon$.
 
-Thus the set of parameters $\pars$ is accepted if $\rho(D,f(\pars))<\epsilon$. In principle, the distance function measures the distance 
-between the statistical summaries of the data and those of the simulation. 
-The distance, in general, can be a vector, with each component representing the distance between one summary statistics of the data (e.g.
-variance) and the same summary statistics of the simulation.
-That is
-\beq 
-\rho(D,X) = \Big[\rho_{1}(S_{1}(D) , S_{1}(X)), ... , \rho_{M}(S_{M}(D) , S_{M}(X)) \Big],
-\eeq 
-where $M$ is the number of statistical summaries used in the inference, $S_{i}$ is the i-th summary statistics, and $\rho_{i}$ is a scalar function that measures the distance between the i-th summary statistics of the data and that of the simulation.
-In this case, the threshold $\epsilon$ is an $M$ dimensional vector.
+\todo{CHH: Paragraph that briefly describes qualitatively what distance is}
+The distance can be a vector with multiple components where each component is 
+a different summary statistic distance between the data and simulation. 
+In that case, the threshold $\epsilon$ in 
+Eq. \ref{eq:abc_condition} will also be a vector with the same dimensions.  
+$\pars$ is accepted if the distance vector is less than the threshold for 
+every component.
 
-We draw a large number of proposals from the prior distribution over the parameter space, 
-and reject the samples for which the distance between the data and the simulation corresponding to that 
-sample is more than a certain threshold $\bm{\epsilon}$. This requirement is satisfied \emph{if and only if}
-every component of the distance vector is less than its corresponding threshold. Note that, each summary statistics 
-has a separate threshold. This procedure is done iteratively until, we are left with a certain number
-of samples that satisfy the distance criteria. The final ensemble approximates the posterior probability distribution over the parameters of the generative model of the data.
+The ABC procedure begins, in the same fashion as rejection sampling, by drawing 
+$\pars$ from the prior distribution $\pi(\pars)$. The simulation is generated from 
+$\pars$ using the forward model, $X = f(\pars)$. Then the distance between 
+the data and simulation, $\rho(D, X)$, is calculated and compared to 
+$\epsilon$. If every component of $\rho(D, X) < \epsilon$, $\pars$ is accepted. 
+This process is repeated until we are left with a sample of $\pars$ that all 
+satisfy the distance criteria. This final ensemble approximates the posterior 
+probability distribution $p(\pars, X|D)$.
 
-For a sufficiently small threshold, this algorithm samples from the posterior probability, $p(\pars | D)$, and for an arbitrarily large threshold, the algorithm samples from the prior $\pi(\pars)$.
-This rejection sampling technique requires repeated sampling from the prior, and a very small threshold to succeed.
+As it is stated, the ABC method poses some practical challenges. If the 
+threshold $\epsilon$ is arbitrarily large, the algorithm essentially 
+samples from the prior $\pi(\pars)$. Therefore a sufficiently small threshold
+is necessary to sample from the posterior probability distribution. However,
+an appropriate value for the threshold is not known \emph{a priori}. Yet, 
+even if an appropriate threshold is selected, a small threshold requires 
+the entire process to be repeated for many draws of $\pars$ from $\pi(\pars)$ 
+until a sufficient sample is acquired. This often presents computation challenges.
 
-In this investigation instead, we use Population Monte Carlo Algorithm (PMC) as our sampling technique.
-PMC is an iterative method that performs rejected sampling from a sequence of distributions $\{p_{1}(\pars),...,p_{T}(\pars)\}$.These distributions are paired with a sequence of decreasing distance thresholds $\{\epsilon_{1} , ...
-, \epsilon_{T}\}$.
-
-At each iteration, $N$ proposals are drawn from a distribution over parameters, and a weight is assigned to each one.
-In the first step, the threshold is set to an arbitrarily large value.
-Therefore the algorithm makes $N$ draws from the prior and assigns equal weights to all the proposals.
-
-In the consequent iterations, the distance threshold is set to a smaller value.
-A proposal is weighted-sampled from the previous sample.
-It is perturbed by a kernel which is set to the covariance of the parameters in the previous sample.
-A simulation is then generated by feeding the perturbed parameter into the forward model.
-The perturbed parameter is then accepted if the distance between the data and the generated simulation is less than the value of the threshold in that iteration.
-This procedure is repeated until $N$ particles are sampled.
-Then the weights are updated according to the kernel, the prior, and the previous set of weights.
+In this paper, we overcome some of the challenges posed by the above ABC method
+by using a Population Monte Carlo (PMC) algorithm as our sampling technique. 
+PMC is an iterative method that performs rejection sampling over a 
+sequence of $\pars$ distributions ($\{p_1(\pars), ..., p_T(\pars)\}$ for 
+$T$ iterations), with a distance threshold that decreases at every iteration of 
+the sequence.
 
@@
 \caption{The procedure for ABC-PMC}
-\begin{algorithmic}[1]
+\begin{algorithmic}[1] \label{alg:abcpmc}
 %\STATE \DATA: D
@@
 
-%\begin{algorithm}[H]
-% \Data{this text}
-% \Result{how to write algorithm with \LaTeX2e }
-% initialization\;
-% \While{not at end of this document}{
-%  read current\;
-%  \eIf{understand}{
-%   go to next section\;
-%   current section becomes this one\;
-%   }{
-%   go back to the beginning of current section\;
-%  }
-% }
-% \caption{How to write algorithms}
-%\end{algorithm}
+As illustrated in Algorithm \ref{alg:abcpmc}, for the first iteration $t = 1$, 
+we begin with an arbitrarily large distance threshold $\epsilon_1$. We 
+draw $\pars^i$ (hereafter referred to as particles) from the prior distribution 
+$\pi(\pars)$. We forward model the simulation $X = f(\pars^i)$, calculate the 
+distance $\rho(D, X)$, compare this distance to $\epsilon_1$, and then 
+accept or reject the $\pars^i$ draw. Because we set $\epsilon_1$ arbitrarily large, 
+the particles essentially sample the prior distribution. This process 
+is repeated until we accept $N$ particles. We then assign equal weights to 
+the $N$ particles, $w_1^i = 1/N$.
 
-\subsection{Forward model of the large scale galaxy populations}\label{sec:halo}
-\subsubsection{halo occupation modeling}
+For subsequent iterations ($t > 1$), first, the distance threshold is set such that
+$\epsilon_{i,t} < \epsilon_{i,t-1}$ for all components $i$. Although there is 
+no general prescriptions, the distance threshold $\epsilon_{i,t}$ can be 
+assigned based on the empirical distribution of the accepted distances in the 
+previous iteration ($t-1$). In \todo{CITE CITE}, for example, the $75th$ percentile 
+of the $t-1$ iteration acceptance distance is set as $\epsilon_t$. We specify
+0our prescription for the distance threshold in Section \todo{REFREF}.
+
+Once $\epsilon_t$ is set, we draw a particle from the previous 
+weighted set of particles ${\pars^i}_{t-1}$. 
+This particle is perturbed by a kernel set to the covariance of 
+${\pars^i}_{t-1}$. Then once again, we generate a simulation from 
+forward modeling $X = f(\pars^i)$, calculate the distance $\rho(X, D)$, 
+compare the distance to new distance threshold ($\epsilon_t$) in order 
+to accept or reject the particle. This process is repeated until we 
+assemble a new $N$ set of particles ${\pars^i}_t$. We then update the 
+particle weights according to the kernel, the prior distribution and the 
+previous set of weights.
+
+\todo{Preliminary remarks about convergence}
+Three conditions can be inspected in order to confirm the convergence of the ABC-PMC algorithm.
+Firstly, we can set a shut-off parameter for the acceptance ratio in each iteration.
+That is, if the acceptance ratio falls bellow the shut-off parameter, the algorithm has converged.
+The second condition is satisfied when the fractional change in the distance threshold becomes smaller than a certain tolerance level.
+The third and the final convergence test is checking whether the derived uncertainties of the infered paramaters have stabilized.
+%In the context of large scale structure, technically, the observed data is the 
+%total angular and redshift positions of galaxies on the sky. However, 
+%in practice, this data is an intractably high dimensional object. As a result, 
+%we use statistical summaries of galaxy positions as our observed data.
+%So the distance metric $\rho(D, X)$ we use measures the closeness of the 
+%statistical summaries of observed galaxy positions versus the simulated galaxy 
+%positions. Common summary statistics used to encapsulate the galaxy position 
+%information include the number density of the galaxy sample, the two-point 
+%correlation function (2PCF), three-point correlation function, and galaxy 
+%group statistics.
+
+%Here, we provide a brief overview of rejection sampling algorithm and how it is used in a simulation-based inference.
+We want to draw samples from a probability distribution $p(\alpha)$.
+Now, let us assume that: (i) there exists an auxiliary density $q(\alpha)$ from which we can directly sample from, and (ii) there exist $M>1$ which is an appropriate bound on $p(\alpha)/q(\alpha)$.
+Given these assumptions, rejection sampling works in the following way.
+We draw a sample $\alpha$ from the auxiliary density $q(\alpha)$.
+Afterwards, we draw a random number $u$ from a uniform distribution between zero and one.
+Then we accept the draw $\alpha$ if the $u$ is less than the ratio $p(\alpha)/Mq(\alpha)$.
+This procedure is repeated until sufficient number of draws are accepted.
+
+%For a sufficiently small threshold, this algorithm samples from the posterior probability, $p(\pars | D)$, and for an arbitrarily large threshold, the algorithm samples from the prior $\pi(\pars)$.
+This rejection sampling technique requires repeated sampling from the prior, and a very small threshold to succeed.
+
+%In this investigation instead, we use Population Monte Carlo Algorithm (PMC) as our sampling technique.
+PMC is an iterative method that performs rejected sampling from a sequence of distributions $\{p_{1}(\pars),...,p_{T}(\pars)\}$.These distributions are paired with a sequence of decreasing distance thresholds $\{\epsilon_{1} , ...
+, \epsilon_{T}\}$.
+
+%At each iteration, $N$ proposals are drawn from a distribution over parameters, and a weight is assigned to each one.
+In the first step, the threshold is set to an arbitrarily large value.
+Therefore the algorithm makes $N$ draws from the prior and assigns equal weights to all the proposals.
+
+%In the consequent iterations, the distance threshold is set to a smaller value.
+A proposal is weighted-sampled from the previous sample.
+It is perturbed by a kernel which is set to the covariance of the parameters in the previous sample.
+A simulation is then generated by feeding the perturbed parameter into the forward model.
+The perturbed parameter is then accepted if the distance between the data and the generated simulation is less than the value of the threshold in that iteration.
+This procedure is repeated until $N$ particles are sampled.
+Then the weights are updated according to the kernel, the prior, and the previous set of weights.
+
+\subsection{Forward model}\label{sec:halo}
+\subsubsection{Halo Occupation Modeling}
 
@@
 \newcommand{\ngalaxy}{\bar{n}_{\mathrm{g}}}
+\newcommand{\gmf}{\mathcal{\zeta}_{\rm g}}
 
@@
 
-\subsubsection{summary statistics of galaxy populations}
+\subsection{summary statistics of galaxy populations}
 
-\subsubsection{distance metrics}
+Computation of the distance between the data and the simulation requires specification and computation of a set of statistical summaries of the data and those of the simulations.
+We conssider the following summaries of the data in our investigation of applying the ABC algorithm to constraining the galaxy-halo connection:
+\begin{itemize}
+\item Galaxy number density $\ngalaxy$: This quantity is the comoving number density of galaxies and is computed by dividing the number of galaxies by the comoving volume.
+This quantity is measured in units of $(\mathrm{Mpc}/\mathrm{h})^{-3}$.
+\item Galaxy-galaxy correlation function $\xigg(r)$: This quantity is the probability in excess of random of finding a pair of galaxies separated by the threee dimensional distance $r$.
+For computation of this statistics, we rely on CITE[landy and Szalay] estimator.
+Following the original notayion of the paper, we refer to data as $D$, and random as $R$.
+The correlation function can then be estimated in the following way:
+\beq
+\xigg(r) = \frac{DD - 2DR + RR}{RR}
+\eeq
+as a function of three-dimensional distance $r$. This quantity is measured in units of $(\mathrm{Mpc}/\mathrm{h})^{-3}$.
+\item Galaxy group multiplicity function $\gmf(N)$: This quantity is defined as the number denisty of galaxy groups in bins of group richness $N$.
+These richness bins can have a width of unity or more.
+In order to link galaxies to each other, we rely on friends-of-friends (Hereafter FoF) group-finder algorithm.
+If both transeverse and line-of-sight separations of a pair of galaxies are smaller than a specified pair of transverse and line-of-sight linking lengths, the two galaxies are assigned to the same group by FoF algorithm.
+Quantitatively, for a pair of galaxies with transverse (line-of-sight) separations of $d_{\perp , ij}$ ($d_{\parallel , ij}$), the linkning condition is satisfied if
+\begin{eqnarray}
+d_{\perp , ij} &<& \ngalaxy ^{-1/3} \; b_{\perp , ij}, \\
+d_{\parallel , ij} &<& \ngalaxy ^{-1/3} \;b_{\parallel , ij},
+\end{eqnarray}
+where $b_{\perp , ij}$ ($b_{\parallel , ij}$) are the maximum projected (line-of-sight) linking  lengths that are normalized to the mean separation between galaxies $\ngalaxy ^{-1/3}$.
+Once the FoF galaxy groups are identified, a histogram of the number of groups in bins of group richness is made.
+Once the bin heights are devided by the comoving volume, the group multiplicity function (as a function of richness bins) is found.
+This quantity is measured in units of $(\mathrm{Mpc}/\mathrm{h})^{-3}$.
 
-\section{Experiments}\label{sec:data}
+\end{itemize}
 
-\subsection{Mock observations}
-
-\newcommand{\gmf}{\mathcal{\zeta}_{\rm g}}
-
+\section{Mock observations}\label{sec:data}
 As a testbed of our method, we create a mock population of galaxies in the following way.
@@
 
-\subsection{Tests}
+%%%%%%%%%%%%%%%%%TEST%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Tests}\label{sec:test}
+
+In order to test the performance of the ABC algorithm in finding constraints over HOD parameters as well as comparing the ABC oncstraints with those found by standard MCMC likelihood analysis, we perform three tests.
+A combination of the statistical summaries of the mock observations is used in each test.
+The first, second, and the third experiment the combinations $\{\ngalaxy , \xigg\}$, $\{\ngalaxy , \gmf\}$, and $\{\ngalaxy, \xigg, \gmf\}$ are used respectively.
+
+The distance metrics used for the ABC analysis are multi-component.
+That is, a distance component is assigned to for every statsitical summary of the data used in a particular test.
+Hereafter, the distance components associated with $\ngalaxy$, $\xigg$, $\gmf$ are denoted by $\rho_{\ngalaxy}$, $\rho_{\xigg}$, and $\rho_{\gmf}$:
+
+\begin{eqnarray}
+\rho_{\ngalaxy} &=& \frac{(\ngalaxy^{d} - \ngalaxy^{m})^{2}}{\sigma^{2}_{\ngalaxy}},\\
+\rho_{\xigg}    &=& \sum_{k} \frac{[\xigg^{d}(r_k) - \xigg^{m}(r_k)]^{2}}{C_{\xi}[k,k]},\\
+\rho_{\gmf}     &=& \sum_{k} \frac{[\gmf^{d}(N_k) - \gmf^{m}(N_k)]^{2}}{C_{\zeta}[k,k]},
+\end{eqnarray}
+where the superscripts $d$ and $m$ denote data and model respectively.
+
+Another major component of any ABC analysis is the choice of prior.
+Uniform prior ranges are chosen for all paramters, except the scatter paramter which has a log-uniform prior between 0.1 and 0.7.
+All the prior ranges are specified in Table ~\ref{tab:prior}.
+
+So far, we have not specified the manner in which we change the threshold values in our ABC-PMC algorithm. 
+In the first iteration, we set the threshold value to $\infty$.
+Before thr seond and the third iteration, we set the the threshold to 50-percentile of the distances of the particles in the previous iterations.
+In the subsequent iterations, we set lower the threshold to the 75-percentile of the disatnces of the particles in the previous iterations.
+
+In order to test the convergence of the algorithm, we inspect all the convergence test methods outlined in Section \ref{sec:abc}.
+That is , we set the shut-off value for the acceptence ratio to 0.001.
+Moreover, we set the fractional change in the threshold values of the consequent iterations to 0.1.
+
+%%%%%%%%%%%%%%%%%%TABLES%%%%%%%%%%%%
+\begin{table*}
+  \label{tab:prior}
+  \caption{{\bf Prior Specifications}. Priors that contain the form [a, b] mean the
+parameter in question is restricted to values within that range. All mass parameters are in unit of $h^{-1}M_\sun$}
+\begin{tabular}{@{}lll}
+\\ \hline \hline
+    HOD Parameter & & Prior \\ \hline
+   $\log M_{0}$   & & Uniform on [10.0, 13.0] \\
+  $\log \sigma_{\rm M}$ & & Log--Uniform on [0.1, 0.7] \\
+  $\log M_{min}$ & &   Uniform on [11.02, 13.02] \\
+  $\alpha$ & & Uniform on [0.8, 1.3] \\
+  $\log M_{1}$ & & Uniform on [13.0, 14.0] \\
+ \hline
+  \end{tabular}
+\end{table*}
+
+%%%%%%%%%%%%%%%%%FIGURES%%%%%%%%%%%%%%
 
 %Plots showing 2pcf, gmf, ... of the mock data.
-%How we compute the covariance matrix from mocks...
-%Talk about error bars. Talk about poisson errors in group %populations .
+
 %--------------------------------------------------------
@@
 $\sigma_{\log M}$, $\log M_{min}$, $\alpha$, and $\log M_1$ using the
-standard Markov Chain Monte Carlo method as described in Section \todo{CITECITE}. 
+standard MCMC method as described in Section \todo{CITECITE}. 
 \todo{THIS IS A SPACE MARKER NOT THE FINAL FIGURE.}}
@@
 \section{Discussion}\label{sec:discussion}
+
 \todo{talk about cosmological inference and why this is good first step for likelihood free inference in large scale ...}
+
+\todo{I'll talk in the discussions that all of these are necessary conditions not suficient.}
+\todo{standard mcmc has the same problem.
+It has two widely used necessary conditions: autocorrelation of the chains, and the derived uncertainty of the infered parameters.
+We'll expand this discussion in the result and discussion sections.
+sounds good?}
 

### 2017-02-24 ccppabc/text/main.tex
updated latex of paper

\documentclass[12pt, preprint]{aastex}
\usepackage[breaklinks,colorlinks, 
urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{mathtools}
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\usepackage{algorithmic,algorithm}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{colorlinks, linkcolor=darkblue, filecolor=darkgreen, urlcolor=darkred, citecolor=darkblue}
\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand{\todo}[1]{{\em \textcolor{red}{ #1}}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\lang}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\vep}{\bm{\epsilon}}
\newcommand{\ep}{\epsilon}
\newcommand{\pars}{\vec{\theta}}
\newcommand{\dev}{\mathrm{d}}

\begin{document}
\title{Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection}
\author{ChangHoon~Hahn\altaffilmark{\dagger1}, Mohammadjavad~Vakili\altaffilmark{\dagger1},
        Kilian~Walsh\altaffilmark{1},
        Andrew~P.~Hearin\altaffilmark{2},
        David~W.~Hogg\altaffilmark{1,3,4,5},
    and~Duncan~Campbell\altaffilmark{6}}
\altaffiltext{$\dagger$}{These authors have contributed equally to the paper: \url{mjvakili@nyu.edu}, \url{chh327@nyu.edu}}
\altaffiltext{1}{Center for Cosmology and Particle Physics, 
                 Department of Physics, New York University, 4 Washington Pl., New York, NY}    
\altaffiltext{2}{Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT}
\altaffiltext{3}{Simons Center for Data Analysis, 160 Fifth Avenue, New York, NY}
\altaffiltext{4}{Center for Data Science, New York University, 726 Broadway, New York, NY}
\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}

\altaffiltext{6}{Department of Astronomy, Yale University, New Haven, CT}

\begin{abstract}
The standard approaches to Bayesian parameter inference in large scale structure (LSS) assume a Gaussian functional form (chi-squared form) for the likelihood. They are also typically restricted to measurements 
such as the two point correlation function. 
Likelihood free inferences such as Approximate Bayesian Computation (ABC) make inference possible without 
assuming any functional form for the likelihood, thereby relaxing the assumptions and restrictions of 
the standard approach. Instead it relies on a forward generative model of the data and a metric
for measuring the distance between the model and data. In this work, we 
demonstrate that ABC is feasible for LSS parameter inference by using it to constrain parameters of the 
halo occupation distribution (HOD) model for populating dark matter halos with galaxies.

Using specific implementation of ABC supplemented with Population Monte Carlo
importance sampling, a generative forward model using HOD, and a distance metric 
based on galaxy number density, two-point correlation function, and galaxy group
multiplicity function, we constrain the HOD parameters of mock observation 
generated from selected ``true'' HOD parameters. The parameter constraints we 
obtain from ABC are consistent with the ``true'' HOD parameters, demonstrating that ABC can be reliably  used for parameter inference in LSS. Furthermore, we compare our ABC constraints to constraints we obtain using a pseudo-likelihood function of Gaussian form with MCMC and find consistent HOD parameter constraints. Ultimately our results suggest that ABC can and should be applied in parameter  inference for LSS analyses. 
\end{abstract}

\section{Introduction}
Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
pro\-ba\-bil\-ities for the data given the theoretical model---%
for combining data from different surveys and performing principled inferences of
the cosmological parameters (\citealt{White:1996aa, Riess:1998aa}). 
Nowhere has this been more true than in cosmic microwave background (CMB) studies,
where it is nearly possible to analytically evaluate a likelihood function that
involves no (or minimal) approximations (\citealt{Oh:1999aa}, \citealt{Wandelt:2004aa},  
\citealt{Eriksen:2004aa}, \citealt{planckI, planckII}).

Fundamentally, the tractability of likelihood functions in cosmology flows from
the fact that the initial conditions are exceedingly close to Gaussian in form 
(\citealt{planck_NG, planck_inflation}),
and that many sources of measurement noise are also Gaussian (\citealt{Knox:1995aa, Leach:2008aa}).
Likelihood functions are easier to write down and evaluate when things are closer 
to Gaussian, so at large scales and in the early universe. Hence likelihood analyses 
are ideally suitable for CMB data.

In large-scale structure (LSS) with galaxies, quasars, and quasar absorption systems as tracers,
formed through nonlinear gravitational evolution and biasing, the likelihood {\em cannot} be Gaussian. 
Even if the initial conditions are perfectly Gaussian, the growth of structure creates non-linearities 
which are non-Gaussian (see \citealt{Bernardeau:2002aa} for a comprehensive review). 
Galaxies form within the density field in some complex manner that is modeled only effectively
(\citealt{Dressler:1980aa, Kaiser:1984aa, Santiago:1992aa, Steidel:1998aa}; see \citealt{somerville15} for a recent review).  
Even if the galaxies were a Poisson sampling of the density field, which they are not (\citealt{Mo:1996aa, Sommerville:2001aa, Casas-Miranda:2002aa}), it would be tremendously difficult to write down even 
an approximate likelihood function (\citealt{devpois}).

The standard approach makes the strong assumption that the likelihood function 
for the data can be approximated by a pseudo-likelihood function that is a Gaussian
probability density in the space of the two-point correlation function estimate. 
It is also typically limited to (density and) two-point correlation 
function (2PCF) measurements, assuming that these measurements constitute 
sufficient statistics for the cosmological parameters. 
As Hogg (in preparation) demonstrates, the assumption of a Gaussian 
pseudo-likelihood function cannot be correct (in detail) at any scale, 
since a correlation function, being related to the 
variance of a continuous field, must satisfy non-trivial positive-definiteness 
requirements. These requirements truncate function space such that the 
likelihood in that function space could never be Gaussian. The failure of this 
assumption becomes more relevant as the correlation function becomes better measured, 
so it is particularly critical on intermediate scales, where neither shot 
noise nor cosmic variance significantly influence the measurement.

Fortunately, these assumptions are not required for cosmological inferences, 
because high-precision cosmological simulations can be used to directly calculate 
LSS observables. Therefore, we can simulate not just the one- or two-point statistics of the galaxies, but also any higher order statistics that might provide additional constraining power on a model. In principle, there is therefore no strict need to rely on these common but specious analysis  assumptions as it is possible to calculate a likelihood function directly from simulation outputs.

Of course, any naive approach to sufficiently simulating the data would be ruinously
expensive. Fortunately, there are principled, (relatively) efficient methods for 
minimizing computation and delivering correct posterior inferences, using only a 
data simulator and some choices about statistics. 
In the present work, we use Approximate Bayesian Computation---ABC---which provides a \emph{rejection sampling} 
framework (\citealt{abcrejectionsampling}) that relaxes the assumptions of the traditional approach.

ABC approximates the posterior probability distribution function (model given the data)
by drawing proposals from the prior over the model parameters, simulating the data from the 
proposals using a forward generative model, and then rejecting the proposals that are beyond 
a certain threshold ``distance'' from the data, based on summary statistics of the data. 
In practice, ABC is used in conjunction with a more efficient sampling operation like 
Population Monte Carlo (PMC; \citealt{smc}). 
PMC initially rejects the proposals from the prior with a relatively large ``distance'' threshold. 
In subsequent steps, the threshold is updated adaptively, and samples from the proposals that have 
passed the previous iteration are subjected to the new, more stringent, threshold criterion (\citealt{abcpmc}). 
In principle, the distance metric can be any positive definite function that compares 
various summary statistics between the data and the simulation.

In the context of astronomy, this approach has been used in a wide range of topics including 
image simulation calibration for wide field surveys (\citealt{abccosmology}),
the study of the morphological properties of galaxies at high redshifts (\citealt{abcmorphology}),
stellar initial mass function modeling (Cisewski et al. in preparation),
and cosmological inference with  
with weak-lensing peak counts (\citealt{abcwl,abcwl2}), Type Ia Supernovae (\citealt{abcsn}), 
and galaxy cluster number counts (\citealt{cosmoabc}).

In order to demonstrate that ABC can be tractably applied to parameter estimation in contemporary 
LSS analyses, we narrow our focus to inferring the parameters of a Halo Occupation Distribution (HOD) 
model. The foundation of HOD predictions is the halo model of LSS, that is, 
collapsed dark matter halos are biased tracers of the underlying cosmic density field 
(\citealt{press74, bond91, cooray_sheth2002}). The HOD specifies how the dark matter 
halos are populated with galaxies by modeling the probability that a given halo hosts 
$N$ galaxies subject to some observational selection criteria 
(\citealt{lemson99, seljak2000,scoccimarro2001,berlind_weinberg2002,zheng2005}).
This statistical prescription for connecting galaxies to halos has been remarkably
successful in reproducing the galaxy clustering, galaxy--galaxy lensing, and other 
observational statistics (\citealt{Rodriguez-Torres:2015aa, miyatake15}), and is a 
useful framework for constraining cosmological parameters (\citealt{vdb03, tinker05, cacciato13, more13}) 
as well as galaxy evolution models (\citealt{conroy09, Tinker:2011aa, leauthaud12, behroozi13, Tinker:2013aa}, 
Walsh et al. in preparation).

More specifically, we limit our scope to a likelihood analysis of HOD model parameter space, 
keeping cosmology fixed. We forward model galaxy survey data by populating pre-built dark 
matter halo catalogs obtained from high resolution N-body simulations (\citealt{bolshoi,multidark}) using 
$\mathtt{Halotools}$\footnote{http://halotools.readthedocs.org} 
(\citealt{Hearin:2016aa}), 
an open-source package for modeling the galaxy-halo connection. 
Equipped with the forward model, we use summary statistics such as 
number density, two-point correlation function, galaxy group multiplicity function (GMF)
to infer HOD parameters using ABC.

In Section \ref{sec:method} we discuss the algorithm of the ABC-PMC prescription we use in our analyses. 
This includes the sampling method itself, the HOD forward model, and the computation of summary statistics. 
Then in Section \ref{sec:mock_obv}, we discuss the mock galaxy catalog, which we treat as observation. 
With the specific choices of ABC-PMC ingredients, which we describe in Section \ref{sec:abcpmc_spec}, 
in Section \ref{sec:abc_results} we present the results of our parameter inference using 
two sets of summary statistics, number density and 2PCF and number density and GMF. 
We also include in our results, analogous parameter constraints from the standard MCMC approach, 
which we compare to ABC results in detail, Section \ref{sec:abcvsmcmc}. Finally, we discuss and 
conclude in Section \ref{sec:discussion}.

\section{Methods}\label{sec:method}
%%%%%%%%%%%%%%%%%%%%% SECTION %%%%%%%%%%%%%%%%%%%%%%%%%%
% Approximate Bayesian Computation  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Approximate Bayesian Computation} \label{sec:abc}
ABC is based on rejection sampling, so we begin this section with a brief overview of 
rejection sampling. Broadly speaking, rejection sampling is a Monte Carlo method 
used to draw samples from a probability distribution, $f(\alpha)$, which is difficult to directly sample. The strategy is to draw samples from an instrumental distribution $g(\alpha)$ that satisfies the condition $f(\alpha) < M g(\alpha)$ for all $\alpha,$ where $M > 1$ is some scalar multiplier. The purpose of the instrumental distribution $g(\alpha)$ is that it is easier to sample than $f(\alpha)$ (see \citealt{bishop} and refernces therein).

In the context of simulation-based inference, 
the ultimate goal is to sample from the joint probability of a
simulation $X$ and parameters $\pars$ given observed data $D$, the
posterior probability distribution. From Bayes rule this posterior 
distribution can be written as 
\beq
p(\pars, X | D) = \frac{p(D|X)p(X|\pars)\pi(\pars)}{\mathcal{Z}}
\eeq
where $\pi(\pars)$ is the prior distribution over the parameters of 
interest and $\mathcal{Z}$ is the evidence, 
\beq
\mathcal{Z} = \int d\pars \; dX\; p(D|X) p(X|\pars) \pi(\pars), 
\eeq
where the domain of the integral is all possible values of $X$ and $\pars$. 
Since $p(\pars, X | D)$ cannot be directly sampled, we use rejection 
sampling with instrumental distribution 
\beq
q(\pars, X) = p(X|\pars) \pi(\pars)
\eeq
and the choice of 
\beq
M = \frac{\mathrm{max}\; p(D|X)}{\mathcal{Z}} > 1.
\eeq
Note that we do not ever need to know $\mathcal{Z}$. 
The choices of $q(\pars, X)$ and $M$ satisfy the condition 
\beq
p(\pars, X | D) < M q(\pars, X)
\eeq
so we can sample $p(\pars, X | D)$ by drawing ${\pars, X}$ from $q(\pars, X)$.
In practice, this is done 
by first drawing $\pars$ from the prior $\pi(\pars)$ and then generating a 
simulation $X = f(\pars)$ via the forward model. Then ${\pars, X}$ 
is accepted if
\beq \label{eq:reject_samp}
\frac{p(\pars, X | D)}{M q(\pars, X)} = \frac{p(D|X)}{\mathrm{max}\;p(D|X)} > u 
\eeq
where $u$ is drawn from $\mathtt{Uniform}[0,1]$. By repeating this rejection sampling process, 
we sample the distribution $p(\pars, X |D)$ with the set of $\pars$ and $X$ that are accepted.

At this stage, ABC distinguishes itself by postulating that $p(D|X)$, 
the probability of observing data $D$ given simulation $X$ 
({\em not} the likelihood), is proportional to the 
probability of the distance between the data and the simulation X being less than 
an arbitrarily small threshold $\epsilon$ 
\beq
p(D|X) \propto p(\rho(D,X)<\epsilon)
\label{eq:abc_condition}
\eeq
where $\rho(D, X)$ is the distance between the data $D$ and simulation $X$. 
Eq. \ref{eq:abc_condition} along with the rejection sampling acceptance criteria 
(Eq. \ref{eq:reject_samp}), leads to the acceptance criteria for ABC: $\pars$ is accepted if $\rho(D, X) < \epsilon$.

The distance function is a positive definite function that measures the closeness 
of the data and the simulation. The distance can be a vector with multiple components 
where each component is a distance between a single summary statistic of the data 
and that of the simulation. In that case, the threshold $\epsilon$ in 
Eq. \ref{eq:abc_condition} will also be a vector with the same dimensions.  
$\pars$ is accepted if the distance vector is less than the threshold vector for 
every component.

The ABC procedure begins, in the same fashion as rejection sampling, by drawing 
$\pars$ from the prior distribution $\pi(\pars)$. The simulation is generated from 
$\pars$ using the forward model, $X = f(\pars)$. Then the distance between 
the data and simulation, $\vec\rho(D, X)$, is calculated and compared to 
$\vec\epsilon$. If $\vec\rho(D, X) < \vec\epsilon$, $\pars$ is accepted. 
This process is repeated until we are left with a sample of $\pars$ that all 
satisfy the distance criteria. This final ensemble approximates the posterior 
probability distribution $p(\pars, X|D)$.

As it is stated, the ABC method poses some practical challenges. If the 
threshold $\epsilon$ is arbitrarily large, the algorithm essentially 
samples from the prior $\pi(\pars)$. Therefore a sufficiently small threshold
is necessary to sample from the posterior probability distribution. However,
an appropriate value for the threshold is not known \emph{a priori}. Yet, 
even if an appropriate threshold is selected, a small threshold requires 
the entire process to be repeated for many draws of $\pars$ from $\pi(\pars)$ 
until a sufficient sample is acquired. This often presents computation challenges.

We overcome some of the challenges posed by the above ABC method
by using a Population Monte Carlo (PMC) algorithm as our sampling technique. 
PMC is an iterative method that performs rejection sampling over a 
sequence of $\pars$ distributions ($\{p_1(\pars), ..., p_T(\pars)\}$ for 
$T$ iterations), with a distance threshold that decreases at each iteration of 
the sequence.

\begin{algorithm} 
\caption{The procedure for ABC-PMC}
\begin{algorithmic}[1] \label{alg:abcpmc}
%\STATE \DATA: D
%\STATE \RESULT: ABC posterior sample of $\pars$
\IF{$t=1:$}
%\STATE $\epsilon_t \gets \infty$
\FOR{$i=1,...,N$}
   \STATE // \emph{This loop can now be done in parallel for all i}
   \WHILE{$\rho(X,D)>\epsilon_t$}
   \STATE $\pars^{*}_{t} \gets \pi(\pars)$
   \STATE $X = f(\pars^{*}_{t})$
   \ENDWHILE
   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
   \STATE $w^{(i)}_{t} \gets 1/N$
\ENDFOR
\ENDIF
\IF{$t=2,...,T:$}
\FOR{$i=1,...,N$}
   \STATE // \emph{This loop can now be done in parallel for all i}
   \WHILE{$\rho(X,D)>\epsilon_t$}
   \STATE Draw $\pars^{*}_{t}$ from $\{\pars_{t-1}\}$ with probabilities $\{w_{t-1}\}$
   \STATE $\pars^{*}_{t} \gets K(\pars^{*}_{t},.)$
   \STATE $X = f(\pars^{*}_{t})$
   \ENDWHILE
   \STATE $\pars^{(i)}_{t} \gets \pars^{*}_{t}$
   \STATE $w^{(i)}_{t} \gets \pi(\pars^{(i)}_{t}) / \big(\sum\limits_{j=1}^{N}w_{t-1}^{(i)}K(\pars^{(j)}_{t-1},\pars^{(i)}_{t}) \big)$
\ENDFOR
\ENDIF
\end{algorithmic}
\end{algorithm}

As illustrated in Algorithm \ref{alg:abcpmc}, for the first iteration $t = 1$, 
we begin with an arbitrarily large distance threshold $\epsilon_1$. We 
draw $\pars$ (hereafter referred to as particles) from the prior distribution 
$\pi(\pars)$. We forward model the simulation $X = f(\pars)$, calculate the 
distance $\rho(D, X)$, compare this distance to $\epsilon_1$, and then 
accept or reject the $\pars$ draw. Because we set $\epsilon_1$ arbitrarily large, 
the particles essentially sample the prior distribution. This process 
is repeated until we accept $N$ particles. We then assign equal weights to 
the $N$ particles: $w_1^i = 1/N$.

For subsequent iterations ($t > 1$) the distance threshold is set such that
$\epsilon_{i,t} < \epsilon_{i,t-1}$ for all components $i$. Although there is 
no general prescription, the distance threshold $\epsilon_{i,t}$ can be 
assigned based on the empirical distribution of the accepted distances of the 
previous iteration, $t-1$. In \citealt{abcsn}, for instance, the threshold of 
the second iteration is set to the $25^\mathrm{th}$ percentile of the distances 
in the first iterations; afterwards in the subsequent iterations, $t$, $\epsilon_{t}$ 
is set to the $50^\mathrm{th}$ percentile of the distances in the previous $t-1$ iteration. 
Alternatively, \citealt{abcwl} set $\epsilon_{t}$ to the median of the distances from 
the previous iteration. In Section \ref{sec:abcatwork}, we describe our prescription 
for the distance threshold, which follows \citealt{abcwl}.

Once $\epsilon_t$ is set, we draw a particle from the previous 
weighted set of particles ${\pars}_{t-1}$. 
This particle is perturbed by a kernel, set to the covariance of ${\pars}_{t-1}$.
Then once again, we generate a simulation by forward modeling $X = f(\pars^i)$, 
calculate the distance $\rho(X, D)$, and compare the distance to the new distance 
threshold ($\epsilon_t$) in order to accept or reject the particle. This process is 
repeated until we assemble a new set of $N$ particles ${\pars}_t$. We then update the 
particle weights according to the kernel, the prior distribution, and the 
previous set of weights, as described in Algorithm \ref{alg:abcpmc}. The 
entire procedure is then repeated for the next iteration, $t+1$.

There are a number of ways to specify the perturbation kernel in the ABC-PMC algorithm. 
A widely used technique is to define the perturbation kernel as a multivariate Gaussian 
centered on the weighted mean of the particle population with a covariance matrix set to 
the covariance of the particle population. This perturbation kernel is often called the 
global multivariate Gaussian kernel. For a thorough discussion of various schemes for 
specifying the perturbation kernel, we refer the reader to \citealt{optimalkernel}.

The iterations continue in the ABC-PMC algorithm until convergence is confirmed. 
One way to ensure convergence is to impose a threshold for the acceptance ratio, 
which is measured in each iteration. The acceptance ratio is the ratio of the number 
of proposals accepted by the distance threshold, to the full number 
of proposed particles at every step. Once the acceptance ratio for 
an iteration falls below the imposed threshold, the algorithm has converged and is
suspended. Another way to ensure convergence is by monitoring the fractional change in 
the distance threshold ($\epsilon_t/\epsilon_{t-1} - 1$)
after each iteration. When the fractional change becomes smaller than some 
specified tolerance level, the algorithm has reached convergence. Another 
convergence criteria, is through the derived uncertainties of the inferred
parameters measured after each iteration. When the uncertainties stabilize 
and show negligible variations, convergence is ensured. In Section \ref{sec:abcpmc_spec} 
we detail the specific convergence criteria used in our analysis.

\subsection{Forward model}\label{sec:forwardmodel}
\subsubsection{Halo Occupation Modeling}

\newcommand{\lcdm}{\Lambda {\rm CDM}}
\newcommand{\dd}{\mathrm{d}}
\newcommand{\mean}[2]{\left\langle#1 \vert {#2}\right\rangle}

%%% HALO OCCUPATION DISTRIBUTION %%%
\newcommand{\ngal}{N_{\mathrm{g}}}
\newcommand{\nsat}{N_\mathrm{s}}
\newcommand{\ncen}{N_\mathrm{c}}
\newcommand{\pnm}[2]{p(#1|#2)}

\newcommand{\mhalo}{M_{\rm h}}
\newcommand{\mvir}{M_\mathrm{vir}}

\newcommand{\dndmvir}{\frac{\dd n}{\dd\mvir}}
\newcommand{\dndmhalo}{\frac{\dd n}{\dd\mhalo}}
\newcommand{\dndmvirprime}{\frac{\dd n}{\dd\mvir'}}

%%% CLUSTERING STATISTICS %%%
\newcommand{\xigg}{\xi_{\mathrm{gg}}}
\newcommand{\xihh}{\xi_{\mathrm{hh}}}
\newcommand{\xiggr}{\xi_{\mathrm{gg}}(r)}
\newcommand{\xiggroneh}{\xi^{1h}_{\mathrm{gg}}(r)}
\newcommand{\xiggrtwoh}{\xi^{2h}_{\mathrm{gg}}(r)}
\newcommand{\ngalaxy}{\bar{n}_{\mathrm{g}}}
\newcommand{\gmf}{\mathcal{\zeta}_{\rm g}}

ABC requires a forward generative model. In large scale structure studies, this implies a model
that is able to generate a galaxy catalog. We then calculate and compare summary statistics of the data and model catalog in an identical fashion
In this section, we describe the forward generative model we use within the framework of the 
halo occupation distribution.

The assumption that galaxies reside in dark matter halos is the bedrock underlying 
all contemporary theoretical predictions for galaxy clustering. The Halo Occupation Distribution 
(HOD) is one of the most widely used approaches to characterizing this galaxy-halo connection. 
The central quantity in the HOD is $\pnm{\ngal}{\mhalo}$, the probability that a halo of mass 
$\mhalo$ hosts $\ngal$ galaxies.

The most common technical methods for estimating the theoretical galaxy 2PCF utilize the 
first two moments of $P$, which contain the necessary information to calculate the one- 
and two-halo terms of the galaxy correlation function:
\begin{eqnarray}
\label{eq:onehaloterm}
1+\xiggroneh \simeq \frac{1}{4\pi{}r^{2}\ngalaxy^{2}}\int\dd\mhalo\dndmhalo\Xi_{\rm gg}(r|\mhalo) \times \mean{\ngal(\ngal-1)}{\mhalo},
\end{eqnarray} and

\begin{eqnarray}
\label{eq:twohaloterm}
\xiggrtwoh \simeq \xi_{\mathrm{mm}}(r)\left[\frac{1}{\ngalaxy}\int\dd\mhalo\dndmhalo \mean{\ngal}{\mhalo}b_{\mathrm{h}}(\mhalo)\right]^{2}
\end{eqnarray}
In Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}), $\ngalaxy$ is the galaxy number density,
$\dd\mathrm{n}/\dd\mhalo$ is the halo mass function, the spatial bias of dark matter halos is 
$b_{\mathrm{h}}(\mhalo),$ and $\xi_{\rm mm}$ is the correlation function of dark matter.  
If we represent the spherically symmetric intra-halo distribution of galaxies by a unit-normalized 
$n_{\rm g}(r),$ then the quantity $\Xi_{\rm gg}(r)$ appearing in the above two equations 
is the convolution of $n_{\rm g}(r)$ with itself. These fitting functions are calibrated using $N$-body 
simulations.

Fitting function techniques, however, require many simplifying assumptions. For example, 
Eqs.~(\ref{eq:onehaloterm}) and (\ref{eq:twohaloterm}) assume that the galaxy 
distribution within a halo is spherically symmetric. These equations 
also face well-known difficulties of properly treating halo exclusion and scale-dependent bias, 
which results in additional inaccuracies commonly exceeding the $10\%$ level (\citealt{vdBosch13}). 
Direct emulation methods have made significant improvements in precision and accuracy in recent 
years (\citealt{coyote2,coyote1}); however, a labor- and computation-intensive interpolation 
exercise must be carried out each time any alternative statistic is explored, which is one of the goals of the present work.

To address these problems, throughout this paper we make no appeal to fitting functions or emulators. 
Instead, we use the $\mathtt{Halotools}$ package to populate dark matter halos with mock galaxies and 
then calculate our summary statistics directly on the resulting galaxy catalog with the same estimators 
that are used on observational data (\citealt{Hearin:2016aa}). Additionally, through our forward modeling approaching, we are
able to explore observables beyond the 2PCF, such as the group multiplicity function, for which 
there is no available fitting function. This framework allows us to use group multiplicity function for providing quantitative constraints on the galaxy-halo connection. In the following section, we will show that using this observable, we can obtain constraints on the HOD parameters comparable to those found from the 2PCF measurements.

For the fiducial HOD used throughout this paper, we use the model described in \citealt{zheng07}. 
The occupation statistics of central galaxies follow a nearest-integer distribution with first 
moment given by 
\begin{equation}
\label{eq:ncen}
\langle N_{\mathrm{cen}}\rangle = \frac{1}{2} \left[ 1 + \mathrm{erf}\left(\frac{\log M - \log M_{\mathrm{min}}}{\sigma_{\log M}}\right)\right].
\end{equation}
Satellite occupation is governed by a Poisson distribution with the mean given by 
\begin{equation}
\label{eq:nsat}
\langle N_{\mathrm{sat}}\rangle =  \langle N_{\mathrm{cen}} \rangle \left(\frac{M-M_{0}}{M_1}\right)^{\alpha}.
\end{equation}
We assume that central galaxies are seated at the exact center of the host dark matter halo and 
are at rest with respect to the halo velocity, defined according to $\mathtt{Rockstar}$ halo finder (\cite{rockstar})
as the mean velocity of the inner $10\%$ of particles in the halo. Satellite galaxies are confined to 
reside within the virial radius following an NFW spatial profile (\citealt{nfw}) with a concentration 
parameter given by the $c(M)$ relation (\citealt{nfw_c(M)}). The peculiar velocity of satellites with 
respect to their host halo is calculated according to the solution of the Jeans equation of an NFW 
profile (\citealt{more2010}). We refer the reader to \cite{hearin15}, 
\cite{Hearin:2016aa}, and \url{http://halotools.readthedocs.io}
for further details.

For the halo catalog of our forward model, we use the publicly available $\mathtt{Rockstar}$ 
(\citealt{rockstar}) halo catalogs of the $\mathtt{MultiDark}$ cosmological $N$-body simulation 
(\citealt{multidark}).\footnote{In particular, we use the {\tt halotools\_alpha\_version2} version of this catalog, made publicly available as part of {\tt Halotools}.} $\mathtt{MultiDark}$ is a collision-less dark-matter only $N$-body simulation. 
The $\Lambda$CDM cosmological parameters of $\mathtt{MultiDark}$ are $\Omega_m = 0.27$, $\Omega_{\Lambda}=0.73$,
$\Omega_{b}=0.042$, $n_{s}=0.95$, $\sigma_{8} = 0.82$, and $h = 0.7$. The gravity solver used in the 
$N$-body simulation is the Adaptive Refinement Tree code (ART; \citealt{art}) run on $2048^3$ particles in a
$1\; h^{-1}\mathrm{Gpc}$ periodic box. $\mathtt{MultiDark}$ particles have a mass of $m_{p} \simeq 8.72 \times 
10^{8} \; h^{-1}M_{\odot}$; the force resolution of the simulation is $\epsilon \simeq 7 h^{-1}$ kpc.

One key detail of our forward generative model is that when we populate the $\mathtt{MultiDark}$ halos 
with galaxies, we do not populate the entire simulation volume. Rather, we divide the volume into a 
grid of $125$ cubic subvolumes, each with side lengths of $200\;h^{-1}\mathrm{Mpc}$. We refer to these 
subvolumes as $\{\mathtt{BOX1}, ..., \mathtt{BOX125}\}$. 
The first subvolume is reserved to generate the mock observations which we describe in Section 
\ref{sec:mock_obv}. When we simulate a galaxy catalog for a given $\pars$ in parameter space, 
we randomly select one of the subvolumes from $\{\mathtt{BOX2}, ..., \mathtt{BOX125}\}$ and then 
populate the halos within this subvolume with galaxies. 
We implement this procedure to account for sample variance within our forward generative model.

\subsection{Summary Statistics}\label{sec:statistics}
One of the key ingredients for parameter inference using ABC, is the distance metric between 
the data and the simulations. In essence, it quantifies how close the simulation 
is to reproducing the data. The data and simulation in our scenario (the HOD framework) 
are galaxy populations and their positions. A direct comparison, which would involve comparing 
the actual galaxy positions of the populations, proves to be difficult. 
Instead, a set of statistical summaries are used to encapsulate the information of the data 
and simulations. These quantities should sufficiently describe the information of the data
and simulations while providing the convenience for comparison. 
For the positions of galaxies, sensible summary statistics, 
which we later use in our analysis, include

\begin{itemize}
\item Galaxy number density, $\ngalaxy$: the comoving number density of galaxies computed  
by dividing the comoving volume of the sample from the total number of galaxies. $\ngalaxy$ is 
measured in units of $(\mathrm{Mpc}/h)^{-3}$.

\item Galaxy two-point correlation function, $\xigg(r)$: a measurement of the excess probability 
of finding a galaxy pair with separation $r$ over an random distribution. To compute $\xigg(rr)$
in our analysis, for computational reasons, we use the Natural estimator (\citealt{peebles80}): 
\beq
\xi(r) = \frac{DD}{RR} - 1, 
\eeq
where $DD$ and $RR$ refer to counts of data-data and random-random pairs.

\item Galaxy group multiplicity function, $\gmf(N)$: the number density of galaxy groups in bins 
of group richness $N$ where group richness is the number of galaxies within a galaxy group. We 
rely on a Friends-of-Friends (hereafter FoF) group-finder algorithm (\citealt{fof}) to identify 
galaxy groups in our galaxy samples. That is, if the separation of a galaxy pair is smaller than a 
specified linking length, the two galaxies are assigned to the same group. The FoF group-finder 
has been used to identify and analyze the galaxy groups in the SDSS main galaxy sample (\cite{groups}). 
For details regarding the group finding algorithm, we refer readers to \cite{fof}.

In this study we set the linking length to be $0.25$ times the mean separation of galaxies which 
is given by $\ngalaxy^{-1/3}$. Once the galaxy groups are identified, we bin them into bins of 
group richness. The total number of groups in each bin is divided by the comoving volume to get 
$\gmf(N)$ --- in units of $(\mathrm{Mpc}/h)^{-3}$.

\end{itemize}

%%%%%%%%%%%%%%%%%%%%% SECTION %%%%%%%%%%%%%%%%%%%%%%%%%%
%  ABC at work 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{ABC at work}\label{sec:abcatwork}
With the methodology and the key components of ABC explained above, here we set out to 
demonstrate how ABC can be used to constrain HOD parameters. We start, in Section \ref{sec:mock_obv} 
by creating our ``observation''. We select a set of HOD parameters which we deem as the ``true'' 
parameters and run it through our forward model producing a catalog of galaxy positions 
which we treat as our observation. Then, in Section \ref{sec:abcpmc_spec}, we explain 
the distance metric and other specific choices we make for the ABC-PMC algorithm. 
Ultimately, we demonstrate the use of ABC in LSS, in Section \ref{sec:abc_results},
where we present the parameter constraints we get from our ABC analyses. 
Lastly, in order to both assess the quality of the ABC-PMC parameter inference and also 
discuss the assumptions of the standard Gaussian likelihood approach, we compare the 
ABC-PMC results to parameter constraints using the standard approach in Section 
\ref{sec:abcvsmcmc}.

%%%%%%%%%%%%%%%%%%%%% SUBSECTION %%%%%%%%%%%%%%%%%%%%%%%
% Mock Observations  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Mock Observations}\label{sec:mock_obv}
In generating our ``observations'', and more generally for our forward model, we adopt 
the HOD model from \cite{zheng07} where the expected number of galaxies populating a 
dark matter halo is governed by Eqs (\ref{eq:ncen}) and (\ref{eq:nsat}). For the parameters 
of the model used to generate the fiducial mock observations, we choose the \cite{zheng07} best-fit HOD parameters for the 
SDSS main galaxy sample with a luminosity threshold $M_{r} = -21$: 
\begin{center}
\renewcommand{\arraystretch}{1.5}
\begin{tabular}{ccccc} \hline \hline 
$\log M_{\rm min} \; \left[ h^{-1} M_{\odot} \right]$ & $\sigma_{\log M}$ & $\log M_{0} \; \left[ h^{-1} M_{\odot} \right]$ & $\log M_{1} \; \left[ h^{-1} M_{\odot} \right]$ & $\alpha$ \\ \hline
$12.79$ & $0.39$ & $11.92$ & $13.94$ & $1.15$ \\ \hline 
\end{tabular} \par
\end{center}
Since these parameters are used to generate the mock observation, they are the parameters
that we ultimately want to recover from our parameter inference. We refer to them as the
true HOD parameters. Plugging them into our forward model (Section \ref{sec:forwardmodel}), 
we generate a catalog of galaxy positions.

For our summary statistics of the catalogs we use: 
the mean number density $\ngalaxy$, the galaxy two-point correlation function $\xigg(r)$, 
and the group multiplicity function $\gmf(N)$. Our mock observation catalog has 
$\ngalaxy = 9.28875 \times 10^{-4} \;h^{-3} \mathrm{Mpc}^3$ and in Figure \ref{fig:mock} 
we plot $\xigg(r)$ (left panel) and $\gmf(N)$ (right panel). The width of the shaded region
represent the square root of the diagonal elements of the summary statistic covariance matrix, 
which is computed as we describe below.

We calculate $\xigg$ using the natural estimator (Section \ref{sec:statistics}) with 
fifteen radial bins. The edges of the first radial bin are $0.15$ and $0.5 \; 
h^{-1}\mathrm{Mpc}$. The bin edges for the next 14 bins are logarithmically-spaced between 
$0.5$ and $20\;h^{-1}\mathrm{Mpc}$. We compute the $\gmf(N)$ as described in Section
\ref{sec:statistics} with nine richness bins where the bin edges are logarithmically-spaced 
between $3$ and $20$. To calculate the covariance matrix, we first run the forward model 
using the true HOD parameters for all $125$ halo catalog subvolumes: 
$\{\mathtt{BOX1}, ..., \mathtt{BOX125}\}$. 
We compute the summary statistics of each subvolume galaxy sample $k$: 
\beq
\mathbf{x}^{(k)}=[\ngalaxy, \; \xigg, \; \gmf],
\eeq
Then we compute the covariance matrix as
\begin{eqnarray} 
\mathrm{C}^\mathrm{sample}_{i,j} &=& 
\frac{1}{N_{\mathrm{mocks}}-1}\sum_{k=1}^{N_{\mathrm{mocks}}} 
\Big[\mathbf{x}^{(k)}_{i}-\overline{\mathbf{x}}_{i}\Big]
\Big[\mathbf{x}^{(k)}_{j}-\overline{\mathbf{x}}_{j}\Big], \label{eq:cov} \\
\mathrm{where} \; \overline{\mathbf{x}}_{i} &=& 
\frac{1}{N_{\mathrm{mocks}}}\sum_{k=1}^{N_{\mathrm{mocks}}} \mathbf{x}^{(k)}_{i}.
\end{eqnarray}

Throughout our ABC-PMC analysis, we treat the $\ngalaxy$, $\xigg(r)$, and $\gmf(N)$ we describe in this 
section as if they were the summary statistics of actual observations. 
However, we benefit from the fact that these observables are generated from mock observations 
using the true HOD parameters of our choice: we can use the true HOD parameters to assess the 
quality of the parameter constraints we obtain from ABC-PMC.

%%%%%%%%%%%%%%%%%%%%% SUBSECTION %%%%%%%%%%%%%%%%%%%%%%%
% Mock Observations  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{ABC-PMC Design} \label{sec:abcpmc_spec}
In Section \ref{sec:abc}, we describe the key components of the ABC algorithm we use in our analysis.
Now, we describe the more specific choices we make within the algorithm: the distance metric, 
the choice of priors, the distance threshold, and the convergence criteria. So far we have 
described three summary statistics: $\ngalaxy$, $\xigg(r)$, and $\gmf(N)$. In order to explore 
the detailed differences in the ABC-PMC parameter constraints based on our choice of summary 
statistics, we run our analysis for two sets of observables: ($\ngalaxy$, $\xigg$) and 
($\ngalaxy$, $\gmf$).

For both analyses, we use a multi-component distance (\citealt{silk12}, Cisewsky et al in preparation). 
Each summary statistic has a distance associated to it: $\rho_{n}$, $\rho_{\xi}$, and $\rho_{\zeta}$. 
We calculate each of these distance components as,   
\begin{eqnarray}
\rho_{n} &=& \frac{\left(\ngalaxy^\mathrm{d} - \ngalaxy^\mathrm{m} \right)^{2}}{\sigma^{2}_{n}}, \label{dn}\\
\rho_{\xi} &=& \sum_{k} \frac{\left[\xigg^\mathrm{d}(r_k) - \xigg^\mathrm{m}(r_k) \right]^{2}}{\sigma_{\xi,k}^{2}}, \label{dxi} \\
\rho_{\zeta} &=& \sum_{k} \frac{\left[\gmf^\mathrm{d}(N_k) - \gmf^\mathrm{m}(N_k) \right]^{2}}{\sigma_{\zeta,k}^{2}}. \label{dgmf}
\end{eqnarray}
The superscripts $\mathrm{d}$ and $\mathrm{m}$ denote the data and model respectively. The data, 
are the observables calculated from the mock observation (Section \ref{sec:mock_obv}). 
$\sigma_{n}^{2}$, $\sigma_{\xi,k}^{2}$, and $\sigma_{\zeta,k}^{2}$ are not the diagonal elements 
of the covariance matrix (\ref{eq:cov}). Instead, they are diagonal elements of the covariance matrix 
$C^\mathrm{ABC}$.

We construct $C^\mathrm{ABC}$ by populating the entire $\mathtt{MultiDark}$ halo catalogs $125$ times 
repeatedly, calculating $\ngalaxy$, $\xigg$, and $\gmf$ for each realization, and then computing 
the covariance associated with these observables across all realizations. We highlight that $C^\mathrm{ABC}$ 
differs from Eq. \ref{eq:cov}, in that it does not populate the $125$ subvolumes but the entire 
$\mathtt{MultiDark}$ simulation and therefore does not incorporate sample variance. The ABC-PMC analysis 
instead accounts for the sample variance through the forward generative model, which populates the subvolumes
in the same manner as the observations. We use $\sigma_{n}^{2}$, $\sigma_{\xi,k}^{2}$, and $\sigma_{\zeta,k}^{2}$ 
to ensure that the distance is not biased to variations of observables on specific radial or richness bin.

For our ABC-PMC analysis using the observables $\ngalaxy$ and $\xigg$, our distance metric 
$\vec\rho = [\rho_n, \rho_\xi]$ while the distance metric for the ABC-PMC analysis using 
the observables $\ngalaxy$ and $\gmf$, is $\vec\rho = [\rho_n, \rho_\zeta]$.

To avoid any complications from the choice for our prior, we select uniform priors over all parameters aside from 
the scatter parameter $\sigma_{\log\;M}$, for which we choose a log-uniform prior. 
We list the range of our prior distributions in Table ~\ref{tab:prior}.

With the distances and priors specified, we now describe the distance thresholds and the 
convergence criteria we impose in our analyses. For the initial iteration, we set distance
thresholds for each distance component to $\infty$. This means, that the initial pool 
$\vec\theta_1$ is simply sampled from the prior distribution we specify above. After the initial 
iteration, the distance threshold is adaptively lowered in subsequent iterations. More 
specifically, we follow the choice of \cite{abcwl} and set the distance threshold 
$\vec\epsilon_t$ to the median of $\vec\rho_{t-1}$, the multi-component distance of the 
previous iteration of particles ($\pars_{t-1}$).

The distance threshold $\vec\epsilon_t$ will progressively decrease. Eventually after a 
sufficient number of iterations, the region of parameter space occupied by $\pars_t$ 
will remain unchanged. As this happens, the acceptance ratio begins to fall significantly. 
When the acceptance ratio drops below $0.001$, our acceptance ratio threshold of choice, 
we deem the ABC-PMC algorithm as converged. In addition to the acceptance ratio threshold 
we impose, we also ensure that distribution of the parameters converges -- another sign 
that the algorithm has converged. Next, we present the results of our ABC-PMC analyses 
using the sets of observables ($\ngalaxy$, $\xigg$) and ($\ngalaxy$, $\gmf$).

%%%%%%%%%%%%%%%%%%TABLES%%%%%%%%%%%%
\begin{table*}
\begin{center}
  \label{tab:prior}
  \caption{{\bf Prior Specifications}: The prior probability distribution 
  and its range for each of the \cite{zheng07} HOD parameters. 
  All mass parameters are in unit of $h^{-1}M_\odot$}
\begin{tabular}{@{}lllll}
\\ \hline 
    HOD Parameter & & Prior & & Range \\ \hline
  $\alpha$ & & Uniform & & [0.8, 1.3] \\
  $\sigma_{\rm \log M}$ & & Log--Uniform & &  [0.1, 0.7] \\
   $\log M_{0}$   & & Uniform & &  [10.0, 13.0] \\
  $\log M_{min}$ & &   Uniform & &  [11.02, 13.02] \\
  $\log M_{1}$ & & Uniform & & [13.0, 14.0] \\
 \hline
  \end{tabular}
\end{center}
\end{table*}

%%%%%%%%%%%%%%%%%%%%%%% SECTION: %%%%%%%%%%%%%%%%%%%%%%%
% SECTION: RESULTS 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Results: ABC}\label{sec:abc_results}
We describe the ABC algorithm in Section \ref{sec:abc} and list the particular choices
we make in the implementation in the previous section. Finally, we demonstrate how the 
ABC algorithm produces parameter constraints and present the results of our ABC analysis -- the 
parameter constraints for the \cite{zheng07} HOD model.

We begin with a qualitative demonstration of the ABC algorithm in Figure \ref{fig:pool_demo},
where we plot the evolution of the ABC $\pars_t$ over the iterations $t = 1$ to $9$, 
in the parameter space of $[\log\mathcal{M}_1, \log\mathcal{M}_{min}]$. The ABC procedure we plot 
in Figure \ref{fig:pool_demo} uses $\bar{n}$ and $\gmf(N)$ for observables, but 
the overall evolution is the same when we use $\bar{n}$ and $\xigg(r)$. The darker and lighter 
contours represent the $68\%$ and $95\%$ confident regions of the posterior distribution over $\pars_t$. For 
reference, we also plot the ``true'' HOD parameter $\pars_\mathrm{true}$ 
(black star) in each of the panels.
The parameter ranges of the panels are equivalent to the ranges of the prior probabilities we 
specify in Table \ref{tab:prior}.

For $t=1$, the initial pool (top left), the distance threshold $\vec\epsilon_1 = [\infty , \infty]$, 
so $\pars_1$ uniformly samples the prior probability over the parameters.
At each subsequent iteration, the threshold is lowered (Section \ref{sec:abcatwork}), so for 
$t < 6$ panels, we note that the parameter spaced occupied by $\pars_t$ dramatically shrinks. 
Eventually when the algorithm begins to converge, $t > 7$, the contours enclosing the $68\%$ and 
$95\%$ confidence interval stabilize. At the final iteration $t=9$ (bottom right), 
the algorithm has converged and we find that 
$\pars_\mathrm{true}$ lies within the $68\%$ confidence interval of the $\pars_{t=9}$ particle 
distribution. This $\pars_t$ distribution at the final iteration represents the 
posterior distribution of the parameters.

To better illustrate the criteria for convergence, in Figure \ref{fig:abc_converge}, 
we plot the evolution of the $\pars_t$ distribution as a function of iteration 
for parameters $\log\mathcal{M}_\mathrm{min}$ (left), $\alpha$ (center), and 
$\log\mathcal{M}_1$ (right). The darker and lighter shaded regions correspond to the 
$68\%$ and $95\%$ confidence levels of the $\pars_t$ distributions. The top panels 
correspond to our ABC results using $(\bar{n}, \gmf)$ as observables and the bottom 
panels correspond to our results using $(\bar{n}, \xigg)$. 
For each of the parameters in both top and bottom panels, we find that the distribution 
does not evolve significantly for $t > 7$. At this point additional iterations in
our ABC algorithm will neither impact the distance threshold $\vec\epsilon_t$ nor 
the posterior distribution of $\pars_t$. We also emphasize that the convergence of the 
parameter distributions coincides with when the acceptance ratio, discussed in Section 
\ref{sec:abcpmc_spec}, crosses the predetermined shut-off value of $0.001$. Based on these criteria, 
our ABC results for both $(\bar{n}, \gmf)$ and $(\bar{n}, \xigg)$ observables have converged.

We present the parameter constraints from the converged ABC analysis in Figure \ref{fig:abc_corner_nbarxi} and 
Figure \ref{fig:abc_corner_nbargmf}. Figure \ref{fig:abc_corner_nbarxi} shows 
the parameter constraints using $\bar{n}$ and $\xigg(r)$ while Figure 
\ref{fig:abc_corner_nbargmf} plots the constraints using $\bar{n}$ and $\gmf(N)$.
For both figures, the diagonal panels plot the posterior distribution of the 
HOD parameters with vertical dashed lines marking the $50\%$ (median) and $68\%$ 
confidence intervals. The off-diagonal panels plot the degeneracy between  
parameter pairs. To determine the accuracy of our ABC parameter constraints, we plot
the ``true'' HOD parameters (black) in each of the panels.
For both sets of observables, our ABC constraints are consistent with the 
``true'' HOD parameters. For $\log\mathcal{M}_0$, $\log\sigma_{\log M}$,
and $\alpha$, the true parameter values lie near the center of the $68\%$ confidence 
interval. For the other parameter, which have much tighter constraints, the true 
parameters lie within the $68\%$ confidence interval.

To further test the ABC results, in Figure \ref{fig:abc_moneyplot}, 
we compare $\xigg(r)$ (left) and $\gmf(N)$ (right) of the mock observations from Section \ref{sec:mock_obv}
to the predictions of the ABC posterior distribution (shaded). The error bars of the mock observations 
represent the square root of the diagonal elements of the covariance matrix (Eq. \ref{eq:cov}) while the 
darker and lighter shaded regions represent the $68\%$ and $95\%$ confidence regions of the ABC posterior 
predictions. In the lower panels, we plot the ratio of the ABC posterior prediction $\xigg(r)$ and $\gmf(N)$ over
the mock observation $\xigg^\mathrm{obvs}(r)$ and $\gmf^\mathrm{obvs}(N)$. Overall, the 
ratio of the $68\%$ confidence region of ABC posterior predictions is consistent with unity 
throughout the $r$ and $N$ range. We observe slight deviations in the $\xigg$ ratio for 
$r > 5\;\mathrm{Mpc}/h$; however, any deviation is within the uncertainties of the mock observations. 
Therefore, the observables drawn from the ABC posterior distributions are in good agreement with the
observables of the mock observation.

The ABC results we obtain using the algorithm of Section \ref{sec:abc} with the choices of
Section \ref{sec:abcpmc_spec} produce parameter constraints that are consistent with 
the ``true'' HOD parameters (Figures \ref{fig:abc_corner_nbarxi} and \ref{fig:abc_corner_nbargmf}).
They also produce observables $\xigg(r)$ and $\gmf(N)$ that are consistent with $\xigg^\mathrm{obvs}$ 
and $\gmf^\mathrm{obvs}$. Thus, through ABC we are able to produce consistent 
parameter constraints. {\em More importantly, we demonstrate that ABC is feasible 
for parameter inference in large scale structure.} % boom mic drop.

%%%%%%%%%%%%%%%%%%%%%%% SECTION: %%%%%%%%%%%%%%%%%%%%%%%
% SECTION: Comparison to MCMC analysis  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison to the Gaussian Likelihood MCMC Analysis} \label{sec:abcvsmcmc}
In order to assess the quality of the parameter inference described in the previous section, 
we compare the ABC-PMC results with the HOD parameter constraints assuming a Gaussian likelihood 
function. The model used for the Gaussian likelihood analysis is different than the forward 
generative model adopted for the ABC-PMC algorithm, to be consistent with the standard approach.
In the ABC analysis, the model accounts for sample variance by randomly sampling a subvolume to be 
populated with galaxies. The Gaussian likelihood analysis assumes a covariance matrix that captures 
the uncertainties from the sample variance, and therefore in the model, we populate the halos of
the entire $\mathtt{MultiDark}$ simulation rather than a subvolume.

We introduce the observable $\mathbf{x}$ as a combination of the summary statistics of the 
galaxy mock catalog used in the inference. When we use $\ngalaxy$ and $\xigg(r)$ as observables, 
$\mathbf{x} = [\ngalaxy , \xigg]$, while $\mathbf{x} = [\ngalaxy , \gmf]$ when we use $\ngalaxy$ and $\gmf(N)$
Based on this notation, we write likelihood function as 
\begin{eqnarray}
-2 \ln \mathcal{L}(\theta | d) &=& \Delta \mathbf{x}^{T}\widehat{C^{-1}}\Delta \mathbf{x} + \ln\Big[(2\pi)^{d}\mathrm{det}(C)\Big], \label{loglike}\\
\Delta \mathbf{x} &=& [\mathbf{x}_{obs} -\mathbf{x}_{mod}],
\end{eqnarray}
where $\Delta \mathbf{x}$ is the difference between the measured $\mathbf{x}_{obs}$ of the mock observations 
and that of the model $\mathbf{x}_{mod}(\mathbb{\theta})$. $d$ is the dimension of the observable $\mathbf{x}$. 
When we use $\mathbf{x} = [\ngalaxy, \xigg]$, $d = 13$. When we use $\mathbf{x} = [\ngalaxy, \gmf]$, $d = 10$.  
$\widehat{C^{-1}}$ is the estimate of the inverse covariance matrix, which we calculate 
following  \cite{hartlap2007}:
\beq
\widehat{C^{-1}} = \frac{N_{\rm mocks}-d - 1}{N_{\rm mocks} -1}\widehat{C}^{-1},
\eeq
where $\widehat{C}$ is the estimated covariance matrix and $N_{\rm mock}$ is the number of mocks used to 
estimate $\widehat{C}$. $N_{\rm mock} = 124$ (Section \ref{sec:mock_obv}). Depending on $\mathbf{x}$, 
$\widehat{C}$ is given by the appropriate block of the covariance matrix Eq. \ref{eq:cov} corresponding 
to the observables. We note that in the covariance matrix the dependence on the HOD parameters is neglected. 
Therefore, the second term in the expression of the log-likelihood (Eq. \ref{loglike}) can be neglected. 
We sample from the posterior distribution of the likelihood given the prior distribution using the 
MCMC sampler $\mathtt{EMCEE}$ (\citealt{emcee}).

In Figure \ref{fig:hist_nbarxi} we compare the ABC-PMC and Gaussian likelihood MCMC 
results for the marginalized posterior PDFs over three parameters of the HOD model $\{\log \mathcal{M}_{\rm min},
\alpha , \log\mathcal{M}_1\}$ using $\ngalaxy$ and $\xigg$ as observables. 
Figure \ref{fig:hist_nbargmf} makes the same comparison when $\ngalaxy$ and $\gmf$ are used in the inference.

Figure \ref{fig:hist_nbarxi} demonstrates in the top panel that constraints obtained from the two 
methods are consistent with the ``true'' HOD parameters represented by the vertical dashed lines. 
The marginalized posteriors over the parameters follow each other very closely. The lower panel of 
Figure \ref{fig:hist_nbarxi} plot the $68\%$ and $95\%$ confidence intervals of the constraints 
derived from the two methods as a box plot. The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ 
is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ 
are slightly larger for our ABC-PMC results.

In Figure \ref{fig:hist_nbargmf}, when we use $\ngalaxy$ and $\gmf$ as observables, both methods 
produce consistent constraints with each other and ``true'' HOD parameters. The lower panel of Figure 
\ref{fig:hist_nbargmf} shows that the uncertainties from the two methods are comparable. The 
ABC-PMC constraints over $\alpha$ is slightly less biased and slightly less precise.

In Figures \ref{fig:cont_nbarxi} and \ref{fig:cont_nbargmf}, we plot the contours enclosing the $68\%$ 
and $95\%$ confidence regions of the posterior probabilities of the two methods. Figure \ref{fig:cont_nbarxi}
uses $\ngalaxy$ and $\xigg$ as observables, while Figure \ref{fig:cont_nbargmf} uses $\ngalaxy$ and $\gmf$ 
as observables. 
In both figures, the ``true'' HOD parameters are plotted as black stars. The overall shape of the contours 
are generally in good agreement with each other. However, the contours for the ABC-PMC method are
more elongated along $\alpha$.

In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 
MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using 
ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, 
accounts for sample variance. In principle, they also have the advantage that they can account for 
sources of systematic uncertainties in the observational data.

Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix
estimate that captures the sample variance of the data. While we are able to do this accurately 
within the scope of the HOD framework, for more general LSS parameter inference situations, it 
is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs,
which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein). 
Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a 
covariance matrix estimate; hence it does not face these problems.

{\em ABC-PMC -- unlike the Gaussian-likelihood approach -- is agnostic about the functional form of the 
underlying distribution of the summary 
statistics} ($\xigg$ and $\gmf$). This may explain why we find less biased constraints from the
ABC-PMC analysis compared to the Gaussian-likelihood analysis, when using $\gmf(N)$ as an observable 
(Figures \ref{fig:hist_nbargmf} and \ref{fig:cont_nbargmf}).

The ABC method of parameter inference has often been viewed as computationally infeasible due to the 
fact that it necessitates a generative forward model for a large number of realizations. However, 
by combining ABC with the PMC sampling method,
%of importance sampling {\bf [APH: the term "importance sampling" has not been defined]}, 
we have a method that efficiently converges 
to give us reliable posteriors of the model parameters. Furthermore, in our analysis, the computational resources 
required for the ABC-PMC inference were comparable to those used in achieving convergence in the 
MCMC Gaussian-Likelihood inference. This is excluding the typically computationally intensive step of 
calculating a covariance matrix estimate. Therefore, we find that ABC-PMC offers a method for parameter
inference in large scale structure studies with a number of advantages we describe above.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Summary and Conclusion  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary and Conclusion}\label{sec:discussion}

%Approximate Bayesian Computation, ABC, is a generative, simulation-based 
%inference that is guaranteed to deliver correct parameter estimation with 
%an appropriate set of summary statistics of the data is selected. 
Approximate Bayesian Computation, ABC, is a generative, simulation-based
inference that can deliver correct parameter estimation with
appropriate choices for its design.
It has the advantage over the standard approach in that it 
does not require explicit knowledge of the likelihood function. It only relies on the ability to simulate
the observed data, accounting for the uncertainties associated with observation and on specifying a metric 
for the distance between the observed data and simulation. When the specification of the likelihood function 
proves to be challenging or when the true underlying distribution of the observable is unknown, 
ABC provides a promising alternative for inference.

The standard approach to large scale structure studies relies on the assumption that the likelihood function 
for the observables -- often two-point correlation function -- given the model has a Gaussian functional form. 
In other words, it assumes that the statistical summaries are Gaussian distributed. In principle to rigorously 
test such an assumption, a large number of realistic simulations would need to be generated in order to examine 
the actual distribution of the observables. This process, however, is prohibitively---both labor and computationally 
---expensive. Therefore, our assumption of a Gaussian likelihood function remains largely unconfirmed and so 
unknown. Fortunately, the framework of ABC permits us to bypass any 
assumptions regarding the distribution of observables. Through ABC, we can provide constraints for our models without making the unexamined assumption of Gaussianity.

With the ultimate goal of demonstrating that ABC is feasible for LSS studies, we use it to 
constrain parameters of the halo occupation distribution, which dictates the galaxy-halo connection. 
We begin by constructing a mock observation of galaxy distribution with a chosen set of ``true'' HOD model 
parameters. Then we attempt to constrain these parameters using ABC. More specifically, in this paper:

\begin{itemize}
\item We provide an explanation of the ABC algorithm and present how Population Monte Carlo can be utilized 
to efficiently reach convergence and estimate the posterior distributions of model parameters. 
We use this ABC-PMC algorithm with a generative forward model built with $\mathtt{Halotools}$, a software 
package for creating catalogs of galaxy positions based on models of the galaxy-halo connection such as the HOD.

\item We choose $\ngalaxy$, $\xigg$ and $\gmf$ as observables and summary statistics of the galaxy position catalogs. 
And for our ABC-PMC algorithm, we specify a multi-component distance metric, uniform priors, 
a median threshold implementation, and an acceptance rate-based convergence criterion.

\item From our specific ABC-PMC method, we obtain parameter constraints that are consistent with the ``true'' 
HOD parameters of our mock observations. Hence we demonstrate that ABC-PMC can be used for parameter inference
in LSS studies.

\item We compare our ABC-PMC parameter constraints to constraints using the standard Gaussian-likelihood
MCMC analysis. The constraints we get from both methods are comparable in accuracy and precision. However, 
for our analysis using $\ngalaxy$ and $\gmf$ in particular, we obtain less biased posterior distributions when
comparing to the ``true'' HOD parameters. 
\end{itemize}

Based on our results, we conclude that ABC-PMC is able to consistently infer parameters in the context
of LSS. We also find that the computation required for our ABC-PMC and standard Gaussian-likelihood 
analyses are comparable. Therefore, with the statistical advantages that ABC offers, we present ABC-PMC 
as an improved alternative for parameter inference.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Mock observation 2PCF and GMF
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{paper_data_observables.pdf}
\caption{The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
function $\gmf(N)$ (right) summary statistics of the mock observations generated from 
the ``true'' HOD parameters described in Section \ref{sec:mock_obv}. The width of the 
shaded region corresponds to the square root of the covariance matrix diagonal 
elements (Eq. \ref{eq:cov}). In our ABC analysis, we treat the $\xigg(r)$ 
and $\gmf(N)$ above as the summary statistics of the observation.}
\label{fig:mock}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Pool Evolution plot 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=0.95\textwidth]{paper_ABC_poolevolution_nbargmf.pdf}
\caption{We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ 
in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
$\bar{n}$ and $\gmf(N)$ are used as observables for the above results. For reference, 
in each panel, we include the ``true'' HOD parameters (black star) listed in 
Section \ref{sec:mock_obv}. The initial distance threshold, $\vec\epsilon_1 = 
[\infty, \infty]$ at $t=1$ (top left) so the $\pars_1$ spans the entire range of the prior distribution, 
which is also the range of the panels. We see for $t < 5$, the parameter space occupied 
by the ABC $\pars_t$ shrinks dramatically. Eventually when the algorithm converges, 
$t > 7$, the parameter space occupied by $\pars_t$ no longer shrinks and their 
distributions represent the posterior distribution of the 
parameters. At $t=9$, the final iteration, the ABC algorithm has converged and 
we find that $\pars_\mathrm{true}$ lies safely within the $68\%$ confidence region.
} 
\label{fig:pool_demo}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Pool Convergence 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{paper_ABCconvergence.pdf}
\caption{We illustrate the convergence of the ABC algorithm through the evolution 
of the ABC particle distribution as a function of iteration for 
parameters $\log \mathcal{M}_\mathrm{min}$ (left), $\alpha$ (center), and 
$\log \mathcal{M}_1$ (right).
The top panel corresponds our ABC results using the observables $(\bar{n}, \gmf(N))$, 
while the lower panel plots  corresponds to the ABC results using $(\bar{n}, \xigg(r))$. 
The distributions of parameters show no significant change after $t > 7$, which suggests
that the ABC algorithm has converged.}
\label{fig:abc_converge}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Corner Plots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbarxi.pdf}
\caption{We present the constraints on the \cite{zheng07} HOD model parameters
obtained from our ABC-PMC analysis using $\bar{n}$ and $\xigg(r)$ as observables.  
The diagonal panels plot the posterior distribution of each HOD parameter with 
vertical dashed lines marking the $50\%$ quantile and $68\%$ confidence intervals of the 
distribution. The off-diagonal panels plot the degeneracies between parameter pairs. 
The range of each panel corresponds to the range of our prior choice. The ``true''
HOD parameters, listed in Section \ref{sec:mock_obv}, are also plotted in each of 
the panels (black). For $\log\mathcal{M}_0$, $\alpha$, and $\sigma_{\log M}$, 
the ``true'' parameter values lie near the center of the $68\%$ confidence interval 
of the posterior distribution. For $\log\mathcal{M}_1$ and $\log\mathcal{M}_\mathrm{min}$, 
which have tight constraints, the ``true'' values lie within the $68\%$ confidence 
interval. Ultimately, the ABC parameter constraints we obtain in our analysis are consistent 
with the ``true'' HOD parameters. 
}
\label{fig:abc_corner_nbarxi}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbargmf.pdf}
\caption{Same as Figure \ref{fig:abc_corner_nbarxi} but for our ABC analysis 
using $\bar{n}$ and $\gmf(N)$ as observables. The ABC parameter constraints we obtain are
consistent with the ``true'' HOD parameters.}
\label{fig:abc_corner_nbargmf}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC Posterior Observable  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=\textwidth]{paper_ABCposterior.pdf}
\caption{We compare the ABC-PMC posterior prediction for the observables $\xigg(r)$ (left) and $\gmf(N)$ (right) (orange; Section \ref{sec:abc_results}) 
to $\xigg(r)$ and $\gmf(N)$ of the mock observation (black) in the top panels.
In the lower panels, we plot the ratio between the ABC-PMC posterior predictions for $\xigg$ and $\gmf$ 
to the mock observation $\xigg^\mathrm{obvs}$ and $\gmf^\mathrm{obvs}$. The darker 
and lighter shaded regions represent the $68\%$ and $95\%$ confidence regions of the posterior predictions, 
respectively. The error-bars represent the square root of the diagonal elements of the error 
covariance matrix (equation \ref{eq:cov}) of the mock observations. Overall, the observables 
drawn from the ABC-PMC posteriors are in good agreement with $\xigg$ and $\gmf$ of the 
mock observations. The lower panels demonstrate that for both observables, the error-bars 
of the mock observations lie within the $68\%$ confidence interval of the ABC-PMC posterior 
predictions.} 
\label{fig:abc_moneyplot}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC vs MCMC Histograms 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbarxi.pdf}
\caption{We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter 
constraints from ABC-PMC (orange) to the constraints from the Gaussian Likelihood MCMC (blue)
using $\ngalaxy$ and $\xigg(r)$ as observables. The \emph{top} panels show the histograms of the 
marginalized posterior PDFs over the parameters. In the \emph{bottom} panels, we include box plots marking
the confidence intervals of the posterior distributions. The boxes represent the $68\%$ 
confidence interval while the ``whiskers'' represent the $95\%$ confidence interval. We also
plot the ``true'' HOD parameters with vertical black dashed line. Marginalized posterior PDFs 
obtained from the two methods are consistent with each other. The ABC-PMC constraints are slightly 
narrower for $\log \mathcal{M}_{\rm min}$, slightly wider for $\log \mathcal{M}_{1}$, and slightly 
less biased for $\alpha$.} 
\label{fig:hist_nbarxi}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbargmf.pdf}
\caption{Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC analysis and the standard Gaussian 
Likelihood MCMC analysis are done using the observables $\ngalaxy$ and $\gmf(N)$. The constraints are 
consistent with the ``true'' HOD parameters and both methods infer the region of allowed values to similar 
precision. The MCMC result for $\alpha$ is slightly more biased compared to ABC-PMC estimate. 
This may stem from the fact that the use of Gaussian-likelihood and its associated assumptions is more 
spurious when modeling the group multiplicity function.}
\label{fig:hist_nbargmf}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% ABC vs MCMC Contours nbar +xi
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure*}
\begin{center}
\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbarxi.pdf}
\caption{We compare the ABC-PMC (orange) and the Gaussian likelihood MCMC (blue)
predictions of the 68\% and 95\% posterior confidence regions over the HOD 
parameters ($\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$) 
using $\ngalaxy$ and $\xigg(r)$ as observables. The ``true'' HOD parameters used 
to create the mock observations are plotted by black stars in each panel. The two approaches are consistent with the true values of the parameters used to generate the data.}
\label{fig:cont_nbarxi}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbargmf.pdf}
\caption{Same as Figure \ref{fig:cont_nbarxi}, but using $\ngalaxy$ and $\gmf(N)$ as observables. Again, both methods accurately estimate the parameter value regions of the true values for the data. The MCMC estimation of $\alpha$ by use of a Gaussian-likelihood is biased when compared with the ABC-PMC contours. This may be due to the fact that the group multiplicity function is particularly unsuited to the use of a Gaussian-likelihood analysis.}
\label{fig:cont_nbargmf}
\end{center}
\end{figure*}

%%%%%%%%%%%%%%%%Acknowledgments%%%%%%%%%%%%%%%%%%%

\section*{Acknowledgments}
We thank Jessie Cisewsky for reading and making valuable comments on the draft. We would also like to thank Michael R. Blanton, Jeremy R. Tinker, Uros Seljak, Layne Price, 
Boris Leidstadt, Alex Malz, Patrick McDonald, and Dan Foreman-Mackey for productive 
and insightful discussions. MV was supported by NSF grant AST-1517237. DWH was supported 
by NSF (grants IIS-1124794 and AST-1517237), NASA (grant NNX12AI50G), and the Moore-Sloan 
Data Science Environment at NYU. KW was supported by NSF grant AST-1211889. Computations 
were performed using computational resources at NYU-HPC. We thank Shenglong Wang, the 
administrator of NYU-HPC computational facility, for his consistent and continuous support 
throughout the development of this project. We would like to thank the organizers of 
the AstroHackWeek 2015 workshop (\url{http://astrohackweek.org/2015/}), 
since the direction and the scope of this investigation was ---to some degree--- initiated 
through discussions in this workshop. Throughout this investigation, we have made use of 
publicly available software packages $\mathtt{emcee}$ and $\mathtt{abcpmc}$. We have also used the publicly available python implementation of the FoF algorithm $\mathtt{pyfof}$ (\url{https://github.com/simongibbons/pyfof}).

\bibliographystyle{yahapj}
\bibliography{ccppabc}
\end{document}

### 2017-02-24 ccppabc/text/main.tex
more updates

@@
 
-In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 
-MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using 
-ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, 
-accounts for sample variance. In principle, they also have the advantage that they can account for 
-sources of systematic uncertainties in the observational data.
+{\bf \color{red} Overall, the constraints from ABC-PMC are consistent with those from 
+the Gaussian pseudo-likelihood MCMC method; however, using ABC-PMC has a number of 
+advantages. ABC-PMC utilizes a forward generative model. Our forward generative model, 
+for instance, accounts for sample variance. In principle, they also have the advantage that they can account for sources of systematic uncertainties in the observational data. 
+}
+
+%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 
+%MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using 
+%ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, 
+%accounts for sample variance. In principle, they also have the advantage that they can account for 
+%sources of systematic uncertainties in the observational data.
 

### 2017-02-24 ccppabc/text/main.tex
some further edits on the draft
ref comment #1 see pdf

@@
 \title{Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection}
-\author{ChangHoon~Hahn\altaffilmark{\dagger1}, Mohammadjavad~Vakili\altaffilmark{\dagger1},
-        Kilian~Walsh\altaffilmark{1},
-        Andrew~P.~Hearin\altaffilmark{2},
-        David~W.~Hogg\altaffilmark{1,3,4,5},
-    and~Duncan~Campbell\altaffilmark{6}}
+\author{ChangHoon~Hahn\altaffilmark{\dagger1}, Mohammadjavad~Vakili\altaffilmark{\dagger1}, Kilian~Walsh\altaffilmark{1}, Andrew~P.~Hearin\altaffilmark{2}, David~W.~Hogg\altaffilmark{1,3,4,5}, and~Duncan~Campbell\altaffilmark{6}}
 \altaffiltext{$\dagger$}{These authors have contributed equally to the paper: \url{mjvakili@nyu.edu}, \url{chh327@nyu.edu}}
@@
 advantages. ABC-PMC utilizes a forward generative model. Our forward generative model, 
-for instance, accounts for sample variance. In principle, they also have the advantage that they can account for sources of systematic uncertainties in the observational data. 
-}
+for instance, accounts for sample variance. Forward generative models also have the 
+advantage that they can account for sources of systematic uncertainties that affect 
+observational data.}
 
@@
 
+{\bf \color{red} CHH: @MJV I'm not a big fan of this entire paragraph. The accuracy of the covariance
+matrix depends on the accuracy of the mock catalogs (a.k.a. the forward model). While it's definitely 
+true that on small scales they break down, forward models in ABC suffers from the same issues.} 
 Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix

### 2017-03-07 ccppabc/text/revision/main.tex
clean up in tex directory

--- 2017-02-24 ccppabc/text/main.tex
@@
-\documentclass[12pt, preprint]{aastex}
-\usepackage[breaklinks,colorlinks, 
-urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
-\usepackage{graphicx}	% For figures
-\usepackage{natbib}	% For citep and citep
-\usepackage{amsmath}	% for \iint
-\usepackage{mathtools}
-\usepackage{bm}
-\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
-\usepackage{hyperref}
+% mnras_template.tex 
+%
+% LaTeX template for creating an MNRAS paper
+%
+% v3.0 released 14 May 2015
+% (version numbers match those of mnras.cls)
+%
+% Copyright (C) Royal Astronomical Society 2015
+% Authors:
+% Keith T. Smith (Royal Astronomical Society)
+
+% Change log
+%
+% v3.0 May 2015
+%    Renamed to match the new package name
+%    Version number matches mnras.cls
+%    A few minor tweaks to wording
+% v1.0 September 2013
+%    Beta testing only - never publicly released
+%    First version: a simple (ish) template for creating an MNRAS paper
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Basic setup. Most papers should leave these options alone.
+\documentclass[fleqn,usenatbib]{mnras}
+
+% MNRAS is set in Times font. If you don't have this installed (most LaTeX
+% installations will be fine) or prefer the old Computer Modern fonts, comment
+% out the following line
+\usepackage{newtxtext,newtxmath}
+% Depending on your LaTeX fonts installation, you might get better results with one of these:
+%\usepackage{mathptmx}
+%\usepackage{txfonts}
+
+% Use vector fonts, so it zooms properly in on-screen viewing software
+% Don't change these lines unless you know what you are doing
+\usepackage[T1]{fontenc}
+\usepackage{ae,aecompl}
+
 \usepackage{algorithmic,algorithm}
-\hypersetup{colorlinks}
-\usepackage{color}
-\usepackage{morefloats}
-\definecolor{darkred}{rgb}{0.5,0,0}
-\definecolor{darkgreen}{rgb}{0,0.5,0}
-\definecolor{darkblue}{rgb}{0,0,0.5}
-\hypersetup{colorlinks, linkcolor=darkblue, filecolor=darkgreen, urlcolor=darkred, citecolor=darkblue}
-\DeclareMathOperator*{\argmax}{arg\,max}
+
+%%%%% AUTHORS - PLACE YOUR OWN PACKAGES HERE %%%%%
+
+% Only include extra packages if you really need them. Common packages are:
+\usepackage{graphicx}	% Including figure files
+\usepackage{amsmath}	% Advanced maths commands
+\usepackage{amssymb}	% Extra maths symbols
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
+
+% Please keep new commands to a minimum, and use \newcommand not \def to avoid
+% overwriting existing commands. Example:
+%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
 
@@
 \newcommand{\dev}{\mathrm{d}}
-
+\newcommand{\mstar}{h^{-1}M_\odot}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%% TITLE PAGE %%%%%%%%%%%%%%%%%%%
+
+% Title of the paper, and the short title which is used in the headers.
+% Keep the title short and informative.
+\title[ABC in large scale structure]{Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection}
+
+% The list of authors, and the short list which is used in the headers.
+% If you need two or more lines of authors, add an extra line using \newauthor
+\author[Hahn \& Vakili et al.]{
+\parbox{\textwidth}{
+ChangHoon~Hahn,$^{1,\ast}$\thanks{E-mail: chh327@nyu.edu}
+Mohammadjavad~Vakili,$^{1,\ast}$\thanks{E-mail: mjvakili@nyu.edu}
+Kilian~Walsh,$^{1}$\\
+Andrew~P.~Hearin,$^{2}$
+David~W.~Hogg,$^{1,3,4,5}$
+and Duncan Campbell,$^{6}$
+}
+  \vspace*{10pt} \\
+% List of institutions
+$*$ These authors have contributed equally to the paper\\ 
+$^{1}$Center for Cosmology and Particle Physics,Department of Physics, New York University, 726 Broadway, New York, NY\\
+$^{2}$Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT\\
+$^{3}$Flatiron institute, 160 Fifth Avenue, New York, NY\\
+$^{4}$Center for Data Science, New York University, 726 Broadway, New York, NY\\
+$^{5}$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany\\
+$^{6}$Department of Astronomy, Yale University, New Haven, CT
+}
+
+% These dates will be filled out by the publisher
+\date{Accepted XXX. Received YYY; in original form ZZZ}
+
+% Enter the current year, for the copyright statements etc.
+\pubyear{2015}
+
+% Don't change these lines
 \begin{document}
-\title{Approximate Bayesian Computation in Large Scale Structure: constraining the galaxy-halo connection}
-\author{ChangHoon~Hahn\altaffilmark{\dagger1}, Mohammadjavad~Vakili\altaffilmark{\dagger1}, Kilian~Walsh\altaffilmark{1}, Andrew~P.~Hearin\altaffilmark{2}, David~W.~Hogg\altaffilmark{1,3,4,5}, and~Duncan~Campbell\altaffilmark{6}}
-\altaffiltext{$\dagger$}{These authors have contributed equally to the paper: \url{mjvakili@nyu.edu}, \url{chh327@nyu.edu}}
-\altaffiltext{1}{Center for Cosmology and Particle Physics, 
-                 Department of Physics, New York University, 4 Washington Pl., New York, NY}    
-\altaffiltext{2}{Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT}
-\altaffiltext{3}{Simons Center for Data Analysis, 160 Fifth Avenue, New York, NY}
-\altaffiltext{4}{Center for Data Science, New York University, 726 Broadway, New York, NY}
-\altaffiltext{5}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany}
-
-\altaffiltext{6}{Department of Astronomy, Yale University, New Haven, CT}
-
+\label{firstpage}
+\pagerange{\pageref{firstpage}--\pageref{lastpage}}
+\maketitle
+
+% Abstract of the paper
 \begin{abstract}
@@
 
+% Select between one and six entries from the list of approved keywords.
+% Don't make up new ones.
+\begin{keywords}
+Large scale structure -- cosmology -- statistic
+\end{keywords}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%% BODY OF PAPER %%%%%%%%%%%%%%%%%%
+
 \section{Introduction}
+
 Cosmology was revolutionized in the 1990s with the introduction of likelihoods---%
@@
 \ref{sec:abcvsmcmc}.
-
 %%%%%%%%%%%%%%%%%%%%% SUBSECTION %%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{tabular}{ccccc} \hline \hline 
-$\log M_{\rm min} \; \left[ h^{-1} M_{\odot} \right]$ & $\sigma_{\log M}$ & $\log M_{0} \; \left[ h^{-1} M_{\odot} \right]$ & $\log M_{1} \; \left[ h^{-1} M_{\odot} \right]$ & $\alpha$ \\ \hline
+$\log M_{\rm min}$ & $\sigma_{\log M}$ & $\log M_{0}$ & $\log M_{1}$ & $\alpha$ \\ \hline
 $12.79$ & $0.39$ & $11.92$ & $13.94$ & $1.15$ \\ \hline 
@@
 the observables $\ngalaxy$ and $\gmf$, is $\vec\rho = [\rho_n, \rho_\zeta]$.
-
-To avoid any complications from the choice for our prior, we select uniform priors over all parameters aside from 
-the scatter parameter $\sigma_{\log\;M}$, for which we choose a log-uniform prior. 
+To avoid any complications from the choice for our prior, we select uniform priors over all parameters aside from the scatter parameter $\sigma_{\log\;M}$, for which we choose a log-uniform prior. 
 We list the range of our prior distributions in Table ~\ref{tab:prior}.
@@
 %%%%%%%%%%%%%%%%%%TABLES%%%%%%%%%%%%
-\begin{table*}
-\begin{center}
+
+\begin{table}
+	\centering
+	\caption{{\bf Prior Specifications}: The prior probability distribution 
+  and its range for each of the \citet{zheng07} HOD parameters. All mass parameters are in unit of $\mstar$}
 	\label{tab:prior}
-  \caption{{\bf Prior Specifications}: The prior probability distribution 
-  and its range for each of the \cite{zheng07} HOD parameters. 
-  All mass parameters are in unit of $h^{-1}M_\odot$}
-\begin{tabular}{@{}lllll}
-\\ \hline 
-    HOD Parameter & & Prior & & Range \\ \hline
-  $\alpha$ & & Uniform & & [0.8, 1.3] \\
-  $\sigma_{\rm \log M}$ & & Log--Uniform & &  [0.1, 0.7] \\
-   $\log M_{0}$   & & Uniform & &  [10.0, 13.0] \\
-  $\log M_{min}$ & &   Uniform & &  [11.02, 13.02] \\
-  $\log M_{1}$ & & Uniform & & [13.0, 14.0] \\
+	\begin{tabular}{lcr} % four columns, alignment for each
+		\hline
+		HOD Parameter & Prior & Range\\
+		\hline
+		$\alpha$ & Uniform & [0.8, 1.3]\\
+		$\sigma_{\rm \log M}$ & Log-Uniform & [0.1, 0.7]\\
+		$\log M_{0}$ & Uniform & [10.0, 13.0]\\
+        $\log M_{min}$ & Uniform & [11.02, 13.02]\\
+        $\log M_{1}$ & Uniform & [13.0, 14.0]\\
 		\hline
 	\end{tabular}
-\end{center}
-\end{table*}
+\end{table}
 
@@
 the overall evolution is the same when we use $\bar{n}$ and $\xigg(r)$. The darker and lighter 
-contours represent the $68\%$ and $95\%$ confident regions of the posterior distribution over $\pars_t$. For 
-reference, we also plot the ``true'' HOD parameter $\pars_\mathrm{true}$ 
+contours represent the $68\%$ and $95\%$ confident regions of the posterior distribution over $\pars_t$. For reference, we also plot the ``true'' HOD parameter $\pars_\mathrm{true}$ 
 (black star) in each of the panels.
@@
 
-\item From our specific ABC-PMC method, we obtain parameter constraints that are consistent with the ``true'' 
-HOD parameters of our mock observations. Hence we demonstrate that ABC-PMC can be used for parameter inference
-in LSS studies.
+\item From our specific ABC-PMC method, we obtain parameter constraints that are consistent with the ``true'' HOD parameters of our mock observations.
+Hence we demonstrate that ABC-PMC can be used for parameter inference in LSS studies.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Mock observation 2PCF and GMF
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{paper_data_observables.pdf}
-\caption{The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
-function $\gmf(N)$ (right) summary statistics of the mock observations generated from 
-the ``true'' HOD parameters described in Section \ref{sec:mock_obv}. The width of the 
-shaded region corresponds to the square root of the covariance matrix diagonal 
-elements (Eq. \ref{eq:cov}). In our ABC analysis, we treat the $\xigg(r)$ 
-and $\gmf(N)$ above as the summary statistics of the observation.}
-\label{fig:mock}
-\end{center}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABC Pool Evolution plot 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\begin{center}
-\includegraphics[width=0.95\textwidth]{paper_ABC_poolevolution_nbargmf.pdf}
-\caption{We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ 
-in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
-$\bar{n}$ and $\gmf(N)$ are used as observables for the above results. For reference, 
-in each panel, we include the ``true'' HOD parameters (black star) listed in 
-Section \ref{sec:mock_obv}. The initial distance threshold, $\vec\epsilon_1 = 
-[\infty, \infty]$ at $t=1$ (top left) so the $\pars_1$ spans the entire range of the prior distribution, 
-which is also the range of the panels. We see for $t < 5$, the parameter space occupied 
-by the ABC $\pars_t$ shrinks dramatically. Eventually when the algorithm converges, 
-$t > 7$, the parameter space occupied by $\pars_t$ no longer shrinks and their 
-distributions represent the posterior distribution of the 
-parameters. At $t=9$, the final iteration, the ABC algorithm has converged and 
-we find that $\pars_\mathrm{true}$ lies safely within the $68\%$ confidence region.
-} 
-\label{fig:pool_demo}
-\end{center}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABC Pool Convergence 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{paper_ABCconvergence.pdf}
-\caption{We illustrate the convergence of the ABC algorithm through the evolution 
-of the ABC particle distribution as a function of iteration for 
-parameters $\log \mathcal{M}_\mathrm{min}$ (left), $\alpha$ (center), and 
-$\log \mathcal{M}_1$ (right).
-The top panel corresponds our ABC results using the observables $(\bar{n}, \gmf(N))$, 
-while the lower panel plots  corresponds to the ABC results using $(\bar{n}, \xigg(r))$. 
-The distributions of parameters show no significant change after $t > 7$, which suggests
-that the ABC algorithm has converged.}
-\label{fig:abc_converge}
-\end{center}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABC Corner Plots
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\begin{center}
-\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbarxi.pdf}
-\caption{We present the constraints on the \cite{zheng07} HOD model parameters
-obtained from our ABC-PMC analysis using $\bar{n}$ and $\xigg(r)$ as observables.  
-The diagonal panels plot the posterior distribution of each HOD parameter with 
-vertical dashed lines marking the $50\%$ quantile and $68\%$ confidence intervals of the 
-distribution. The off-diagonal panels plot the degeneracies between parameter pairs. 
-The range of each panel corresponds to the range of our prior choice. The ``true''
-HOD parameters, listed in Section \ref{sec:mock_obv}, are also plotted in each of 
-the panels (black). For $\log\mathcal{M}_0$, $\alpha$, and $\sigma_{\log M}$, 
-the ``true'' parameter values lie near the center of the $68\%$ confidence interval 
-of the posterior distribution. For $\log\mathcal{M}_1$ and $\log\mathcal{M}_\mathrm{min}$, 
-which have tight constraints, the ``true'' values lie within the $68\%$ confidence 
-interval. Ultimately, the ABC parameter constraints we obtain in our analysis are consistent 
-with the ``true'' HOD parameters. 
-}
-\label{fig:abc_corner_nbarxi}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbargmf.pdf}
-\caption{Same as Figure \ref{fig:abc_corner_nbarxi} but for our ABC analysis 
-using $\bar{n}$ and $\gmf(N)$ as observables. The ABC parameter constraints we obtain are
-consistent with the ``true'' HOD parameters.}
-\label{fig:abc_corner_nbargmf}
-\end{center}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABC Posterior Observable  
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\begin{center}
-\includegraphics[width=\textwidth]{paper_ABCposterior.pdf}
-\caption{We compare the ABC-PMC posterior prediction for the observables $\xigg(r)$ (left) and $\gmf(N)$ (right) (orange; Section \ref{sec:abc_results}) 
-to $\xigg(r)$ and $\gmf(N)$ of the mock observation (black) in the top panels.
-In the lower panels, we plot the ratio between the ABC-PMC posterior predictions for $\xigg$ and $\gmf$ 
-to the mock observation $\xigg^\mathrm{obvs}$ and $\gmf^\mathrm{obvs}$. The darker 
-and lighter shaded regions represent the $68\%$ and $95\%$ confidence regions of the posterior predictions, 
-respectively. The error-bars represent the square root of the diagonal elements of the error 
-covariance matrix (equation \ref{eq:cov}) of the mock observations. Overall, the observables 
-drawn from the ABC-PMC posteriors are in good agreement with $\xigg$ and $\gmf$ of the 
-mock observations. The lower panels demonstrate that for both observables, the error-bars 
-of the mock observations lie within the $68\%$ confidence interval of the ABC-PMC posterior 
-predictions.} 
-\label{fig:abc_moneyplot}
-\end{center}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABC vs MCMC Histograms 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\begin{center}
-\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbarxi.pdf}
-\caption{We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter 
-constraints from ABC-PMC (orange) to the constraints from the Gaussian Likelihood MCMC (blue)
-using $\ngalaxy$ and $\xigg(r)$ as observables. The \emph{top} panels show the histograms of the 
-marginalized posterior PDFs over the parameters. In the \emph{bottom} panels, we include box plots marking
-the confidence intervals of the posterior distributions. The boxes represent the $68\%$ 
-confidence interval while the ``whiskers'' represent the $95\%$ confidence interval. We also
-plot the ``true'' HOD parameters with vertical black dashed line. Marginalized posterior PDFs 
-obtained from the two methods are consistent with each other. The ABC-PMC constraints are slightly 
-narrower for $\log \mathcal{M}_{\rm min}$, slightly wider for $\log \mathcal{M}_{1}$, and slightly 
-less biased for $\alpha$.} 
-\label{fig:hist_nbarxi}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbargmf.pdf}
-\caption{Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC analysis and the standard Gaussian 
-Likelihood MCMC analysis are done using the observables $\ngalaxy$ and $\gmf(N)$. The constraints are 
-consistent with the ``true'' HOD parameters and both methods infer the region of allowed values to similar 
-precision. The MCMC result for $\alpha$ is slightly more biased compared to ABC-PMC estimate. 
-This may stem from the fact that the use of Gaussian-likelihood and its associated assumptions is more 
-spurious when modeling the group multiplicity function.}
-\label{fig:hist_nbargmf}
-\end{center}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABC vs MCMC Contours nbar +xi
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-\begin{center}
-\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbarxi.pdf}
-\caption{We compare the ABC-PMC (orange) and the Gaussian likelihood MCMC (blue)
-predictions of the 68\% and 95\% posterior confidence regions over the HOD 
-parameters ($\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$) 
-using $\ngalaxy$ and $\xigg(r)$ as observables. The ``true'' HOD parameters used 
-to create the mock observations are plotted by black stars in each panel. The two approaches are consistent with the true values of the parameters used to generate the data.}
-\label{fig:cont_nbarxi}
-\end{center}
-\end{figure*}
-
-\begin{figure*}
-\begin{center}
-\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbargmf.pdf}
-\caption{Same as Figure \ref{fig:cont_nbarxi}, but using $\ngalaxy$ and $\gmf(N)$ as observables.
-Again, both methods accurately estimate the parameter value regions of the true values for the data.
-The MCMC estimation of $\alpha$ by use of a Gaussian-likelihood is biased when compared with the ABC-PMC contours.
-This may be due to the fact that the group multiplicity function is particularly unsuited to the use of a Gaussian-likelihood analysis.}
-\label{fig:cont_nbargmf}
-\end{center}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%Acknowledgments%%%%%%%%%%%%%%%%%%%
-
-\section*{Acknowledgments}
+\section*{Acknowledgements}
+
 We thank Jessie Cisewsky for reading and making valuable comments on the draft. We would also like to thank Michael R. Blanton, Jeremy R. Tinker, Uros Seljak, Layne Price, 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%% FIGURES %%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Mock observation 2PCF and GMF
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=\textwidth]{paper_data_observables.pdf}
+\caption{\label{fig:mock} The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
+function $\gmf(N)$ (right) summary statistics of the mock observations generated from 
+the ``true'' HOD parameters described in Section \ref{sec:mock_obv}. The width of the 
+shaded region corresponds to the square root of the covariance matrix diagonal 
+elements (Eq. \ref{eq:cov}). In our ABC analysis, we treat the $\xigg(r)$ 
+and $\gmf(N)$ above as the summary statistics of the observation.}
+%\end{center}
+\end{figure*}
+%%%%%%%%%%%%%%%%%% End of FIGURE %%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABC Pool Evolution plot 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=0.95\textwidth]{paper_ABC_poolevolution_nbargmf.pdf}
+\caption{\label{fig:pool_demo} We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
+$\bar{n}$ and $\gmf(N)$ are used as observables for the above results. For reference, 
+in each panel, we include the ``true'' HOD parameters (black star) listed in 
+Section \ref{sec:mock_obv}. The initial distance threshold, $\vec\epsilon_1 = 
+[\infty, \infty]$ at $t=1$ (top left) so the $\pars_1$ spans the entire range of the prior distribution, 
+which is also the range of the panels. We see for $t < 5$, the parameter space occupied 
+by the ABC $\pars_t$ shrinks dramatically. Eventually when the algorithm converges, 
+$t > 7$, the parameter space occupied by $\pars_t$ no longer shrinks and their 
+distributions represent the posterior distribution of the 
+parameters. At $t=9$, the final iteration, the ABC algorithm has converged and 
+we find that $\pars_\mathrm{true}$ lies safely within the $68\%$ confidence region.} 
+%\end{center}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABC Pool Convergence 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=\textwidth]{paper_ABCconvergence.pdf}
+\caption{\label{fig:abc_converge} We illustrate the convergence of the ABC algorithm through the evolution of the ABC particle distribution as a function of iteration for 
+parameters $\log \mathcal{M}_\mathrm{min}$ (left), $\alpha$ (center), and 
+$\log \mathcal{M}_1$ (right).
+The top panel corresponds our ABC results using the observables $(\bar{n}, \gmf(N))$, 
+while the lower panel plots  corresponds to the ABC results using $(\bar{n}, \xigg(r))$. 
+The distributions of parameters show no significant change after $t > 7$, which suggests
+that the ABC algorithm has converged.}
+%\end{center}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABC Corner Plots
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbarxi.pdf}
+\caption{\label{fig:abc_corner_nbarxi} We present the constraints on the \citet{zheng07} HOD model parameters obtained from our ABC-PMC analysis using $\bar{n}$ and $\xigg(r)$ as observables.  
+The diagonal panels plot the posterior distribution of each HOD parameter with 
+vertical dashed lines marking the $50\%$ quantile and $68\%$ confidence intervals of the 
+distribution. The off-diagonal panels plot the degeneracies between parameter pairs. 
+The range of each panel corresponds to the range of our prior choice. The ``true''
+HOD parameters, listed in Section \ref{sec:mock_obv}, are also plotted in each of 
+the panels (black). For $\log\mathcal{M}_0$, $\alpha$, and $\sigma_{\log M}$, 
+the ``true'' parameter values lie near the center of the $68\%$ confidence interval 
+of the posterior distribution. For $\log\mathcal{M}_1$ and $\log\mathcal{M}_\mathrm{min}$, 
+which have tight constraints, the ``true'' values lie within the $68\%$ confidence 
+interval. Ultimately, the ABC parameter constraints we obtain in our analysis are consistent with the ``true'' HOD parameters.}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbargmf.pdf}
+\caption{\label{fig:abc_corner_nbargmf}Same as Figure \ref{fig:abc_corner_nbarxi} but for our ABC analysis using $\bar{n}$ and $\gmf(N)$ as observables. The ABC parameter constraints we obtain are
+consistent with the ``true'' HOD parameters.}
+%\end{center}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABC Posterior Observable  
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=\textwidth]{paper_ABCposterior.pdf}
+\caption{\label{fig:abc_moneyplot} We compare the ABC-PMC posterior prediction for the observables $\xigg(r)$ (left) and $\gmf(N)$ (right) (orange; Section \ref{sec:abc_results}) 
+to $\xigg(r)$ and $\gmf(N)$ of the mock observation (black) in the top panels.
+In the lower panels, we plot the ratio between the ABC-PMC posterior predictions for $\xigg$ and $\gmf$ 
+to the mock observation $\xigg^\mathrm{obvs}$ and $\gmf^\mathrm{obvs}$. The darker 
+and lighter shaded regions represent the $68\%$ and $95\%$ confidence regions of the posterior predictions, 
+respectively. The error-bars represent the square root of the diagonal elements of the error 
+covariance matrix (equation \ref{eq:cov}) of the mock observations. Overall, the observables 
+drawn from the ABC-PMC posteriors are in good agreement with $\xigg$ and $\gmf$ of the 
+mock observations. The lower panels demonstrate that for both observables, the error-bars 
+of the mock observations lie within the $68\%$ confidence interval of the ABC-PMC posterior 
+predictions.} 
+%\end{center}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABC vs MCMC Histograms 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbarxi.pdf}
+\caption{\label{fig:hist_nbarxi} We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter constraints from ABC-PMC (orange) to the constraints from the Gaussian Likelihood MCMC (blue) using $\ngalaxy$ and $\xigg(r)$ as observables.
+The \emph{top} panels show the histograms of the marginalized posterior PDFs over the parameters.
+In the \emph{bottom} panels, we include box plots marking the confidence intervals of the posterior distributions.
+The boxes represent the $68\%$ confidence interval while the ``whiskers'' represent the $95\%$ confidence interval.
+We also
+plot the ``true'' HOD parameters with vertical black dashed line. Marginalized posterior PDFs 
+obtained from the two methods are consistent with each other. The ABC-PMC constraints are slightly 
+narrower for $\log \mathcal{M}_{\rm min}$, slightly wider for $\log \mathcal{M}_{1}$, and slightly 
+less biased for $\alpha$.} 
+%\end{center}
+\end{figure*}
+
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbargmf.pdf}
+\caption{\label{fig:hist_nbargmf} Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC analysis and the standard Gaussian 
+Likelihood MCMC analysis are done using the observables $\ngalaxy$ and $\gmf(N)$. The constraints are 
+consistent with the ``true'' HOD parameters and both methods infer the region of allowed values to similar 
+precision. The MCMC result for $\alpha$ is slightly more biased compared to ABC-PMC estimate. 
+This may stem from the fact that the use of Gaussian-likelihood and its associated assumptions is more 
+spurious when modeling the group multiplicity function.}
+%\end{center}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% ABC vs MCMC Contours nbar +xi
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbarxi.pdf}
+\caption{\label{fig:cont_nbarxi}
+ We compare the ABC-PMC (orange) and the Gaussian likelihood MCMC (blue)
+predictions of the 68\% and 95\% posterior confidence regions over the HOD 
+parameters ($\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$) 
+using $\ngalaxy$ and $\xigg(r)$ as observables. The ``true'' HOD parameters used 
+to create the mock observations are plotted by black stars in each panel. The two approaches are consistent with the true values of the parameters used to generate the data.}
+%\end{center}
+\end{figure*}
+
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbargmf.pdf}
+\caption{ \label{fig:cont_nbargmf} Same as Figure \ref{fig:cont_nbarxi}, but using $\ngalaxy$ and $\gmf(N)$ as observables.
+Again, both methods accurately estimate the parameter value regions of the true values for the data.
+The MCMC estimation of $\alpha$ by use of a Gaussian-likelihood is biased when compared with the ABC-PMC contours.
+This may be due to the fact that the group multiplicity function is particularly unsuited to the use of a Gaussian-likelihood analysis.}
+%\end{center}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
+
 \bibliographystyle{yahapj}
 \bibliography{ccppabc}
+%\clearpage
+% The best way to enter references is to use BibTeX:
+
+%\bibliographystyle{mnras}
+%\bibliography{example} % if your bibtex file is called example.bib
+
+% Alternatively you could enter them by hand, like this:
+% This method is tedious and prone to error if you have lots of references
+%\begin{thebibliography}{99}
+%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
+%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
+%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
+%Others S., 2012, Journal of Interesting Stuff, 17, 198
+%\end{thebibliography}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
+
+%\appendix
+
+%\section{Some extra material}
+
+%If you want to present additional material which would interrupt the flow of the main paper,
+%it can be placed in an Appendix which appears after the list of references.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% Don't change these lines
+%\bsp	% typesetting comment
+\label{lastpage}
 \end{document}
+
+% End of mnras_template.tex

### 2017-03-07 ccppabc/text/revision/main.tex
some edits

@@
-% mnras_template.tex 
-%
-% LaTeX template for creating an MNRAS paper
-%
-% v3.0 released 14 May 2015
-% (version numbers match those of mnras.cls)
-%
-% Copyright (C) Royal Astronomical Society 2015
-% Authors:
-% Keith T. Smith (Royal Astronomical Society)
-
-% Change log
-%
-% v3.0 May 2015
-%    Renamed to match the new package name
-%    Version number matches mnras.cls
-%    A few minor tweaks to wording
-% v1.0 September 2013
-%    Beta testing only - never publicly released
-%    First version: a simple (ish) template for creating an MNRAS paper
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Basic setup. Most papers should leave these options alone.
 \documentclass[fleqn,usenatbib]{mnras}
-
 % MNRAS is set in Times font. If you don't have this installed (most LaTeX
@@
 %%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
-
-% Please keep new commands to a minimum, and use \newcommand not \def to avoid
-% overwriting existing commands. Example:
-%\newcommand{\pcm}{\,cm$^{-2}$}	% per cm-squared
 
@@
 \newcommand{\mstar}{h^{-1}M_\odot}
+
+\definecolor{dred}{rgb}{0.75,0,0}
+\definecolor{darkred}{rgb}{0.5,0,0}
+\definecolor{darkgreen}{rgb}{0,0.5,0}
 
@@
 
-{\bf \color{red} Overall, the constraints from ABC-PMC are consistent with those from 
+%%%
+{\bf \color{dred} 
+Overall, the constraints from ABC-PMC are consistent with those from 
 the Gaussian pseudo-likelihood MCMC method; however, using ABC-PMC has a number of 
@@
 advantage that they can account for sources of systematic uncertainties that affect 
-observational data.}
+observational data.
+}
+%%%
 
@@
 
-{\bf \color{red} CHH: @MJV I'm not a big fan of this entire paragraph. The accuracy of the covariance
+%%%
+{\bf \color{darkgreen} 
+CHH: @MJV I'm not a big fan of this entire paragraph. The accuracy of the covariance
 matrix depends on the accuracy of the mock catalogs (a.k.a. the forward model). While it's definitely 
-true that on small scales they break down, forward models in ABC suffers from the same issues.} 
+true that on small scales they break down, forward models in ABC suffers from the same issues.
+} 
+%%%
 Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix
@@
 %\begin{center}
-\includegraphics[width=\textwidth]{paper_data_observables.pdf}
+\includegraphics[width=\textwidth]{figs/paper_data_observables.pdf}
 \caption{\label{fig:mock} The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
@@
 %\begin{center}
-\includegraphics[width=0.95\textwidth]{paper_ABC_poolevolution_nbargmf.pdf}
+\includegraphics[width=0.95\textwidth]{figs/paper_ABC_poolevolution_nbargmf.pdf}
 \caption{\label{fig:pool_demo} We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
@@
 %\begin{center}
-\includegraphics[width=\textwidth]{paper_ABCconvergence.pdf}
+\includegraphics[width=\textwidth]{figs/paper_ABCconvergence.pdf}
 \caption{\label{fig:abc_converge} We illustrate the convergence of the ABC algorithm through the evolution of the ABC particle distribution as a function of iteration for 
@@
 %\begin{center}
-\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbarxi.pdf}
+\includegraphics[width=0.85\textwidth]{figs/paper_ABCcorner_nbarxi.pdf}
 \caption{\label{fig:abc_corner_nbarxi} We present the constraints on the \citet{zheng07} HOD model parameters obtained from our ABC-PMC analysis using $\bar{n}$ and $\xigg(r)$ as observables.  
@@
 %\begin{center}
-\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbargmf.pdf}
+\includegraphics[width=0.85\textwidth]{figs/paper_ABCcorner_nbargmf.pdf}
 \caption{\label{fig:abc_corner_nbargmf}Same as Figure \ref{fig:abc_corner_nbarxi} but for our ABC analysis using $\bar{n}$ and $\gmf(N)$ as observables. The ABC parameter constraints we obtain are
@@
 %\begin{center}
-\includegraphics[width=\textwidth]{paper_ABCposterior.pdf}
+\includegraphics[width=\textwidth]{figs/paper_ABCposterior.pdf}
 \caption{\label{fig:abc_moneyplot} We compare the ABC-PMC posterior prediction for the observables $\xigg(r)$ (left) and $\gmf(N)$ (right) (orange; Section \ref{sec:abc_results}) 
@@
 %\begin{center}
-\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbarxi.pdf}
+\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbarxi.pdf}
 \caption{\label{fig:hist_nbarxi} We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter constraints from ABC-PMC (orange) to the constraints from the Gaussian Likelihood MCMC (blue) using $\ngalaxy$ and $\xigg(r)$ as observables.
@@
 %\begin{center}
-\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbargmf.pdf}
+\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbargmf.pdf}
 \caption{\label{fig:hist_nbargmf} Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC analysis and the standard Gaussian 
@@
 %\begin{center}
-\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbarxi.pdf}
+\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_contour_nbarxi.pdf}
 \caption{\label{fig:cont_nbarxi}
@@
 %\begin{center}
-\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbargmf.pdf}
+\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_contour_nbargmf.pdf}
 \caption{ \label{fig:cont_nbargmf} Same as Figure \ref{fig:cont_nbarxi}, but using $\ngalaxy$ and $\gmf(N)$ as observables.

### 2017-03-08 ccppabc/text/revision/main.tex
directory clean up; significant edits to Section 3.4 to generally address Ref Report #1; see green text for issues requiring attention

@@
 \usepackage{amssymb}	% Extra maths symbols
+\usepackage{xcolor}	% Extra maths symbols
 
@@
 %%%%% AUTHORS - PLACE YOUR OWN COMMANDS HERE %%%%%
+
+\definecolor{dred}{rgb}{0.75,0.,0.}
+\definecolor{darkred}{rgb}{0.5,0.,0.}
+\definecolor{darkgreen}{rgb}{0.,0.5,0.}
 
@@
 \newcommand{\mstar}{h^{-1}M_\odot}
-
-\definecolor{dred}{rgb}{0.75,0,0}
-\definecolor{darkred}{rgb}{0.5,0,0}
-\definecolor{darkgreen}{rgb}{0,0.5,0}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison to the Gaussian Likelihood MCMC Analysis} \label{sec:abcvsmcmc}
+\subsection{Comparison to the Gaussian {\color{dred} Pseudo - Likelihood} MCMC Analysis} \label{sec:abcvsmcmc}
 In order to assess the quality of the parameter inference described in the previous section, 
-we compare the ABC-PMC results with the HOD parameter constraints assuming a Gaussian likelihood 
+we compare the ABC-PMC results with the HOD parameter constraints 
+{\bf \color{dred} from}
+assuming a Gaussian likelihood 
 function. The model used for the Gaussian likelihood analysis is different than the forward 
 generative model adopted for the ABC-PMC algorithm, to be consistent with the standard approach.
+
 In the ABC analysis, the model accounts for sample variance by randomly sampling a subvolume to be 
-populated with galaxies. The Gaussian likelihood analysis assumes a covariance matrix that captures 
-the uncertainties from the sample variance, and therefore in the model, we populate the halos of
-the entire $\mathtt{MultiDark}$ simulation rather than a subvolume.
-
-We introduce the observable $\mathbf{x}$ as a combination of the summary statistics of the 
-galaxy mock catalog used in the inference. When we use $\ngalaxy$ and $\xigg(r)$ as observables, 
-$\mathbf{x} = [\ngalaxy , \xigg]$, while $\mathbf{x} = [\ngalaxy , \gmf]$ when we use $\ngalaxy$ and $\gmf(N)$
-Based on this notation, we write likelihood function as 
+populated with galaxies. 
+{\bf \color{dred} %%%
+Instead, in the Gaussian pseudo-likelihood analysis, the covariance matrix is assumed to capture the
+uncertainties from sample variance. Hence, in the model for the Gaussian pseudo-likelihood analysis, 
+we populate halos of the {\em entire} $\mathtt{MultiDark}$ simulation rather than a subvolume.
+We describe the Gaussian pseudo-likelihood analysis below.
+}
+%The Gaussian likelihood analysis assumes a covariance matrix that captures the uncertainties from the sample variance, and therefore in the model, we populate the halos of the entire $\mathtt{MultiDark}$ simulation rather than a subvolume.
+
+{\bf \color{dred}
+To write down the Gaussian pseudo-likelihood, we first introduce the vector $\mathbf{x}$: 
+a combination of the summary statistics (observables) for a galaxy catalog. 
+When we use $\ngalaxy$ and $\xigg(r)$ as observables in the analysis: $\mathbf{x} = [\ngalaxy , \xigg]$;
+when we use $\ngalaxy$ and $\gmf(N)$ as observables in the analysis: $\mathbf{x} = [\ngalaxy , \gmf]$.
+Based on this notation, we can write pseudo-likelihood function as 
 \begin{eqnarray}
--2 \ln \mathcal{L}(\theta | d) &=& \Delta \mathbf{x}^{T}\widehat{C^{-1}}\Delta \mathbf{x} + \ln\Big[(2\pi)^{d}\mathrm{det}(C)\Big], \label{loglike}\\
+-2 \ln \mathcal{L}(\theta | d) &=& \Delta \mathbf{x}^{T}\widehat{C^{-1}}\Delta \mathbf{x} + \ln\Big[(2\pi)^{d}\mathrm{det}(C)\Big], \label{loglike}
+\end{eqnarray}
+where 
+\begin{eqnarray}
 \Delta \mathbf{x} &=& [\mathbf{x}_{obs} -\mathbf{x}_{mod}], 
 \end{eqnarray}
-where $\Delta \mathbf{x}$ is the difference between the measured $\mathbf{x}_{obs}$ of the mock observations 
-and that of the model $\mathbf{x}_{mod}(\mathbb{\theta})$. $d$ is the dimension of the observable $\mathbf{x}$. 
-When we use $\mathbf{x} = [\ngalaxy, \xigg]$, $d = 13$. When we use $\mathbf{x} = [\ngalaxy, \gmf]$, $d = 10$.  
-$\widehat{C^{-1}}$ is the estimate of the inverse covariance matrix, which we calculate 
-following  \cite{hartlap2007}:
-\beq
-\widehat{C^{-1}} = \frac{N_{\rm mocks}-d - 1}{N_{\rm mocks} -1}\widehat{C}^{-1},
-\eeq
-where $\widehat{C}$ is the estimated covariance matrix and $N_{\rm mock}$ is the number of mocks used to 
-estimate $\widehat{C}$. $N_{\rm mock} = 124$ (Section \ref{sec:mock_obv}). Depending on $\mathbf{x}$, 
-$\widehat{C}$ is given by the appropriate block of the covariance matrix Eq. \ref{eq:cov} corresponding 
-to the observables. We note that in the covariance matrix the dependence on the HOD parameters is neglected. 
-Therefore, the second term in the expression of the log-likelihood (Eq. \ref{loglike}) can be neglected. 
-We sample from the posterior distribution of the likelihood given the prior distribution using the 
-MCMC sampler $\mathtt{EMCEE}$ (\citealt{emcee}).
-
-In Figure \ref{fig:hist_nbarxi} we compare the ABC-PMC and Gaussian likelihood MCMC 
-results for the marginalized posterior PDFs over three parameters of the HOD model $\{\log \mathcal{M}_{\rm min},
-\alpha , \log\mathcal{M}_1\}$ using $\ngalaxy$ and $\xigg$ as observables. 
-Figure \ref{fig:hist_nbargmf} makes the same comparison when $\ngalaxy$ and $\gmf$ are used in the inference.
-
-Figure \ref{fig:hist_nbarxi} demonstrates in the top panel that constraints obtained from the two 
-methods are consistent with the ``true'' HOD parameters represented by the vertical dashed lines. 
-The marginalized posteriors over the parameters follow each other very closely. The lower panel of 
-Figure \ref{fig:hist_nbarxi} plot the $68\%$ and $95\%$ confidence intervals of the constraints 
-derived from the two methods as a box plot. The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ 
-is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ 
-are slightly larger for our ABC-PMC results.
-
-In Figure \ref{fig:hist_nbargmf}, when we use $\ngalaxy$ and $\gmf$ as observables, both methods 
-produce consistent constraints with each other and ``true'' HOD parameters. The lower panel of Figure 
-\ref{fig:hist_nbargmf} shows that the uncertainties from the two methods are comparable. The 
-ABC-PMC constraints over $\alpha$ is slightly less biased and slightly less precise.
+the difference between $\mathbf{x}_{obs}$, measured from the mock observation, 
+and $\mathbf{x}_{mod}(\mathbb{\theta})$ measured from the mock catalog generated 
+from the model with parameters $\theta$ .
+$d$ here is the dimension of $\mathbf{x}$ (for $\mathbf{x} = [\ngalaxy, \xigg]$, $d = 13$; 
+for $\mathbf{x} = [\ngalaxy, \gmf]$, $d = 10$).  
+$\widehat{C^{-1}}$ is the inverse covariance matrix, which we estimate following \cite{hartlap2007}:
+\begin{eqnarray}
+\widehat{C^{-1}} = \frac{N_{\rm mocks}-d - 1}{N_{\rm mocks} -1} \; \widehat{C}^{-1}.
+\end{eqnarray}
+$\widehat{C}$ is the estimated covariance matrix, calculated using the corresponding 
+$\mathbf{x}$ block of the covariance matrix from Eq.~\ref{eq:cov}, and $N_{\rm mock}$ 
+is the number of mocks used for the estimation ($N_{\rm mock} = 124$; see Section \ref{sec:mock_obv}).
+We note that in $\widehat{C}$ the dependence on the HOD parameters is neglected, 
+so the second term in the expression of Eq.~\ref{loglike} can be neglected. 
+Finally, using this pseudo-likelihood, we sample from the posterior distribution 
+given the prior distribution using the MCMC sampler $\mathtt{emcee}$ (\citealt{emcee}). 
+}
+
+{\bf \color{dred}
+In Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, we compare the results from
+ABC-PMC and Gaussian pseudo-likelihood MCMC analyses using $[\ngalaxy, \xigg]$ and 
+$[\ngalaxy, \gmf]$ as observables respectively. The top panels in each figure 
+compares the marginalized posterior PDFs for three parameters of the HOD model: 
+$\{\log \mathcal{M}_{\rm min}, \alpha , \log\mathcal{M}_1\}$. The lower panels in each 
+figure compares the $68\%$ and $95\%$ confidence intervals of the constraints 
+derived from the two inference methods as a box plot. The ``true'' HOD parameters
+are marked by vertical dashed lines in each panel. 
+}
+
+{\bf \color{dred}
+In Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, for each of the parameters, 
+the marginalized posteriors from both inference methods are comparable. Similarly, 
+the lower panels of the figures reveal that the constraints derived for each parameter
+from the two methods are also consistent, with minor discrepancies. Furthermore, we 
+find that both methods produce constraints consistent with the ``true'' HOD parameters.
+}
+{\bf \color{darkgreen}
+CHH: A few sentences discussing the width of the distributions for each of the panels
+have been commented out. Not sure if it's meaningful to include such slight discrepancies. 
+@MJV: Could you take a look at what's written above and see if you feel that it's sufficient?  
+}
+%The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ are slightly larger for our ABC-PMC results.
+In Figure \ref{fig:hist_nbargmf}, when we use $\ngalaxy$ and $\gmf$ as observables, both methods produce consistent constraints with each other and ``true'' HOD parameters.
+The lower panel of Figure \ref{fig:hist_nbargmf} shows that the uncertainties from the two methods are comparable.
+The ABC-PMC constraints over $\alpha$ is slightly less biased and slightly less precise.
 
@@
 \includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbarxi.pdf}
-\caption{\label{fig:hist_nbarxi} We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter constraints from ABC-PMC (orange) to the constraints from the Gaussian Likelihood MCMC (blue) using $\ngalaxy$ and $\xigg(r)$ as observables.
-The \emph{top} panels show the histograms of the marginalized posterior PDFs over the parameters.
-In the \emph{bottom} panels, we include box plots marking the confidence intervals of the posterior distributions.
-The boxes represent the $68\%$ confidence interval while the ``whiskers'' represent the $95\%$ confidence interval.
-We also
-plot the ``true'' HOD parameters with vertical black dashed line. Marginalized posterior PDFs 
+\caption{\label{fig:hist_nbarxi} 
+{\bf \color{dred}
+We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter 
+constraints from ABC-PMC (orange) to constraints from the Gaussian pseudo-ikelihood MCMC (blue) 
+using $\ngalaxy$ and $\xigg(r)$ as observables. The \emph{top} panels compares the two methods' 
+marginalized posterior PDFs over the parameters. In the \emph{bottom} panels, we include box 
+plots marking the confidence intervals of the posterior distributions. The boxes represent the 
+$68\%$ confidence interval while the ``whiskers'' represent the $95\%$ confidence interval. We mark 
+the ``true'' HOD parameters with vertical black dashed line. The marginalized posterior PDFs 
+}
 obtained from the two methods are consistent with each other. The ABC-PMC constraints are slightly 

### 2017-03-08 ccppabc/text/response.md

Dear Editor,

This was a very constructive report and we have made revision to the paper
accordingly. We think the comments have led to improvement in quality of the paper.

ChangHooh Hahn & Mohammadjavad Vakili (on behalf of the authors):

-----
We are grateful to the constructive and helpful comments of the 
anonymous referee on our manuscript. They have led to improvements in the paper.  
In what follows, the referee comments are indented and our responses are not indented:
-----

> a deeper discussion on why the ABC approach should be preferred in this particular context 
(one or two phrases on this should be included in the abstract).

>in section 3.4 it is said: "Furthermore, the Gaussian-likelihood approach relies on constructing an accurate covariance
matrix estimate that captures the sample variance of the data. While we are able to
do this accurately within the scope of the HOD framework, for more general LSS parameter
inference situations, it is both labor and computationally expensive and dependent on the
accuracy of simulated mock catalogs, which are known to be unreliable on small scales (see
Heitmann et al. 2008; Chuang et al. 2015 and references therein). Since ABC-PMC utilizes
a forward model to account for sample variance, it does not depend on a covariance matrix
estimate; hence it does not face these problems."
this might lead to confusion. Do the authors believe that ABC presents no advantage  
in the specific case studied in this paper?

> if the Gaussian hypothesis underlying the standard approach are so unrealistic 
as stated in the text, why results are still consistent with the ABC ones? 
how likely it is that we will face a real data scenario which might lead to significantly different results?

> is it possible to simulate an extreme data situation where the results from 
the standard analysis are not consistent with the ABC ones? if so,  how realistic it is?

> what caveats one would face in trying to  apply this kind of analysis in real data? 
How can they affect the final results and how can they be circumvented?

### 2017-03-08 ccppabc/text/revision/main.tex
some text to address issue 3 (referee's extreme data comment)
section 3.4

@@
 }
+
+{\bf \color{dred}
+It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC.
+Let us consider a dataset---from a galaxy survey---consisting of objects with different data qualities, signal-to-noise ratios, and systematic effects.
+This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift.
+Another example could be a catalog with a singal-to-noise ratio cut.
+}
+
+{\bf \color{dred}
+In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest.
+Hierarchical models can modify the likelihood functions through by taking into account the selection effects.
+Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the data that incorporates signal-to-noise ratio cuts, incompleteness, etc.
+}{\bf \color{darkgreen}
+MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ?
+}
 %%%
@@
 %%%
+{\bf \color{darkgreen}
+MJV: @CHH I think this paragraph is misleading since both ABC and Gaussian likelihood covariances suffer from inaccuracy of the simulations.
+I say we remove this paragraph entirely.
+It does not add any clarify anything.
+It just causes confusion!
+}
 {\bf \color{darkgreen} 

### 2017-03-08 ccppabc/text/revision/main.tex
Further edits to Section 3.4 to address Ref Report #1

--- 2017-03-08 ccppabc/text/revision/main.tex
@@
 
-In Figures \ref{fig:cont_nbarxi} and \ref{fig:cont_nbargmf}, we plot the contours enclosing the $68\%$ 
-and $95\%$ confidence regions of the posterior probabilities of the two methods. Figure \ref{fig:cont_nbarxi}
-uses $\ngalaxy$ and $\xigg$ as observables, while Figure \ref{fig:cont_nbargmf} uses $\ngalaxy$ and $\gmf$ 
-as observables. 
-In both figures, the ``true'' HOD parameters are plotted as black stars. The overall shape of the contours 
-are generally in good agreement with each other. However, the contours for the ABC-PMC method are
-more elongated along $\alpha$.
-
-%%%
 {\bf \color{dred}
-Overall, the constraints from ABC-PMC are consistent with those from 
+In Figures \ref{fig:cont_nbarxi} and \ref{fig:cont_nbargmf}, we plot the contours 
+enclosing the $68\%$ and $95\%$ confidence regions of the posterior probabilities of 
+the two methods using $[\ngalaxy, \xigg]$ and $[\ngalaxy, \gmf]$ as observables 
+respectively. In both figures, we mark the ``true'' HOD parameters (black star). The 
+overall shape of the contours are in agreement with each other. However, we note
+that the contours for the ABC-PMC method are more extended along $\alpha$. 
+}
+
+%%% RefReport #1 
+%%% CHH (3/8/2017): Made significant changes to the text in order to better emphasize
+%%%      the advantages of ABC over MCMC
+{\bf \color{dred} 
+Overall, the HOD parameter constraints from ABC-PMC are consistent with those from 
 the Gaussian pseudo-likelihood MCMC method; however, using ABC-PMC has a number of 
-advantages. ABC-PMC utilizes a forward generative model. Our forward generative model, 
-for instance, accounts for sample variance. Forward generative models also have the 
+advantages. For instance, ABC-PMC utilizes a forward generative model. Our forward 
+generative model accounts for sample variance. On the other hand, the Gaussian 
+pseudo-likelihood approach, as mentioned earlier this section, does not account 
+for sample variance in the model and relies on the covariance matrix estimate to 
+capture the sample variance of the data. 
+}
+
+{\bf \color{dred} 
+Accurate estimation of the covariance matrix 
+in LSS, however, faces a number of challenges. It is both labor and computationally 
+expensive and dependent on the accuracy of simulated mock catalogs, known to be 
+unreliable on small scales~(see \citealt{cosmiccode,nifty} and references therein). 
+In fact, as \cite{Sellentin:2016a} points out, using estimates of the covariance 
+matrix in the Gaussian psuedo-likelihood approach become further problematic. Even 
+when inferring parameters from a Gaussian-distributed data set, using covariance 
+matrix estimates rather than the {\em true} covariance matrix leads to a likelihood 
+function that is {\em no longer} Gaussian.
+ABC-PMC does not depend on a covariance matrix estimate; hence, it does not face 
+these problems. In principle, forward models of the ABC-PMC method also have the 
 advantage that they can account for sources of systematic uncertainties that affect 
-observational data.
+observations.
 }
-%%%
-
-%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 
-%MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using 
-%ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, 
-%accounts for sample variance. In principle, they also have the advantage that they can account for 
-%sources of systematic uncertainties in the observational data.
-
-%%%
+%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision.
+However, we emphasize some advantages of using ABC-PMC.
+ABC-PMC utilizes a forward generative model.
+Our forward generative model, for instance, accounts for sample variance.
+In principle, they also have the advantage that they can account for sources of systematic uncertainties in the observational data.
+%Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix estimate that captures the sample variance of the data.
+While we are able to do this accurately within the scope of the HOD framework, for more general LSS parameter inference situations, it is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs, which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein).
+Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a covariance matrix estimate; hence it does not face these problems.
+
+Furthermore, {\em ABC-PMC -- unlike the Gaussian pseudo-likelihood approach -- 
+is agnostic about the functional form of the underlying distribution of the 
+summary statistics} 
+($\xigg$ and $\gmf$). 
 {\bf \color{darkgreen}
-CHH: @MJV I'm not a big fan of this entire paragraph. The accuracy of the covariance
-matrix depends on the accuracy of the mock catalogs (a.k.a. the forward model). While it's definitely 
-true that on small scales they break down, forward models in ABC suffers from the same issues.
+CHH: Need a sentence here that reiterates that the Gaussian functional form 
+assumption is incorrect and especially incorrect for the GMF observable case.
+Also, for us to say the following sentence, we need to talk about it more explicitly 
+when discussing Figures 7,8,9, and 10. 
 }
-%%%
-Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix
-estimate that captures the sample variance of the data. While we are able to do this accurately 
-within the scope of the HOD framework, for more general LSS parameter inference situations, it 
-is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs,
-which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein). 
-Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a 
-covariance matrix estimate; hence it does not face these problems.
-
-{\em ABC-PMC -- unlike the Gaussian-likelihood approach -- is agnostic about the functional form of the 
-underlying distribution of the summary 
-statistics} ($\xigg$ and $\gmf$). This may explain why we find less biased constraints from the
+This may explain why we find less biased constraints from the
 ABC-PMC analysis compared to the Gaussian-likelihood analysis, when using $\gmf(N)$ as an observable 

### 2017-03-08 ccppabc/text/revision/main.tex
struggling with merge conflict still

@@
 }
-This may explain why we find less biased constraints from the
+
+{\bf \color{dred}
+It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC.
+Let us consider a dataset---from a galaxy survey---consisting of objects with different data qualities, signal-to-noise ratios, and systematic effects.
+This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift.
+Another example could be a catalog with a singal-to-noise ratio cut.
+}
+
+{\bf \color{dred}
+In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest.
+Hierarchical models can modify the likelihood functions through by taking into account the selection effects.
+Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the data that incorporates signal-to-noise ratio cuts, incompleteness, etc.
+}{\bf \color{darkgreen}
+MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ?
+}
+%%%
+
+%In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 
+%MCMC method both in terms of accuracy and precision. However, we emphasize some advantages of using 
+%ABC-PMC. ABC-PMC utilizes a forward generative model. Our forward generative model, for instance, 
+%accounts for sample variance. In principle, they also have the advantage that they can account for 
+%sources of systematic uncertainties in the observational data.
+
+%%%
+{\bf \color{darkgreen}
+MJV: @CHH I think this paragraph is misleading since both ABC and Gaussian likelihood covariances suffer from inaccuracy of the simulations.
+I say we remove this paragraph entirely.
+It does not add any clarify anything.
+It just causes confusion!
+}
+{\bf \color{darkgreen} 
+CHH: @MJV I'm not a big fan of this entire paragraph. The accuracy of the covariance
+matrix depends on the accuracy of the mock catalogs (a.k.a. the forward model). While it's definitely 
+true that on small scales they break down, forward models in ABC suffers from the same issues.
+} 
+%%%
+Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix
+estimate that captures the sample variance of the data. While we are able to do this accurately 
+within the scope of the HOD framework, for more general LSS parameter inference situations, it 
+is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs,
+which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein). 
+Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a 
+covariance matrix estimate; hence it does not face these problems.
+
+{\em ABC-PMC -- unlike the Gaussian-likelihood approach -- is agnostic about the functional form of the 
+underlying distribution of the summary 
+statistics} ($\xigg$ and $\gmf$). This may explain why we find less biased constraints from the
+>>>>>>> afd43473184a6a4637881fb3504a38a867b27226
 ABC-PMC analysis compared to the Gaussian-likelihood analysis, when using $\gmf(N)$ as an observable 

### 2017-03-08 ccppabc/text/revision/main.tex
Ref Report #1 has (I believe) been settled

@@
 
-%%%
-{\bf \color{darkgreen}
-MJV: @CHH I think this paragraph is misleading since both ABC and Gaussian likelihood covariances suffer from inaccuracy of the simulations.
-I say we remove this paragraph entirely.
-It does not add any clarify anything.
-It just causes confusion!
-}
-{\bf \color{darkgreen} 
-CHH: @MJV I'm not a big fan of this entire paragraph. The accuracy of the covariance
-matrix depends on the accuracy of the mock catalogs (a.k.a. the forward model). While it's definitely 
-true that on small scales they break down, forward models in ABC suffers from the same issues.
-} 
-%%%
-Furthermore, the Gaussian-likelihood  approach relies on constructing an accurate covariance matrix
-estimate that captures the sample variance of the data. While we are able to do this accurately 
-within the scope of the HOD framework, for more general LSS parameter inference situations, it 
-is both labor and computationally expensive and dependent on the accuracy of simulated mock catalogs,
-which are known to be unreliable on small scales (see \citealt{cosmiccode,nifty} and references therein). 
-Since ABC-PMC utilizes a forward model to account for sample variance, it does not depend on a 
-covariance matrix estimate; hence it does not face these problems.
-
-{\em ABC-PMC -- unlike the Gaussian-likelihood approach -- is agnostic about the functional form of the 
-underlying distribution of the summary 
-statistics} ($\xigg$ and $\gmf$). This may explain why we find less biased constraints from the
->>>>>>> afd43473184a6a4637881fb3504a38a867b27226
-ABC-PMC analysis compared to the Gaussian-likelihood analysis, when using $\gmf(N)$ as an observable 
-(Figures \ref{fig:hist_nbargmf} and \ref{fig:cont_nbargmf}).
-
 The ABC method of parameter inference has often been viewed as computationally infeasible due to the 

### 2017-03-08 ccppabc/text/revision/main.tex
Edits on Figure captions for 7,8, 9, 10

@@
 the ``true'' HOD parameters with vertical black dashed line. The marginalized posterior PDFs 
-}
-obtained from the two methods are consistent with each other. The ABC-PMC constraints are slightly 
-narrower for $\log \mathcal{M}_{\rm min}$, slightly wider for $\log \mathcal{M}_{1}$, and slightly 
-less biased for $\alpha$.} 
+obtained from the two methods are consistent with each other. The ABC-PMC and Gaussian pseudo-likelihood
+constraints are generally consistent for $\log \mathcal{M}_{\rm min}$ and $\log \mathcal{M}_{1}$. 
+The ABC-PMC constraint for $\alpha$ is slightly less biased and has slightly larger uncertainty
+then the constraint from Gaussian pseudo-likelihood analysis.} 
+}
 %\end{center}
@@
 \includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbargmf.pdf}
-\caption{\label{fig:hist_nbargmf} Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC analysis and the standard Gaussian 
-Likelihood MCMC analysis are done using the observables $\ngalaxy$ and $\gmf(N)$. The constraints are 
-consistent with the ``true'' HOD parameters and both methods infer the region of allowed values to similar 
-precision. The MCMC result for $\alpha$ is slightly more biased compared to ABC-PMC estimate. 
-This may stem from the fact that the use of Gaussian-likelihood and its associated assumptions is more 
+\caption{\label{fig:hist_nbargmf} 
+{\bf \color{dred}
+Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC
+analysis and the Gaussian pseudo-likelihood MCMC analysis use $\ngalaxy$ and $\gmf(N)$ as 
+observables. Both methods derive constraints consistent with the ``true'' HOD parameters 
+and infer the region of allowed values to similar precision. We note that the MCMC constraint 
+on $\alpha$ is slightly more biased compared to ABC-PMC estimate. This discrepancy may stem 
+from the fact that the use of Gaussian pseudo-likelihood and its associated assumptions is more 
 spurious when modeling the group multiplicity function.}
+}
 %\end{center}
@@
 \caption{\label{fig:cont_nbarxi}
- We compare the ABC-PMC (orange) and the Gaussian likelihood MCMC (blue)
+{\bf \color{dred}
+We compare the ABC-PMC (orange) and the Gaussian pseudo-likelihood MCMC (blue)
 predictions of the 68\% and 95\% posterior confidence regions over the HOD 
 parameters ($\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$) 
-using $\ngalaxy$ and $\xigg(r)$ as observables. The ``true'' HOD parameters used 
-to create the mock observations are plotted by black stars in each panel. The two approaches are consistent with the true values of the parameters used to generate the data.}
+using $\ngalaxy$ and $\xigg(r)$ as observables. In each panel, the black star
+represents the ``true'' HOD parameters used to generate the mock observations. 
+Both inference methods derive confidence regions consistent with the ``true'' 
+HOD parameters.
+}
+}
 %\end{center}
@@
 \includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_contour_nbargmf.pdf}
-\caption{ \label{fig:cont_nbargmf} Same as Figure \ref{fig:cont_nbarxi}, but using $\ngalaxy$ and $\gmf(N)$ as observables.
-Again, both methods accurately estimate the parameter value regions of the true values for the data.
-The MCMC estimation of $\alpha$ by use of a Gaussian-likelihood is biased when compared with the ABC-PMC contours.
-This may be due to the fact that the group multiplicity function is particularly unsuited to the use of a Gaussian-likelihood analysis.}
+\caption{ \label{fig:cont_nbargmf} 
+Same as Figure \ref{fig:cont_nbarxi}, but using $\ngalaxy$ and $\gmf(N)$ as 
+observables. Again, the confidence regions derived from both methods are 
+consistent with the ``true'' HOD parameters used to generate the mock 
+observations. The confidence region of $\alpha$ from the Gaussian 
+pseudo-likelood method is biased compared to the ABC-PMC contours. 
+This may be due to the fact that the true likelihood function that 
+describes $\gmf(N)$ deviates significantly from the assumed Gaussian 
+functional form.}
+}
 %\end{center}

### 2017-03-08 ccppabc/text/revision/main.tex
reworded the selection effect stuff! removed hierarchical models

@@
 
-{\bf \color{dred}
-It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC.
+{\bf \color{dred} It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC.
 Let us consider a dataset---from a galaxy survey---consisting of objects with different data qualities, signal-to-noise ratios, and systematic effects.
 This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift.
-Another example could be a catalog with a singal-to-noise ratio cut.
-}
-
-{\bf \color{dred}
-In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest.
-Hierarchical models can modify the likelihood functions through by taking into account the selection effects.
-Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the data that incorporates signal-to-noise ratio cuts, incompleteness, etc.
-}{\bf \color{darkgreen}
-MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ?
-}
+Another example could be a catalog with a signal-to-noise ratio cut.}  {\bf \color{dred} In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest.
+Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.
+%Hierarchical models can modify the likelihood functions through by %taking into account the selection effects. 
+Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the simulated data that incorporates signal-to-noise ratio cuts, incompleteness, systematic uncertainties etc.
+}{
+\bf \color{darkgreen} MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ? }
 %%%
@@
 functional form.}
-}
 %\end{center}

### 2017-03-08 ccppabc/text/revision/main.tex
Edited text descriptions of Figure 7 and 8. Moved a few figures around

--- 2017-03-08 ccppabc/text/revision/main.tex
@@
 $\gmf(N)$ --- in units of $(\mathrm{Mpc}/h)^{-3}$. 
-
 \end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Figure: Mock observation 2PCF and GMF
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=\textwidth]{figs/paper_data_observables.pdf}
+\caption{\label{fig:mock} The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
+function $\gmf(N)$ (right) summary statistics of the mock observations generated from 
+the ``true'' HOD parameters described in Section \ref{sec:mock_obv}. The width of the 
+shaded region corresponds to the square root of the covariance matrix diagonal 
+elements (Eq. \ref{eq:cov}). In our ABC analysis, we treat the $\xigg(r)$ 
+and $\gmf(N)$ above as the summary statistics of the observation.}
+%\end{center}
+\end{figure*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Figure: ABC Pool Evolution plot 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=0.8\textwidth]{figs/paper_ABC_poolevolution_nbargmf.pdf}
+\caption{\label{fig:pool_demo} We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
+$\bar{n}$ and $\gmf(N)$ are used as observables for the above results. For reference, 
+in each panel, we include the ``true'' HOD parameters (black star) listed in 
+Section \ref{sec:mock_obv}. The initial distance threshold, $\vec\epsilon_1 = 
+[\infty, \infty]$ at $t=1$ (top left) so the $\pars_1$ spans the entire range of the prior distribution, 
+which is also the range of the panels. We see for $t < 5$, the parameter space occupied 
+by the ABC $\pars_t$ shrinks dramatically. Eventually when the algorithm converges, 
+$t > 7$, the parameter space occupied by $\pars_t$ no longer shrinks and their 
+distributions represent the posterior distribution of the 
+parameters. At $t=9$, the final iteration, the ABC algorithm has converged and 
+we find that $\pars_\mathrm{true}$ lies safely within the $68\%$ confidence region.} 
+%\end{center}
+\end{figure*}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{ABC at work}\label{sec:abcatwork}
@@
 \ref{sec:abcvsmcmc}.
+
 %%%%%%%%%%%%%%%%%%%%% SUBSECTION %%%%%%%%%%%%%%%%%%%%%%%
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Figure: ABC Pool Convergence 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{figure*}
+%\begin{center}
+\includegraphics[width=\textwidth]{figs/paper_ABCconvergence.pdf}
+\caption{\label{fig:abc_converge} We illustrate the convergence of the ABC algorithm through the evolution of the ABC particle distribution as a function of iteration for 
+parameters $\log \mathcal{M}_\mathrm{min}$ (left), $\alpha$ (center), and 
+$\log \mathcal{M}_1$ (right).
+The top panel corresponds our ABC results using the observables $(\bar{n}, \gmf(N))$, 
+while the lower panel plots  corresponds to the ABC results using $(\bar{n}, \xigg(r))$. 
+The distributions of parameters show no significant change after $t > 7$, which suggests
+that the ABC algorithm has converged.}
+%\end{center}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%% SUBSECTION %%%%%%%%%%%%%%%%%%%%%%%
-% Mock Observations  
+% ABC-PMC Design 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 ABC-PMC and Gaussian pseudo-likelihood MCMC analyses using $[\ngalaxy, \xigg]$ and 
-$[\ngalaxy, \gmf]$ as observables respectively. The top panels in each figure 
+$[\ngalaxy, \gmf]$ as observables, respectively. The top panels in each figure 
 compares the marginalized posterior PDFs for three parameters of the HOD model: 
-$\{\log \mathcal{M}_{\rm min}, \alpha , \log\mathcal{M}_1\}$. The lower panels in each 
+$\{\log \mathcal{M}_{\rm min}, \alpha, \log\mathcal{M}_1\}$. The lower panels in each 
 figure compares the $68\%$ and $95\%$ confidence intervals of the constraints 
@@
 {\bf \color{dred}
-In Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, for each of the parameters, 
-the marginalized posteriors from both inference methods are comparable. Similarly, 
-the lower panels of the figures reveal that the constraints derived for each parameter
-from the two methods are also consistent, with minor discrepancies. Furthermore, we 
-find that both methods produce constraints consistent with the ``true'' HOD parameters.
+In both Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, the marginalized 
+posteriors for each of the parameters from both inference methods are comparable 
+and consistent with the ``true'' HOD parameters. However, we note that there are 
+minor discrepancies between the maringalized posterior distributions. In particular, 
+the distribution for $\alpha$ derived from ABC-PMC is less biased than the $\alpha$ 
+constraints from the Gaussian pseudo-likelihood approach. 
 }
 {\bf \color{darkgreen}
-CHH: A few sentences discussing the width of the distributions for each of the panels
-have been commented out. Not sure if it's meaningful to include such slight discrepancies. 
 @MJV: Could you take a look at what's written above and see if you feel that it's sufficient? 
@@
 %%%%%%%%%%%%%%%%%%%% FIGURES %%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Mock observation 2PCF and GMF
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-%\begin{center}
-\includegraphics[width=\textwidth]{figs/paper_data_observables.pdf}
-\caption{\label{fig:mock} The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
-function $\gmf(N)$ (right) summary statistics of the mock observations generated from 
-the ``true'' HOD parameters described in Section \ref{sec:mock_obv}. The width of the 
-shaded region corresponds to the square root of the covariance matrix diagonal 
-elements (Eq. \ref{eq:cov}). In our ABC analysis, we treat the $\xigg(r)$ 
-and $\gmf(N)$ above as the summary statistics of the observation.}
-%\end{center}
-\end{figure*}
-%%%%%%%%%%%%%%%%%% End of FIGURE %%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABC Pool Evolution plot 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-%\begin{center}
-\includegraphics[width=0.95\textwidth]{figs/paper_ABC_poolevolution_nbargmf.pdf}
-\caption{\label{fig:pool_demo} We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
-$\bar{n}$ and $\gmf(N)$ are used as observables for the above results. For reference, 
-in each panel, we include the ``true'' HOD parameters (black star) listed in 
-Section \ref{sec:mock_obv}. The initial distance threshold, $\vec\epsilon_1 = 
-[\infty, \infty]$ at $t=1$ (top left) so the $\pars_1$ spans the entire range of the prior distribution, 
-which is also the range of the panels. We see for $t < 5$, the parameter space occupied 
-by the ABC $\pars_t$ shrinks dramatically. Eventually when the algorithm converges, 
-$t > 7$, the parameter space occupied by $\pars_t$ no longer shrinks and their 
-distributions represent the posterior distribution of the 
-parameters. At $t=9$, the final iteration, the ABC algorithm has converged and 
-we find that $\pars_\mathrm{true}$ lies safely within the $68\%$ confidence region.} 
-%\end{center}
-\end{figure*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% ABC Pool Convergence 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{figure*}
-%\begin{center}
-\includegraphics[width=\textwidth]{figs/paper_ABCconvergence.pdf}
-\caption{\label{fig:abc_converge} We illustrate the convergence of the ABC algorithm through the evolution of the ABC particle distribution as a function of iteration for 
-parameters $\log \mathcal{M}_\mathrm{min}$ (left), $\alpha$ (center), and 
-$\log \mathcal{M}_1$ (right).
-The top panel corresponds our ABC results using the observables $(\bar{n}, \gmf(N))$, 
-while the lower panel plots  corresponds to the ABC results using $(\bar{n}, \xigg(r))$. 
-The distributions of parameters show no significant change after $t > 7$, which suggests
-that the ABC algorithm has converged.}
-%\end{center}
-\end{figure*}
 
@@
 \includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_contour_nbargmf.pdf}
-\caption{ \label{fig:cont_nbargmf} 
+\caption{\label{fig:cont_nbargmf} 
+{\bf \color{dred}
 Same as Figure \ref{fig:cont_nbarxi}, but using $\ngalaxy$ and $\gmf(N)$ as 
@@
 %%%%%%%%%%%%%%%%%%%% REFERENCES %%%%%%%%%%%%%%%%%%
-
-\bibliographystyle{yahapj}
-\bibliography{ccppabc}
 %\clearpage
 % The best way to enter references is to use BibTeX:
-
-%\bibliographystyle{mnras}
-%\bibliography{example} % if your bibtex file is called example.bib
-
-% Alternatively you could enter them by hand, like this:
-% This method is tedious and prone to error if you have lots of references
-%\begin{thebibliography}{99}
-%\bibitem[\protect\citeauthoryear{Author}{2012}]{Author2012}
-%Author A.~N., 2013, Journal of Improbable Astronomy, 1, 1
-%\bibitem[\protect\citeauthoryear{Others}{2013}]{Others2013}
-%Others S., 2012, Journal of Interesting Stuff, 17, 198
-%\end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%% APPENDICES %%%%%%%%%%%%%%%%%%%%%
-
-%\appendix
-
-%\section{Some extra material}
-
-%If you want to present additional material which would interrupt the flow of the main paper,
-%it can be placed in an Appendix which appears after the list of references.
-
+\bibliographystyle{mnras}%yahapj}
+\bibliography{ccppabc}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \end{document}
-
-% End of mnras_template.tex

### 2017-03-08 ccppabc/text/revision/main.tex
slightly reworded the first parafgraph of the abstract

@@
 \begin{abstract}
-The standard approaches to Bayesian parameter inference in large scale structure (LSS) assume a Gaussian functional form (chi-squared form) for the likelihood.
-They are also typically restricted to measurements
-such as the two point correlation function. 
-Likelihood free inferences such as Approximate Bayesian Computation (ABC) make inference possible without 
-assuming any functional form for the likelihood, thereby relaxing the assumptions and restrictions of 
-the standard approach. Instead it relies on a forward generative model of the data and a metric
-for measuring the distance between the model and data. In this work, we 
-demonstrate that ABC is feasible for LSS parameter inference by using it to constrain parameters of the 
-halo occupation distribution (HOD) model for populating dark matter halos with galaxies.
+{\bf\color{darkgreen} MJV: @CHH: Since the referee asked us to emphasize the advantage of ABC in the abstract with a few phrases, I slightly reworded the first paragraph!
+Could you take a look at this?}
+
+{\bf \color{dred}The standard approaches to Bayesian parameter inference in large scale structure (LSS) rely on a Gaussian functional form (chi-squared form) for the likelihood.They are typically restricted to measurements such as the two point correlation function.
+Furthermore, they may not correctly predict the effects of uncertainties on the form of the likelihood and parameter constraints.}
+
+{\bf \color{dred} Likelihood free inferences such as Approximate Bayesian Computation (ABC) make inference possible without making any likelihood assumption, thereby relaxing the assumptions and restrictions of the standard approach.} Instead it relies on a forward generative model of the data and a metric for measuring the distance between the model and data.
+In this work, we demonstrate that ABC is feasible for LSS parameter inference by using it to constrain parameters of the halo occupation distribution (HOD) model for populating dark matter halos with galaxies.
 
@@
 describes $\gmf(N)$ deviates significantly from the assumed Gaussian 
-functional form.}
+functional form.}}
 %\end{center}

### 2017-03-08 ccppabc/text/revision/main.tex
More edits to Section 3.4. This time directed towards Ref Report Comment 2 (issue #15)

--- 2017-03-08 ccppabc/text/revision/main.tex
@@
 
+{\bf \color{dred}
 Furthermore, {\em ABC-PMC -- unlike the Gaussian pseudo-likelihood approach -- 
 is agnostic about the functional form of the underlying distribution of the 
-summary statistics} 
-($\xigg$ and $\gmf$). 
+summary statistics} (\emph{e.g.} $\xigg$ and $\gmf$). 
+}
 {\bf \color{darkgreen}
@@
 assumption is incorrect and especially incorrect for the GMF observable case.
-Also, for us to say the following sentence, we need to talk about it more explicitly 
-when discussing Figures 7,8,9, and 10. 
+}
+{\bf \color{dred}
+In the case where the observable $\gmf(N)$ is used as a summary statistic
+of a galaxy catalog, assuming a Gaussian functional form for the likelihood
+would clearly misreprenset the true likelihood function. In fact, this 
+discrepancy may explain the less biased constraints from the ABC-PMC analysis 
+compared to the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
 }
@@
 functional form.}
+}
 %\end{center}

### 2017-03-08 ccppabc/text/revision/main.tex

--- 2017-03-08 ccppabc/text/revision/main.tex
@@
 describes $\gmf(N)$ deviates significantly from the assumed Gaussian 
-functional form.}}
+functional form.}
+}
 %\end{center}

### 2017-03-08 ccppabc/text/revision/main.tex
more section 3.4 edits
Added a sentence to explain advantage of ABC when there are systematic uncertainties!
Some minor corrections including separating the ***extreme data situation*** paragraph into two since it was long!

@@
 }
+{\bf \color{darkgreen}
+@CHH: This looks sufficient!
+}
 %The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ are slightly larger for our ABC-PMC results.
@@
 observations. 
-}
+} {\bf \color{darkgreen} MJV: Added this sentence to emphasize why ABC is advantageous when systematic are present.
+@CHH: Can you take a look at this?} {\bf \color{dred} The systematic uncertainties may affect the likelihood in an unknown way, but their impact can be simulated and marginalized out in the forward generative model of the ABC-PMC method.}
 %In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision.
@@
 would clearly misreprenset the true likelihood function. In fact, this 
-discrepancy may explain the less biased constraints from the ABC-PMC analysis 
+discrepancy may explain the less biased constraints on $\alpha$ from the ABC-PMC analysis 
 compared to the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
-}
+}{\bf \color{darkgreen} @CHH : added $\alpha$ to the previous sentence since the other constraints our consistent!}
 
@@
 This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift.
-Another example could be a catalog with a signal-to-noise ratio cut.}  {\bf \color{dred} In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest.
+Another example could be a catalog with a signal-to-noise ratio cut.}
+
+{\bf \color{dred} In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest.
 Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.

### 2017-03-09 ccppabc/text/revision/main.tex
edits to Section 3.4; see Issue #21

@@
 }
-{\bf \color{darkgreen}
-@MJV: Could you take a look at what's written above and see if you feel that it's sufficient? 
-}
-{\bf \color{darkgreen}
-@CHH: This looks sufficient!
-}
 %The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ are slightly larger for our ABC-PMC results.
@@
 matrix estimates rather than the {\em true} covariance matrix leads to a likelihood 
-function that is {\em no longer} Gaussian.
-ABC-PMC does not depend on a covariance matrix estimate; hence, it does not face 
-these problems. In principle, forward models of the ABC-PMC method also have the 
-advantage that they can account for sources of systematic uncertainties that affect 
-observations. 
-} {\bf \color{darkgreen} MJV: Added this sentence to emphasize why ABC is advantageous when systematic are present.
-@CHH: Can you take a look at this?} {\bf \color{dred} The systematic uncertainties may affect the likelihood in an unknown way, but their impact can be simulated and marginalized out in the forward generative model of the ABC-PMC method.}
+function that is {\em no longer} Gaussian. ABC-PMC does not depend on a covariance 
+matrix estimate; hence, it does not face these problems.
+}
+
+{\bf \color{dred}
+In addition to not requiring accurate covariance matrix estimates, forward models 
+of the ABC-PMC method, in principle, also have the advantage that they can account 
+for sources of systematic uncertainties that affect observations. All observations 
+suffer from significant systematic effects which are often difficult to correct. 
+For instance, in SDSS-III BOSS~\citep{Dawson:2013a}, fiber collisions and redshift
+failures siginifcantly bias measurements and analysis of observables such as 
+$\xigg$ or the galaxy powerspectrum~\citep{Ross:2012aa, Guo:2012a, Hahn:2017a}. 
+In parameter inference, these systematics can affect the likelihood, and thus any 
+analysis that requires writing down the likelihood, in unknown ways. With a forward generative model of the ABC-PMC 
+method, the systematics can be simulated and marginalized out to achieve unbiased 
+constraints. 
+} 
+{\bf \color{darkgreen} 
+CHH: @MJV I incorporated the sentence into the paragraph. Can you check it out
+and confirm it's fine?
+} 
 %In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision.

### 2017-03-09 ccppabc/text/revision/main.tex
Significant edits to @MJV's addition to address Ref Report comment 2,3 Issues #15 and #17; *edits are incomplete

@@
 CHH: Need a sentence here that reiterates that the Gaussian functional form 
-assumption is incorrect and especially incorrect for the GMF observable case.
+assumption is incorrect for 2PCF.
 }
@@
 of a galaxy catalog, assuming a Gaussian functional form for the likelihood
-would clearly misreprenset the true likelihood function. In fact, this 
-discrepancy may explain the less biased constraints on $\alpha$ from the ABC-PMC analysis 
-compared to the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
-}{\bf \color{darkgreen} @CHH : added $\alpha$ to the previous sentence since the other constraints our consistent!}
-
-{\bf \color{dred} It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC.
+clearly misrepresents the true likelihood function. This incorrect likelihood, 
+may explain why the constraints on $\alpha$ are less biased for the ABC-PMC analysis 
+than the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
+}
+
+{\bf \color{dred} 
+Although in our comparison using simple mock observations, we find generally 
+consistent parameter constraints from both the ABC-PMC analysis and the standard 
+Gaussian pseudo-likelihood analysis, more realistic scenarios face mmany factors
+that can generate inconsistencies. Consider a typical galaxy catalog from 
+observations for LSS. These catalogs consist of objects with different data 
+qualities, signal-to-noise ratios, and systematic effects. For example, catalogs 
+are often incomplete beyond some luminosity/redshift or have some threshold 
+signal-to-noise ratio cut imposed on them.
+}
+
+%It is possible to consider a scenario in which the results from the standard analysis are inconsistent with the results from the ABC.
 Let us consider a dataset---from a galaxy survey---consisting of objects with different data qualities, signal-to-noise ratios, and systematic effects.
 This could be a catalog of galaxies which is incomplete beyond a certain luminosity or redshift.
-Another example could be a catalog with a signal-to-noise ratio cut.}
-
-{\bf \color{dred} In standard likelihood analyses, these selection effects, if not accounted for, can lead to biases in the inferred parameters of interest.
-Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.
-%Hierarchical models can modify the likelihood functions through by %taking into account the selection effects. 
-Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the simulated data that incorporates signal-to-noise ratio cuts, incompleteness, systematic uncertainties etc.
+Another example could be a catalog with a signal-to-noise ratio cut.
+{\bf \color{dred} 
+These selection effects, coupled with the systematic effects mentioned above, 
+make correctly predicting the functional form of the likelihood difficult. 
+In the standard Gaussian pseudo-likelihood analysis, and other analysis that
+depend on writing down a likelihood function, these effects can siginficantly 
+bias the inferred parameters of interest. In these situations, employing
+ABC equipped with a generative forward model that incorproates selection and 
+systematic effects may produce less biased parameter constraints. 
+}
+%Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.
+%Alternatively, one could alleviate this issue by employing ABC equipped with a generative forward model of the simulated data that incorporates signal-to-noise ratio cuts, incompleteness, systematic uncertainties etc.
 }{
-\bf \color{darkgreen} MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ? }
+%\bf \color{darkgreen} MJV: @CHH Can you read the two paragraphs above and see if they address referee's concern about the extreme data situation that could lead to discrepancy ? }
 %%%
-
 %In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood 

### 2017-03-09 ccppabc/text/revision/main.tex
continuing edits on Section 3.4

@@
 consistent parameter constraints from both the ABC-PMC analysis and the standard 
-Gaussian pseudo-likelihood analysis, more realistic scenarios face mmany factors
+Gaussian pseudo-likelihood analysis, more realistic scenarios present many factors
 that can generate inconsistencies. Consider a typical galaxy catalog from 
-observations for LSS. These catalogs consist of objects with different data 
+LSS observations. These catalogs consist of objects with different data 
 qualities, signal-to-noise ratios, and systematic effects. For example, catalogs 
@@
 {\bf \color{dred} 
-These selection effects, coupled with the systematic effects mentioned above, 
-make correctly predicting the functional form of the likelihood difficult. 
+These selection effects, coupled with the systematic effects earlier this section, 
+make correctly predicting the functional form of the likelihood intractable.
 In the standard Gaussian pseudo-likelihood analysis, and other analysis that
-depend on writing down a likelihood function, these effects can siginficantly 
-bias the inferred parameters of interest. In these situations, employing
+require writing down a likelihood function, these effects can significantly 
+bias the inferred parameter constraints. In these situations, employing
 ABC equipped with a generative forward model that incorproates selection and 

### 2017-03-09 ccppabc/text/revision/main.tex
Edits to address Ref Report comment 2, 3 (issues #15 and #17); @mjvakili see issue #22

@@
 is agnostic about the functional form of the underlying distribution of the 
-summary statistics} (\emph{e.g.} $\xigg$ and $\gmf$). 
+summary statistics} (\emph{e.g.} $\xigg$ and $\gmf$). As we explain throughout 
+the paper, the likelihood function in LSS {\em cannot} be Gaussian. For $\xigg$, 
+the correlation function must satisfy non-trivial positive-definiteness requirements 
+and hence the Gaussian pseudo-likelihood function assumption is not correct 
+in detail. In the case $\gmf(N)$ is used as a summary statistic
+of a galaxy catalog, assuming a Gaussian functional form for the likelihood
+clearly misrepresents the true likelihood function. In fact, this incorrect 
+likelihood, may explain why the constraints on $\alpha$ are less biased for 
+the ABC-PMC analysis than the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
 }
 {\bf \color{darkgreen}
-CHH: Need a sentence here that reiterates that the Gaussian functional form 
-assumption is incorrect for 2PCF.
-}
-{\bf \color{dred}
-In the case where the observable $\gmf(N)$ is used as a summary statistic
-of a galaxy catalog, assuming a Gaussian functional form for the likelihood
-clearly misrepresents the true likelihood function. This incorrect likelihood, 
-may explain why the constraints on $\alpha$ are less biased for the ABC-PMC analysis 
-than the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
+CHH: @MJV, I parrot a lot of arguments made in the intro without providing much extra 
+'meat'. Let me know what you think. 
 }

### 2017-03-09 ccppabc/text/revision/main.tex
outlines for what needs to be added at the end of Section 3.4. in order to address Ref Report comment #4 (issue #18)

@@
 inference in large scale structure studies with a number of advantages we describe above. 
+{\bf \color{darkgreen}
+Need discussion of applying ABC to real data. 
+\begin{itemize}
+\item Computational tractability caveats. The forward models in real data scenarios would have to be much bigger. 
+\item They would also have to be better than the ones that are currently used for covariance matrix estimation. 
+\item Complexity of systematic effects (mention DESI) 
+\item End with how there's still hope because there's developments in fast simulations
+\end{itemize}
+}
 

### 2017-03-10 ccppabc/text/revision/main.tex
Tried to address Ref Report comment 4 (issue #18); edited the abstract a bit

--- 2017-03-09 ccppabc/text/revision/main.tex
@@
 \begin{abstract}
-{\bf\color{darkgreen} MJV: @CHH: Since the referee asked us to emphasize the advantage of ABC in the abstract with a few phrases, I slightly reworded the first paragraph!
-Could you take a look at this?}
-
-{\bf \color{dred}The standard approaches to Bayesian parameter inference in large scale structure (LSS) rely on a Gaussian functional form (chi-squared form) for the likelihood.They are typically restricted to measurements such as the two point correlation function.
-Furthermore, they may not correctly predict the effects of uncertainties on the form of the likelihood and parameter constraints.}
-
-{\bf \color{dred} Likelihood free inferences such as Approximate Bayesian Computation (ABC) make inference possible without making any likelihood assumption, thereby relaxing the assumptions and restrictions of the standard approach.} Instead it relies on a forward generative model of the data and a metric for measuring the distance between the model and data.
-In this work, we demonstrate that ABC is feasible for LSS parameter inference by using it to constrain parameters of the halo occupation distribution (HOD) model for populating dark matter halos with galaxies.
+{\bf \color{dred}
+Standard approaches to Bayesian parameter inference in large scale 
+structure assume a Gaussian functional form (chi-squared form) 
+for the likelihood. They are also limited to analyzing measurements of   
+the two-point correlation function. These restrictions, in detail, cannot 
+be correct. Likelihood free inferences such as Approximate Bayesian Computation (ABC) 
+relax these restrictions and make inference possible without making any 
+assumptions on the likelihood.
+} 
+Instead it relies on a forward generative model of the data and a metric 
+for measuring the distance between the model and data. In this work, we 
+demonstrate that ABC is feasible for LSS parameter inference by using it 
+to constrain parameters of the halo occupation distribution (HOD) model 
+for populating dark matter halos with galaxies.
 
@@
 
-The ABC method of parameter inference has often been viewed as computationally infeasible due to the 
-fact that it necessitates a generative forward model for a large number of realizations. However, 
-by combining ABC with the PMC sampling method,
-%of importance sampling {\bf [APH: the term "importance sampling" has not been defined]}, 
-we have a method that efficiently converges 
-to give us reliable posteriors of the model parameters. Furthermore, in our analysis, the computational resources 
-required for the ABC-PMC inference were comparable to those used in achieving convergence in the 
-MCMC Gaussian-Likelihood inference. This is excluding the typically computationally intensive step of 
-calculating a covariance matrix estimate. Therefore, we find that ABC-PMC offers a method for parameter
-inference in large scale structure studies with a number of advantages we describe above.
+{\bf \color{dred}
+Despite the advantages of ABC, one obstactle for adopting it to parameter 
+inference has been the computational costs of generative forward models, a 
+key element of ABC. By combining ABC with the PMC sampling method, however, 
+ABC-PMC efficiently converges to give reliable posterior parameter constraints. 
+In fact, in our analysis, the total computational resources required for the 
+ABC-PMC analysis were {\em comparable} to the computational resources used 
+for the Gaussian pseudo-likelihood analysis with MCMC sampling.
+}
+
+{\bf \color{dred}
+Applying ABC-PMC beyond the analysis in this work, to broader LSS analyses 
+imposes some caveats. In this work, we focus on the galaxy-halo connection, so
+our generative forward model populates halos with galaxies. LSS analyses 
+for inferring cosmological parameters would require generating halos by 
+running cosmological simulations. The forward models also need to accurately 
+model the observation systematic effects of the latest observations. Hence, 
+accurate generative forward models in LSS analyses demand improvements in simulations 
+and significant computational resources in order to infer unbiased parameter 
+constraints. Recent cosmology simulations show promising improvements 
+in both accuracy and speed~\citep[\emph{e.g.}][]{fastpm}. Such developements
+will be crucial for applying ABC-PMC to broader LSS analyses and exploiting 
+the significant advantages that ABC-PMC offers.
+}
+
+%the ABC method of parameter inference has often been viewed as computationally infeasible due to the fact that it necessitates a generative forward model for a large number of realizations.
+However, by combining ABC with the PMC sampling method, we have a method that efficiently converges to give us reliable posteriors of the model parameters.
+Furthermore, in our analysis, the computational resources required for the ABC-PMC inference were comparable to those used in achieving convergence in the MCMC Gaussian-Likelihood inference.
+This is excluding the typically computationally intensive step of calculating a covariance matrix estimate.
+Therefore, we find that ABC-PMC offers a method for parameter inference in large scale structure studies with a number of advantages we describe above.
+%{\bf \color{darkgreen}
+%\begin{itemize}
+%\item distance metric 
+%\item End with how there's still hope because there's developments in fast simulations
+%\end{itemize}
+%}
 

### 2017-03-13 ccppabc/text/revision/main.tex
2PCF restriction sentence removed from abstract
Issue #17 (Ref Report comment 3) dealt with

@@
 structure assume a Gaussian functional form (chi-squared form) 
-for the likelihood. They are also limited to analyzing measurements of   
-the two-point correlation function. These restrictions, in detail, cannot 
+for the likelihood. This assumption, in detail, cannot %They are also limited to analyzing measurements of the two-point correlation function. 
 be correct. Likelihood free inferences such as Approximate Bayesian Computation (ABC) 
@@
 These selection effects, coupled with the systematic effects earlier this section, 
-make correctly predicting the functional form of the likelihood intractable.
-In the standard Gaussian pseudo-likelihood analysis, and other analysis that
-require writing down a likelihood function, these effects can significantly 
+make correctly predicting the likelihood intractable. In the standard Gaussian 
+pseudo-likelihood analysis, and other analysis that require writing down a 
+likelihood function, these effects can significantly 
 bias the inferred parameter constraints. In these situations, employing

### 2017-03-13 ccppabc/text/revision/main.tex
Issue #22 addressed; darkgreen text commented out.

@@
 } 
-{\bf \color{darkgreen} 
-CHH: @MJV I incorporated the sentence into the paragraph. Can you check it out
-and confirm it's fine?
-} 
+%{\bf \color{darkgreen} CHH: @MJV I incorporated the sentence into the paragraph. Can you check it out and confirm it's fine?  } 
 %In general, we find that the constraints from ABC-PMC are consistent with those from the Gaussian-likelihood MCMC method both in terms of accuracy and precision.
@@
 and hence the Gaussian pseudo-likelihood function assumption is not correct 
-in detail. In the case $\gmf(N)$ is used as a summary statistic
-of a galaxy catalog, assuming a Gaussian functional form for the likelihood
-clearly misrepresents the true likelihood function. In fact, this incorrect 
+in detail. In the case of $\gmf(N)$, assuming a Gaussian functional form for 
+the likelihood, which in reality is more likely Poisson, misrepresents the true 
+likelihood function. In fact, this incorrect 
 likelihood, may explain why the constraints on $\alpha$ are less biased for 
@@
 }
-{\bf \color{darkgreen}
-CHH: @MJV, I parrot a lot of arguments made in the intro without providing much extra 
-'meat'. Let me know what you think. 
-}
+%{\bf \color{darkgreen} CHH: @MJV, I parrot a lot of arguments made in the intro without providing much extra 'meat'. Let me know what you think.  }
 

### 2017-03-14 ccppabc/text/revision/main.tex
very minor change in the abstract
modified data & address

@@
 $^{3}$Flatiron institute, 160 Fifth Avenue, New York, NY\\
-$^{4}$Center for Data Science, New York University, 726 Broadway, New York, NY\\
+$^{4}$Center for Data Science, New York University, 60 Fifth Ave, New York, NY 10011\\
 $^{5}$Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17, D-69117 Heidelberg, Germany\\
@@
 % Enter the current year, for the copyright statements etc.
-\pubyear{2015}
+\pubyear{2017}
 
@@
 assumptions on the likelihood.
-} 
-Instead it relies on a forward generative model of the data and a metric 
+} Instead ABC relies on a forward generative model of the data and a metric 
 for measuring the distance between the model and data. In this work, we 

### 2017-03-14 ccppabc/text/revision/clean_manuscript.tex
clean manuscript

--- 2017-03-14 ccppabc/text/revision/main.tex
@@
 $*$ These authors have contributed equally to the paper\\ 
-$^{1}$Center for Cosmology and Particle Physics,Department of Physics, New York University, 726 Broadway, New York, NY\\
+$^{1}$Center for Cosmology and Particle Physics, Department of Physics, New York University, 726 Broadway, New York, NY\\
 $^{2}$Yale Center for Astronomy \& Astrophysics, Yale University, New Haven, CT\\
@@
 \begin{abstract}
-{\bf \color{dred}
+
 Standard approaches to Bayesian parameter inference in large scale 
@@
 relax these restrictions and make inference possible without making any 
-assumptions on the likelihood.
-} Instead ABC relies on a forward generative model of the data and a metric 
+assumptions on the likelihood. Instead ABC relies on a forward generative model of the data and a metric 
 for measuring the distance between the model and data. In this work, we 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{Comparison to the Gaussian {\color{dred} Pseudo - Likelihood} MCMC Analysis} \label{sec:abcvsmcmc}
+\subsection{Comparison to the Gaussian Pseudo-Likelihood MCMC Analysis} \label{sec:abcvsmcmc}
 In order to assess the quality of the parameter inference described in the previous section, 
-we compare the ABC-PMC results with the HOD parameter constraints 
-{\bf \color{dred} from}
+we compare the ABC-PMC results with the HOD parameter constraints from
 assuming a Gaussian likelihood 
@@
 In the ABC analysis, the model accounts for sample variance by randomly sampling a subvolume to be 
-populated with galaxies. 
-{\bf \color{dred} %%%
-Instead, in the Gaussian pseudo-likelihood analysis, the covariance matrix is assumed to capture the
+populated with galaxies. Instead, in the Gaussian pseudo-likelihood analysis, the covariance matrix is assumed to capture the
 uncertainties from sample variance. Hence, in the model for the Gaussian pseudo-likelihood analysis, 
@@
 We describe the Gaussian pseudo-likelihood analysis below.
-}
+
 %The Gaussian likelihood analysis assumes a covariance matrix that captures the uncertainties from the sample variance, and therefore in the model, we populate the halos of the entire $\mathtt{MultiDark}$ simulation rather than a subvolume.
 
-{\bf \color{dred}
 To write down the Gaussian pseudo-likelihood, we first introduce the vector $\mathbf{x}$: 
@@
 given the prior distribution using the MCMC sampler $\mathtt{emcee}$ (\citealt{emcee}).
-}
-
-{\bf \color{dred}
+
 In Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, we compare the results from
@@
 are marked by vertical dashed lines in each panel.
-}
-
-{\bf \color{dred}
+
 In both Figures~\ref{fig:hist_nbarxi} and \ref{fig:hist_nbargmf}, the marginalized 
@@
 constraints from the Gaussian pseudo-likelihood approach.
-}
+
 %The uncertainty over the parameter $\log \mathcal{M}_\mathrm{min}$ is slightly smaller while the uncertainties of the parameters $\alpha$ and $\log M_{1}$ are slightly larger for our ABC-PMC results.
@@
 
-{\bf \color{dred}
 In Figures \ref{fig:cont_nbarxi} and \ref{fig:cont_nbargmf}, we plot the contours 
@@
 that the contours for the ABC-PMC method are more extended along $\alpha$.
-}
 
@@
 %%%      the advantages of ABC over MCMC
-{\bf \color{dred} 
+
 Overall, the HOD parameter constraints from ABC-PMC are consistent with those from 
@@
 capture the sample variance of the data.
-}
-
-{\bf \color{dred} 
+
 Accurate estimation of the covariance matrix 
@@
 matrix estimate; hence, it does not face these problems.
-}
-
-{\bf \color{dred}
+
 In addition to not requiring accurate covariance matrix estimates, forward models 
@@
 constraints.
-} 
+
 %{\bf \color{darkgreen} CHH: @MJV I incorporated the sentence into the paragraph. Can you check it out and confirm it's fine?  } 
@@
 
-{\bf \color{dred}
 Furthermore, {\em ABC-PMC -- unlike the Gaussian pseudo-likelihood approach -- 
@@
 the ABC-PMC analysis than the Gaussian-likelihood analysis in \ref{fig:cont_nbargmf}.
-}
+
 %{\bf \color{darkgreen} CHH: @MJV, I parrot a lot of arguments made in the intro without providing much extra 'meat'. Let me know what you think.  }
 
-{\bf \color{dred} 
 Although in our comparison using simple mock observations, we find generally 
@@
 signal-to-noise ratio cut imposed on them.
-}
 
@@
 Another example could be a catalog with a signal-to-noise ratio cut.
-{\bf \color{dred} 
+
 These selection effects, coupled with the systematic effects earlier this section, 
@@
 systematic effects may produce less biased parameter constraints.
-}
+
 %Methods that rely on a likelihood function may not correctly predict how the functional form of the likelihood is impacted by the presence of possible selection effects and systematic uncertainties.
@@
 
-{\bf \color{dred}
 Despite the advantages of ABC, one obstactle for adopting it to parameter 
@@
 for the Gaussian pseudo-likelihood analysis with MCMC sampling.
-}
-
-{\bf \color{dred}
+
 Applying ABC-PMC beyond the analysis in this work, to broader LSS analyses 
@@
 the significant advantages that ABC-PMC offers.
-}
 
@@
 \caption{\label{fig:hist_nbarxi} 
-{\bf \color{dred}
 We compare the $\log \mathcal{M}_{\rm min}$, $\alpha$, and $\log \mathcal{M}_{1}$ parameter 
@@
 then the constraint from Gaussian pseudo-likelihood analysis.} 
-}
 %\end{center}
@@
 \caption{\label{fig:hist_nbargmf} 
-{\bf \color{dred}
 Same as Figure \ref{fig:hist_nbarxi}, but both the ABC-PMC
@@
 spurious when modeling the group multiplicity function.}
-}
 %\end{center}
@@
 \caption{\label{fig:cont_nbarxi}
-{\bf \color{dred}
 We compare the ABC-PMC (orange) and the Gaussian pseudo-likelihood MCMC (blue)
@@
 Both inference methods derive confidence regions consistent with the ``true'' 
-HOD parameters.
-}
-}
+HOD parameters.}
 %\end{center}
@@
 \caption{\label{fig:cont_nbargmf} 
-{\bf \color{dred}
 Same as Figure \ref{fig:cont_nbarxi}, but using $\ngalaxy$ and $\gmf(N)$ as 

### 2017-04-10 ccppabc/text/revision/clean_manuscript.tex
some minor changes made for arxiv repacement submission

@@
 %\begin{center}
-\includegraphics[width=\textwidth]{figs/paper_data_observables.pdf}
+\includegraphics[width=\textwidth]{paper_data_observables.pdf}
 \caption{\label{fig:mock} The two-point correlation function $\xigg(r)$ (left) and group multiplicity 
@@
 %\begin{center}
-\includegraphics[width=0.8\textwidth]{figs/paper_ABC_poolevolution_nbargmf.pdf}
+\includegraphics[width=0.8\textwidth]{paper_ABC_poolevolution_nbargmf.pdf}
 \caption{\label{fig:pool_demo} We demonstrate the evolution of the ABC particles, $\pars_t$, over iterations $t = 1$ to $9$ in the $\log \mathcal{M}_{min}$ and $\log \mathcal{M}_1$ parameter space. 
@@
 %\begin{center}
-\includegraphics[width=\textwidth]{figs/paper_ABCconvergence.pdf}
+\includegraphics[width=\textwidth]{paper_ABCconvergence.pdf}
 \caption{\label{fig:abc_converge} We illustrate the convergence of the ABC algorithm through the evolution of the ABC particle distribution as a function of iteration for 
@@
 %\begin{center}
-\includegraphics[width=0.85\textwidth]{figs/paper_ABCcorner_nbarxi.pdf}
+\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbarxi.pdf}
 \caption{\label{fig:abc_corner_nbarxi} We present the constraints on the \citet{zheng07} HOD model parameters obtained from our ABC-PMC analysis using $\bar{n}$ and $\xigg(r)$ as observables.  
@@
 %\begin{center}
-\includegraphics[width=0.85\textwidth]{figs/paper_ABCcorner_nbargmf.pdf}
+\includegraphics[width=0.85\textwidth]{paper_ABCcorner_nbargmf.pdf}
 \caption{\label{fig:abc_corner_nbargmf}Same as Figure \ref{fig:abc_corner_nbarxi} but for our ABC analysis using $\bar{n}$ and $\gmf(N)$ as observables. The ABC parameter constraints we obtain are
@@
 %\begin{center}
-\includegraphics[width=\textwidth]{figs/paper_ABCposterior.pdf}
+\includegraphics[width=\textwidth]{paper_ABCposterior.pdf}
 \caption{\label{fig:abc_moneyplot} We compare the ABC-PMC posterior prediction for the observables $\xigg(r)$ (left) and $\gmf(N)$ (right) (orange; Section \ref{sec:abc_results}) 
@@
 %\begin{center}
-\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbarxi.pdf}
+\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbarxi.pdf}
 \caption{\label{fig:hist_nbarxi} 
@@
 %\begin{center}
-\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_nbargmf.pdf}
+\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_nbargmf.pdf}
 \caption{\label{fig:hist_nbargmf} 
@@
 %\begin{center}
-\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_contour_nbarxi.pdf}
+\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbarxi.pdf}
 \caption{\label{fig:cont_nbarxi}
@@
 %\begin{center}
-\includegraphics[width=1.\textwidth]{figs/paper_ABCvsMCMC_contour_nbargmf.pdf}
+\includegraphics[width=1.\textwidth]{paper_ABCvsMCMC_contour_nbargmf.pdf}
 \caption{\label{fig:cont_nbargmf} 
